---
title: "CP_04_differential_abundance_Summary_Results - Upset plots"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
   
   
# 4.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(metamisc) #for phyloseq_sep_variable (and more?)
packageVersion("metamisc")
library(DESeq2)
packageVersion("DESeq2")
library(zinbwave)
packageVersion("zinbwave")
library(UpSetR)
packageVersion("UpSetR")
library(tuple)
packageVersion("tuple")
library(ggpubr)
packageVersion("ggpubr")
library(stringr)
packageVersion("stringr")
library(pheatmap)
packageVersion("pheatmap")
library(ggVennDiagram)
packageVersion("ggVennDiagram")
```


# 4.4 Upset ASVs detected by each DA test
## 4.4.1 Combining DA tests per accession at ASV level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Acc_Bac.RData")

## From each data frame, extract the ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASVs <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_ASVs <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))

## use upset function. make sure to have names of the objects inside the lists!
Acc_Bac_upset_6DAtests <- mapply(function(s,t,w,x,y,z){
  input<-list(s,t,w,x,y,z)
  names(input)<- c("Std_DESeq_LRT",
                   "Std_DESeq_Wald", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output <- upset(fromList(input), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
return(output)
  },
  s = Acc_Bac_DESeqLRT_ASVs,
  t = Acc_Bac_DESeqWald_ASVs,
  w = Acc_Bac_zinbLRT_ASVs,
  x = Acc_Bac_zinbWald_ASVs,
  y = Acc_Bac_ancomWZ_ASVs,
  z = Acc_Bac_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

Acc_Bac_upset_6DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 4.4.2 Combining DA tests per domestication level at ASV level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_Dom_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_Dom_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbWald_Dom_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbLRT_Dom_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_Dom_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Dom_Bac.RData")

## From each data frame, extract the ASV IDs caught by each test
Dom_Bac_DESeqLRT_ASVs <- lapply(sigtab_stddds2_LRT_Dom_Bac, function(x) rownames(x))
Dom_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_Dom_Bac, function(x) rownames(x))
Dom_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_Dom_Bac, function(x) rownames(x))
Dom_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_Dom_Bac, function(x) rownames(x))
Dom_Bac_ancomWZ_ASVs <- lapply(fdr_ancomWZ_Dom_Bac, function(x) as.vector(x[,'Species']))
Dom_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_Dom_Bac, function(x) as.vector(x[,'Species']))

## use upset function. make sure to have names of the objects inside the lists!
Dom_Bac_upset_6DAtests <- mapply(function(s,t,w,x,y,z){
  input<-list(s,t,w,x,y,z)
  names(input)<- c("Std_DESeq_LRT",
                   "Std_DESeq_Wald", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output <- upset(fromList(input), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
return(output)
  },
  s = Dom_Bac_DESeqLRT_ASVs,
  t = Dom_Bac_DESeqWald_ASVs,
  w = Dom_Bac_zinbLRT_ASVs,
  x = Dom_Bac_zinbWald_ASVs,
  y = Dom_Bac_ancomWZ_ASVs,
  z = Dom_Bac_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

Dom_Bac_upset_6DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 4.4.3 DA test from All data at ASV level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbWald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbLRT_All_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_All_Bac.RData")

## From each data frame, extract the ASV IDs caught by each test
All_Bac_DESeqLRT_ASVs <- rownames(sigtab_stddds2_LRT_All_Bac)
All_Bac_DESeqWald_ASVs <- rownames(sigtab_stddds2_Wald_All_Bac)
All_Bac_zinbLRT_ASVs <- rownames(sigtab_zinbLRT_All_Bac)
All_Bac_zinbWald_ASVs <- rownames(sigtab_zinbWald_All_Bac)
All_Bac_ancomWZ_ASVs <- as.vector(fdr_ancomWZ_All_Bac[,'Species'])
All_Bac_ancomNoZ_ASVs <- as.vector(fdr_ancomNoZ_All_Bac[,'Species'])

## use upset function. make sure to have names of the objects inside the lists!
input <- list(All_Bac_DESeqLRT_ASVs, All_Bac_DESeqWald_ASVs, All_Bac_zinbLRT_ASVs, All_Bac_zinbWald_ASVs, All_Bac_ancomWZ_ASVs, All_Bac_ancomNoZ_ASVs)
names(input) <- c("Std_DESeq_LRT",
                 "Std_DESeq_Wald", # adds names to the list of tests
                 "zinb_DESeq_LRT",
                 "zinb_DESeq_Wald",
                 "ancomWZ", 
                 "ancomNoZ")

All_Bac_upset_5DAtests <- upset(fromList(input), order.by = "freq", nsets = 12, nintersects = 100, set_size.show = TRUE) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
All_Bac_upset_5DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

<!-- ## 4.4.1 Combining DA tests per accession at Class level -->
<!-- ```{r} -->
<!-- ## Load outputs DESeq -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_stddds2_Wald_Acc_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_stddds2_LRT_Acc_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_zinbWald_Acc_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_zinbLRT_Acc_Bac.RData") -->

<!-- ## Load output Ancom -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ClassLevel/fdr_ancomWZ_Acc_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ClassLevel/fdr_ancomNoZ_Acc_Bac.RData") -->

<!-- ## From each data frame, extract the Class IDs caught by each test -->
<!-- Acc_Bac_DESeqLRT_Class <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x)) -->
<!-- Acc_Bac_DESeqLRT_Class <- lapply(Acc_Bac_DESeqLRT_Class, function(x) x[lengths(x) > 0]) -->
<!-- Acc_Bac_DESeqLRT_Class <- Acc_Bac_DESeqLRT_Class[length(Acc_Bac_DESeqLRT_Class) > 0] -->

<!-- Acc_Bac_DESeqWald_Class <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x)) -->
<!-- Acc_Bac_zinbLRT_Class <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x)) -->
<!-- Acc_Bac_zinbWald_Class <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x)) -->
<!-- Acc_Bac_ancomWZ_Class <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species'])) -->
<!-- Acc_Bac_ancomNoZ_Class <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species'])) -->

<!-- Alltests <- list(Acc_Bac_DESeqLRT_Class,Acc_Bac_DESeqWald_Class, Acc_Bac_zinbLRT_Class, Acc_Bac_zinbWald_Class, Acc_Bac_ancomWZ_Class, Acc_Bac_ancomNoZ_Class) -->
<!-- Alltests_clean <- lapply(Alltests, function(x) x[lengths(x) > 0]) -->
<!-- upset((Alltests_clean)) -->

<!-- ## use upset function. make sure to have names of the objects inside the lists! -->
<!-- Acc_Bac_upset_6DAtests <- mapply(function(s,t,w,x,y,z){ -->
<!--   input<-list(s,t,w,x,y,z) -->
<!--   names(input)<- c("Std_DESeq_LRT", -->
<!--                    "Std_DESeq_Wald", # adds names to the list of tests -->
<!--                    "zinb_DESeq_LRT", -->
<!--                    "zinb_DESeq_Wald", -->
<!--                    "ancomWZ",  -->
<!--                    "ancomNoZ") -->
<!-- output <- upset(fromList(input), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL -->
<!-- return(output) -->
<!--   }, -->
<!--   s = Alltests_clean[1], -->
<!--   t = Alltests_clean[2], -->
<!--   w = Alltests_clean[3], -->
<!--   x = Alltests_clean[4], -->
<!--   y = Alltests_clean[5], -->
<!--   z = Alltests_clean[6],  -->
<!--   SIMPLIFY = FALSE) -->

<!-- Acc_Bac_upset_6DAtests -->
<!-- ``` -->
<!-- ```{r} -->
<!-- # Clean environment -->
<!-- rm(list = ls()) -->
<!-- ``` -->

<!-- ## 4.4.4 DA test from All data at Class level -->
<!-- ```{r} -->
<!-- ## Load outputs DESeq -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_stddds2_Wald_All_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_stddds2_LRT_All_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_zinbWald_All_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ClassLevel/sigtab_zinbLRT_All_Bac.RData") -->

<!-- ## Load output Ancom -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ClassLevel/fdr_ancomWZ_All_Bac.RData") -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ClassLevel/fdr_ancomNoZ_All_Bac.RData") -->

<!-- ## From each data frame, extract the Class IDs caught by each test -->
<!-- All_Bac_DESeqLRT_Class <- rownames(sigtab_stddds2_LRT_All_Bac) -->
<!-- All_Bac_DESeqWald_Class <- rownames(sigtab_stddds2_Wald_All_Bac) -->
<!-- All_Bac_zinbLRT_Class <- rownames(sigtab_zinbLRT_All_Bac) -->
<!-- All_Bac_zinbWald_Class <- rownames(sigtab_zinbWald_All_Bac) -->
<!-- All_Bac_ancomWZ_Class <- as.vector(fdr_ancomWZ_All_Bac[,'Species']) -->
<!-- All_Bac_ancomNoZ_Class <- as.vector(fdr_ancomNoZ_All_Bac[,'Species']) -->

<!-- ## use upset function. make sure to have names of the objects inside the lists! -->
<!-- input <- list(All_Bac_DESeqLRT_Class, All_Bac_DESeqWald_Class, All_Bac_zinbLRT_Class, All_Bac_zinbWald_Class, All_Bac_ancomWZ_Class, All_Bac_ancomNoZ_Class) -->
<!-- names(input) <- c("Std_DESeq_LRT", -->
<!--                  "Std_DESeq_Wald", # adds names to the list of tests -->
<!--                  "zinb_DESeq_LRT", -->
<!--                  "zinb_DESeq_Wald", -->
<!--                  "ancomWZ",  -->
<!--                  "ancomNoZ") -->

<!-- All_Bac_upset_5DAtests <- upset(fromList(input), order.by = "freq", nsets = 12, nintersects = 100, set_size.show = TRUE) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL -->
<!-- All_Bac_upset_5DAtests -->
<!-- ``` -->
<!-- ```{r} -->
<!-- # Clean environment -->
<!-- rm(list = ls()) -->
<!-- ``` -->


# 4.5 Upsetplot comparing taxa per Accession
## 4.5.1 At ASV level
### Total DA ASVs
```{r, fig.width=12, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV.RData")

upset(fromList(Acc_Bac_NoDup_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### ASVs detected by two tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_TwoTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### ASVs detected by three tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_ThreeTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### ASVs detected by four tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FourTimes_DA_ASV.RData")
Acc_Bac_FourTimes_DA_ASV
```

### ASVs detected by five tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FiveTimes_DA_ASV.RData")
Acc_Bac_FiveTimes_DA_ASV
```
```{r}
# Clean environment
rm(list = ls())
```


## 4.5.2 At Class level
### Total DA Classes
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_Bac_NoDup_DA_Class.RData")

upset(fromList(Acc_Bac_NoDup_DA_Class), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Classes detected by two tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_Bac_TwoTimes_DA_Class.RData")

upset(fromList(Acc_Bac_TwoTimes_DA_Class), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Classes detected by three tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_Bac_ThreeTimes_DA_Class.RData")

upset(fromList(Acc_Bac_ThreeTimes_DA_Class), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Classes detected by four tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_Bac_FourTimes_DA_Class.RData")

upset(fromList(Acc_Bac_FourTimes_DA_Class), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Classes detected by five tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_Bac_FiveTimes_DA_Class.RData")
Acc_Bac_FiveTimes_DA_Class
```
```{r}
# Clean environment
rm(list = ls())
```


## 4.5.3 Phylum level
### Total DA Phyla
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_Bac_NoDup_DA_Phylum.RData")

upset(fromList(Acc_Bac_NoDup_DA_Phylum), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```
```{r, fig.width=6}
#upset(fromList(Acc_Bac_NoDup_DA_Phylum), order.by = "freq", nsets = 20, nintersects = 100, text.scale = 1.5)
```

### Phyla detected by two tests
```{r, fig.height=8}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_Bac_TwoTimes_DA_Phylum.RData")

upset(fromList(Acc_Bac_TwoTimes_DA_Phylum), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Phyla detected by three tests


# 4.6 Upsetplot comparing taxa by Domestication
## 4.6.1 At ASV level
### Total DA ASVs
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Dom_Bac_NoDup_DA_ASV.RData")

upset(fromList(Dom_Bac_NoDup_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```
```{r, fig.width=6}
#upset(fromList(Dom_Bac_NoDup_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, text.scale = 1.5)
```

### ASVs detected by two tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Dom_Bac_TwoTimes_DA_ASV.RData")

upset(fromList(Dom_Bac_TwoTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### ASVs detected by three tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Dom_Bac_ThreeTimes_DA_ASV.RData")
Dom_Bac_ThreeTimes_DA_ASV
```
```{r}
# Clean environment
rm(list = ls())
```


## 4.6.2 At Class level
### Total DA Classes
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Dom_Bac_NoDup_DA_Class.RData")
Dom_Bac_NoDup_DA_Class
```
```{r}
# Clean environment
rm(list = ls())
```

## 4.6.3 At Phylum level
### Total DA Phyla
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Dom_Bac_NoDup_DA_Phylum.RData")
Dom_Bac_NoDup_DA_Phylum
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.7 Upset plot all ASVs that may be responsible for resistance
### Load data
```{r}
# Acc ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# All data ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")

# VL and HM with high log2fold change
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/HM_lfc_1.0_ancomWZ_DA_ASV.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/HM_lfc_1.5_ancomWZ_DA_ASV.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/VL_lfc_1.0_ancomWZ_DA_ASV.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/VL_lfc_1.5_ancomWZ_DA_ASV.RData")
```

### Making list of DA ASVs
```{r}
# List with all selected DA ASVs
Selction_DA_ASVs <- list(All_twoDAtests = All_Bac_TwoTimes_DA_ASV,
                         HM_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$HM,
                         VL_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$VL,
                         HM_lfc_1.0 = HM_lfc_1.0_ancomWZ_DA_ASV,
                         HM_lfc_1.5 = HM_lfc_1.5_ancomWZ_DA_ASV,
                         VL_lfc_1.0 = VL_lfc_1.0_ancomWZ_DA_ASV,
                         VL_lfc_1.5 = VL_lfc_1.5_ancomWZ_DA_ASV)

# Save list with selected ASVs
save(Selction_DA_ASVs, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Selction_DA_ASVs.RData")

# List subsets
Comb_DA_ASVs_HM_VL <- list(HM_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$HM,
                           VL_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$VL,
                           HM_lfc_1.0 = HM_lfc_1.0_ancomWZ_DA_ASV,
                           HM_lfc_1.5 = HM_lfc_1.5_ancomWZ_DA_ASV,
                           VL_lfc_1.0 = VL_lfc_1.0_ancomWZ_DA_ASV,
                           VL_lfc_1.5 = VL_lfc_1.5_ancomWZ_DA_ASV)
Comb_DA_ASVs_lfc <- list(HM_lfc_1.0 = HM_lfc_1.0_ancomWZ_DA_ASV,
                         HM_lfc_1.5 = HM_lfc_1.5_ancomWZ_DA_ASV,
                         VL_lfc_1.0 = VL_lfc_1.0_ancomWZ_DA_ASV,
                         VL_lfc_1.5 = VL_lfc_1.5_ancomWZ_DA_ASV)
DA_ASVs_HM_VL_All <- list(HM_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$HM,
                          VL_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$VL,
                          All_twoDAtests = All_Bac_TwoTimes_DA_ASV)
```

### Upset plot
```{r}
upset(fromList(Comb_DA_ASVs_HM_VL), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
upset(fromList(Comb_DA_ASVs_lfc), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
upset(fromList(DA_ASVs_HM_VL_All), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.9 Venn diagram DA
## Accession
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DifferentialAbundance/SummaryFiles/Acc_Bac_conca_NoAn_nodup.RData")

simple_Venn <- ggVennDiagram(Acc_Bac_conca_NoAn_nodup)
print(simple_Venn)

# ASVs in both VL adn HM
ASV_VL_HM <- unlist(Acc_Bac_conca_NoAn_nodup[c("HM", "VL")])
Duplicates_DA_VL_HM <- ASV_VL_HM[duplicated(ASV_VL_HM)]
print(Duplicates_DA_VL_HM)

save(Duplicates_DA_VL_HM, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DifferentialAbundance/SummaryFiles/Duplicates_DA_VL_HM.RData")
```
