---
title: "CP_04_differential_abundance_ANCOM - Less filtered data and outliers removed - ASV level"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
- [Github page Ancom with FAQ including an explanation about what intercepts are](https://github.com/FrederickHuangLin/ANCOMBC)   
- [Info on Ancom](https://www.bioconductor.org/packages/release/bioc/vignettes/ANCOMBC/inst/doc/ANCOMBC.html)   
- [And more info on Ancom](https://microbiome.github.io/course_2022_turku/differential-abundance-analysis.html#ancom-bc)   
   
   
# 4.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(metamisc) #for phyloseq_sep_variable (and more?)
packageVersion("metamisc")
library(DESeq2)
packageVersion("DESeq2")
library(zinbwave)
packageVersion("zinbwave")
library(ANCOMBC)
packageVersion("ANCOMBC")
#library(rmeta) #rusher321/rmeta https://github.com/rusher321/rmeta/
#packageVersion("rmeta") 
library(dplyr)
packageVersion("dplyr")
```

# 4.2 ANCOM-BC ASV level
## 4.2.1 ANCOM including structural zeroes detection
### Accession
#### Run ANCOM
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA_Acc.RData")

ancomWZ_Bac_Acc <- lapply(CP_unnormalized_bac_ps_ForDA_Acc, function (x)
                           ancombc(data = x,
                             #tax_level = "Family",
                             formula = "Cat_treatment",
                             group = "Cat_treatment",
                             p_adj_method = "fdr",
                             prv_cut = 0.1, #default, taxa that are in less than the fraction present are discarded
                             lib_cut = 0, # threshold for filtering samples based on library size, 0 has no filtering
                             struc_zero = TRUE,
                             neg_lb = TRUE,
                             tol = 1e-5,
                             max_iter = 100,
                             conserve = TRUE,
                             alpha = 0.05,
                             global = FALSE))

save(ancomWZ_Bac_Acc, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
```

#### Organize and filter results
```{r}
ancomWZ_res_df_Acc_Bac <- lapply(ancomWZ_Bac_Acc, function (x)
  data.frame(
  Species = x$res$lfc$taxon, # ASVs
  lfc = unlist(x$res$lfc["Cat_treatmentMb"]),
  se = unlist(x$res$se["Cat_treatmentMb"]),
  W = unlist(x$res$W["Cat_treatmentMb"]), # test statistic
  p_val = unlist(x$res$p_val["Cat_treatmentMb"]),
  q_val = unlist(x$res$q_val["Cat_treatmentMb"]), # adjusted p-values
  diff_abn = unlist(x$res$diff_abn["Cat_treatmentMb"]))) # true if q value is less than alpha

fdr_ancomWZ_Acc_Bac <- lapply(ancomWZ_res_df_Acc_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.05)))
fdr_ancomWZ_Acc_Bac_0.001 <- lapply(ancomWZ_res_df_Acc_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.001)))
fdr_ancomWZ_Acc_Bac_0.0001 <- lapply(ancomWZ_res_df_Acc_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.0001)))

head(fdr_ancomWZ_Acc_Bac$VL)

# Number of differential ASV at different alphas
length(fdr_ancomWZ_Acc_Bac$VL$q_val)
length(fdr_ancomWZ_Acc_Bac_0.001$VL$q_val)
length(fdr_ancomWZ_Acc_Bac_0.0001$VL$q_val)

save(fdr_ancomWZ_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_Acc_Bac.RData")
```
```{r}
# Acc_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
# fdr_ancomWZ_Acc_Bac$VL[,'Species']
# fdr_ancomWZ_Acc_Bac$VL$Species
# 
# head(ancomWZ_Bac_Acc$VL$res$diff_abn, 20)
# summary(ancomWZ_Bac_Acc$VL$res$diff_abn)
# summary(ancomWZ_Bac_Acc$VL$res$q_val)
# length(ancomWZ_Bac_Acc$VL$res$q_val$`(Intercept)`)
# 
# head(fdr_ancomWZ_Acc_Bac$VL, 20)
# summary(fdr_ancomWZ_Acc_Bac$VL)
# 
# length(ancomWZ_Bac_Acc$VL$res$diff_abn)
# length(ancomWZ_res_df_Acc_Bac$VL$q_val)

# Clean environment
rm(list = ls())
```

#### Run ANCOM
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA_Dom.RData")

ancomWZ_Bac_Dom <- lapply(CP_unnormalized_bac_ps_ForDA_Dom, function (x)
                           ancombc(data = x,
                             #tax_level = "Family",
                             formula = "Accession + Cat_treatment",
                             group = "Cat_treatment",
                             p_adj_method = "fdr",
                             prv_cut = 0.1, #default, taxa that are in less than the fraction present are discarded
                             lib_cut = 0, # threshold for filtering samples based on library size, 0 has no filtering
                             struc_zero = TRUE,
                             neg_lb = TRUE,
                             tol = 1e-5,
                             max_iter = 100,
                             conserve = TRUE,
                             alpha = 0.05,
                             global = FALSE))

save(ancomWZ_Bac_Dom, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_Dom.RData")
```

#### Organize and filter results
```{r}
ancomWZ_res_df_Dom_Bac <- lapply(ancomWZ_Bac_Dom, function (x)
  data.frame(
  Species = x$res$lfc$taxon, # ASVs
  lfc = unlist(x$res$lfc["Cat_treatmentMb"]),
  se = unlist(x$res$se["Cat_treatmentMb"]),
  W = unlist(x$res$W["Cat_treatmentMb"]), # test statistic
  p_val = unlist(x$res$p_val["Cat_treatmentMb"]),
  q_val = unlist(x$res$q_val["Cat_treatmentMb"]), # adjusted p-values
  diff_abn = unlist(x$res$diff_abn["Cat_treatmentMb"]))) # true if q value is less than alpha

fdr_ancomWZ_Dom_Bac <- lapply(ancomWZ_res_df_Dom_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.05)))
fdr_ancomWZ_Dom_Bac_0.001 <- lapply(ancomWZ_res_df_Dom_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.001)))
fdr_ancomWZ_Dom_Bac_0.0001 <- lapply(ancomWZ_res_df_Dom_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.0001)))

head(fdr_ancomWZ_Dom_Bac$VL)

# Number of differential ASV at different alphas
length(fdr_ancomWZ_Dom_Bac$VL$q_val)
length(fdr_ancomWZ_Dom_Bac_0.001$VL$q_val)
length(fdr_ancomWZ_Dom_Bac_0.0001$VL$q_val)

save(fdr_ancomWZ_Dom_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_Dom_Bac.RData")
```
```{r}
rm(list = ls())
```

### All data
#### Run Ancom
```{r}
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA.RData")

ancomWZ_Bac_All <- ancombc(data = CP_unnormalized_bac_ps_ForDA,
                           #tax_level = "Family",
                           formula = "Accession + Cat_treatment",
                           group = "Cat_treatment",
                           p_adj_method = "fdr",
                           prv_cut = 0.1, #default, taxa that are in less than the fraction present are discarded
                           lib_cut = 0, # threshold for filtering samples based on library size, 0 has no filtering
                           struc_zero = TRUE,
                           neg_lb = TRUE,
                           tol = 1e-5,
                           max_iter = 100,
                           conserve = TRUE,
                           alpha = 0.05,
                           global = FALSE)

save(ancomWZ_Bac_All, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_All.RData")
```

#### Organize and filter results
```{r}
ancomWZ_res_df_All_Bac <- data.frame(
    Species = ancomWZ_Bac_All$res$lfc$taxon, # ASVs
    lfc = unlist(ancomWZ_Bac_All$res$lfc["Cat_treatmentMb"]),
    se = unlist(ancomWZ_Bac_All$res$se["Cat_treatmentMb"]),
    W = unlist(ancomWZ_Bac_All$res$W["Cat_treatmentMb"]), # test statistic
    p_val = unlist(ancomWZ_Bac_All$res$p_val["Cat_treatmentMb"]),
    q_val = unlist(ancomWZ_Bac_All$res$q_val["Cat_treatmentMb"]), # adjusted p-values
    diff_abn = unlist(ancomWZ_Bac_All$res$diff_abn["Cat_treatmentMb"]))

fdr_ancomWZ_All_Bac <- ancomWZ_res_df_All_Bac %>% dplyr::filter(q_val < 0.05)
fdr_ancomWZ_All_Bac_0.001 <- ancomWZ_res_df_All_Bac %>% dplyr::filter(q_val < 0.001)
fdr_ancomWZ_All_Bac_0.0001 <- ancomWZ_res_df_All_Bac %>% dplyr::filter(q_val < 0.0001)

head(fdr_ancomWZ_All_Bac)

# Number of differential ASV at different alphas
length(fdr_ancomWZ_All_Bac$q_val)
length(fdr_ancomWZ_All_Bac_0.001$q_val)
length(fdr_ancomWZ_All_Bac_0.0001$q_val)

save(fdr_ancomWZ_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_All_Bac.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


## 4.2.1 ANCOM WITHOUT structural zeroes detection
### Accession
#### Run ANCOM
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA_Acc.RData")

ancomNoZ_Bac_Acc <- lapply(CP_unnormalized_bac_ps_ForDA_Acc, function (x)
                           ancombc(data = x,
                             #tax_level = "Family",
                             formula = "Cat_treatment",
                             group = "Cat_treatment",
                             p_adj_method = "fdr",
                             prv_cut = 0.1, #default, taxa that are in less than the fraction present are discarded
                             lib_cut = 0, # threshold for filtering samples based on library size, 0 has no filtering
                             struc_zero = FALSE,
                             neg_lb = TRUE,
                             tol = 1e-5,
                             max_iter = 100,
                             conserve = TRUE,
                             alpha = 0.05,
                             global = FALSE))

save(ancomNoZ_Bac_Acc, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomNoZ_Bac_Acc.RData")
```

#### Organize and filter results
```{r}
ancomNoZ_res_df_Acc_Bac <- lapply(ancomNoZ_Bac_Acc, function (x)
    data.frame(
      Species = x$res$lfc$taxon, # ASVs
      lfc = unlist(x$res$lfc["Cat_treatmentMb"]),
      se = unlist(x$res$se["Cat_treatmentMb"]),
      W = unlist(x$res$W["Cat_treatmentMb"]), # test statistic
      p_val = unlist(x$res$p_val["Cat_treatmentMb"]),
      q_val = unlist(x$res$q_val["Cat_treatmentMb"]), # adjusted p-values
      diff_abn = unlist(x$res$diff_abn["Cat_treatmentMb"]))) # true if q value is less than alpha

fdr_ancomNoZ_Acc_Bac <- lapply(ancomNoZ_res_df_Acc_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.05)))
fdr_ancomNoZ_Acc_Bac_0.001 <- lapply(ancomNoZ_res_df_Acc_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.001)))
fdr_ancomNoZ_Acc_Bac_0.0001 <- lapply(ancomNoZ_res_df_Acc_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.0001)))

head(fdr_ancomNoZ_Acc_Bac$VL)

# Number of differential ASV at different alphas
length(fdr_ancomNoZ_Acc_Bac$VL$q_val)
length(fdr_ancomNoZ_Acc_Bac_0.001$VL$q_val)
length(fdr_ancomNoZ_Acc_Bac_0.0001$VL$q_val)

save(fdr_ancomNoZ_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Acc_Bac.RData")

# Clean environment
rm(list = ls())
```

### Domestication
#### Run ANCOM
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA_Dom.RData")

ancomNoZ_Bac_Dom <- lapply(CP_unnormalized_bac_ps_ForDA_Dom, function (x)
                           ancombc(data = x,
                             #tax_level = "Family",
                             formula = "Accession + Cat_treatment",
                             group = "Cat_treatment",
                             p_adj_method = "fdr",
                             prv_cut = 0.1, #default, taxa that are in less than the fraction present are discarded
                             lib_cut = 0, # threshold for filtering samples based on library size, 0 has no filtering
                             struc_zero = FALSE,
                             neg_lb = TRUE,
                             tol = 1e-5,
                             max_iter = 100,
                             conserve = TRUE,
                             alpha = 0.05,
                             global = FALSE))

save(ancomNoZ_Bac_Dom, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomNoZ_Bac_Dom.RData")
```

#### Organize and filter results
```{r}
ancomNoZ_res_df_Dom_Bac <- lapply(ancomNoZ_Bac_Dom, function (x)
    data.frame(
      Species = x$res$lfc$taxon, # ASVs
      lfc = unlist(x$res$lfc["Cat_treatmentMb"]),
      se = unlist(x$res$se["Cat_treatmentMb"]),
      W = unlist(x$res$W["Cat_treatmentMb"]), # test statistic
      p_val = unlist(x$res$p_val["Cat_treatmentMb"]),
      q_val = unlist(x$res$q_val["Cat_treatmentMb"]), # adjusted p-values
      diff_abn = unlist(x$res$diff_abn["Cat_treatmentMb"]))) # true if q value is less than alpha

fdr_ancomNoZ_Dom_Bac <- lapply(ancomNoZ_res_df_Dom_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.05)))
fdr_ancomNoZ_Dom_Bac_0.001 <- lapply(ancomNoZ_res_df_Dom_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.001)))
fdr_ancomNoZ_Dom_Bac_0.0001 <- lapply(ancomNoZ_res_df_Dom_Bac, function(x)
  (x %>% dplyr::filter(q_val < 0.0001)))

head(fdr_ancomNoZ_Dom_Bac$VL)

# Number of differential ASV at different alphas
length(fdr_ancomNoZ_Dom_Bac$VL$q_val)
length(fdr_ancomNoZ_Dom_Bac_0.001$VL$q_val)
length(fdr_ancomNoZ_Dom_Bac_0.0001$VL$q_val)

save(fdr_ancomNoZ_Dom_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Dom_Bac.RData")

# Clean environment
rm(list = ls())
```


### All data
#### Run Ancom
```{r}
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA.RData")

ancomNoZ_Bac_All <- ancombc(data = CP_unnormalized_bac_ps_ForDA,
                           #tax_level = "Family",
                           formula = "Accession + Cat_treatment",
                           group = "Cat_treatment",
                           p_adj_method = "fdr",
                           prv_cut = 0.1, #default, taxa that are in less than the fraction present are discarded
                           lib_cut = 0, # threshold for filtering samples based on library size, 0 has no filtering
                           struc_zero = FALSE,
                           neg_lb = TRUE,
                           tol = 1e-5,
                           max_iter = 100,
                           conserve = TRUE,
                           alpha = 0.05,
                           global = FALSE)

save(ancomNoZ_Bac_All, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomNoZ_Bac_All.RData")
```

#### Organize and filter results
```{r}
ancomNoZ_res_df_All_Bac <- data.frame(
    Species = ancomNoZ_Bac_All$res$lfc$taxon, # ASVs
    lfc = unlist(ancomNoZ_Bac_All$res$lfc["Cat_treatmentMb"]),
    se = unlist(ancomNoZ_Bac_All$res$se["Cat_treatmentMb"]),
    W = unlist(ancomNoZ_Bac_All$res$W["Cat_treatmentMb"]), # test statistic
    p_val = unlist(ancomNoZ_Bac_All$res$p_val["Cat_treatmentMb"]),
    q_val = unlist(ancomNoZ_Bac_All$res$q_val["Cat_treatmentMb"]), # adjusted p-values
    diff_abn = unlist(ancomNoZ_Bac_All$res$diff_abn["Cat_treatmentMb"]))

fdr_ancomNoZ_All_Bac <- ancomNoZ_res_df_All_Bac %>% dplyr::filter(q_val < 0.05)
fdr_ancomNoZ_All_Bac_0.001 <- ancomNoZ_res_df_All_Bac %>% dplyr::filter(q_val < 0.001)
fdr_ancomNoZ_All_Bac_0.0001 <- ancomNoZ_res_df_All_Bac %>% dplyr::filter(q_val < 0.0001)

head(fdr_ancomNoZ_All_Bac)

# Number of differential ASV at different alphas
length(fdr_ancomNoZ_All_Bac$q_val)
length(fdr_ancomNoZ_All_Bac_0.001$q_val)
length(fdr_ancomNoZ_All_Bac_0.0001$q_val)

save(fdr_ancomNoZ_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_All_Bac.RData")

# Clean environment
rm(list = ls())
```