---
title: "CP 01_loading_and_pre_processing - for differential abundance"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

# 0.0 Introduction
This script is for loading and pre-processing microbiome data that is output form the DADA2 [Ernakovich pipeline](https://github.com/ErnakovichLab/dada2_ernakovichlab) as describes in our own [GitHub page](https://github.com/Marcelara/PSF_Anunna). In addition, the ASVs ware assigned to taxonomies with Qiime2 as this results in a higher resolution. Here, the data will be further processed to be able to analyse the data.  

A nice [tutorial](https://www.nicholas-ollberding.com/post/introduction-to-phyloseq/) to learn how to investigate data in a phyloseq object.

This and the next script are originally made by Pedro Beschoren da Costa for the [MeJA_Pilot](https://github.com/PedroBeschoren/MeJA_Pilot) and modified to fit my data.

# 1.0 load all libraries, custom functions, setup environemnt
```{r warning=FALSE, message=FALSE}
# Set working directory
# setwd("")

# load libraries 
library(tidyr)
library(ggplot2)
library(phyloseq)
library(vegan)
library(metagenomeSeq)

# custom functions related to data loading and decontamination
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/RScripts/Functions/functions_loading_and_decontamination.R")

# increases memory limit used by R
# memory.limit(size = 350000) #This function is no longer supported by Windows
```


# 1.1 load microbiome data
### Swap samples
Check whether this is needed.

## 1.1.1 - Loading 16S data 
All samples were sequenced well.

#### Load OTU table
```{r}
# Loads the OTU table, immediately saving it as a phyloseq OTU table (lighter than a df) (from DADA2)
raw_bac_otutab <- otu_table(object = read.table(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Microbiome/seqtab_final.txt", 
                                                header = TRUE,
                                                row.names = 1, # first column has row names (ASV names)
                                                check.names = FALSE), # prevents "X" to be added to column names, such as X49_16S,
                             taxa_are_rows = TRUE)
head(raw_bac_otutab)
```

#### Load Taxonomy table
```{r}
# Updated taxonomy tale (from Qiime2)
sklearn_bac_taxtab <- read.table(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Microbiome/taxonomy.tsv", 
                                 header = TRUE,
                                 sep = "\t",
                                 row.names = 1, # first column has row names (ASV names)
                                 check.names = FALSE) # prevents "X" to be added to column names, such as X49_16S,

# adjusts number and name of columns
sklearn_bac_taxtab <- separate(data = sklearn_bac_taxtab,
                               col = Taxon,
                               into = c("Kingdom", 
                                       "Phylum", 
                                       "Class", 
                                       "Order", 
                                       "Family",
                                       "Genus", 
                                       "Species"),
                               sep = ";")

head(sklearn_bac_taxtab) #Separated well and Feature ID is a row name
#sklearn_bac_taxtab[c(13, 16, 71, 74, 75, 77, 78), ] #NAs are put in places where no classification was present. A subset of these samples is shown here.

# Change from d__Bacteria to k__bacteria, matching fungi data set
sklearn_bac_taxtab$Kingdom <- gsub("d__", "k__", sklearn_bac_taxtab$Kingdom)

head(sklearn_bac_taxtab["Kingdom"])

# Saves taxa as phyloseq object
sklearn_bac_taxtab <- tax_table(object = as.matrix(sklearn_bac_taxtab))

# Change name and remove old object
raw_bac_taxtab <- sklearn_bac_taxtab
rm(sklearn_bac_taxtab)
```

#### Load sequence table
```{r}
# Loads the representative sequences table, immediately saving it as a phyloseq refseq table (lighter than a df) (from DADA2)
raw_bac_refseq <- refseq(physeq = Biostrings::readDNAStringSet(filepath = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Microbiome/repset.fasta", 
                                                               use.names = TRUE)) 
taxa_names(raw_bac_refseq) <- gsub(" .*", "", taxa_names(raw_bac_refseq)) # drops taxonomy from ASV names
```

#### Loading metadata
```{r}
# Loads the mapping tile, immediately saving it as a phyloseq metadata table (lighter than a df)
raw_bac_metadata <- sample_data(object = read.table(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Metadata FP/Meta_Data_CPandFP.txt", 
                                                    header = TRUE,
                                                    sep = "\t",
                                                    row.names = 1, # first column has row names (ASV names)
                                                    check.names = FALSE)) # prevents "X" to be added to column names, such as X49_16S,
```

#### Build phyloseq object
```{r}
# build the main 16S phyloseq object by putting all these phyloseq-class objects (OTU table, tax table, ref seq, metadata) into a single phyloseq object (ps).
raw_bac_ps <- merge_phyloseq(raw_bac_otutab,
                           raw_bac_taxtab,
                           raw_bac_refseq,
                           raw_bac_metadata)

# remove old objects to reduce memory use
rm(raw_bac_otutab, raw_bac_taxtab, raw_bac_refseq, raw_bac_metadata)

# change names from ASV to bASV (bacterial ASV), so they can be distinguished from fungal ASVs.
taxa_names(raw_bac_ps) <- paste("b", taxa_names(raw_bac_ps), sep = "")
tax_table(raw_bac_ps) <- gsub(" ", "", tax_table(raw_bac_ps)) # drops a space character from taxa names. I don't know how that character ended up in there [Kris: I do not see the space]

# let's check the imported objects. Often errors will arise from typos when filling up the data sheets 
otu_table(raw_bac_ps)[1:10,1:10]
sample_data(raw_bac_ps)[1:10,1:10]
tax_table(raw_bac_ps)[1:10,1:6]
refseq(raw_bac_ps)

# run garbage collection after creating large objects (shows a report of memory usages)
gc()
```

#### Split CP and FP
```{r}
# subset phyloseq FP
FP_raw_bac_ps <- subset_samples(raw_bac_ps, Phase_PSF == "FP")
table(sample_data(FP_raw_bac_ps)[ , 1])

# save phyloseq object FP
save(FP_raw_bac_ps, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_raw_bac_ps.RData")

# change phase of blank samples from FP to CP to add them to the CP phyloseq object as well
sample_data(raw_bac_ps)[sample_data(raw_bac_ps)$Soil_conditioning == "B", "Phase_PSF"]
sample_data(raw_bac_ps)[sample_data(raw_bac_ps)$Soil_conditioning == "B", "Phase_PSF"] <- rep("CP", 5)
sample_data(raw_bac_ps)[sample_data(raw_bac_ps)$Soil_conditioning == "B", "Phase_PSF"]

# subset phyloseq CP
CP_raw_bac_ps <- subset_samples(raw_bac_ps, Phase_PSF == "CP")
table(sample_data(CP_raw_bac_ps)[ , 1])

# save phyloseq object FP
#save(CP_raw_bac_ps, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_raw_bac_ps.RData")

# remove original and FP phyloseq objects
rm(raw_bac_ps, FP_raw_bac_ps)
```


**Now we continue with first processing the CP**


# 1.2 - Remove bad ASVs
Here we will remove sequences associated with the host, background prokatyotes, low abundance ASVs, poorly identified sequences, and other bad ASVs.

## 1.2.1 - Remove non-bacterial sequences & filter ASVs
Note that the input data has been pre-filtered in the DADA2 pipeline to remove any ASVs that occur less than 3 times in the data set. This was necessary because the dada2-associated taxonomy assignment tools could not assign taxonomies to the full data set with 1 GB ram on the HPC.

### Here we investigate the length of the sequences
```{r}
hist(as.data.frame(refseq(CP_raw_bac_ps)@ranges)$width, breaks = 300, main = "Raw reads CP bac")
summary(as.data.frame(refseq(CP_raw_bac_ps)@ranges)$width) #summary on length of the reeds
ntaxa(CP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(CP_raw_bac_ps)) #shows number of reeds
```

### Here we get rid of sequences shorter than 380bp
```{r}
sum(as.data.frame(refseq(CP_raw_bac_ps)@ranges)$width < 380) # count number of reads smaller than 380bp
sum(as.data.frame(refseq(CP_raw_bac_ps)@ranges)$width > 380)

CP_raw_bac_ps <- prune_taxa(taxa = as.data.frame(refseq(CP_raw_bac_ps)@ranges)$width > 380,
                       x = CP_raw_bac_ps)
ntaxa(CP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(CP_raw_bac_ps)) #shows number of reeds
```

### Keeps only ASVs identified as bacterial
```{r}
CP_raw_bac_ps <- subset_taxa(CP_raw_bac_ps, Kingdom == "k__Bacteria")
ntaxa(CP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(CP_raw_bac_ps)) #shows number of reeds
```

### Remove Salinibacter ASV(s)
```{r}
CP_raw_bac_ps #original data set
physeq_OnlySal <- subset_taxa(CP_raw_bac_ps, Genus == "g__Salinibacter")
physeq_OnlySal
CP_raw_bac_ps <- subset_taxa(CP_raw_bac_ps, Genus != "g__Salinibacter" | is.na(Genus)) # remove Salinibacter genus but keep undefined genesis (NAs)
CP_raw_bac_ps #data set without salinibacter

rm(physeq_OnlySal)
```

### Define library sizes as metadata before filtering
```{r}
CP_raw_bac_ps@sam_data$library_sizes_prefiltering <- sample_sums(CP_raw_bac_ps)
hist(sample_sums(CP_raw_bac_ps), breaks = 20)
```

### Removes taxa having less than 8 reads across all samples, as VSEARCH standard
Turned off to run Ancom properly
```{r}
# otu_table(CP_raw_bac_ps) <- otu_table(CP_raw_bac_ps)[which (rowSums(otu_table(CP_raw_bac_ps)) > 7),]
# ntaxa(CP_raw_bac_ps) #shows total number of ASVs
# summary(sample_sums(CP_raw_bac_ps)) #shows number of reeds
```

### Remove ASV occurring in less than three samples
We remove ASVs that occur in less than 3 samples.
```{r}
filter <- phyloseq::genefilter_sample(CP_raw_bac_ps, filterfun_sample(function(x) x > 0), A = 3) 
CP_raw_bac_ps <- prune_taxa(filter, CP_raw_bac_ps)
ntaxa(CP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(CP_raw_bac_ps)) #shows number of reeds
```

### Check and remove plant-host contamination
```{r}
### define plastid, mitochondria and host plant contamination ps objects
Mitochondria_ps <- subset_taxa(CP_raw_bac_ps, Family == "f__Mitochondria" | Family == "Mitochondria")
Plastid_ps <- subset_taxa(CP_raw_bac_ps, Order == "o__Chloroplast" | Order == "Chloroplast") 
host_plant_ps <- merge_phyloseq(Mitochondria_ps, Plastid_ps)

### quick histogram showing plant DNA contamination
hist(sample_sums(host_plant_ps)/sample_sums(CP_raw_bac_ps)*100, breaks = 50)
summary(sample_sums(host_plant_ps)/sample_sums(CP_raw_bac_ps)*100)

summary(sample_sums(Mitochondria_ps)/sample_sums(CP_raw_bac_ps)*100)
summary(sample_sums(Plastid_ps)/sample_sums(CP_raw_bac_ps)*100)
```

Between 0 and 1% of the ASVs per sample is contaminated with mitochondrial or chloroplast DNA.

```{r}
### define host plant 16S contamination as metadata (add the info the the metadata sheet)
CP_raw_bac_ps@sam_data$Mitochondria_reads <- sample_sums(Mitochondria_ps)
CP_raw_bac_ps@sam_data$Plastid_reads <- sample_sums(Plastid_ps)
CP_raw_bac_ps@sam_data$Host_DNA_n_reads <- sample_sums(host_plant_ps)
CP_raw_bac_ps@sam_data$Host_DNA_contamination_pct <- sample_sums(host_plant_ps)/sample_sums(CP_raw_bac_ps)*100

### remove plant host sequences (plastid and mitochondrial DNA) 
CP_raw_bac_ps <- remove_Chloroplast_Mitochondria(CP_raw_bac_ps)
```

### Check library size
```{r}
# add library sizes as part of metadata
sample_data(CP_raw_bac_ps)$library_size <- sample_sums(CP_raw_bac_ps)

#check library size distribution
hist(sample_data(CP_raw_bac_ps)$library_size, breaks = 20)
summary(sample_data(CP_raw_bac_ps)$library_size) #shows the number of reeds

# remove samples with library size = 0  (non-informative; failed PCR/sequencing)
#CP_raw_bac_ps<-subset_samples(CP_raw_bac_ps, library_size >0) #not applicable

# remove ps objects
rm(Mitochondria_ps, Plastid_ps, host_plant_ps)

# run garbage collection after creating large objects
gc()
```


# 1.3 - Decontaminate phyloseq objects
The decontam package will use blank DNA samples to remove possible contaminants. Here we use a single custom function to run decontamination, generate plots, and return a clean phyloseq object

We will also check the number of reads in a blank sample (average +SD) so we can compare those blank samples with libraries with low number of reads. samples that cannot be distinguished from a blank (in terms of library size) 

### Decontaminate
```{r}
CP_unnormalized_bac_ps <- decontaminate_and_plot(CP_raw_bac_ps)

# phyloseq object without contamination
CP_unnormalized_bac_ps[1]

# Number of ASVs assigned as contamination
CP_unnormalized_bac_ps[5]

# ASVs assigned as contamination
CP_unnormalized_bac_ps[4]
```

### Check blanks
Let's look into how many reads our blank had (average + sd) so we can compare those to libraries with very low number of reads.
```{r}
blank_reads_bac <- sort(sample_sums(subset_samples(CP_raw_bac_ps, Soil_conditioning == "B")))
plot(blank_reads_bac)

#clean up environment to free memory
rm(CP_raw_bac_ps)
```

### Check phyloseq object
After checking the decontamination plots and reports, we can remove them and focus on the phyloseq object.
```{r}
CP_unnormalized_bac_ps <- unlist(CP_unnormalized_bac_ps[[1]])
```

#### Check library size Bacteria
Now let's check the lower end of our library sizes.
```{r}
sort(sample_sums(CP_unnormalized_bac_ps))
plot(sort(sample_sums(CP_unnormalized_bac_ps)), xlab = "Number of Reads")
```

No samples have a smaller library size than 6000 so no samples will be removed.

#### Remove samples with a number of reads similar to blanks
```{r}
CP_unnormalized_bac_ps <- subset_samples(CP_unnormalized_bac_ps, library_size > 6000)
```

### Set factors
```{r}
# set Cat_treatment as factor, and then order it properly
CP_unnormalized_bac_ps@sam_data$Cat_treatment <- factor(CP_unnormalized_bac_ps@sam_data$Cat_treatment, levels = c("Co", "Mb"))

# set Accession as factor, and then order it properly
CP_unnormalized_bac_ps@sam_data$Accession <- factor(CP_unnormalized_bac_ps@sam_data$Accession, levels = c("OH", "DD", "HE", "KI", "VL", "CD", "RI", "KT", "MC", "HM", "IT1", "GO1"))

# set Domestication as factor, and then order it properly
CP_unnormalized_bac_ps@sam_data$Domestication <- factor(CP_unnormalized_bac_ps@sam_data$Domestication, levels = c("Wild", "Cultivated"))

# set Batch as factor
CP_unnormalized_bac_ps@sam_data$Batch <- factor(CP_unnormalized_bac_ps@sam_data$Batch)
```


# 1.4 - Plot library sizes per treatment and check NAs and uncultured ASVs
### Check number of reeds
```{r}
# let's check the number of reads in a couple histograms
hist(sample_sums(CP_unnormalized_bac_ps), breaks = 50)
summary(sample_sums(CP_unnormalized_bac_ps))
```

### Check library size
```{r}
# check plot for some minimum library sizes on all samples

ggplot(data = sample_data(CP_unnormalized_bac_ps), 
           mapping = aes(x = library_size, y = Accession, fill = Cat_treatment)) +
      geom_jitter() +
      geom_boxplot() +
      scale_y_discrete(limits = rev) +
      geom_vline(xintercept = 30000, color="green") +
      geom_vline(xintercept = 25000, color="yellow") +
      geom_vline(xintercept = 20000, color="blue") +
      geom_vline(xintercept = 15000, color="red") +
      labs(x = "library size",
           y = "Accession") #+
      #theme(axis.text.x = element_text(angle = 20, vjust = 0.5, hjust=1))
```

### Check plant-host contamination
```{r}
# check bacterial 16S plant host contamination
ggplot(data = sample_data(CP_unnormalized_bac_ps), 
           mapping = aes(x = Host_DNA_contamination_pct, y = Accession, fill = Cat_treatment)) +
      geom_jitter() +
      geom_boxplot() +
      scale_y_discrete(limits = rev) +
      # geom_vline(xintercept = 30000, color="green") +
      # geom_vline(xintercept = 25000, color="yellow") +
      # geom_vline(xintercept = 20000, color="blue") +
      #geom_vline(xintercept = 15000, color="red") +
      labs(x = "Plant host DNA contamination",
           y = "Accession") #+
      #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

### check percentage of NA in taxonomy
```{r}
### check_n_taxa_NA_percentage(ps_object, ntaxa))
check_n_taxa_NA_percentage(CP_unnormalized_bac_ps, 100) # top 100 taxa
check_n_taxa_NA_percentage(CP_unnormalized_bac_ps, 1000)
check_n_taxa_NA_percentage(CP_unnormalized_bac_ps, 10000) # top 10,000 taxa
#A higher the number of taxa does not make a difference
```

### Check uncultured and unassigned taxa
Out commanded lines mean that type is not present in the data set.
```{r}
#subset_taxa(CP_unnormalized_bac_ps, Kingdom == "Unassigned") #No hits as I sub-setted the data with only k__Bacteria before

# subset_taxa(CP_unnormalized_bac_ps, Class == "c__uncultured" | Class == "c__unidentified")
# subset_taxa(CP_unnormalized_bac_ps, Class == "c__uncultured")
# subset_taxa(CP_unnormalized_bac_ps, Class == "c__unidentified") #not in data set

# subset_taxa(c__unidentified, Family == "f__uncultured" | Family == "f__unidentified")
# subset_taxa(CP_unnormalized_bac_ps, Family == "f__uncultured")
# subset_taxa(CP_unnormalized_bac_ps, Family == "f__unidentified") #not in data set

# subset_taxa(c__unidentified, Genus == "g__uncultured" | Genus == "g__unidentified")
# subset_taxa(CP_unnormalized_bac_ps, Genus == "g__uncultured")
# subset_taxa(CP_unnormalized_bac_ps, Genus == "g__unidentified") #not in data set
```

There are no uncultured or unidentified samples.

# 1.5 Remove samples?
## Rarefraction curve
```{r}
# finally, a rarefaction curve
a <- rarecurve(t(as.data.frame(otu_table(CP_unnormalized_bac_ps))), 
          label = FALSE, 
          step = 2000,
          main="Rarefaction", ylab = "Number of ASVs", xlab = "Number of DNA reads",
          abline(v = 10000, col="red", lwd=3, lty=2))
rm(a)

table(sample_data(CP_unnormalized_bac_ps)$Accession,
      sample_data(CP_unnormalized_bac_ps)$Soil_conditioning,
      sample_data(CP_unnormalized_bac_ps)$Cat_treatment)

sort(sample_sums(CP_unnormalized_bac_ps))[1:6]
sort(sample_sums(CP_unnormalized_bac_ps))[112:118]
```

#### Remove sample
Outliers with low library size (check per accession) are removed here.   
Samples with low library size (lowest 10% in t-distr.) of that accession and removed: C085 (5%), C256 (4%), C326 (0.7%), C375 (7%), C379 (9%), C429 (1%)    
Not low and not removed: C105 (56%), C217 (51%), C425 (34%), C438 (36%), C480 (21%)    
Not removed as well: C110 (51%), C380 (22%)   
   
```{r}
Outliers <- c("C085", "C256", "C326", "C375", "C379", "C429")
CP_unnormalized_bac_ps_ForDA <- subset_samples(CP_unnormalized_bac_ps, !sample_names(CP_unnormalized_bac_ps) %in% Outliers)
```
#### Remove CTRL samples

```{r}
Outliers2 <- c("CTRL1", "CTRL2", "CTRL3", "CTRL4", "CTRL5", "CTRL6", "CTRL7", "CTRL8", "CTRL9", "CTRL10")
CP_unnormalized_bac_ps_ForDA <- subset_samples(CP_unnormalized_bac_ps, !sample_names(CP_unnormalized_bac_ps) %in% Outliers2)
```

#### Rename NAs tax table to Unclassfied
```{r}
tax <- tax_table(CP_unnormalized_bac_ps_ForDA)
tax[is.na(tax[, "Genus"]), "Genus"] <- "g__Unclassified"
tax[is.na(tax[, "Family"]), "Family"] <- "f__Unclassified"
tax[is.na(tax[, "Order"]), "Order"] <- "o__Unclassified"
tax[is.na(tax[, "Class"]), "Class"] <- "c__Unclassified"
tax[is.na(tax[, "Phylum"]), "Phylum"] <- "p__Unclassified"
tax_table(CP_unnormalized_bac_ps_ForDA) <- tax
```

#### Split data on accessions and domestication
```{r}
CP_unnormalized_bac_ps_ForDA_Acc <- metagMisc::phyloseq_sep_variable(CP_unnormalized_bac_ps_ForDA, variable = "Accession")
CP_unnormalized_bac_ps_ForDA_Dom <- metagMisc::phyloseq_sep_variable(CP_unnormalized_bac_ps_ForDA, variable = "Domestication")
```


# 1.6 export filtered unnormalized ps object
### Save RData
let's save these normalized objects as RData so they can be loaded in other scripts.
```{r}
save(CP_unnormalized_bac_ps_ForDA, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA.RData")
save(CP_unnormalized_bac_ps_ForDA_Acc, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA_Acc.RData")
save(CP_unnormalized_bac_ps_ForDA_Dom, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA_Dom.RData")
```
