---
title: "FP_05_differential_abundance_Summary_Results - Upset plots"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b635.html)
   
   
# 5.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(metamisc) #for phyloseq_sep_variable (and more?)
packageVersion("metamisc")
library(DESeq2)
packageVersion("DESeq2")
library(zinbwave)
packageVersion("zinbwave")
library(UpSetR)
packageVersion("UpSetR")
library(tuple)
packageVersion("tuple")
library(ggpubr)
packageVersion("ggpubr")
library(stringr)
packageVersion("stringr")
library(pheatmap)
packageVersion("pheatmap")
library(ggVennDiagram)
packageVersion("ggVennDiagram")
```


# 5.4 Upset ASVs detected by each DA test
## 5.4.1 Combining DA tests per accession at ASV level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Acc_Bac.RData")

## From each data frame, extract the ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASVs <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_ASVs <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))

## use upset function. make sure to have names of the objects inside the lists!
Acc_Bac_upset_6DAtests <- mapply(function(s,t,w,x,y,z){
  input<-list(s,t,w,x,y,z)
  names(input)<- c("Std_DESeq_LRT",
                   "Std_DESeq_Wald", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output <- upset(fromList(input), order.by = "freq", nsets = 20, nintersects = 100) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
return(output)
  },
  s = Acc_Bac_DESeqLRT_ASVs,
  t = Acc_Bac_DESeqWald_ASVs,
  w = Acc_Bac_zinbLRT_ASVs,
  x = Acc_Bac_zinbWald_ASVs,
  y = Acc_Bac_ancomWZ_ASVs,
  z = Acc_Bac_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

Acc_Bac_upset_6DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.4.2 DA test from All data at ASV level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbWald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbLRT_All_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomWZ_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomNoZ_All_Bac.RData")

## From each data frame, extract the ASV IDs caught by each test
All_Bac_DESeqLRT_ASVs <- rownames(sigtab_stddds2_LRT_All_Bac)
All_Bac_DESeqWald_ASVs <- rownames(sigtab_stddds2_Wald_All_Bac)
All_Bac_zinbLRT_ASVs <- rownames(sigtab_zinbLRT_All_Bac)
All_Bac_zinbWald_ASVs <- rownames(sigtab_zinbWald_All_Bac)
All_Bac_ancomWZ_ASVs <- as.vector(fdr_ancomWZ_All_Bac[,'Species'])
All_Bac_ancomNoZ_ASVs <- as.vector(fdr_ancomNoZ_All_Bac[,'Species'])

## use upset function. make sure to have names of the objects inside the lists!
input <- list(All_Bac_DESeqLRT_ASVs, All_Bac_DESeqWald_ASVs, All_Bac_zinbLRT_ASVs, All_Bac_zinbWald_ASVs, All_Bac_ancomWZ_ASVs, All_Bac_ancomNoZ_ASVs)
names(input) <- c("Std_DESeq_LRT",
                 "Std_DESeq_Wald", # adds names to the list of tests
                 "zinb_DESeq_LRT",
                 "zinb_DESeq_Wald",
                 "ancomWZ", 
                 "ancomNoZ")

All_Bac_upset_5DAtests <- upset(fromList(input), order.by = "freq", nsets = 12, nintersects = 100) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
All_Bac_upset_5DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.4.3 Combining DA tests per accession at Class level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ClassLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ClassLevel/fdr_ancomNoZ_Acc_Bac.RData")

## From each data frame, extract the Class IDs caught by each test
Acc_Bac_DESeqLRT_Class <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
#Acc_Bac_DESeqLRT_Class <- lapply(Acc_Bac_DESeqLRT_Class, function(x) x[lengths(x) > 0])
#Acc_Bac_DESeqLRT_Class <- Acc_Bac_DESeqLRT_Class[length(Acc_Bac_DESeqLRT_Class) > 0]

Acc_Bac_DESeqWald_Class <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_Class <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_Class <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_Class <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_Class <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))

Alltests <- list(Acc_Bac_DESeqLRT_Class,Acc_Bac_DESeqWald_Class, Acc_Bac_zinbLRT_Class, Acc_Bac_zinbWald_Class, Acc_Bac_ancomWZ_Class, Acc_Bac_ancomNoZ_Class)
names(Alltests) <- c("Acc_Bac_DESeqLRT_Class", "Acc_Bac_DESeqWald_Class", "Acc_Bac_zinbLRT_Class", "Acc_Bac_zinbWald_Class", "Acc_Bac_ancomWZ_Class", "Acc_Bac_ancomNoZ_Class")
#Alltests_clean <- lapply(Alltests, function(x) x[lengths(x) > 0])

# print length of each object in list
lapply(Alltests, function(x) {lapply(x, function(y) length(y))})

#upset((Alltests_clean))

# ## use upset function. make sure to have names of the objects inside the lists!
# Acc_Bac_upset_6DAtests <- mapply(function(s,t,w,x,y,z){
#   input<-list(s,t,w,x,y,z)
#   names(input)<- c("Std_DESeq_LRT",
#                    "Std_DESeq_Wald", # adds names to the list of tests
#                    "zinb_DESeq_LRT",
#                    "zinb_DESeq_Wald",
#                    "ancomWZ",
#                    "ancomNoZ")
# output <- upset(fromList(input), order.by = "freq", nsets = 20, nintersects = 100) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
# return(output)
#   },
#   s = Alltests_clean[1],
#   t = Alltests_clean[2],
#   w = Alltests_clean[3],
#   x = Alltests_clean[4],
#   y = Alltests_clean[5],
#   z = Alltests_clean[6],
#   SIMPLIFY = FALSE)
# 
# Acc_Bac_upset_6DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.4.4 DA test from All data at Class level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_stddds2_Wald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_stddds2_LRT_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_zinbWald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_zinbLRT_All_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ClassLevel/fdr_ancomWZ_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ClassLevel/fdr_ancomNoZ_All_Bac.RData")

## From each data frame, extract the Class IDs caught by each test
All_Bac_DESeqLRT_Class <- rownames(sigtab_stddds2_LRT_All_Bac)
All_Bac_DESeqWald_Class <- rownames(sigtab_stddds2_Wald_All_Bac)
All_Bac_zinbLRT_Class <- rownames(sigtab_zinbLRT_All_Bac)
All_Bac_zinbWald_Class <- rownames(sigtab_zinbWald_All_Bac)
All_Bac_ancomWZ_Class <- as.vector(fdr_ancomWZ_All_Bac[,'Species'])
All_Bac_ancomNoZ_Class <- as.vector(fdr_ancomNoZ_All_Bac[,'Species'])

Alltests <- list(All_Bac_DESeqLRT_Class, All_Bac_DESeqWald_Class, All_Bac_zinbLRT_Class, All_Bac_zinbWald_Class, All_Bac_ancomWZ_Class, All_Bac_ancomNoZ_Class)
names(Alltests) <- c("All_Bac_DESeqLRT_Class", "All_Bac_DESeqWald_Class", "All_Bac_zinbLRT_Class", "All_Bac_zinbWald_Class", "All_Bac_ancomWZ_Class", "All_Bac_ancomNoZ_Class")

# print length of each object in list
lapply(Alltests, function(x) {length(x)})

# ## use upset function. make sure to have names of the objects inside the lists!
# input <- list(All_Bac_DESeqLRT_Class, All_Bac_DESeqWald_Class, All_Bac_zinbLRT_Class, All_Bac_zinbWald_Class, All_Bac_ancomWZ_Class, All_Bac_ancomNoZ_Class)
# names(input) <- c("Std_DESeq_LRT",
#                  "Std_DESeq_Wald", # adds names to the list of tests
#                  "zinb_DESeq_LRT",
#                  "zinb_DESeq_Wald",
#                  "ancomWZ",
#                  "ancomNoZ")
# 
# All_Bac_upset_5DAtests <- upset(fromList(input), order.by = "freq", nsets = 12, nintersects = 100) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
# All_Bac_upset_5DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.4.5 Combining DA tests per accession at Phylum level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_PhylumLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_PhylumLevel/fdr_ancomNoZ_Acc_Bac.RData")

## From each data frame, extract the Phylum IDs caught by each test
Acc_Bac_DESeqLRT_Phylum <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
#Acc_Bac_DESeqLRT_Phylum <- lapply(Acc_Bac_DESeqLRT_Phylum, function(x) x[lengths(x) > 0])
#Acc_Bac_DESeqLRT_Phylum <- Acc_Bac_DESeqLRT_Phylum[length(Acc_Bac_DESeqLRT_Phylum) > 0]

Acc_Bac_DESeqWald_Phylum <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_Phylum <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_Phylum <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_Phylum <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_Phylum <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))

Alltests <- list(Acc_Bac_DESeqLRT_Phylum,Acc_Bac_DESeqWald_Phylum, Acc_Bac_zinbLRT_Phylum, Acc_Bac_zinbWald_Phylum, Acc_Bac_ancomWZ_Phylum, Acc_Bac_ancomNoZ_Phylum)
#Alltests_clean <- lapply(Alltests, function(x) x[lengths(x) > 0])
names(Alltests) <- c("Acc_Bac_DESeqLRT_Phylum", "Acc_Bac_DESeqWald_Phylum", "Acc_Bac_zinbLRT_Phylum", "Acc_Bac_zinbWald_Phylum", "Acc_Bac_ancomWZ_Phylum", "Acc_Bac_ancomNoZ_Phylum")

# print length of each object in list
lapply(Alltests, function(x) {lapply(x, function(y) length(y))})

#upset((Alltests_clean))

# ## use upset function. make sure to have names of the objects inside the lists!
# Acc_Bac_upset_6DAtests <- mapply(function(s,t,w,x,y,z){
#   input<-list(s,t,w,x,y,z)
#   names(input)<- c("Std_DESeq_LRT",
#                    "Std_DESeq_Wald", # adds names to the list of tests
#                    "zinb_DESeq_LRT",
#                    "zinb_DESeq_Wald",
#                    "ancomWZ",
#                    "ancomNoZ")
# output <- upset(fromList(input), order.by = "freq", nsets = 20, nintersects = 100) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
# return(output)
#   },
#   s = Alltests_clean[1],
#   t = Alltests_clean[2],
#   w = Alltests_clean[3],
#   x = Alltests_clean[4],
#   y = Alltests_clean[5],
#   z = Alltests_clean[6],
#   SIMPLIFY = FALSE)
# 
# Acc_Bac_upset_6DAtests
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.4.6 DA test from All data at Phylum level
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_stddds2_Wald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_stddds2_LRT_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_zinbWald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_zinbLRT_All_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_PhylumLevel/fdr_ancomWZ_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_PhylumLevel/fdr_ancomNoZ_All_Bac.RData")

## From each data frame, extract the Phylum IDs caught by each test
All_Bac_DESeqLRT_Phylum <- rownames(sigtab_stddds2_LRT_All_Bac)
All_Bac_DESeqWald_Phylum <- rownames(sigtab_stddds2_Wald_All_Bac)
All_Bac_zinbLRT_Phylum <- rownames(sigtab_zinbLRT_All_Bac)
All_Bac_zinbWald_Phylum <- rownames(sigtab_zinbWald_All_Bac)
All_Bac_ancomWZ_Phylum <- as.vector(fdr_ancomWZ_All_Bac[,'Species'])
All_Bac_ancomNoZ_Phylum <- as.vector(fdr_ancomNoZ_All_Bac[,'Species'])

Alltests <- list(All_Bac_DESeqLRT_Phylum,All_Bac_DESeqWald_Phylum, All_Bac_zinbLRT_Phylum, All_Bac_zinbWald_Phylum, All_Bac_ancomWZ_Phylum, All_Bac_ancomNoZ_Phylum)
names(Alltests) <- c("All_Bac_DESeqLRT_Phylum", "All_Bac_DESeqWald_Phylum", "All_Bac_zinbLRT_Phylum", "All_Bac_zinbWald_Phylum", "All_Bac_ancomWZ_Phylum", "All_Bac_ancomNoZ_Phylum")

# print length of each object in list
lapply(Alltests, function(x) {length(x)})

# ## use upset function. make sure to have names of the objects inside the lists!
# input <- list(All_Bac_DESeqLRT_Phylum, All_Bac_DESeqWald_Phylum, All_Bac_zinbLRT_Phylum, All_Bac_zinbWald_Phylum, All_Bac_ancomWZ_Phylum, All_Bac_ancomNoZ_Phylum)
# names(input) <- c("Std_DESeq_LRT",
#                  "Std_DESeq_Wald", # adds names to the list of tests
#                  "zinb_DESeq_LRT",
#                  "zinb_DESeq_Wald",
#                  "ancomWZ",
#                  "ancomNoZ")
# 
# All_Bac_upset_5DAtests <- upset(fromList(input), order.by = "freq", nsets = 12, nintersects = 100) # text.scale = 3, set_size.angles = 0, set_size.show = FALSE, set_size.numbers_size = NULL
# All_Bac_upset_5DAtests
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.5 Upsetplot comparing ASVs by Accessions at ASV level
## 5.5.1 Total DA ASVs
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV.RData")

upset(fromList(Acc_Bac_NoDup_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100)
```
```{r, fig.width=6}
#upset(fromList(Acc_Bac_NoDup_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100, text.scale = 1.5)
```

## 5.5.2 ASVs detected by two tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_TwoTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100)
```

## 5.5.3 ASVs detected by three tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_ThreeTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100)
```

## 5.5.4 ASVs detected by four tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FourTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_FourTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100)
```

## 5.5.5 ASVs detected by five tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FiveTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_FiveTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100)
```

## 5.5.6 ASVs detected by six tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_SixTimes_DA_ASV.RData")

upset(fromList(Acc_Bac_SixTimes_DA_ASV), order.by = "freq", nsets = 20, nintersects = 100)
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.6 Upsetplot comparing Classes by Accessions at Class level
## 5.6.1 Total DA Classes
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_NoDup_Merged_DA_Class.RData")

Acc_NoDup_Merged_DA_Class <- lapply(Acc_NoDup_Merged_DA_Class, function(x) {
  x <- x$Class
})

upset(fromList(Acc_NoDup_Merged_DA_Class), order.by = "freq", nsets = 20, nintersects = 100)
```
```{r, fig.width=6}
#upset(fromList(Acc_Bac_NoDup_DA_Class), order.by = "freq", nsets = 20, nintersects = 100, text.scale = 1.5)
```

## 5.6.2 Classes detected by two tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_TwoTimes_Merged_DA_Class.RData")

Acc_TwoTimes_Merged_DA_Class <- lapply(Acc_TwoTimes_Merged_DA_Class, function(x) {
  x <- x$Class
})

upset(fromList(Acc_TwoTimes_Merged_DA_Class), order.by = "freq", nsets = 20, nintersects = 100)
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.7 Upsetplot comparing Phyla by Accessions at Phylum level
## 5.7.1 Total DA Phyla
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_NoDup_Merged_DA_Phylum.RData")

upset(fromList(Acc_NoDup_Merged_DA_Phylum), order.by = "freq", nsets = 20, nintersects = 100)
```
```{r, fig.width=6}
#upset(fromList(Acc_Bac_NoDup_DA_Phylum), order.by = "freq", nsets = 20, nintersects = 100, text.scale = 1.5)
```

## 5.7.2 Phyla detected by two tests
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_TwoTimes_Merged_DA_Phylum.RData")

upset(fromList(Acc_TwoTimes_Merged_DA_Phylum), order.by = "freq", nsets = 20, nintersects = 100)
```
```{r}
# Clean environment
rm(list = ls())
```

# 5.8 Upset plot all ASVs that may be responsible for resistance
### Load data
```{r}
# Acc ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# All data ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")

# VL and HM with high log2fold change
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/HM_lfc_2.5_ancomWZ_DA_ASV.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/HM_lfc_3_ancomWZ_DA_ASV.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/VL_lfc_2.5_ancomWZ_DA_ASV.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/VL_lfc_3_ancomWZ_DA_ASV.RData")
```

### Making list of DA ASVs
```{r}
# List with all selected DA ASVs
Selction_DA_ASVs <- list(All_twoDAtests = All_Bac_TwoTimes_DA_ASV,
                         HM_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$HM,
                         VL_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$VL,
                         HM_lfc_2.5 = HM_lfc_2.5_ancomWZ_DA_ASV,
                         HM_lfc_3 = HM_lfc_3_ancomWZ_DA_ASV,
                         VL_lfc_2.5 = VL_lfc_2.5_ancomWZ_DA_ASV,
                         VL_lfc_3 = VL_lfc_3_ancomWZ_DA_ASV)

# Save list with selected ASVs
save(Selction_DA_ASVs, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Selction_DA_ASVs.RData")

# List subsets
Comb_DA_ASVs_HM_VL <- list(HM_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$HM,
                           VL_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$VL,
                           HM_lfc_2.5 = HM_lfc_2.5_ancomWZ_DA_ASV,
                           HM_lfc_3 = HM_lfc_3_ancomWZ_DA_ASV,
                           VL_lfc_2.5 = VL_lfc_2.5_ancomWZ_DA_ASV,
                           VL_lfc_3 = VL_lfc_3_ancomWZ_DA_ASV)
Comb_DA_ASVs_lfc <- list(HM_lfc_2.5 = HM_lfc_2.5_ancomWZ_DA_ASV,
                         HM_lfc_3 = HM_lfc_3_ancomWZ_DA_ASV,
                         VL_lfc_2.5 = VL_lfc_2.5_ancomWZ_DA_ASV,
                         VL_lfc_3 = VL_lfc_3_ancomWZ_DA_ASV)
DA_ASVs_HM_VL_All <- list(HM_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$HM,
                          VL_twoDAtests = Acc_Bac_TwoTimes_DA_ASV$VL,
                          All_twoDAtests = All_Bac_TwoTimes_DA_ASV)
```

### Upset plot
```{r}
upset(fromList(Comb_DA_ASVs_HM_VL), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
upset(fromList(Comb_DA_ASVs_lfc), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
upset(fromList(DA_ASVs_HM_VL_All), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```


# 5.9 Venn diagram DA
### Accession
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DifferentialAbundance/SummaryFiles/Acc_Bac_conca_NoAn_nodup.RData")

simple_Venn <- ggVennDiagram(Acc_Bac_conca_NoAn_nodup)
print(simple_Venn)

# ASVs in both VL adn HM
ASV_VL_HM <- unlist(Acc_Bac_conca_NoAn_nodup[c("HM", "VL")])
Duplicates_DA_VL_HM <- ASV_VL_HM[duplicated(ASV_VL_HM)]
print(Duplicates_DA_VL_HM)

save(Duplicates_DA_VL_HM, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DifferentialAbundance/SummaryFiles/Duplicates_DA_VL_HM.RData")
```


