---
title: "FP_05_differential_abundance_DESeq - Less filtered data and outliers removed - Run at ASV level"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Base tutorial on DESeq2](https://joey711.github.io/phyloseq-extensions/DESeq2.html)   
- [Elaborate tutorial on DESeq2](https://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html)   
- [Setting contrasts in DESeq2](https://www.atakanekiz.com/technical/a-guide-to-designs-and-contrasts-in-DESeq2/)   
- [More information on contrast](https://www.biostars.org/p/9597417/)   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b635.html)   
   
   
# 5.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(metamisc) #for phyloseq_sep_variable (and more?)
packageVersion("metamisc")
library(DESeq2)
packageVersion("DESeq2")
library(zinbwave)
packageVersion("zinbwave")
```

### Load data without unplanted CTRL and Co caterpillar treatment
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean_Acc.RData")
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

#### Remove Batch 1
```{r}
FP_unnormalized_bac_ps_ForDA_clean_Acc <- lapply(FP_unnormalized_bac_ps_ForDA_clean_Acc, function(x) subset_samples(x, Batch != 1))
FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Batch != 1)
```


### For ZINBWAve, remove taxa with a total sum of 0
#### Seperate to each accession
```{r}
lapply(FP_unnormalized_bac_ps_ForDA_clean_Acc, function(x) summary(taxa_sums(x)))

Acc_unnorm_bac_ps_l_filtered <- lapply(FP_unnormalized_bac_ps_ForDA_clean_Acc, function(x)
         prune_taxa(taxa_sums(x)>0, x))

save(Acc_unnorm_bac_ps_l_filtered, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/Acc_unnorm_bac_ps_l_filtered.RData")
```
At this moment, filtered and unfiltered object are the same.

#### All data together
```{r}
summary(taxa_sums(FP_unnormalized_bac_ps_ForDA_clean))

All_unnorm_bac_ps_filtered <- prune_taxa(taxa_sums(FP_unnormalized_bac_ps_ForDA_clean)>0, FP_unnormalized_bac_ps_ForDA_clean)

save(All_unnorm_bac_ps_filtered, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/All_unnorm_bac_ps_filtered.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.1 DeSeq and ZINBDeSeq
## 5.1.1 Convert phyloseq object to DeSeq object with lists
```{r}
# Accession unfiltered
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean_Acc.RData")

FP_unnormalized_bac_ps_ForDA_clean_Acc <- lapply(FP_unnormalized_bac_ps_ForDA_clean_Acc, function(x) subset_samples(x, Batch != 1))
dds_Acc_unnorm_bac_ps_l_unfilt <- lapply(FP_unnormalized_bac_ps_ForDA_clean_Acc, function(x) phyloseq_to_deseq2(x, ~ Soil_conditioning))

save(dds_Acc_unnorm_bac_ps_l_unfilt, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_Acc_unnorm_bac_ps_l_unfilt.RData")
rm(FP_unnormalized_bac_ps_ForDA_clean_Acc, dds_Acc_unnorm_bac_ps_l_unfilt)

# Accession filtered
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/Acc_unnorm_bac_ps_l_filtered.RData")
dds_Acc_unnorm_bac_ps_l_filtered <- lapply(Acc_unnorm_bac_ps_l_filtered, function(x) phyloseq_to_deseq2(x, ~ Soil_conditioning))
save(dds_Acc_unnorm_bac_ps_l_filtered, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_Acc_unnorm_bac_ps_l_filtered.RData")
rm(Acc_unnorm_bac_ps_l_filtered, dds_Acc_unnorm_bac_ps_l_filtered)

# All data unfiltered
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Batch != 1)
dds_All_unnorm_bac_ps_unfilt <- phyloseq_to_deseq2(FP_unnormalized_bac_ps_ForDA_clean, ~ Accession + Soil_conditioning)

save(dds_All_unnorm_bac_ps_unfilt, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_All_unnorm_bac_ps_unfilt.RData")
rm(FP_unnormalized_bac_ps_ForDA_clean, dds_All_unnorm_bac_ps_unfilt)

# All data filtered
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/All_unnorm_bac_ps_filtered.RData")
dds_All_unnorm_bac_ps_filtered <- phyloseq_to_deseq2(All_unnorm_bac_ps_filtered, ~ Accession + Soil_conditioning)
save(dds_All_unnorm_bac_ps_filtered, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_All_unnorm_bac_ps_filtered.RData")
rm(All_unnorm_bac_ps_filtered, dds_All_unnorm_bac_ps_filtered)
```

### Checking which fitType fits best
Testing with one accession
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_Acc_unnorm_bac_ps_l_unfilt.RData")

DDS <- estimateSizeFactors(dds_Acc_unnorm_bac_ps_l_unfilt$GO1)
par <- estimateDispersions(DDS, fitType = "parametric")
loc <- estimateDispersions(DDS, fitType = "local")
mea <- estimateDispersions(DDS, fitType = "mean")
plotDispEsts(par, main = "dispEst: parametric")
plotDispEsts(loc, main = "dispEst: local")
plotDispEsts(mea, main = "dispEst: mean")

# residual <- mcols(DDS)$dispGeneEst - mcols(DDS)$dispFit
# plotMA(residual)

rm(DDS, par, loc, mea)
```
The plot of local looks best.

## 5.1.2 DESeq with Wald test
### Per Accession
#### Run DESeq object
```{r}
# Standard DESeq
stddds2_Wald_Acc_Bac <- lapply(dds_Acc_unnorm_bac_ps_l_unfilt, function(x) 
                  DESeq(x, 
                  test = "Wald",  
                  minReplicatesForReplace = Inf, # not replacing outliers
                  fitType = "local")) 

# Print results for GO1
res <- results(stddds2_Wald_Acc_Bac$GO1, alpha = 0.1)
res
summary(res)

# Number of p-values smaller than 0.1
sum(res$padj < 0.1, na.rm = TRUE)

# Plot output: blue dots have p-values below 0.1
plotMA(res)

# Save DESeq object
save(stddds2_Wald_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/stddds2_Wald_Acc_Bac.RData")
```

#### Creat data frame and shrink data
```{r}
# Data frame generation
# "Adds shrunken log2 fold changes (LFC) and SE to a results table from DESeq run without LFC shrinkage"
lfcres_stddds2_Wald_Acc_Bac <- lapply(stddds2_Wald_Acc_Bac, function(x) lfcShrink(x, coef = "Soil_conditioning_Mb_vs_Co", type = "normal")) #type is type of shrinkage estimator

# Save file
save(lfcres_stddds2_Wald_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/lfcres_stddds2_Wald_Acc_Bac.RData")
```

#### Filter p-values
```{r}
#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_Wald_Acc_Bac <- lapply(lfcres_stddds2_Wald_Acc_Bac, function(x) x[which(x$padj < alpha), ])
sigtab_stddds2_Wald_Acc_Bac

# Safe file
save(sigtab_stddds2_Wald_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_Acc_Bac.RData")

# Clean environment
rm(list = ls())
```

### All data
#### Run DESeq object
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_All_unnorm_bac_ps_unfilt.RData")

# Standard DESeq
stddds2_Wald_All_Bac <- DESeq(dds_All_unnorm_bac_ps_unfilt, 
                              test = "Wald",
                              minReplicatesForReplace = Inf, # not replacing outliers
                              fitType = "local")

# stddds2_Wald_All_Bac_S <- DESeq(dds_All_unnorm_bac_ps_unfilt_S,
#                               test = "Wald",
#                               minReplicatesForReplace = Inf, # not replacing outliers
#                               fitType = "local")

# Print results
resultsNames(stddds2_Wald_All_Bac) # Show different contrasts, last one is default in result output
res <- results(stddds2_Wald_All_Bac, alpha = 0.1)
#res <- results(stddds2_Wald_All_Bac, alpha = 0.1, contrast = c("Soil_conditioning", "Co", "Mb")) # specifically setting a contrast
#res <- results(stddds2_Wald_All_Bac, alpha = 0.1, contrast = list("Soil_conditioning_Mb_vs_Co")) # alternative to specifically set a contrast
# I can compare CD and RI for instance by setting contrast like: contrast = list("Accession_CD_vs_VL", "Accession_RI_vs_VL")
res
summary(res)

# resultsNames(stddds2_Wald_All_Bac_S) # Show different contrasts, last one is default in result output
# res_S <- results(stddds2_Wald_All_Bac_S, alpha = 0.1)
# res_S
# summary(res_S)

# Number of p-values smaller than 0.1
sum(res$padj < 0.1, na.rm = TRUE)

# sum(res_S$padj < 0.1, na.rm = TRUE)

# Plot output: blue dots have p-values below 0.1
plotMA(res)

# Save DESeq object
save(stddds2_Wald_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/stddds2_Wald_All_Bac.RData")
```

#### Shrink data
```{r}
# Data frame generation
# "Adds shrunken log2 fold changes (LFC) and SE to a results table from DESeq run without LFC shrinkage"
lfcres_stddds2_Wald_All_Bac <- lfcShrink(stddds2_Wald_All_Bac, coef = "Soil_conditioning_Mb_vs_Co", type = "normal") #type is type of shrinkage estimator

# Save file
save(lfcres_stddds2_Wald_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/lfcres_stddds2_Wald_All_Bac.RData")
```

#### Filter p-values
```{r}
#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_Wald_All_Bac <- lfcres_stddds2_Wald_All_Bac[which(lfcres_stddds2_Wald_All_Bac$padj < alpha), ]
sigtab_stddds2_Wald_All_Bac

# Safe file
save(sigtab_stddds2_Wald_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_All_Bac.RData")

# Clean environment
rm(list = ls())
```

## 5.1.3 DESeq with LRT test
### Per Accession
#### Run DESeq object
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_Acc_unnorm_bac_ps_l_unfilt.RData")

# Standard DESeq
stddds2_LRT_Acc_Bac <- lapply(dds_Acc_unnorm_bac_ps_l_unfilt, function(x) 
                  DESeq(x, 
                  test = "LRT",
                  reduced = ~1,
                  minReplicatesForReplace = Inf, # not replacing outliers
                  fitType = "local")) 

# Print results for GO1
res <- results(stddds2_LRT_Acc_Bac$GO1, alpha = 0.1)
res
summary(res)

# Number of p-values smaller than 0.1
sum(res$padj < 0.1, na.rm = TRUE)

# Plot output: blue dots have p-values below 0.1
plotMA(res)

# Save DESeq object
save(stddds2_LRT_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/stddds2_LRT_Acc_Bac.RData")
```

#### Creat data frame and shrink data
```{r}
# Data frame generation
# "Adds shrunken log2 fold changes (LFC) and SE to a results table from DESeq run without LFC shrinkage"
lfcres_stddds2_LRT_Acc_Bac <- lapply(stddds2_LRT_Acc_Bac, function(x) lfcShrink(x,  coef = "Soil_conditioning_Mb_vs_Co", type = "normal")) #type is type of shrinkage estimator

# Save file
save(lfcres_stddds2_LRT_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/lfcres_stddds2_LRT_Acc_Bac.RData")
```

#### Filter p-values
```{r}
#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_LRT_Acc_Bac <- lapply(lfcres_stddds2_LRT_Acc_Bac, function(x) x[which(x$padj < alpha), ])
sigtab_stddds2_LRT_Acc_Bac

# Safe file
save(sigtab_stddds2_LRT_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_Acc_Bac.RData")

# Clean environment
rm(list = ls())
```

### All data
#### Run DESeq object
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_All_unnorm_bac_ps_unfilt.RData")

# Standard DESeq
stddds2_LRT_All_Bac <- DESeq(dds_All_unnorm_bac_ps_unfilt, 
                             test = "LRT",
                             reduced = ~Accession,
                             minReplicatesForReplace = Inf, # not replacing outliers
                             fitType = "local")

# Print results
resultsNames(stddds2_LRT_All_Bac)
res <- results(stddds2_LRT_All_Bac, alpha = 0.1)
res
summary(res)

# Number of p-values smaller than 0.1
sum(res$padj < 0.1, na.rm = TRUE)

# Plot output: blue dots have p-values below 0.1
plotMA(res)

# Save DESeq object
save(stddds2_LRT_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/stddds2_LRT_All_Bac.RData")
```

#### Shrink data
```{r}
# Data frame generation
# "Adds shrunken log2 fold changes (LFC) and SE to a results table from DESeq run without LFC shrinkage"
lfcres_stddds2_LRT_All_Bac <- lfcShrink(stddds2_LRT_All_Bac, coef = "Soil_conditioning_Mb_vs_Co", type = "normal") #type is type of shrinkage estimator

# Save file
save(lfcres_stddds2_LRT_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/lfcres_stddds2_LRT_All_Bac.RData")
```

#### Filter p-values
```{r}
#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_LRT_All_Bac <- lfcres_stddds2_LRT_All_Bac[which(lfcres_stddds2_LRT_All_Bac$padj < alpha), ]
sigtab_stddds2_LRT_All_Bac

# Safe file
save(sigtab_stddds2_LRT_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_All_Bac.RData")

# Clean environment
rm(list = ls())
```

## 5.1.4 ZINBDESeq
### 5.1.4.1 Calculate weights
#### Per accession
```{r}
# library("scran") DESeq can compute the size factors already from the weights output of zinbwave

load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_Acc_unnorm_bac_ps_l_filtered.RData")

# Calculate observational weights with ZINBwave
zw_Acc_Bac <- lapply(dds_Acc_unnorm_bac_ps_l_filtered, function(x)
                                zinbwave(x,
                               epsilon = 1e10,
                               verbose = TRUE,
                               K = 0,
                               observationalWeights = TRUE,
                               BPPARAM = BiocParallel::SerialParam()))

save(zw_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/zw_Acc_Bac.RData")

# Check that weights have been calculated for each assay in each sample
assay(zw_Acc_Bac$GO1, "weights")[1:20, 1:10]

# Clean environment
rm(list = ls())
```
 
#### All data
```{r}
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds_All_unnorm_bac_ps_filtered.RData")

# Calculate observational weights with ZINBwave
zw_All_Bac <- zinbwave(dds_All_unnorm_bac_ps_filtered,
                       epsilon = 1e10,
                       verbose = TRUE,
                       K = 0,
                       observationalWeights = TRUE,
                       BPPARAM = BiocParallel::SerialParam())

save(zw_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/zw_All_Bac.RData")

# Check that weights have been calculated for each assay in each sample
assay(zw_All_Bac, "weights")[1:20, 1:10]

# Clean environment
rm(list = ls())
```

### 5.1.4.2 Estimate size factors and run DeSeq
#### Accession
```{r}
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/zw_Acc_Bac.RData")

# Compute size factors with DeSeq
dds2_zw_Acc_Bac <- lapply(zw_Acc_Bac, 
                          function(x) DESeqDataSet(x, design = ~ Soil_conditioning))
dds2_zw_Acc_Bac <- lapply(dds2_zw_Acc_Bac, 
                            function(x) estimateSizeFactors(x, type = "poscounts"))

# Save data
save(dds2_zw_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_zw_Acc_Bac.RData")

# DeSeq with LRT test
dds2_LRT_zinb_Acc_Bac <- lapply(dds2_zw_Acc_Bac, function(x) 
                  DESeq(x, 
                  test = "LRT", 
                  reduced = ~1,
                  sfType = "poscounts", 
                  minmu = 1e-6, 
                  minReplicatesForReplace = Inf,
                  fitType = "local"))

# Save output
save(dds2_LRT_zinb_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_LRT_zinb_Acc_Bac.RData")

# DeSeq with Wald test test
dds2_Wald_zinb_Acc_Bac <- lapply(dds2_zw_Acc_Bac, function(x) 
                  DESeq(x, 
                  test = "Wald",
                  sfType = "poscounts",
                  useT = TRUE,
                  minmu = 1e-6, 
                  minReplicatesForReplace = Inf,
                  fitType = "local")) 

save(dds2_Wald_zinb_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_Wald_zinb_Acc_Bac.RData")

# Clean environment
rm(list = ls())
```

#### All accession
```{r}
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/zw_All_Bac.RData")

# Compute size factors with DeSeq
dds2_zw_All_Bac <- DESeqDataSet(zw_All_Bac, design = ~ Accession + Soil_conditioning)
dds2_zw_All_Bac <- estimateSizeFactors(dds2_zw_All_Bac, type = "poscounts")

# Save data
save(dds2_zw_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_zw_All_Bac.RData")

# DeSeq with LRT test
dds2_LRT_zinb_All_Bac <-  DESeq(dds2_zw_All_Bac, 
                                  test = "LRT", 
                                  reduced = ~ Accession,
                                  sfType = "poscounts", 
                                  minmu = 1e-6, 
                                  minReplicatesForReplace = Inf,
                                  fitType = "local")

# Save output
save(dds2_LRT_zinb_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_LRT_zinb_All_Bac.RData")

# DeSeq with Wald test test
dds2_Wald_zinb_All_Bac <- DESeq(dds2_zw_All_Bac, 
                                test = "Wald",
                                sfType = "poscounts",
                                useT = TRUE,
                                minmu = 1e-6, 
                                minReplicatesForReplace = Inf,
                                fitType = "local")

save(dds2_Wald_zinb_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_Wald_zinb_All_Bac.RData")

# Clean environment
rm(list = ls())
```

### 5.1.4.3 Dataframe generation
```{r}
alpha <- 0.05
```

#### Accession
```{r}
# LRT
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_LRT_zinb_Acc_Bac.RData")
res_zinbLRT_Acc_Bac <- lapply(dds2_LRT_zinb_Acc_Bac, function(x) results(x, contrast = c("Soil_conditioning", "Mb", "Co")))
save(res_zinbLRT_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/res_zinbLRT_Acc_Bac.RData")

## Filter to only p adjusted under alpha
sigtab_zinbLRT_Acc_Bac <- lapply(res_zinbLRT_Acc_Bac, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbLRT_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbLRT_Acc_Bac.RData")
sigtab_zinbLRT_Acc_Bac
rm(dds2_LRT_zinb_Acc_Bac, sigtab_zinbLRT_Acc_Bac, res_zinbLRT_Acc_Bac)

# Wald
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_Wald_zinb_Acc_Bac.RData")
res_zinbWald_Acc_Bac <- lapply(dds2_Wald_zinb_Acc_Bac, function(x) results(x, contrast = c("Soil_conditioning", "Mb", "Co")))
save(res_zinbWald_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/res_zinbWald_Acc_Bac.RData")

## Filter to only p adjusted under alpha
sigtab_zinbWald_Acc_Bac <- lapply(res_zinbWald_Acc_Bac, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbWald_Acc_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbWald_Acc_Bac.RData")
sigtab_zinbWald_Acc_Bac
rm(dds2_Wald_zinb_Acc_Bac, res_zinbWald_Acc_Bac, sigtab_zinbWald_Acc_Bac)
```

#### All data
```{r}
# LRT
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_LRT_zinb_All_Bac.RData")
res_zinbLRT_All_Bac <- results(dds2_LRT_zinb_All_Bac, contrast = c("Soil_conditioning", "Mb", "Co"))
save(res_zinbLRT_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/res_zinbLRT_All_Bac.RData")

## Filter to only p adjusted under alpha
sigtab_zinbLRT_All_Bac <- res_zinbLRT_All_Bac[which(res_zinbLRT_All_Bac$padj < alpha), ]
save(sigtab_zinbLRT_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbLRT_All_Bac.RData")
sigtab_zinbLRT_All_Bac
rm(dds2_LRT_zinb_All_Bac, res_zinbLRT_All_Bac, sigtab_zinbLRT_All_Bac)

# Wald
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/dds2_Wald_zinb_All_Bac.RData")
res_zinbWald_All_Bac <- results(dds2_Wald_zinb_All_Bac, contrast = c("Soil_conditioning", "Mb", "Co"))
save(res_zinbWald_All_Bac,file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/res_zinbWald_All_Bac.RData")

## Filter to only p adjusted under alpha
sigtab_zinbWald_All_Bac <- res_zinbWald_All_Bac[which(res_zinbWald_All_Bac$padj < alpha), ]
save(sigtab_zinbWald_All_Bac, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbWald_All_Bac.RData")
sigtab_zinbWald_All_Bac
rm(dds2_Wald_zinb_All_Bac, res_zinbWald_All_Bac, sigtab_zinbWald_All_Bac)
```
