---
title: "FP_05_differential_abundance_Summary_Results"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b635.html)
   
   
# 5.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(metamisc) #for phyloseq_sep_variable (and more?)
packageVersion("metamisc")
library(DESeq2)
packageVersion("DESeq2")
library(zinbwave)
packageVersion("zinbwave")
library(UpSetR)
packageVersion("UpSetR")
library(tuple)
packageVersion("tuple")
library(ggpubr)
packageVersion("ggpubr")
library(stringr)
packageVersion("stringr")
library(pheatmap)
packageVersion("pheatmap")
library(ggVennDiagram)
packageVersion("ggVennDiagram")
```


# 5.1 Relative abundance comparing ASVs HM and VL
## ASVs detected by two DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# Extracting overlapping ASVs HM and VL
Acc_Bac_HM_VL_DA_ASV <- c(Acc_Bac_TwoTimes_DA_ASV$VL, Acc_Bac_TwoTimes_DA_ASV$HM)
Acc_Bac_NumbOcc_DA_ASV <- table(Acc_Bac_HM_VL_DA_ASV)
Acc_Bac_HMVLoverlap_DA_ASV <- names(Acc_Bac_NumbOcc_DA_ASV[Acc_Bac_NumbOcc_DA_ASV == 2])
Acc_Bac_HMVLoverlap_DA_ASV

# Remove DA ASVs that are shared with GO1
Acc_Bac_HM_VL_GO1_DA_ASV <- c(Acc_Bac_TwoTimes_DA_ASV$VL, Acc_Bac_TwoTimes_DA_ASV$HM, Acc_Bac_TwoTimes_DA_ASV$GO1)
Acc_Bac_NumbOcc_DA_ASV <- table(Acc_Bac_HM_VL_GO1_DA_ASV)
Acc_Bac_HMVLGO1overlap_DA_ASV <- names(Acc_Bac_NumbOcc_DA_ASV[Acc_Bac_NumbOcc_DA_ASV == 3])
Acc_Bac_HMVLGO1overlap_DA_ASV
Acc_Bac_HMVLoverlap_DA_ASV <- Acc_Bac_HMVLoverlap_DA_ASV[! Acc_Bac_HMVLoverlap_DA_ASV %in% Acc_Bac_HMVLGO1overlap_DA_ASV]

# Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA_VLHM <- prune_taxa(Acc_Bac_HMVLoverlap_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean_RA)
ps_RA_VLHM <- subset_samples(ps_RA_VLHM, Accession == "HM" | Accession == "VL")
tax_table(ps_RA_VLHM)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA_VLHM <- psmelt(ps_RA_VLHM)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA_VLHM$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA_VLHM$Genus[na_genus] <- paste0("Uncl_", ps_RA_VLHM$Class[na_genus])
ps_RA_VLHM$ASV_Genus <- paste(ps_RA_VLHM$OTU, ps_RA_VLHM$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA_VLHM,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA_VLHM <- merge(ps_RA_VLHM,
                    otu_means[, c("OTU", "difference")],
                    by = "OTU",
                    all.x = TRUE)
ps_RA_VLHM$Genus <- with(ps_RA_VLHM, reorder(Genus, difference))
ps_RA_VLHM$OTU <- with(ps_RA_VLHM, reorder(OTU, -difference))
```

```{r, fig.height=12, fig.width=19}
ggplot(ps_RA_VLHM, 
       aes(x = Accession, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Accession), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Accession") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "right")
```

```{r, fig.height=6, fig.width=8}
ggplot(ps_RA_VLHM, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


## ASVs detected by three DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

# Extracting overlapping ASVs HM and VL
Acc_Bac_HM_VL_DA_ASV <- c(Acc_Bac_ThreeTimes_DA_ASV$VL, Acc_Bac_ThreeTimes_DA_ASV$HM)
Acc_Bac_NumbOcc_DA_ASV <- table(Acc_Bac_HM_VL_DA_ASV)
Acc_Bac_HMVLoverlap_DA_ASV <- names(Acc_Bac_NumbOcc_DA_ASV[Acc_Bac_NumbOcc_DA_ASV == 2])
Acc_Bac_HMVLoverlap_DA_ASV

# Remove DA ASVs that are shared with GO1
Acc_Bac_HM_VL_GO1_DA_ASV <- c(Acc_Bac_ThreeTimes_DA_ASV$VL, Acc_Bac_ThreeTimes_DA_ASV$HM, Acc_Bac_ThreeTimes_DA_ASV$GO1)
Acc_Bac_NumbOcc_DA_ASV <- table(Acc_Bac_HM_VL_GO1_DA_ASV)
Acc_Bac_HMVLGO1overlap_DA_ASV <- names(Acc_Bac_NumbOcc_DA_ASV[Acc_Bac_NumbOcc_DA_ASV == 3])
Acc_Bac_HMVLGO1overlap_DA_ASV
Acc_Bac_HMVLoverlap_DA_ASV <- Acc_Bac_HMVLoverlap_DA_ASV[! Acc_Bac_HMVLoverlap_DA_ASV %in% Acc_Bac_HMVLGO1overlap_DA_ASV]

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA_VLHM <- prune_taxa(Acc_Bac_HMVLoverlap_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean_RA)
ps_RA_VLHM <- subset_samples(ps_RA_VLHM, Accession == "HM" | Accession == "VL")
tax_table(ps_RA_VLHM)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA_VLHM <- psmelt(ps_RA_VLHM)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA_VLHM$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA_VLHM$Genus[na_genus] <- paste0("Uncl_", ps_RA_VLHM$Class[na_genus])
ps_RA_VLHM$ASV_Genus <- paste(ps_RA_VLHM$OTU, ps_RA_VLHM$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA_VLHM,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA_VLHM <- merge(ps_RA_VLHM,
                    otu_means[, c("OTU", "difference")],
                    by = "OTU",
                    all.x = TRUE)
ps_RA_VLHM$Genus <- with(ps_RA_VLHM, reorder(Genus, difference))
ps_RA_VLHM$OTU <- with(ps_RA_VLHM, reorder(OTU, -difference))
```

```{r, fig.height=4, fig.width=10}
ggplot(ps_RA_VLHM, 
       aes(x = Accession, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Accession), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Accession") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "right")
```

```{r, fig.height=3}
ggplot(ps_RA_VLHM, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.2 Relative abundance of DA ASVs all data
## ASVs detected by two DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA <- prune_taxa(All_Bac_TwoTimes_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean_RA)
tax_table(ps_RA)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA <- psmelt(ps_RA)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA$Genus[na_genus] <- paste0("Uncl_", ps_RA$Class[na_genus])
ps_RA$ASV_Genus <- paste(ps_RA$OTU, ps_RA$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA <- merge(ps_RA,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA$Genus <- with(ps_RA, reorder(Genus, difference))
ps_RA$OTU <- with(ps_RA, reorder(OTU, -difference))
```

```{r, fig.height=12, fig.width=20}
ggplot(ps_RA, 
       aes(x = Accession, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Accession), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Accession") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = c(0.9, 0.1))
```

```{r, fig.height=8, fig.width=12}
ggplot(ps_RA, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "none")
```

```{r, fig.height=5, fig.width=12}
ggplot(ps_RA, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


## ASVs detected by three DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_ThreeTimes_DA_ASV.RData")

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA <- prune_taxa(All_Bac_ThreeTimes_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean_RA)
tax_table(ps_RA)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA <- psmelt(ps_RA)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA$Genus[na_genus] <- paste0("Uncl_", ps_RA$Class[na_genus])
ps_RA$ASV_Genus <- paste(ps_RA$OTU, ps_RA$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA <- merge(ps_RA,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA$Genus <- with(ps_RA, reorder(Genus, difference))
ps_RA$OTU <- with(ps_RA, reorder(OTU, -difference))
```

```{r, fig.height=6, fig.width=8}
ggplot(ps_RA, 
       aes(x = Accession, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Accession), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Accession") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "right")
```

```{r, fig.height=6, fig.width=6}
ggplot(ps_RA, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "none")
```

```{r, fig.height=4}
ggplot(ps_RA, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```

```{r}
# Clean environment
rm(list = ls())
```


# 5.3 Relative abundance of DA ASVs HM
## ASVs detected by two DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA <- prune_taxa(Acc_Bac_TwoTimes_DA_ASV$HM, FP_unnormalized_bac_ps_ForDA_clean_RA)
ps_RA <- subset_samples(ps_RA, Accession == "HM")
tax_table(ps_RA)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA <- psmelt(ps_RA)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA$Genus[na_genus] <- paste0("Uncl_", ps_RA$Class[na_genus])
ps_RA$ASV_Genus <- paste(ps_RA$OTU, ps_RA$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA <- merge(ps_RA,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA$Genus <- with(ps_RA, reorder(Genus, difference))
ps_RA$OTU <- with(ps_RA, reorder(OTU, -difference))
```

```{r, fig.height=30, fig.width=30}
ggplot(ps_RA, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 1, color = "black"),
        axis.title = element_text(size = 3),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 5),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "none")
```

```{r, fig.height=50, fig.width=20}
ggplot(ps_RA, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


## ASVs detected by three DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA <- prune_taxa(Acc_Bac_ThreeTimes_DA_ASV$HM, FP_unnormalized_bac_ps_ForDA_clean_RA)
ps_RA <- subset_samples(ps_RA, Accession == "HM")
tax_table(ps_RA)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA <- psmelt(ps_RA)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA$Genus[na_genus] <- paste0("Uncl_", ps_RA$Class[na_genus])
ps_RA$ASV_Genus <- paste(ps_RA$OTU, ps_RA$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA <- merge(ps_RA,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA$Genus <- with(ps_RA, reorder(Genus, difference))
ps_RA$OTU <- with(ps_RA, reorder(OTU, -difference))
```

```{r, fig.height=10, fig.width=12}
ggplot(ps_RA, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 4, color = "black"),
        axis.title = element_text(size = 15),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 5),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "none")
```

```{r, fig.height=25, fig.width=15}
ggplot(ps_RA, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


## ASVs detected filtered based on log2fold change
ASVs detected by two tests and ancomWZ used for log2fold change

### Data preperation
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")

# Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# number of ASVs with large log2Fold change
summary(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb)
length(ancomWZ_Bac_Acc$HM$res$lfc[abs(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) > 1, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$HM$res$lfc[(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) > 2, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$HM$res$lfc[(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) < -2, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$HM$res$lfc[(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) > 2.5, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$HM$res$lfc[(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) < -2.5, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$HM$res$lfc[(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) > 3, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$HM$res$lfc[(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) < -3, "Soil_conditioningMb" ])

# ASVs with a log2fold change larger than 2.5 or smaller than -2.5
ancomWZ_Bac_Acc_lfc_2.5 <- ancomWZ_Bac_Acc$HM$res$lfc[abs(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) > 2.5, ]
HM_lfc_2.5_ancomWZ_DA_ASV <- ancomWZ_Bac_Acc_lfc_2.5$taxon
save(HM_lfc_2.5_ancomWZ_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/HM_lfc_2.5_ancomWZ_DA_ASV.RData")

# ASVs with a log2fold change larger than 3 or smaller than -3
ancomWZ_Bac_Acc_lfc_3 <- ancomWZ_Bac_Acc$HM$res$lfc[abs(ancomWZ_Bac_Acc$HM$res$lfc$Soil_conditioningMb) > 3, ]
HM_lfc_3_ancomWZ_DA_ASV <- ancomWZ_Bac_Acc_lfc_3$taxon
save(HM_lfc_3_ancomWZ_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/HM_lfc_3_ancomWZ_DA_ASV.RData")

# Subset phyloseq object
ps_RA <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean_RA, Accession == "HM")
ps_RA_lfc_2.5 <- prune_taxa(ancomWZ_Bac_Acc_lfc_2.5$taxon, ps_RA)
ps_RA_lfc_3 <- prune_taxa(ancomWZ_Bac_Acc_lfc_3$taxon, ps_RA)
tax_table(ps_RA_lfc_2.5)[ , 2:6]
tax_table(ps_RA_lfc_3)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA_lfc_2.5 <- psmelt(ps_RA_lfc_2.5)
ps_RA_lfc_3 <- psmelt(ps_RA_lfc_3)

# Identify rows where Genus is NA
na_genus_2.5 <- is.na(ps_RA_lfc_2.5$Genus)
na_genus_3 <- is.na(ps_RA_lfc_3$Genus)

# Replace those NAs with "Unclassified_" + Class name for 2.5
ps_RA_lfc_2.5$Genus[na_genus_2.5] <- paste0("Uncl_", ps_RA_lfc_2.5$Class[na_genus_2.5])
ps_RA_lfc_2.5$ASV_Genus <- paste(ps_RA_lfc_2.5$OTU, ps_RA_lfc_2.5$Genus, sep = "_")

## Change order of ASVs depending on difference between soil treatments for 2.5
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA_lfc_2.5,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA_lfc_2.5 <- merge(ps_RA_lfc_2.5,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA_lfc_2.5$Genus <- with(ps_RA_lfc_2.5, reorder(Genus, difference))
ps_RA_lfc_2.5$OTU <- with(ps_RA_lfc_2.5, reorder(OTU, -difference))

# Replace those NAs with "Unclassified_" + Class name for 3
ps_RA_lfc_3$Genus[na_genus_3] <- paste0("Uncl_", ps_RA_lfc_3$Class[na_genus_3])
ps_RA_lfc_3$ASV_Genus <- paste(ps_RA_lfc_3$OTU, ps_RA_lfc_3$Genus, sep = "_")

## Change order of ASVs depending on difference between soil treatments for 3
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA_lfc_3,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA_lfc_3 <- merge(ps_RA_lfc_3,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA_lfc_3$Genus <- with(ps_RA_lfc_3, reorder(Genus, difference))
ps_RA_lfc_3$OTU <- with(ps_RA_lfc_3, reorder(OTU, -difference))
```

### Grapth lfc of 2.5
```{r, fig.height=10, fig.width=12}
ggplot(ps_RA_lfc_2.5, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 7, color = "black"),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 8),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.position = c(0.9, 0.1))
```

```{r, fig.height=10, fig.width=12}
ggplot(ps_RA_lfc_2.5, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```

### Grapth lfc of 3
```{r, fig.height=7, fig.width=10}
ggplot(ps_RA_lfc_3, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 7, color = "black"),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 8),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.position = c(0.9, 0.1))
```

```{r, fig.height=6, fig.width=10}
ggplot(ps_RA_lfc_3, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.4 Relative abundance of DA ASVs VL
## ASVs detected by two DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA <- prune_taxa(Acc_Bac_TwoTimes_DA_ASV$VL, FP_unnormalized_bac_ps_ForDA_clean_RA)
ps_RA <- subset_samples(ps_RA, Accession == "VL")
tax_table(ps_RA)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA <- psmelt(ps_RA)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA$Genus[na_genus] <- paste0("Uncl_", ps_RA$Class[na_genus])
ps_RA$ASV_Genus <- paste(ps_RA$OTU, ps_RA$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA <- merge(ps_RA,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA$Genus <- with(ps_RA, reorder(Genus, difference))
ps_RA$OTU <- with(ps_RA, reorder(OTU, -difference))
```

```{r, fig.height=12, fig.width=15}
ggplot(ps_RA, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 1, color = "black"),
        axis.title = element_text(size = 3),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 5),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "none")
```

```{r, fig.height=35, fig.width = 12}
ggplot(ps_RA, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```

## ASVs detected by three DA tests
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

## Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# Subset phyloseq object
ps_RA <- prune_taxa(Acc_Bac_ThreeTimes_DA_ASV$VL, FP_unnormalized_bac_ps_ForDA_clean_RA)
ps_RA <- subset_samples(ps_RA, Accession == "VL")
tax_table(ps_RA)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA <- psmelt(ps_RA)

# Identify rows where Genus is NA
na_genus <- is.na(ps_RA$Genus)

# Replace those NAs with "Unclassified_" + Class name
ps_RA$Genus[na_genus] <- paste0("Uncl_", ps_RA$Class[na_genus])
ps_RA$ASV_Genus <- paste(ps_RA$OTU, ps_RA$Genus, sep = "_")

#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "g__", "") # remove g__
#ps_RA_VLHM$Genus <- str_replace_all(ps_RA_VLHM$Genus, "_c__", "\n") # remove c__
#ps_RA_VLHM[ps_RA_VLHM$OTU == "bASV_831", "OTU_Genus"] <- "bASV_831 Burkholderia- *"

## Change order of ASVs depending on difference between soil treatments
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA <- merge(ps_RA,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA$Genus <- with(ps_RA, reorder(Genus, difference))
ps_RA$OTU <- with(ps_RA, reorder(OTU, -difference))
```

```{r, fig.height=6, fig.width=10}
ggplot(ps_RA, 
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 5, color = "black"),
        axis.title = element_text(size = 10),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 8),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 15),
        legend.position = "none")
```

```{r, fig.height=12, fig.width=16}
ggplot(ps_RA, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```


## ASVs detected filtered based on log2fold change
ASVs detected by two tests and ancomWZ used for log2fold change

### Data preperation
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")

# Calculate relative abundance
FP_unnormalized_bac_ps_ForDA_clean_RA <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))

# number of ASVs with large log2Fold change
summary(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb)
length(ancomWZ_Bac_Acc$VL$res$lfc[abs(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) > 1, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$VL$res$lfc[(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) > 2, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$VL$res$lfc[(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) < -2, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$VL$res$lfc[(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) > 2.5, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$VL$res$lfc[(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) < -2.5, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$VL$res$lfc[(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) > 3, "Soil_conditioningMb" ])
length(ancomWZ_Bac_Acc$VL$res$lfc[(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) < -3, "Soil_conditioningMb" ])

# ASVs with a log2fold change larger than 2.5 or smaller than -2.5
ancomWZ_Bac_Acc_lfc_2.5 <- ancomWZ_Bac_Acc$VL$res$lfc[abs(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) > 2.5, ]
VL_lfc_2.5_ancomWZ_DA_ASV <- ancomWZ_Bac_Acc_lfc_2.5$taxon
save(VL_lfc_2.5_ancomWZ_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/VL_lfc_2.5_ancomWZ_DA_ASV.RData")


# ASVs with a log2fold change larger than 3 or smaller than -3
ancomWZ_Bac_Acc_lfc_3 <- ancomWZ_Bac_Acc$VL$res$lfc[abs(ancomWZ_Bac_Acc$VL$res$lfc$Soil_conditioningMb) > 3, ]
VL_lfc_3_ancomWZ_DA_ASV <- ancomWZ_Bac_Acc_lfc_3$taxon
save(VL_lfc_3_ancomWZ_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/VL_lfc_3_ancomWZ_DA_ASV.RData")

# Subset phyloseq object
ps_RA <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean_RA, Accession == "VL")
ps_RA_lfc_2.5 <- prune_taxa(ancomWZ_Bac_Acc_lfc_2.5$taxon, ps_RA)
ps_RA_lfc_3 <- prune_taxa(ancomWZ_Bac_Acc_lfc_3$taxon, ps_RA)
tax_table(ps_RA_lfc_2.5)[ , 2:6]
tax_table(ps_RA_lfc_3)[ , 2:6]

## Phyloseq object to data frame with relative abundance, meta data and taxonomy
ps_RA_lfc_2.5 <- psmelt(ps_RA_lfc_2.5)
ps_RA_lfc_3 <- psmelt(ps_RA_lfc_3)

# Identify rows where Genus is NA
na_genus_2.5 <- is.na(ps_RA_lfc_2.5$Genus)
na_genus_3 <- is.na(ps_RA_lfc_3$Genus)

# Replace those NAs with "Unclassified_" + Class name for 2.5
ps_RA_lfc_2.5$Genus[na_genus_2.5] <- paste0("Uncl_", ps_RA_lfc_2.5$Class[na_genus_2.5])
ps_RA_lfc_2.5$ASV_Genus <- paste(ps_RA_lfc_2.5$OTU, ps_RA_lfc_2.5$Genus, sep = "_")

## Change order of ASVs depending on difference between soil treatments for 2.5
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA_lfc_2.5,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA_lfc_2.5 <- merge(ps_RA_lfc_2.5,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA_lfc_2.5$Genus <- with(ps_RA_lfc_2.5, reorder(Genus, difference))
ps_RA_lfc_2.5$OTU <- with(ps_RA_lfc_2.5, reorder(OTU, -difference))

# Replace those NAs with "Unclassified_" + Class name for 3
ps_RA_lfc_3$Genus[na_genus_3] <- paste0("Uncl_", ps_RA_lfc_3$Class[na_genus_3])
ps_RA_lfc_3$ASV_Genus <- paste(ps_RA_lfc_3$OTU, ps_RA_lfc_3$Genus, sep = "_")

## Change order of ASVs depending on difference between soil treatments for 3
otu_means <- aggregate(Abundance ~ OTU + Soil_conditioning,
                       data = ps_RA_lfc_3,
                       FUN = mean)
otu_means <- reshape(otu_means, # Change configuration df
                     timevar = "Soil_conditioning",
                     idvar = "OTU",
                     direction = "wide")
otu_means$difference <- otu_means$Abundance.Mb - otu_means$Abundance.Co
ps_RA_lfc_3 <- merge(ps_RA_lfc_3,
                otu_means[, c("OTU", "difference")],
                by = "OTU",
                all.x = TRUE)
ps_RA_lfc_3$Genus <- with(ps_RA_lfc_3, reorder(Genus, difference))
ps_RA_lfc_3$OTU <- with(ps_RA_lfc_3, reorder(OTU, -difference))
```

### Grapth lfc of 2.5
```{r, fig.height=7, fig.width=9.5}
ggplot(ps_RA_lfc_2.5,
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 7, color = "black"),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 8),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.position = c(0.9, 0.1))
```

```{r, fig.height=7, fig.width=8}
ggplot(ps_RA_lfc_2.5, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```

### Grapth lfc of 3
```{r, fig.height=5, fig.width=7}
ggplot(ps_RA_lfc_3,
       aes(x = Soil_conditioning, y = Abundance, fill = Soil_conditioning)) +
  facet_wrap(~ ASV_Genus, scales = "free_y"#,
             #labeller = as_labeller(otu_labels)
             ) +
  stat_boxplot(aes(y = Abundance, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.3,    # horizontal jitter (x-direction)
                                              jitter.height = 0, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point", 
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment",
                    labels = c("CTRL", ~italic("M. brassicae"))) +
  labs(y = "Relative Abundance",
       x = "Soil Conditioning") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 5, hjust = 0),
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 8),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.position = "right")
```

```{r, fig.height=4, fig.width=8}
ggplot(ps_RA_lfc_3, 
       aes(y = ASV_Genus, x = Abundance, fill = Soil_conditioning)) +
  stat_boxplot(aes(x = Abundance, y = ASV_Genus), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               position = position_dodge(0.75),
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0,    # horizontal jitter (x-direction)
                                              #jitter.height = 0.2, # vertical jitter (y-direction, since flipped)
                                              dodge.width = 0.75),   # match boxplot dodge width
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(aes(group = Soil_conditioning),
               fun = mean, 
               geom = "point",
               position = position_dodge(0.75),
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white dimond
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +   # extra space (%) below and above min and max data point
  expand_limits(y = 0) + # start y axis at 0
  scale_fill_manual(values = c("#56B4E9", "#009E73"),
                    name = "Soil treatment") + # Manually change the colors of the bars. You need to define fill in second line of the plot script to 
  labs(y = "Differetnial Abundant ASVs",
       x = "Relative Abundance") + 
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```
```{r}
# Clean environment
rm(list = ls())
```