---
title: "FP 01_loading_and_pre_processing - for Differential Abundance"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

# 0.0 Introduction
This script is for loading and pre-processing microbiome data that is output form the DADA2 [Ernakovich pipeline](https://github.com/ErnakovichLab/dada2_ernakovichlab) as describes in our own [GitHub page](https://github.com/Marcelara/PSF_Anunna). In addition, the ASVs ware assigned to taxonomies with Qiime2 as this results in a higher resolution. Here, the data will be further processed to be able to analyse the data.  

A nice [tutorial](https://www.nicholas-ollberding.com/post/introduction-to-phyloseq/) to learn how to investigate data in a phyloseq object.

This and the next script are originally made by Pedro Beschoren da Costa for the [MeJA_Pilot](https://github.com/PedroBeschoren/MeJA_Pilot) and modified to fit my data.

# 1.0 load all libraries, custom functions, setup environemnt
```{r warning=FALSE, message=FALSE}
R.version$version.string # prints R version

# load libraries 
library(tidyr)
packageVersion("tidyr")
library(ggplot2)
packageVersion("ggplot2")
library(phyloseq)
packageVersion("phyloseq")
library(vegan)
packageVersion("vegan")
library(metagenomeSeq)
packageVersion("metagenomeSeq")
packageVersion("metagMisc")

# custom functions related to data loading and decontamination
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/RScripts/Functions/functions_loading_and_decontamination.R")

# increases memory limit used by R
# memory.limit(size = 350000) #This function is no longer supported by Windows
```


# 1.1 load microbiome data
See script of CP. Here we load the phyloseq object of the FP that is created in that script.

```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_raw_bac_ps.RData")
```


# 1.2 - Remove bad ASVs
Here we will remove sequences associated with the host, background prokatyotes, low abundance ASVs, poorly identified sequences, and other bad ASVs.

## 1.2.1 - Remove non-bacterial sequences & filter ASVs
Note that the input data has been pre-filtered in the DADA2 pipeline to remove any ASVs that occur less than 3 times in the data set. This was necessary because the dada2-associated taxonomy assignment tools could not assign taxonomies to the full data set with 1 GB ram on the HPC.

### Here we investigate the length of the sequences
```{r}
hist(as.data.frame(refseq(FP_raw_bac_ps)@ranges)$width, breaks = 300, main = "Raw reads CP bac")
summary(as.data.frame(refseq(FP_raw_bac_ps)@ranges)$width) #summary on length of the reeds
ntaxa(FP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(FP_raw_bac_ps)) #shows number of reeds
```

### Here we get rid of sequences shorter than 380bp
```{r}
sum(as.data.frame(refseq(FP_raw_bac_ps)@ranges)$width < 380) # count number of reads smaller than 380bp
sum(as.data.frame(refseq(FP_raw_bac_ps)@ranges)$width > 380)

FP_raw_bac_ps <- prune_taxa(taxa = as.data.frame(refseq(FP_raw_bac_ps)@ranges)$width > 380,
                       x = FP_raw_bac_ps)
ntaxa(FP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(FP_raw_bac_ps)) #shows number of reeds
```

### Keeps only ASVs identified as bacterial
```{r}
FP_raw_bac_ps <- subset_taxa(FP_raw_bac_ps, Kingdom == "k__Bacteria")
ntaxa(FP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(FP_raw_bac_ps)) #shows number of reeds
```

### Remove Salinibacter ASV(s)
```{r}
FP_raw_bac_ps #original data set
physeq_OnlySal <- subset_taxa(FP_raw_bac_ps, Genus == "g__Salinibacter")
physeq_OnlySal
FP_raw_bac_ps <- subset_taxa(FP_raw_bac_ps, Genus != "g__Salinibacter" | is.na(Genus)) # remove Salinibacter genus but keep undefined genesis (NAs)
FP_raw_bac_ps #data set without salinibacter

rm(physeq_OnlySal)
```

### Define library sizes as metadata before filtering
```{r}
FP_raw_bac_ps@sam_data$library_sizes_prefiltering <- sample_sums(FP_raw_bac_ps)
hist(sample_sums(FP_raw_bac_ps), breaks = 50)
```

### Removes taxa having less than 8 reads across all samples, as VSEARCH standard
<!-- ```{r} -->
<!-- otu_table(FP_raw_bac_ps) <- otu_table(FP_raw_bac_ps)[which (rowSums(otu_table(FP_raw_bac_ps)) > 7),] -->
<!-- ntaxa(FP_raw_bac_ps) #shows total number of ASVs -->
<!-- summary(sample_sums(FP_raw_bac_ps)) #shows number of reeds -->
<!-- ``` -->

### Remove ASV occurring in less than three samples
We remove ASVs that occur in less than 3 samples.
```{r}
filter <- phyloseq::genefilter_sample(FP_raw_bac_ps, filterfun_sample(function(x) x > 0), A = 3)
FP_raw_bac_ps <- prune_taxa(filter, FP_raw_bac_ps)
ntaxa(FP_raw_bac_ps) #shows total number of ASVs
summary(sample_sums(FP_raw_bac_ps)) #shows number of reeds
```

### Check and remove plant-host contamination
```{r}
### define plastid, mitochondria and host plant contamination ps objects
Mitochondria_ps <- subset_taxa(FP_raw_bac_ps, Family == "f__Mitochondria" | Family == "Mitochondria")
Plastid_ps <- subset_taxa(FP_raw_bac_ps, Order == "o__Chloroplast" | Order == "Chloroplast") 
host_plant_ps <- merge_phyloseq(Mitochondria_ps, Plastid_ps)

### quick histogram showing plant DNA contamination
hist(sample_sums(host_plant_ps)/sample_sums(FP_raw_bac_ps)*100, breaks = 50)
summary(sample_sums(host_plant_ps)/sample_sums(FP_raw_bac_ps)*100)

summary(sample_sums(Mitochondria_ps)/sample_sums(FP_raw_bac_ps)*100)
summary(sample_sums(Plastid_ps)/sample_sums(FP_raw_bac_ps)*100)
```

Between 0 and 2% of the ASVs per sample is contaminated with mitochondrial or chloroplast DNA.

```{r}
### define host plant 16S contamination as metadata (add the info the the metadata sheet)
FP_raw_bac_ps@sam_data$Mitochondria_reads <- sample_sums(Mitochondria_ps)
FP_raw_bac_ps@sam_data$Plastid_reads <- sample_sums(Plastid_ps)
FP_raw_bac_ps@sam_data$Host_DNA_n_reads <- sample_sums(host_plant_ps)
FP_raw_bac_ps@sam_data$Host_DNA_contamination_pct <- sample_sums(host_plant_ps)/sample_sums(FP_raw_bac_ps)*100

### remove plant host sequences (plastid and mitochondrial DNA) 
FP_raw_bac_ps <- remove_Chloroplast_Mitochondria(FP_raw_bac_ps)
```

### Check library size
```{r}
# add library sizes as part of metadata
sample_data(FP_raw_bac_ps)$library_size <- sample_sums(FP_raw_bac_ps)

#check library size distribution
hist(sample_data(FP_raw_bac_ps)$library_size, breaks = 50)
summary(sample_data(FP_raw_bac_ps)$library_size) #shows the number of reeds

# remove samples with library size = 0  (non-informative; failed PCR/sequencing)
#FP_raw_bac_ps<-subset_samples(FP_raw_bac_ps, library_size >0) #not applicable

# remove ps objects
rm(Mitochondria_ps, Plastid_ps, host_plant_ps)

# run garbage collection after creating large objects
gc()
```


# 1.3 - Decontaminate phyloseq objects
The decontam package will use blank DNA samples to remove possible contaminants. Here we use a single custom function to run decontamination, generate plots, and return a clean phyloseq object

We will also check the number of reads in a blank sample (average +SD) so we can compare those blank samples with libraries with low number of reads. samples that cannot be distinguished from a blank (in terms of library size) 

### Decontaminate
```{r}
FP_unnormalized_bac_ps <- decontaminate_and_plot(FP_raw_bac_ps)

# phyloseq object without contamination
FP_unnormalized_bac_ps[1]

# Number of ASVs assigned as contamination
FP_unnormalized_bac_ps[5]

# ASVs assigned as contamination
FP_unnormalized_bac_ps[4]
```

### Check blanks
Let's look into how many reads our blank had (average + sd) so we can compare those to libraries with very low number of reads.
```{r}
blank_reads_bac <- sort(sample_sums(subset_samples(FP_raw_bac_ps, Soil_conditioning == "B")))
plot(blank_reads_bac)

#clean up environment to free memory
rm(FP_raw_bac_ps)
```

### Check phyloseq object
After checking the decontamination plots and reports, we can remove them and focus on the phyloseq object.
```{r}
FP_unnormalized_bac_ps <- unlist(FP_unnormalized_bac_ps[[1]])
```

#### Check library size Bacteria
Now let's check the lower end of our library sizes.
```{r}
sort(sample_sums(FP_unnormalized_bac_ps))
plot(sort(sample_sums(FP_unnormalized_bac_ps)), xlab = "Number of Reads")
```

<!-- #### Remove samples with a number of reads similar to blanks -->
<!-- ```{r} -->
<!-- FP_unnormalized_bac_ps <- subset_samples(FP_unnormalized_bac_ps, library_size > 6000) -->
<!-- ``` -->

### Set factors
```{r}
# set Soil_conditioning as factor, and then order it properly
FP_unnormalized_bac_ps@sam_data$Soil_conditioning <- factor(FP_unnormalized_bac_ps@sam_data$Soil_conditioning, levels = c("Co", "Mb", "S"))

# set Cat_treatment as factor, and then order it properly
FP_unnormalized_bac_ps@sam_data$Cat_treatment <- factor(FP_unnormalized_bac_ps@sam_data$Cat_treatment, levels = c("Co", "Mb"))

# set Accession as factor, and then order it properly
FP_unnormalized_bac_ps@sam_data$Accession <- factor(FP_unnormalized_bac_ps@sam_data$Accession, levels = c("VL", "CD", "RI", "HM", "GO1", "Un"))

# set Batch as factor
FP_unnormalized_bac_ps@sam_data$Batch <- factor(FP_unnormalized_bac_ps@sam_data$Batch)
```


# 1.4 - Plot library sizes per treatment and check NAs and uncultured ASVs
### Check number of reeds
```{r}
# let's check the number of reads in a couple histograms
hist(sample_sums(FP_unnormalized_bac_ps), breaks = 50)
summary(sample_sums(FP_unnormalized_bac_ps))
```

### Check library size
```{r}
# check plot for some minimum library sizes on all samples

ggplot(data = sample_data(FP_unnormalized_bac_ps), 
           mapping = aes(x = library_size, y = Accession, fill = Soil_conditioning)) +
      geom_jitter() +
      geom_boxplot() +
      scale_y_discrete(limits = rev) +
      geom_vline(xintercept = 30000, color="green") +
      geom_vline(xintercept = 25000, color="yellow") +
      geom_vline(xintercept = 20000, color="blue") +
      geom_vline(xintercept = 15000, color="red") +
      labs(x = "library size",
           y = "Accession") #+
      #theme(axis.text.x = element_text(angle = 20, vjust = 0.5, hjust=1))

# # In case I want to do a subset of the data
# lapply(FP_unnormalized_bac_ps, function(x)
#     ggplot(data = sample_data(subset_samples(x, Compartment == "Rhizosphere")), 
#            mapping = aes(x = library_size, y = Soil_Slurry_Treatment, color = Soil_Slurry_Treatment, fill = Phase)) +
#     geom_jitter() +
#     geom_boxplot() +
#     scale_y_discrete(limits = rev) +
#     geom_vline(xintercept = 30000, color="green") +
#     geom_vline(xintercept = 25000, color="yellow") +
#     geom_vline(xintercept = 20000, color="blue") +
#     geom_vline(xintercept = 15000, color="red") +
#     labs(x = "library size",
#          y = "Soil slurry treatment and phase") +
#     theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)))
```

### Check plant-host contamination
```{r}
# check bacterial 16S plant host contamination
ggplot(data = sample_data(FP_unnormalized_bac_ps), 
           mapping = aes(x = Host_DNA_contamination_pct, y = Accession, fill = Soil_conditioning)) +
      geom_jitter() +
      geom_boxplot() +
      scale_y_discrete(limits = rev) +
      # geom_vline(xintercept = 30000, color="green") +
      # geom_vline(xintercept = 25000, color="yellow") +
      # geom_vline(xintercept = 20000, color="blue") +
      #geom_vline(xintercept = 15000, color="red") +
      labs(x = "Plant host DNA contamination",
           y = "Accession") #+
      #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

### check percentage of NA in taxonomy
```{r}
### check_n_taxa_NA_percentage(ps_object, ntaxa))
check_n_taxa_NA_percentage(FP_unnormalized_bac_ps, 100) # top 100 taxa
check_n_taxa_NA_percentage(FP_unnormalized_bac_ps, 1000)
check_n_taxa_NA_percentage(FP_unnormalized_bac_ps, 10000) # top 10,000 taxa
#A higher the number of taxa does not make a difference
```

### Check uncultured and unassigned taxa
Out commanded lines mean that type is not present in the data set.
```{r}
#subset_taxa(FP_unnormalized_bac_ps, Kingdom == "Unassigned") #No hits as I sub-setted the data with only k__Bacteria before

# subset_taxa(FP_unnormalized_bac_ps, Class == "c__uncultured" | Class == "c__unidentified")
# subset_taxa(FP_unnormalized_bac_ps, Class == "c__uncultured")
# subset_taxa(FP_unnormalized_bac_ps, Class == "c__unidentified") #not in data set

# subset_taxa(c__unidentified, Family == "f__uncultured" | Family == "f__unidentified")
# subset_taxa(FP_unnormalized_bac_ps, Family == "f__uncultured")
# subset_taxa(FP_unnormalized_bac_ps, Family == "f__unidentified") #not in data set

# subset_taxa(c__unidentified, Genus == "g__uncultured" | Genus == "g__unidentified")
# subset_taxa(FP_unnormalized_bac_ps, Genus == "g__uncultured")
# subset_taxa(FP_unnormalized_bac_ps, Genus == "g__unidentified") #not in data set
```

There are no uncultured or unidentified samples.

# 1.5 Remove samples?
## Rarefraction curve
```{r}
# finally, a rarefaction curve
a <- rarecurve(t(as.data.frame(otu_table(FP_unnormalized_bac_ps))), 
          label = FALSE, 
          step = 2000,
          main="Rarefaction", ylab = "Number of ASVs", xlab = "Number of DNA reads",
          abline(v = 10000, col="red", lwd=3, lty=2))
rm(a)

table(sample_data(FP_unnormalized_bac_ps)$Accession,
      sample_data(FP_unnormalized_bac_ps)$Soil_conditioning,
      sample_data(FP_unnormalized_bac_ps)$Cat_treatment)

sort(sample_sums(FP_unnormalized_bac_ps))[1:6]
sort(sample_sums(FP_unnormalized_bac_ps))[112:118]
```

#### Remove sample
```{r}
FP_unnormalized_bac_ps_cut <- subset_samples(FP_unnormalized_bac_ps, sample_names(FP_unnormalized_bac_ps) != "F371")
FP_unnormalized_bac_ps_ForDA <- subset_samples(FP_unnormalized_bac_ps, sample_names(FP_unnormalized_bac_ps) != "F178")
```

#### Remove Unplanted CTRL and Co caterpillar treatment
```{r}
FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA, Accession != "Un")
FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Cat_treatment != "Co")
```

#### Rename NAs tax table to Unclassfied
```{r}
tax <- tax_table(FP_unnormalized_bac_ps_ForDA_clean)
tax[is.na(tax[, "Genus"]), "Genus"] <- "g__Unclassified"
tax[is.na(tax[, "Family"]), "Family"] <- "f__Unclassified"
tax[is.na(tax[, "Order"]), "Order"] <- "o__Unclassified"
tax[is.na(tax[, "Class"]), "Class"] <- "c__Unclassified"
tax[is.na(tax[, "Phylum"]), "Phylum"] <- "p__Unclassified"
tax_table(FP_unnormalized_bac_ps_ForDA_clean) <- tax
```

#### Split data on accessions
```{r}
FP_unnormalized_bac_ps_ForDA_clean_Acc <- metagMisc::phyloseq_sep_variable(FP_unnormalized_bac_ps_ForDA_clean, variable = "Accession")
```

#### Remove Batch 1
```{r}
FP_unnormalized_bac_ps_ForDA_clean_Acc_b2 <- lapply(FP_unnormalized_bac_ps_ForDA_clean_Acc, function(x) subset_samples(x, Batch != 1))
FP_unnormalized_bac_ps_ForDA_clean_b2 <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Batch != 1)
```

# 1.6 export filtered unnormalized ps object
### Save RData
let's save these normalized objects as RData so they can be loaded in other scripts.
```{r}
save(FP_unnormalized_bac_ps_ForDA, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA.RData")
save(FP_unnormalized_bac_ps_ForDA_clean, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
save(FP_unnormalized_bac_ps_ForDA_clean_Acc, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean_Acc.RData")
save(FP_unnormalized_bac_ps_ForDA_clean_b2, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean_b2.RData")
save(FP_unnormalized_bac_ps_ForDA_clean_Acc_b2, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean_Acc_b2.RData")
```

# Data ready for analysis!
