---
title: "C and N content - 12 accession experiment FP"
author: "Kris"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---
# Introduction

   

# Start
## Import pre-made functions
```{r, warning = FALSE, message = FALSE}
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R") #Functions for comparing distributions, models and model assumptions
```
    
## Libraries
```{r, warning = FALSE, message = FALSE}
R.version$version.string # prints R version

library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(ggh4x) #for coloring strip facet wrap
packageVersion("ggh4x")
library(plyr) #for summary table
packageVersion("plyr")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")
```

## Loading data and processing
#### Loading data
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Metadata FP/Meta_Data_FP.csv")
```

#### Remove NAs
```{r}
data <- data[complete.cases(data[ , "C_content"]),]
```

#### Remove RI
```{r}
data <- data[data$Accession != "RI", ]
```

#### Defining factors
```{r}
data$Soil_conditioning <- as.factor(data$Soil_conditioning)
data$Accession <- factor(data$Accession, levels = c("VL", "CD", "HM", "GO1"))
```

# Data analysis
## Data exploration
```{r}
str(data[c(1:8, 51:54)])
head(data[c(1:8, 51:54)], 12)
summary(data[c(1:8, 51:54)]) #check for outliers
table(data$Batch, data$Accession, data$Soil_conditioning)
table(data$Accession, data$Soil_conditioning)
data[data$Accession == "GO1", c(1:3, 6, 51:53)] # Three samples of GO1 Co are from batch 1. There does not seem to be a batch effect.
``` 

## Summary table
### Summary table C
```{r}
ddply(data, c("Accession", "Soil_conditioning"), summarise,
      N = sum(!is.na(C_content)), #number of samples
      Mis = sum(is.na(C_content)), #number of missing values
      Mean = round(mean(C_content, na.rm=T),3), #mean
      Median = round(median(C_content, na.rm=T),3), #median
      SD = round(sd(C_content, na.rm=T),3), #standard deviation
      SE = round(sd(C_content, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval

ddply(data, c("Accession"), summarise,
      N = sum(!is.na(C_content)), #number of samples
      Mis = sum(is.na(C_content)), #number of missing values
      Mean = round(mean(C_content, na.rm=T),3), #mean
      Median = round(median(C_content, na.rm=T),3), #median
      SD = round(sd(C_content, na.rm=T),3), #standard deviation
      SE = round(sd(C_content, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval

ddply(data, c("Soil_conditioning"), summarise,
      N = sum(!is.na(C_content)), #number of samples
      Mis = sum(is.na(C_content)), #number of missing values
      Mean = round(mean(C_content, na.rm=T),3), #mean
      Median = round(median(C_content, na.rm=T),3), #median
      SD = round(sd(C_content, na.rm=T),3), #standard deviation
      SE = round(sd(C_content, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval

```
    
### Summary table N
```{r}
ddply(data, c("Accession", "Soil_conditioning"), summarise,
      N = sum(!is.na(N_content)), #number of samples
      Mis = sum(is.na(N_content)), #number of missing values
      Mean = round(mean(N_content, na.rm=T),3), #mean
      Median = round(median(N_content, na.rm=T),3), #median
      SD = round(sd(N_content, na.rm=T),3), #standard deviation
      SE = round(sd(N_content, na.rm=T) / sqrt(N),5), #standard error
      LNI = round(Mean - (2*SE),3), #lower confidence interval
      HNI = round(Mean + (2*SE),3)) # upper confidence interval

ddply(data, c("Accession"), summarise,
      N = sum(!is.na(N_content)), #number of samples
      Mis = sum(is.na(N_content)), #number of missing values
      Mean = round(mean(N_content, na.rm=T),3), #mean
      Median = round(median(N_content, na.rm=T),3), #median
      SD = round(sd(N_content, na.rm=T),3), #standard deviation
      SE = round(sd(N_content, na.rm=T) / sqrt(N),5), #standard error
      LNI = round(Mean - (2*SE),3), #lower confidence interval
      HNI = round(Mean + (2*SE),3)) # upper confidence interval

ddply(data, c("Soil_conditioning"), summarise,
      N = sum(!is.na(N_content)), #number of samples
      Mis = sum(is.na(N_content)), #number of missing values
      Mean = round(mean(N_content, na.rm=T),3), #mean
      Median = round(median(N_content, na.rm=T),3), #median
      SD = round(sd(N_content, na.rm=T),3), #standard deviation
      SE = round(sd(N_content, na.rm=T) / sqrt(N),5), #standard error
      LNI = round(Mean - (2*SE),3), #lower confidence interval
      HNI = round(Mean + (2*SE),3)) # upper confidence interval
```

### Summary table C_N_ratio
```{r}
ddply(data, c("Accession", "Soil_conditioning"), summarise,
      N = sum(!is.na(C_N_ratio)), #number of samples
      Mis = sum(is.na(C_N_ratio)), #number of missing values
      Mean = round(mean(C_N_ratio, na.rm=T),3), #mean
      Median = round(median(C_N_ratio, na.rm=T),3), #median
      SD = round(sd(N, na.rm=T),3), #standard deviation
      SE = round(sd(N, na.rm=T) / sqrt(N),5), #standard error
      LNI = round(Mean - (2*SE),3), #lower confidence interval
      HNI = round(Mean + (2*SE),3)) # upper confidence interval

ddply(data, c("Accession"), summarise,
      N = sum(!is.na(C_N_ratio)), #number of samples
      Mis = sum(is.na(C_N_ratio)), #number of missing values
      Mean = round(mean(C_N_ratio, na.rm=T),3), #mean
      Median = round(median(C_N_ratio, na.rm=T),3), #median
      SD = round(sd(N, na.rm=T),3), #standard deviation
      SE = round(sd(N, na.rm=T) / sqrt(N),5), #standard error
      LNI = round(Mean - (2*SE),3), #lower confidence interval
      HNI = round(Mean + (2*SE),3)) # upper confidence interval

ddply(data, c("Soil_conditioning"), summarise,
      N = sum(!is.na(C_N_ratio)), #number of samples
      Mis = sum(is.na(C_N_ratio)), #number of missing values
      Mean = round(mean(C_N_ratio, na.rm=T),3), #mean
      Median = round(median(C_N_ratio, na.rm=T),3), #median
      SD = round(sd(N, na.rm=T),3), #standard deviation
      SE = round(sd(N, na.rm=T) / sqrt(N),5), #standard error
      LNI = round(Mean - (2*SE),3), #lower confidence interval
      HNI = round(Mean + (2*SE),3)) # upper confidence interval
```


# Statistical analysis
## Testing model
### Making function
```{r}
lmTest <- function(x) {
  lm1 <- glmmTMB(data = data, formula = x ~ Accession * Soil_conditioning, family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm1)
  Output[[2]] <- Anova(lm1)
  names(Output) <- c("LM - Summary", "LM - Anova")
  print(Output)
}

lmlogTest <- function(x) {
  lm2 <- glmmTMB(data = data, formula = log(x) ~ Accession * Soil_conditioning, family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm2)
  Output[[2]] <- Anova(lm2)
  names(Output) <- c("LM with log transformed data - Summary", "LM with log transfomred data - Anova")
  print(Output)
}

lmSqrtTest <- function(x) {
  lm2 <- glmmTMB(data = data, formula = sqrt(x) ~ Accession * Soil_conditioning, family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm2)
  Output[[2]] <- Anova(lm2)
  names(Output) <- c("LM with sqrt transformed data - Summary", "LM with sqrt transfomred data - Anova")
  print(Output)
}

glmTest <- function(x) {
  glm1 <- glmmTMB(data = data, formula = (x) ~ Accession * Soil_conditioning, family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm1)
  Output[[2]] <- Anova(glm1)
  names(Output) <- c("GLM gamma - Summary", "GLM gamma Anova")
  print(Output)
}

glmlogTest <- function(x) {
  glm2 <- glmmTMB(data = data, formula = log(x) ~ Accession * Soil_conditioning, family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm2)
  Output[[2]] <- Anova(glm2)
  names(Output) <- c("GLM gamma with log transformed data - Summary", "GLM gamma with log transformed data - Anova")
  print(Output)
}

glmSqrtTest <- function(x) {
  glm2 <- glmmTMB(data = data, formula = sqrt(x) ~ Accession * Soil_conditioning, family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm2)
  Output[[2]] <- Anova(glm2)
  names(Output) <- c("GLM gamma with sqrt transformed data - Summary", "GLM gamma with sqrt transformed data - Anova")
  print(Output)
}
```

### Model output
```{r, results='hide'}
Col_number <- c(53:55) # columns of which I want statistical analysis
List_Stats_Leaf_CP <- list() # make an empty list to store output

for (i in Col_number) {
  List_Stats_Leaf_CP[[length(List_Stats_Leaf_CP) + 1]] <- lmTest(x = data[ , i])
  List_Stats_Leaf_CP[[length(List_Stats_Leaf_CP) + 1]] <- lmlogTest(x = data[ , i])
  List_Stats_Leaf_CP[[length(List_Stats_Leaf_CP) + 1]] <- lmSqrtTest(x = data[ , i])
  List_Stats_Leaf_CP[[length(List_Stats_Leaf_CP) + 1]] <- glmTest(x = data[ , i])
  List_Stats_Leaf_CP[[length(List_Stats_Leaf_CP) + 1]] <- glmlogTest(x = data[ , i])
  List_Stats_Leaf_CP[[length(List_Stats_Leaf_CP) + 1]] <- glmSqrtTest(x = data[ , i])

  print(List_Stats_Leaf_CP)
}
```
```{r}
names(List_Stats_Leaf_CP) <- paste(rep(colnames(data[53:55]), each = 6), rep(c(1:6), times = 3), sep = "")
print(List_Stats_Leaf_CP)
```

### Extract p-values
```{r}
stats_l <- list(N_content = List_Stats_Leaf_CP$N_content1$`LM - Anova`, 
                C_content = List_Stats_Leaf_CP$C_content5$`GLM gamma with log transformed data - Anova`,
                C_N_ratio = List_Stats_Leaf_CP$C_N_ratio6$`GLM gamma with sqrt transformed data - Anova`)

stats_df_l <- lapply(names(stats_l), function(x) {
  df <- as.data.frame(stats_l[[x]])
  df$Analysis <- x
  df$Factor <- rownames(stats_l[[x]])
  return(df)
})

Stat_CN_acc <- do.call(rbind, stats_df_l)[c(4,5,1:3)]
Stat_CN_acc

write.csv(Stat_CN_acc, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_CN_acc.csv")
```


## Model assumptions
### Making fuction
```{r, fig.width=10, fig.height=6}
lmFunc <- function(x) {
  lm <- glmmTMB(data = data, formula = x ~ Accession * Soil_conditioning, family = gaussian(link = "identity"))
}

lmlogFunc <- function(x) {
  lm <- glmmTMB(data = data, formula = log(x) ~ Accession * Soil_conditioning, family = gaussian(link = "identity"))
}

lmSqrtFunc <- function(x) {
  lm <- glmmTMB(data = data, formula = sqrt(x) ~ Accession * Soil_conditioning, family = gaussian(link = "identity"))
}

glmFunc <- function(x) {
  glm <- glmmTMB(data = data, formula = x ~ Accession * Soil_conditioning, family = Gamma(link = "log"))
}

glmlogFunc <- function(x) {
  glm <- glmmTMB(data = data, formula = log(x) ~ Accession * Soil_conditioning, family = Gamma(link = "log"))
}

glmSqrtFunc <- function(x) {
  glm <- glmmTMB(data = data, formula = sqrt(x) ~ Accession * Soil_conditioning, family = Gamma(link = "log"))
}
```

### Model output
```{r, fig.width=10, fig.height=7}
List_Asump_Leaf_CP <- list()

for (i in 1:length(Col_number)) {
  print(c(colnames(data)[Col_number[i]], "Linear model"))
    List_Asump_Leaf_CP[[length(List_Asump_Leaf_CP) + 1]] <- DHARMa.sum(lmFunc(x = data[ , Col_number[i]]))
  print(c(colnames(data)[Col_number[i]], "Linear model with log transformed data"))
  List_Asump_Leaf_CP[[length(List_Asump_Leaf_CP) + 1]] <- DHARMa.sum(lmlogFunc(x = data[ , Col_number[i]]))
  print(c(colnames(data)[Col_number[i]], "Linear model with sqrt transformed data"))
  List_Asump_Leaf_CP[[length(List_Asump_Leaf_CP) + 1]] <- DHARMa.sum(lmSqrtFunc(x = data[ , Col_number[i]]))
  print(c(colnames(data)[Col_number[i]], "GLM with gamma distribution "))
  List_Asump_Leaf_CP[[length(List_Asump_Leaf_CP) + 1]] <- DHARMa.sum(glmFunc(x = data[ , Col_number[i]]))
  print(c(colnames(data)[Col_number[i]], "GLM with gamma distribution and log transformed data"))
  List_Asump_Leaf_CP[[length(List_Asump_Leaf_CP) + 1]] <- DHARMa.sum(glmlogFunc(x = data[ , Col_number[i]]))
  print(c(colnames(data)[Col_number[i]], "GLM with gamma distribution and sqrt transformed data"))
  List_Asump_Leaf_CP[[length(List_Asump_Leaf_CP) + 1]] <- DHARMa.sum(glmSqrtFunc(x = data[ , Col_number[i]]))
}

names(List_Asump_Leaf_CP) <- rep(colnames(data[53:55]), each = 6)
```

### Post hoc test
```{r}
# N
emmeans(lmFunc(x = data[ , 53]), pairwise ~ Accession)
emmN <- cld(emmeans(lmFunc(x = data[ , 53]), pairwise ~ Accession), Letters = letters)
emmN

# C
emmeans(glmlogFunc(x = data[ , 54]), pairwise ~ Accession)
emmC <- cld(emmeans(glmlogFunc(x = data[ , 54]), pairwise ~ Accession), Letters = letters, decreasing = TRUE)
emmC

# c/N
emmeans(glmSqrtFunc(x = data[ , 55]), pairwise ~ Accession)
emmCN <- cld(emmeans(glmSqrtFunc(x = data[ , 55]), pairwise ~ Accession), Letters = letters, decreasing = TRUE)
emmCN
```

### Extracting ratios on log scale gamma distribution
only for C/N ratio
```{r}
# C/N
# Define function to extract ratios on log scale
EMM_mod <- emmeans(glmFunc(x = data[ , 55]), pairwise ~ Soil_conditioning | Accession, adjust = "tukey", type = "link")
EMM_mod_df <- as.data.frame(summary(EMM_mod$contrasts, infer = TRUE))

# Define factors
EMM_mod_df$Accession <- factor(EMM_mod_df$Accession, levels = c("VL", "CD", "HM", "GO1"))
EMM_mod_df
```

# Kruskal Wallis Test 
## Define funciton
```{r}
Kr_test <- function(x, factor) {
  kruskal.test(data = data, x ~ factor)
}
```

## Printing output
```{r}
fac <- c(2:3)
comb <- data.frame(Fac = rep(fac, each = 3),
                   Col_number = rep(Col_number, times = 2))

List_KrStats_Leaf_CP <- list() # make an empty list to store output

for (i in 1:length(comb$Col_number)) {
  List_KrStats_Leaf_CP[[i]] <- Kr_test(x = data[ , comb$Col_number[i]], 
                                       factor = data[ , comb$Fac[i]])
}
```
```{r}
names(List_KrStats_Leaf_CP) <- paste(colnames(data[comb$Fac]), 
                                     colnames(data[comb$Col_number]))
print(List_KrStats_Leaf_CP)
```

## Pairwise comparisons
### Define function
```{r}
Pw_Kr_test <- function(x, factor) {
  pairwise.wilcox.test(x, factor, p.adjust.method = "BH")
}
```

### Printing output
```{r, warning=FALSE}
List_PWKrStats_Leaf_CP <- list() # make an empty list to store output

for (i in 1:length(comb$Col_number)) {
  List_PWKrStats_Leaf_CP[[i]] <- Pw_Kr_test(x = data[ , comb$Col_number[i]], 
                                            factor = data[ , comb$Fac[i]])
}
```
```{r}
names(List_PWKrStats_Leaf_CP) <- paste(colnames(data[comb$Fac]), 
                                       colnames(data[comb$Col_number]))
print(List_PWKrStats_Leaf_CP)
```

## Plots
```{r}
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#0072B2", 1), rep("#D55E00", 4))),
                        text_x = elem_list_text(colour = c(rep("white", 1), rep("black", 4))))

Stats_C <- paste0("GLM: Accession p < 0.001; Soil p = 0.332; \nInteraction p = 0.868")
Stats_N <- paste0("LM: Accession p < 0.001; Soil p = 0.725; \nInteraction p = 0.494")
Stats_CN <- paste0("GLM: Accession p < 0.001; Soil p = 0.805; \nInteraction p = 0.269")

# C
# Significant letters in right order for stats Accession
emmC <- emmC[order(emmC$Accession), ]

# Alternate letters and blanks
Blank <- rep("", 4)
Letters_C <- c()

for (i in 1:4) {
  Letters_C[length(Letters_C) + 1] <- emmC$.group[i]
  Letters_C[length(Letters_C) + 1] <- Blank[i]
}

# Remove spaces around letters
Letters_C <- gsub(pattern = " ", replacement = "", x = Letters_C)
Letters_C

DataSum_C <- data.frame(Accession = rep(emmC$Accession, each = 2),
                        Soil_conditioning = rep(c("Co", "Mb"), 2),
                        Letters = Letters_C)

# N
# Significant letters in right order for stats Accession
emmN <- emmN[order(emmN$Accession), ]

# Alternate letters and blanks
Blank <- rep("", 4)
Letters_N <- c()

for (i in 1:4) {
  Letters_N[length(Letters_N) + 1] <- emmN$.group[i]
  Letters_N[length(Letters_N) + 1] <- Blank[i]
}

# Remove spaces around letters
Letters_N <- gsub(pattern = " ", replacement = "", x = Letters_N)
Letters_N

DataSum_N <- data.frame(Accession = rep(emmN$Accession, each = 2),
                        Soil_conditioning = rep(c("Co", "Mb"), 2),
                        Letters = Letters_N)

# C/N
# Significant letters in right order for stats Accession
emmCN <- emmCN[order(emmCN$Accession), ]

# Alternate letters and blanks
Blank <- rep("", 4)
Letters_CN <- c()

for (i in 1:4) {
  Letters_CN[length(Letters_CN) + 1] <- emmCN$.group[i]
  Letters_CN[length(Letters_CN) + 1] <- Blank[i]
}

# Remove spaces around letters
Letters_CN <- gsub(pattern = " ", replacement = "", x = Letters_CN)
Letters_CN

DataSum_CN <- data.frame(Accession = rep(emmCN$Accession, each = 2),
                        Soil_conditioning = rep(c("Co", "Mb"), 2),
                        Letters = Letters_CN)
```

### Plot data C
```{r, fig.width=4, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = C_content, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = C_content, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 43, xend = 2.2, yend = 43)) +
  geom_text(data = DataSum_C,
            aes(label = Letters_C, y = 45.5, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$C_content*1.15))) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("C content leaves (%)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats_C) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 4.2,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_C_content.svg", "PF_C_content.png"))
```

#### Not starting at 0
```{r, fig.width=4, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = C_content, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = C_content, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 41.3, xend = 2.2, yend = 41.3)) +
  geom_text(data = DataSum_C,
            aes(label = Letters_C, y = 41.6, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  #scale_y_continuous(#expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     #limits = c(0, max(data$C_content*1.15))) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("C content leaves (%)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats_C) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3.9,
         height = 3.9,
         units = "cm",
         dpi = 300),
  x = c("PF_C_content2.svg", "PF_C_content2.png"))
```

### Plot data N
```{r, fig.width=4, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = N_content, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = N_content, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 3.9, xend = 2.2, yend = 3.9)) +
  geom_text(data = DataSum_N,
            aes(label = Letters_N, y = 4.05, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$N_content*1.15))) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("N content leaves (%)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats_N) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),        
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3.7,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_N_content.svg", "PF_N_content.png"))
```

### Plot data C_N_ratio
```{r, fig.width=4, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = C_N_ratio, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = C_N_ratio, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 35, xend = 2.2, yend = 35)) +
  geom_text(data = DataSum_CN,
            aes(label = Letters_CN, y = 37, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$C_N_ratio*1.15))) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("C/N ratio leaves (%)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats_CN) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), 
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3.7,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_CN_ratio.svg", "PF_CN_ratio.png"))

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3.7,
         height = 3.5,
         units = "cm",
         dpi = 300),
  x = c("PF_CN_ratio_small.svg", "PF_CN_ratio_small.png"))
```

### Figure PSF effect C/N ratio
```{r, fig.width=5, fig.height=3}
ggplot(data = EMM_mod_df,
       mapping = aes(y = estimate, x = Accession)) + #Defines the data
  geom_hline(yintercept = 0, 
             colour = "gray30") +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), 
                color = c(rep("#0072B2", 1), rep("#D55E00", 3)),
                linewidth = 1,
                width = 0.3) +
  geom_point(aes(y = estimate, x = Accession),
             size = 2) +
  # geom_text(data = stats_PSF,
  #           aes(label = Stats, y = 0.065, x = Accession), # fill out the significant letters
  #           size = 8,
  #           #hjust = 0.5,
  #           #position = position_dodge(2)
  #           ) +
  # scale_y_continuous(breaks = seq(-0.03, 0.06, 0.03),
  #                    limits = c(-0.05, 0.07)) + #steps on the y-axis
  ggtitle(Stats_CN) +
  ylab("PSF ratio C/N content") +    #y-label with Mb in italic
  xlab("Accession") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3,
         height = 3,
         units = "cm",
         dpi = 300),
  x = c("PF_PSF_ratio_CN-ratio.svg", "PF_PSF_ratio_CN-ratio.png"))
```