---
title: "12 accession Exp - Caterpillar weight FP"
author: "Kris"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---
  
# Description

   
# Start
### Import pre-made functions
```{r, warning = FALSE, message = FALSE}
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R") #Functions for comparing distributions, models and model assumptions
#source("C:/Users/K/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R")
```
  
### Libraries
```{r, warning = FALSE, message = FALSE}
R.version$version.string # prints R version

library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(ggpubr) #graphs
packageVersion("ggpubr")
library(patchwork) # for wrap_plot
packageVersion("patchwork")
library(plyr) #for summary table
packageVersion("plyr")
library(ggh4x) #for coloring strip facet wrap
packageVersion("ggh4x")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")
library(lme4) #for glm with random factor as alternative to glmmTMB
packageVersion("lme4")
library(DHARMa) #for model assumptions
packageVersion("DHARMa")
library(cowplot) #for printing legend
packageVersion("cowplot")
```
  
## Loading data and processing
### Loading data  
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Metadata FP/Meta_Data_FP.csv")
```
  
### Removing NAs and sterile treatment 
```{r}
data <- data[complete.cases(data$Mean_Weight_Cat), ]
data <- data[data$Soil_conditioning != "S", ]
```
  
### Remove Batch 1 for all accessions
```{r}
# cn <- c(which(data$Accession == "VL" & data$Batch == 1), which(data$Accession == "HM" & data$Batch == 1))
# data <- data[ -cn, ]
data <- data[data$Batch != 1, ]
```
  
### Defining factors  
```{r}
data$Soil_conditioning <- as.factor(data$Soil_conditioning)
data$Batch <- as.factor(data$Batch)
data$Accession <- factor(data$Accession, levels = c("OH", "DD", "HE", "KI", "VL", "CD", "RI", "KT", "MC", "HM", "IT1", "GO1"))
```
  
# Exploring data
```{r}
str(data)
head(data, 12)
summary(data) #check for outliers
table(data$Block_g, data$Accession, data$Soil_conditioning)
```

## Summary table
### Accession
```{r}
DataSum <- ddply(data, c("Accession", "Soil_conditioning"), summarize,
            N = sum(!is.na(Mean_Weight_Cat)), #number of samples
            Mis = sum(is.na(Mean_Weight_Cat)), #number of missing values
            Mean = round(mean(Mean_Weight_Cat, na.rm=T),3), #mean
            Median = round(median(Mean_Weight_Cat, na.rm=T),3), #median
            SD = round(sd(Mean_Weight_Cat, na.rm=T),3), #standard deviation
            SE = round(sd(Mean_Weight_Cat, na.rm=T) / sqrt(N),5), #standard error
            LCI = round(Mean - (2*SE),3), #lower confidence interval
            HCI = round(Mean + (2*SE),3)) # upper confidence interval
print(DataSum)
```

## Plot 
```{r, fig.width=8, fig.height=4}
# Defining color for strip in facet wrap
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#E69F00", 5), rep("#D55E00", 7))))

ggplot(data = data,
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Accession, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean,
               geom = "point",
               size = 1.5,
               color = "white",
               shape = 18) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, max(data$Mean_Weight_Cat*1.1))) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Dry caterpillar weight") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```

## Plot all accessions individually
### Without batch
#### Defining function
```{r}
Graph_Acc2 <- function(x){
ggplot(data = data[data$Accession == x, ],
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "black", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, max(data[data$Accession == x, "Mean_Weight_Cat"])*1.1)) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab(paste("Dry caterpillar weight on", x)) +    #y-label with Mb in italic
  xlab("Soil treatment") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
}
```

#### Making plots with for loop
```{r}
Acc <- c("OH", "DD", "HE", "KI", "VL", "CD", "RI", "KT", "MC", "HM", "IT1", "GO1")
List_Graph_Acc2 <- list() # make an empty list to store the plots

for (i in 1:length(Acc)) {
  List_Graph_Acc2[[i]] <- Graph_Acc2(x = Acc[i]) # store plot in list
}
```

#### Print plots
```{r, fig.height=7, fig.width=12}
wrap_plots(List_Graph_Acc2, ncol = 6) #, axis_titles = 'collect_x'
```


# Statistical analysis all accessions together
### Model
```{r}
lm1 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Accession*Soil_conditioning, family = gaussian(link = "identity"))
lm2 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Accession*Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
lm3 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat) ~ Accession*Soil_conditioning, family = gaussian(link = "identity"))
lm4 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat) ~ Accession*Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
lm5 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Accession*Soil_conditioning, family = gaussian(link = "identity"))
lm6 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Accession*Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))

glm1 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Accession*Soil_conditioning, family = Gamma(link = "log"))
glm2 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Accession*Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
glm3 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat + 1) ~ Accession*Soil_conditioning, family = Gamma(link = "log"))
glm4 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat + 1) ~ Accession*Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
glm5 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Accession*Soil_conditioning, family = Gamma(link = "log"))
glm6 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Accession*Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
```

### Model output
```{r}
# summary(lm1)
# Anova(lm1)
# summary(lm2)
# Anova(lm2)
# summary(lm3)
# Anova(lm3)
# summary(lm4)
# Anova(lm4)
# summary(lm5)
# Anova(lm5)
summary(lm6)
an_lm6 <- Anova(lm6)
an_lm6
Stat_catweight_acc <- as.data.frame(an_lm6)
write.csv(Stat_catweight_acc, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_catweight_acc.csv")

# summary(glm1)
# Anova(glm1)
# summary(glm2)
# Anova(glm2)
# summary(glm3)
# Anova(glm3)
# summary(glm4)
# Anova(glm4)
# summary(glm5)
# Anova(glm5)
# summary(glm6)
# Anova(glm6)
```

### Model assumptions
```{r, fig.width=10, fig.height=7}
DHARMa.sum(Model = lm1)
DHARMa.sum(Model = lm2)
DHARMa.sum(Model = lm3)
DHARMa.sum(Model = lm4)
DHARMa.sum(Model = lm5)
DHARMa.sum(Model = lm6)

DHARMa.sum(Model = glm1)
DHARMa.sum(Model = glm2)
DHARMa.sum(Model = glm3)
DHARMa.sum(Model = glm4)
DHARMa.sum(Model = glm5)
DHARMa.sum(Model = glm6)
```

### Post hoc test
```{r}
#EMMA1 <- emmeans(lm6, pairwise ~ Accession, adjust = "fdr", type = "response")
EMMA2 <- emmeans(lm6, pairwise ~ Accession, adjust = "tukey", type = "response") # more stringed
EMA <- cld(EMMA2, Letters = letters)
EMA
```
   
https://aosmith.rbind.io/2019/03/25/getting-started-with-emmeans/ (Nice website!)   
https://cran.r-project.org/web/packages/emmeans/vignettes/comparisons.html#contents   
https://cran.r-project.org/web/packages/emmeans/vignettes/basics.html#contents   
https://people.math.ethz.ch/~meier/teaching/anova/contrasts-and-multiple-testing.html#multiple-testing   
help("contrast-methods")   
help("summary.emmGrid") describes all adjustment methods #"holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", and "none" and more).   
    
# Statistical analysis per accession
## Testing model
### Making function
```{r}
lmTest <- function(x) {
  lm1 <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm1)
  Output[[2]] <- Anova(lm1)
  names(Output) <- c("LMM - Summary", "LMM - Anova")
  print(Output)
}

lmlogTest <- function(x) {
  lm2 <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm2)
  Output[[2]] <- Anova(lm2)
  names(Output) <- c("LMM with log transformed data - Summary", "LMM with log transfomred data - Anova")
  print(Output)
}

lmSqrtTest <- function(x) {
  lm2 <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm2)
  Output[[2]] <- Anova(lm2)
  names(Output) <- c("LMM with sqrt transformed data - Summary", "LMM with sqrt transfomred data - Anova")
  print(Output)
}

glmTest <- function(x) {
  glm1 <- glmmTMB(data = data[data$Accession == x, ], formula = (Mean_Weight_Cat) ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm1)
  Output[[2]] <- Anova(glm1)
  names(Output) <- c("GLMM gamma - Summary", "GLMM gamma Anova")
  print(Output)
}

glmlogTest <- function(x) {
  glm2 <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm2)
  Output[[2]] <- Anova(glm2)
  names(Output) <- c("GLMM gamma with log transformed data - Summary", "GLMM gamma with log transformed data - Anova")
  print(Output)
}

glmSqrtTest <- function(x) {
  glm2 <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm2)
  Output[[2]] <- Anova(glm2)
  names(Output) <- c("GLMM gamma with sqrt transformed data - Summary", "GLMM gamma with sqrt transformed data - Anova")
  print(Output)
}


lmTest2 <- function(x) {
  lm1 <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning, family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm1)
  Output[[2]] <- Anova(lm1)
  names(Output) <- c("LM - Summary", "LM - Anova")
  print(Output)
}

lmlogTest2 <- function(x) {
  lm2 <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning, family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm2)
  Output[[2]] <- Anova(lm2)
  names(Output) <- c("LM with log transformed data - Summary", "LM with log transfomred data - Anova")
  print(Output)
}

lmSqrtTest2 <- function(x) {
  lm2 <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning, family = gaussian(link = "identity"))
  Output <- list()
  Output[[1]] <- summary(lm2)
  Output[[2]] <- Anova(lm2)
  names(Output) <- c("LM with sqrt transformed data - Summary", "LM with sqrt transfomred data - Anova")
  print(Output)
}

glmTest2 <- function(x) {
  glm1 <- glmmTMB(data = data[data$Accession == x, ], formula = (Mean_Weight_Cat) ~ Soil_conditioning, family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm1)
  Output[[2]] <- Anova(glm1)
  names(Output) <- c("GLM gamma - Summary", "GLM gamma Anova")
  print(Output)
}

glmlogTest2 <- function(x) {
  glm2 <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning, family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm2)
  Output[[2]] <- Anova(glm2)
  names(Output) <- c("GLM gamma with log transformed data - Summary", "GLM gamma with log transformed data - Anova")
  print(Output)
}

glmSqrtTest2 <- function(x) {
  glm2 <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning, family = Gamma(link = "log"))
  Output <- list()
  Output[[1]] <- summary(glm2)
  Output[[2]] <- Anova(glm2)
  names(Output) <- c("GLM gamma with sqrt transformed data - Summary", "GLM gamma with sqrt transformed data - Anova")
  print(Output)
}
```

### Model output
```{r, results='hide'}
List_Stats_Acc <- list() # make an empty list to store output

for (i in 1:length(Acc)) {
  List_Stats_Acc[[(i)*12-11]] <- lmTest2(x = Acc[i])
  List_Stats_Acc[[(i)*12-10]] <- lmTest(x = Acc[i])
  List_Stats_Acc[[(i)*12-9]] <- lmlogTest2(x = Acc[i])
  List_Stats_Acc[[(i)*12-8]] <- lmlogTest(x = Acc[i])
  List_Stats_Acc[[(i)*12-7]] <- lmSqrtTest2(x = Acc[i])
  List_Stats_Acc[[(i)*12-6]] <- lmSqrtTest(x = Acc[i])
  List_Stats_Acc[[(i)*12-5]] <- glmTest2(x = Acc[i])
  List_Stats_Acc[[(i)*12-4]] <- glmTest(x = Acc[i])
  List_Stats_Acc[[(i)*12-3]] <- glmlogTest2(x = Acc[i])
  List_Stats_Acc[[(i)*12-2]] <- glmlogTest(x = Acc[i])
  List_Stats_Acc[[(i)*12-1]] <- glmSqrtTest2(x = Acc[i])
  List_Stats_Acc[[(i)*12]] <- glmSqrtTest(x = Acc[i])
}
```
```{r}
names(List_Stats_Acc) <- paste(rep(Acc, each = 12), rep(1:12), sep = "_")
print(List_Stats_Acc)
```

### Extract p-values
#### Best fitting models
```{r}
stats_l <- list(List_Stats_Acc$OH_6$`LMM with sqrt transfomred data - Anova`, 
                List_Stats_Acc$DD_4$`LMM with log transfomred data - Anova`, 
                List_Stats_Acc$HE_10$`GLMM gamma with log transformed data - Anova`, 
                List_Stats_Acc$KI_12$`GLMM gamma with sqrt transformed data - Anova`, 
                List_Stats_Acc$VL_4$`LMM with log transfomred data - Anova`, 
                List_Stats_Acc$CD_12$`GLMM gamma with sqrt transformed data - Anova`, 
                List_Stats_Acc$RI_10$`GLMM gamma with log transformed data - Anova`, 
                List_Stats_Acc$KT_10$`GLMM gamma with log transformed data - Anova`, 
                List_Stats_Acc$MC_6$`LMM with sqrt transfomred data - Anova`, 
                List_Stats_Acc$HM_8$`GLMM gamma Anova`, 
                List_Stats_Acc$IT1_8$`GLMM gamma Anova`, 
                List_Stats_Acc$GO1_10$`GLMM gamma with log transformed data - Anova`)
names(stats_l) <- Acc

stats_df_l <- lapply(names(stats_l), function(x) {
  df <- as.data.frame(stats_l[[x]])
  df$Accession <- x
  df$Factor <- rownames(stats_l[[x]])
  return(df)
})

Stat_catweight_peracc <- do.call(rbind, stats_df_l)[c(4,5,1:3)]
Stat_catweight_peracc

#write.csv(Stat_catweight_peracc, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_catweight_peracc.csv")
```

#### Models with gamma distribution
These models fit also very well for all accessions. This distribution is used for calculating PSF ratios.   
```{r}
names(List_Stats_Acc) <- rep(Acc, each = 12)

stats_l2 <- lapply(Acc, function(x) {
  List_Stats_Acc[[x]]$`GLMM gamma`
})

names(stats_l2) <- Acc

stats_df_l2 <- lapply(names(stats_l2), function(x) {
  df <- as.data.frame(stats_l[[x]])
  df$Accession <- x
  df$Factor <- rownames(stats_l[[x]])
  return(df)
})

Stat_catweight_peracc <- do.call(rbind, stats_df_l)[c(4,5,1:3)]
Stat_catweight_peracc

write.csv(Stat_catweight_peracc, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_catweight_peracc.csv")
```


## Model assumptions
### Making fuction
```{r, fig.width=10, fig.height=6}
lmFunc <- function(x) {
  lm <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
}

lmlogFunc <- function(x) {
  lm <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
}

lmSqrtFunc <- function(x) {
  lm <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))
}

glmFunc <- function(x) {
  glm <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
}

glmlogFunc <- function(x) {
  glm <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
}

glmSqrtFunc <- function(x) {
  glm <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
}

lmFunc2 <- function(x) {
  lm <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning, family = gaussian(link = "identity"))
}

lmlogFunc2 <- function(x) {
  lm <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning, family = gaussian(link = "identity"))
}

lmSqrtFunc2 <- function(x) {
  lm <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning, family = gaussian(link = "identity"))
}

glmFunc2 <- function(x) {
  glm <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning, family = Gamma(link = "log"))
}

glmlogFunc2 <- function(x) {
  glm <- glmmTMB(data = data[data$Accession == x, ], formula = log(Mean_Weight_Cat + 2) ~ Soil_conditioning, family = Gamma(link = "log"))
}

glmSqrtFunc2 <- function(x) {
  glm <- glmmTMB(data = data[data$Accession == x, ], formula = sqrt(Mean_Weight_Cat) ~ Soil_conditioning, family = Gamma(link = "log"))
}
```

### Model output all models
```{r, fig.width=10, fig.height=7}
List_Asump_Acc <- list()

for (i in 1:length(Acc)) {
  print(c(Acc[i], "Linear model"))
  List_Asump_Acc[[i*12-11]] <- DHARMa.sum(lmFunc2(x = Acc[i]))
  print(c(Acc[i], "Linear mixed model"))
  List_Asump_Acc[[i*12-10]] <- DHARMa.sum(lmFunc(x = Acc[i]))
  print(c(Acc[i], "Linear model with log transformed data"))
  List_Asump_Acc[[i*12-9]] <- DHARMa.sum(lmlogFunc2(x = Acc[i]))
  print(c(Acc[i], "Linear mixed model with log transformed data"))
  List_Asump_Acc[[i*12-8]] <- DHARMa.sum(lmlogFunc(x = Acc[i]))
  print(c(Acc[i], "Linear model with sqrt transformed data"))
  List_Asump_Acc[[i*12-7]] <- DHARMa.sum(lmSqrtFunc2(x = Acc[i]))
  print(c(Acc[i], "Linear mixed model with sqrt transformed data"))
  List_Asump_Acc[[i*12-6]] <- DHARMa.sum(lmSqrtFunc(x = Acc[i]))
  print(c(Acc[i], "GLM with gamma distribution"))
  List_Asump_Acc[[i*12-5]] <- DHARMa.sum(glmFunc2(x = Acc[i]))
  print(c(Acc[i], "GLMM with gamma distribution"))
  List_Asump_Acc[[i*12-4]] <- DHARMa.sum(glmFunc(x = Acc[i]))
  print(c(Acc[i], "GLM with gamma distribution and log transformed data"))
  List_Asump_Acc[[i*12-3]] <- DHARMa.sum(glmlogFunc2(x = Acc[i]))
  print(c(Acc[i], "GLMM with gamma distribution and log transformed data"))
  List_Asump_Acc[[i*12-2]] <- DHARMa.sum(glmlogFunc(x = Acc[i]))
  print(c(Acc[i], "GLM with gamma distribution and sqrt transformed data"))
  List_Asump_Acc[[i*12-1]] <- DHARMa.sum(glmSqrtFunc2(x = Acc[i]))
  print(c(Acc[i], "GLMM with gamma distribution and sqrt transformed data"))
  List_Asump_Acc[[i*12]] <- DHARMa.sum(glmSqrtFunc(x = Acc[i]))
}

names(List_Asump_Acc) <- rep(Acc, each = 12)
```

### Model output log scale models
```{r, fig.width=10, fig.height=7}
List_Asump_Acc_log <- list()

for (i in 1:length(Acc)) {
  print(c(Acc[i], "Linear model with log transformed data"))
  List_Asump_Acc_log[[i*4-3]] <- DHARMa.sum(lmlogFunc2(x = Acc[i]))
  print(c(Acc[i], "Linear mixed model with log transformed data"))
  List_Asump_Acc_log[[i*4-2]] <- DHARMa.sum(lmlogFunc(x = Acc[i]))
  print(c(Acc[i], "GLM with gamma distribution"))
  List_Asump_Acc_log[[i*4-1]] <- DHARMa.sum(glmFunc2(x = Acc[i]))
  print(c(Acc[i], "GLMM with gamma distribution"))
  List_Asump_Acc_log[[i*4]] <- DHARMa.sum(glmFunc(x = Acc[i]))
}

names(List_Asump_Acc_log) <- rep(Acc, each = 4)
```

## Extracting ratios on log scale gamma distribution
```{r}
# Define function to extract ratios on log scale
PSF_func <- function(x) {
  glmm <- glmmTMB(data = data[data$Accession == x, ], formula = Mean_Weight_Cat ~ Soil_conditioning + (1|Block_g), family = Gamma(link = "log"))
  EMM_mod <- emmeans(glmm, pairwise ~ Soil_conditioning, adjust = "tukey", type = "link")
  EMM_mod_df <- as.data.frame(summary(EMM_mod$contrasts, infer = TRUE))
  #EMM_mod_est <- EMM_mod_df$estimate
  return(EMM_mod_df)
}

# Extract ratios on log scale
PSF_ratio <- lapply(Acc, function(x) {
  df <- PSF_func(x)
  df$Accession <- x
  return(df)
})

# combine list into one data frame
PSF_ratio_df <- do.call(rbind, PSF_ratio)

# Define factors
PSF_ratio_df$Accession <- factor(PSF_ratio_df$Accession, levels = Acc)
PSF_ratio_df
```


# Final figure
### Data preperation
```{r, fig.width=10, fig.height=4}
# Alter color facet wrap label
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#0072B2", 5), rep("#D55E00", 7))),
                        text_x = elem_list_text(colour = c(rep("white", 5), rep("black", 7))))

# Statistical info main effects
Stats <- paste0("\nLMM: Accession p < 0.001; Soil treatment p = 0.815; Interaction p = 0.212")

# Significant letters in right order for stats Accession
EMA <- EMA[order(EMA$Accession), ]

# Alternate letters and blanks
Blank <- rep("", 12)
Letters <- c()

for (i in 1:12) {
  Letters[length(Letters) + 1] <- EMA$.group[i]
  Letters[length(Letters) + 1] <- Blank[i]
}

# Remove spaces around letters
Letters <- gsub(pattern = " ", replacement = "", x = Letters)
Letters

DataSum <- data.frame(Accession = rep(EMA$Accession, each = 2),
                      Soil_conditioning = rep(c("Co", "Mb"), 12),
                      Letters = Letters)

# Prepare stars and dots for statistical differences soil conditioning
DataSum$lab2 <- c(rep("", 8), ".\n", rep("", 5), "*", rep("", 3), "*", rep("", 5))
```

### Figure effect of soil conditioning
```{r, fig.width=10, fig.height=4}
ggplot(data = data,
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Accession, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  # geom_segment(aes(x = 0.8, y = 86, xend = 2.2, yend = 86)) +
  # geom_text(data = DataSum,
  #           aes(label = lab, y = 90), # fill out the significant letters
  #           size = 4,
  #           hjust = 0.5,
  #           #position = position_dodge(0.5)
  #          ) +
  geom_text(data = DataSum,
            aes(label = lab2, y = 24), # fill out the significant letters
            size = 8,
            hjust = -1.1,
            #position = position_dodge(0.5)
           ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, 26.8)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Dry caterpillar weight (mg)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

### Differences between accessions
```{r, fig.width=10, fig.height=4}
ggplot(data = data,
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Accession, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 25, xend = 2.2, yend = 25)) +
  geom_text(data = DataSum,
            aes(label = Letters, y = 26.5, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, 26.8)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Dry caterpillar weight (mg)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

### Differences between accessions and effect of soil conditioning
```{r, fig.width=10, fig.height=4}
ggplot(data = data,
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Accession, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 26.5, xend = 2.2, yend = 26.5)) +
  geom_text(data = DataSum,
            aes(label = Letters, y = 28, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  geom_text(data = DataSum,
            aes(label = lab2, y = 23.5), # fill out the significant letters
            size = 8,
            hjust = -1.1,
            #position = position_dodge(0.5)
           ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     breaks = seq(0, 25, 5), #steps on the y-axis
                     limits = c(0, 29.2)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Dry caterpillar weight (mg)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 10,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_CatWeigh_Acc.svg", "PF_CatWeigh_Acc.png"))
```

### Figure PSF effect
```{r, fig.width=5, fig.height=3}
ggplot(data = PSF_ratio_df,
       mapping = aes(y = estimate, x = Accession)) + #Defines the data
  geom_hline(yintercept = 0, 
             colour = "gray30") +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), 
                color = c(rep("#0072B2", 5), rep("#D55E00", 7)),
                linewidth = 1,
                width = 0.3) +
  geom_point(aes(y = estimate, x = Accession),
             size = 2) +
  geom_text(data = DataSum,
            aes(label = lab2, y = 0.85), # fill out the significant letters
            size = 8) +
  ylab("PSF ratio caterpillar weight") +    #y-label with Mb in italic
  xlab("Accession") +
  ggtitle("\nGLMM (Table S8)") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 6.5,
         height = 3,
         units = "cm",
         dpi = 300),
  x = c("PF_PSF_ratio_CatWeigh_Acc.svg", "PF_PSF_ratio_CatWeigh_Acc.png"))
```