---
title: "12 accession Exp - Caterpillar weight FP - Wild vs Domesticated"
author: "Kris"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---
  
# Description

   
# Start
### Import pre-made functions
```{r, warning = FALSE, message = FALSE}
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R") #Functions for comparing distributions, models and model assumptions
#source("C:/Users/K/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R")
```
  
### Libraries
```{r, warning = FALSE, message = FALSE}
R.version$version.string # prints R version

library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(ggpubr) #graphs
packageVersion("ggpubr")
library(patchwork) # for wrap_plot
packageVersion("patchwork")
library(plyr) #for summary table
packageVersion("plyr")
library(ggh4x) #for coloring strip facet wrap
packageVersion("ggh4x")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")
library(lme4) #for glm with random factor as alternative to glmmTMB
packageVersion("lme4")
library(DHARMa) #for model assumptions
packageVersion("DHARMa")
library(cowplot) #for printing legend
packageVersion("cowplot")
```
  
## Loading data and processing
### Loading data  
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Metadata FP/Meta_Data_FP.csv")
```
  
### Removing NAs and sterile treatment 
```{r}
data <- data[complete.cases(data$Mean_Weight_Cat), ]
data <- data[data$Soil_conditioning != "S", ]
```
  
### Remove Batch 1 for all accessions
```{r}
# cn <- c(which(data$Domestication == "VL" & data$Batch == 1), which(data$Domestication == "HM" & data$Batch == 1))
# data <- data[ -cn, ]
data <- data[data$Batch != 1, ]
```
  
### Defining factors  
```{r}
data$Soil_conditioning <- as.factor(data$Soil_conditioning)
data$Batch <- as.factor(data$Batch)
data$Domestication <- factor(data$Domestication, levels = c("Wild", "Cultivated"))
```
  
# Exploring data
```{r}
str(data)
head(data, 12)
summary(data) #check for outliers
table(data$Block_g, data$Domestication, data$Soil_conditioning)
```


## Summary table
### Domestication
```{r}
DataSum <- ddply(data, c("Domestication", "Soil_conditioning"), summarize,
            N = sum(!is.na(Mean_Weight_Cat)), #number of samples
            Mis = sum(is.na(Mean_Weight_Cat)), #number of missing values
            Mean = round(mean(Mean_Weight_Cat, na.rm=T),3), #mean
            Median = round(median(Mean_Weight_Cat, na.rm=T),3), #median
            SD = round(sd(Mean_Weight_Cat, na.rm=T),3), #standard deviation
            SE = round(sd(Mean_Weight_Cat, na.rm=T) / sqrt(N),5), #standard error
            LCI = round(Mean - (2*SE),3), #lower confidence interval
            HCI = round(Mean + (2*SE),3)) # upper confidence interval
print(DataSum)

ddply(data, c("Domestication"), summarize,
            N = sum(!is.na(Mean_Weight_Cat)), #number of samples
            Mis = sum(is.na(Mean_Weight_Cat)), #number of missing values
            Mean = round(mean(Mean_Weight_Cat, na.rm=T),3), #mean
            Median = round(median(Mean_Weight_Cat, na.rm=T),3), #median
            SD = round(sd(Mean_Weight_Cat, na.rm=T),3), #standard deviation
            SE = round(sd(Mean_Weight_Cat, na.rm=T) / sqrt(N),5), #standard error
            LCI = round(Mean - (2*SE),3), #lower confidence interval
            HCI = round(Mean + (2*SE),3)) # upper confidence interval
```

## Plot
```{r, fig.width=8}
# Defining color for strip in facet wrap
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#E69F00", 1), rep("#D55E00", 1))))

ggplot(data = data,
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Domestication, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean,
               geom = "point",
               size = 1.5,
               color = "white",
               shape = 18) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, max(data$Mean_Weight_Cat*1.1))) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Dry caterpillar weight") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```


# Statistical analysis
### Model
```{r}
lm1 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning, family = gaussian(link = "identity"))
lm2 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning + (1|Accession), family = gaussian(link = "identity"))
lm2b <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning + (1|Block_g) + (1|Accession), family = gaussian(link = "identity"))
lm2n <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ (Domestication/Accession)*Soil_conditioning + (1|Block_g), family = gaussian(link = "identity"))

lm3 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat) ~ Domestication*Soil_conditioning, family = gaussian(link = "identity"))
lm4 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat) ~ Domestication*Soil_conditioning + (1|Accession), family = gaussian(link = "identity"))
lm4b <- glmmTMB(data = data, formula = log(Mean_Weight_Cat) ~ Domestication*Soil_conditioning + (1|Accession) + (1|Block_g), family = gaussian(link = "identity"))
lm5 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Domestication*Soil_conditioning, family = gaussian(link = "identity"))
lm6 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Domestication*Soil_conditioning + (1|Accession), family = gaussian(link = "identity"))
lm6b <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Domestication*Soil_conditioning + (1|Accession) + (1|Block_g), family = gaussian(link = "identity"))

glm1 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning, family = Gamma(link = "log"))
glm2 <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning + (1|Accession), family = Gamma(link = "log"))
glm2b <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning + (1|Accession) + (1|Block_g), family = Gamma(link = "log"))
glm3 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat + 1) ~ Domestication*Soil_conditioning, family = Gamma(link = "log"))
glm4 <- glmmTMB(data = data, formula = log(Mean_Weight_Cat + 1) ~ Domestication*Soil_conditioning + (1|Accession), family = Gamma(link = "log"))
glm4b <- glmmTMB(data = data, formula = log(Mean_Weight_Cat + 1) ~ Domestication*Soil_conditioning + (1|Accession) + (1|Block_g), family = Gamma(link = "log"))
glm5 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Domestication*Soil_conditioning, family = Gamma(link = "log"))
glm6 <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Domestication*Soil_conditioning + (1|Accession), family = Gamma(link = "log"))
glm6b <- glmmTMB(data = data, formula = sqrt(Mean_Weight_Cat) ~ Domestication*Soil_conditioning + (1|Accession) + (1|Block_g), family = Gamma(link = "log"))
```

### Model output
```{r}
# summary(lm1)
# Anova(lm1)
# summary(lm2)
# Anova(lm2)
# summary(lm3)
# Anova(lm3)
# summary(lm4)
# Anova(lm4)
# summary(lm5)
# Anova(lm5)
summary(lm6)
an_lm6 <- Anova(lm6)
an_lm6
Stat_catweight_dom <- as.data.frame(an_lm6)
write.csv(Stat_catweight_dom, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_catweight_dom.csv")

# summary(glm1)
# Anova(glm1)
# summary(glm2)
# Anova(glm2)
summary(glm2b)
an_glm2b <- Anova(glm2b)
an_glm2b
Stat_catweight_dom_glm <- as.data.frame(an_glm2b)
write.csv(Stat_catweight_dom_glm, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_catweight_dom_glm.csv")

# summary(glm3)
# Anova(glm3)
# summary(glm4)
# Anova(glm4)
# summary(glm5)
# Anova(glm5)
# summary(glm6)
# Anova(glm6)
```

### Model assumptions
```{r, fig.width=10, fig.height=7}
DHARMa.sum(Model = lm1)
DHARMa.sum(Model = lm2)
DHARMa.sum(Model = lm2b)
DHARMa.sum(Model = lm3)
DHARMa.sum(Model = lm4)
DHARMa.sum(Model = lm4b)
DHARMa.sum(Model = lm5)
DHARMa.sum(Model = lm6)
DHARMa.sum(Model = lm6b)

DHARMa.sum(Model = glm1)
DHARMa.sum(Model = glm2)
DHARMa.sum(Model = glm2b)
DHARMa.sum(Model = glm3)
DHARMa.sum(Model = glm4)
DHARMa.sum(Model = glm4b)
DHARMa.sum(Model = glm5)
DHARMa.sum(Model = glm6)
DHARMa.sum(Model = glm6b)
```

## Extracting ratios on log scale
### glmm not transformed
```{r}
# Define model
glm2b <- glmmTMB(data = data, formula = Mean_Weight_Cat ~ Domestication*Soil_conditioning + (1|Accession) + (1|Block_g), family = Gamma(link = "log"))

# Extract ratios on log scale
EMM_mod_glmm_link <- emmeans(glm2b, pairwise ~ Soil_conditioning | Domestication, adjust = "tukey", type = "link")
PSF_ratio_glmm_link <- summary(EMM_mod_glmm_link$contrasts, infer = TRUE)

## Add domestication
PSF_ratio_glmm_link <- as.data.frame(PSF_ratio_glmm_link)
PSF_ratio_glmm_link$Domestication <- c("Wild", "Cultivated")
PSF_ratio_glmm_link$Domestication <- factor(PSF_ratio_glmm_link$Domestication, levels = c("Wild", "Cultivated"))
PSF_ratio_glmm_link
```

# Final figure
### Data preperation
```{r, fig.width=10, fig.height=4}
# Alter color facet wrap label
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#0072B2", 1), rep("#D55E00", 1))),
                                                     text_x = elem_list_text(colour = c(rep("white", 1), rep("black", 1))))

DataSum <- data.frame(Domestication = rep(c("Wild", "Cultivated"), each = 2),
                      Soil_conditioning = rep(c("Co", "Mb"), 2),
                      lab = c("a", "", "b", ""))
DataSum$Domestication <- factor(DataSum$Domestication, levels = c("Wild", "Cultivated"))

# Prepare statistical informatin for title graph
Stats <- paste0("LMM: Dom p < ", 0.001, "; Soil p = ", 0.845, ";\nInteraction p = ", 0.678)
```

### Figure
```{r, fig.width = 3.5, fig.height=4}
ggplot(data = data,
       mapping = aes(y = Mean_Weight_Cat, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Mean_Weight_Cat, x = Soil_conditioning), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Domestication, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 26.5, xend = 2.2, yend = 26.5)) +
  geom_text(data = DataSum,
            aes(label = lab, y = 28, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  # geom_text(data = DataSum,
  #           aes(label = lab2, y = 80), # fill out the significant letters
  #           size = 8,
  #           hjust = -1.1,
  #           #position = position_dodge(0.5)
  #          ) +
  scale_y_continuous(expand = c(0, 0),   #start graph at y = min and x = min
                     breaks = seq(0, 25, 5), #steps on the y-axis
                     limits = c(0, 29.2)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("Dry caterpillar weight (mg)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10), #, hjust = 0
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3.2,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_CatWeight_Dom.svg", "PF_CatWeight_Dom.png"))
```

### Figure PSF effect
```{r, fig.width=2.3, fig.height=3}
p <- ggplot(data = PSF_ratio_glmm_link,
       mapping = aes(y = estimate, x = Domestication)) + #Defines the data
  geom_hline(yintercept = 0, 
             colour = "gray30") +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                color = c(rep("#0072B2", 1), rep("#D55E00", 1)),
                linewidth = 1,
                width = 0.3) +
  geom_point(aes(y = estimate, x = Domestication),
             size = 2) +
  #scale_color_manual(values = c("Wild" = "#0072B2", "Cultivated" = "#D55E00")) +  
  #scale_y_continuous(#expand = c(0, 0),        #start graph at y = min and x = min
    #                 breaks = seq(-0.5, 0.75, 0.25), #steps on the y-axis
     #                limits = c(-0.5, 0.75)) +   #y limits
  # geom_text(data = DataSum,
  #           aes(label = lab2, y = 0.7), # fill out the significant letters
  #           size = 8) +
  ylab("PSF ratio caterpillar weight") +    #y-label with Mb in italic
  xlab("Domestication") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
print(p)

mapply(function(x)
  ggsave(plot = last_plot(),
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 2.8,
         height = 3,
         units = "cm",
         dpi = 300),
  x = c("PF_PSF_ratio_CatWeigh_Dom.svg", "PF_PSF_ratio_CatWeigh_Dom.png"))
```

#### Get legend
```{r, fig.width=1.1, fig.height=0.9}
legend_dom <- get_legend(p + theme(legend.position = "right"))
ggdraw(legend_dom)

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 1.1,
         height = 0.9,
         units = "cm",
         dpi = 300),
  x = c("PF_PSF_ratio_CatWeight_Dom_legend.svg", "PF_PSF_ratio_CatWeigh_Dom_legend.png"))
```
