---
title: "Leaf damage - Pilot 1 - Determining infestation number CP"
author: "Kris"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---

# Introduction

   

# Start
## Libraries
```{r, warning = FALSE, message = FALSE}
R.version$version.string # prints R version

library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(ggh4x) #for coloring strip facet wrap
packageVersion("ggh4x")
library(plyr) #for summary table
packageVersion("plyr")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")
```

## Loading data and processing
#### Loading data
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Pilot 1 Determining infestation nr CP/Pilot1_Leafdamage_DataSummary.csv")
```

#### Remove NAs
```{r}
data <- data[complete.cases(data[ , "Consumed_Leaf_Area"]),]
```

#### Defining factors
```{r}
data$Accession <- factor(data$Accession, levels = c("DD", "OH", "HE", "CD", "KT", "KI", "MC", "SA", "GO1", "IT1", "IT2", "RI", "VL", "GO2", "BS", "CY"))
```

```{r}
summary(data$Consumed_Leaf_Area)
table(data$Accession)
```


# Determine inductin number CP 12 accession experiment
### Preparing data set
```{r}
mean_fc <- mean(data$Fr_Consumed) #calculating the overall mean of all accessions together of the fraction of plants that is consumed 
data$Change <- data$Fr_Consumed / mean_fc #calculate the fold difference of the fraction eaten per plant relative to the overall mean
```

#### Summary table
```{r}
sumtab <- cbind(aggregate(x = data$Consumed_Leaf_Area, by = list(data$Accession), FUN = mean),
                aggregate(x = data$Fr_Consumed, by = list(data$Accession), FUN = mean)$x,
                aggregate(x = data$Change, by = list(data$Accession), FUN = mean)$x)
colnames(sumtab) <- c("Accession", "CLA", "FCLA", "Change")
sumtab
```

### Exploring fold difference in induciton number
```{r}
summary(data$Change)
plot(Change ~ Accession, data = data, las = 2)
abline(h = 1)
```

### Exploring difference in induction number
```{r}
data_S <- aggregate(x = data$Change, by = list(data$Accession), FUN = mean) #Make as summary of the data per accession to determine order in graphs below
colnames(data_S) <- c("Accession", "Mean_ch")

data_S$Catnr4 <- (1/data_S$Mean_ch) * 4 #calculating the number of caterpillars on a plant to have get similar feeding damage on all accessions. In the future experiments, on average 4 caterpillars will be introduced
data_S$Catnr4r <- round(data_S$Catnr4, digits = 0) #round the number of caterpillars

print(data_S)
```

### Calculate induction number by starting with means of change per accession
```{r}
plot(Catnr4r ~ Accession, data = data_S, las = 2,  ylab = "Number of caterpillars", ylim = c(0, max(Catnr4r)+3))
j <- 0
invisible(sapply(data_S$Catnr4r,
                 function(i) {
                   j <<- j + 1
                   text(j, i+2, paste(i))})) #plotting the number of caterpillars above the line
```


# Plotting data
### Removing accessions not tested in 12 accession experiment
```{r}
unique(data$Accession)
Acc_rm <- c("BS", "GO2", "IT2", "SA")
data <- data[!data$Accession %in% Acc_rm, ]
data_S <- data_S[!data_S$Accession %in% Acc_rm, ]
data_S$Accession <- droplevels(data_S$Accession)
```

### Plot induction number
```{r}
plot(Catnr4r ~ Accession, data = data_S, las = 2,  ylab = "Number of caterpillars", ylim = c(0, max(Catnr4r)+3))
j <- 0
invisible(sapply(data_S$Catnr4r,
                 function(i) {
                   j <<- j + 1
                   text(j, i+2, paste(i))})) #plotting the number of caterpillars above the line
```

### Defining factors
```{r}
# data$Accession <- factor(data$Accession, levels = c("DD", "OH", "HE", "KI", "CD", "KT", "MC", "IT1", "GO1", "RI", "VL", "CY"))
# data_S$Accession <- factor(data_S$Accession, levels = c("DD", "OH", "HE", "KI", "CD", "KT", "MC", "IT1", "GO1", "RI", "VL", "CY"))
```

### Plot consumed leaf area
```{r, fig.width=10, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = Consumed_Leaf_Area, x = Accession, fill = Accession)) + #Defines the data
  stat_boxplot(aes(y = Consumed_Leaf_Area, x = Accession),  
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  #geom_segment(aes(x = 0.8, y = 1.4, xend = 2.2, yend = 1.4)) +
  # geom_text(data = Test,
  #           aes(label = Letters, y = 1.2), # fill out the significant letters
  #           size = 4,
  #           hjust = 0.5) +
  # geom_vline(xintercept = c(0.4, 2.6), color = "gray60") +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$Consumed_Leaf_Area)*1.15)) +   #y limits
  scale_fill_manual(values = c(rep("#0072B2", 4), "#D55E00", "#0072B2", rep("#D55E00", 6))) +
  ylab(bquote("Leaf damage " ~(cm^2))) +    #y-label with Mb in italic
  xlab("Accession") +
  # ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

### Plot fraction consumed leaf area of total leaf area
```{r, fig.width=10, fig.height=4}
p <- ggplot(data = data, 
       mapping = aes(y = Fr_Consumed, x = Accession, fill = Accession)) + #Defines the data
  stat_boxplot(aes(y = Fr_Consumed, x = Accession),  
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  #geom_segment(aes(x = 0.8, y = 1.4, xend = 2.2, yend = 1.4)) +
  # geom_text(data = Test,
  #           aes(label = Letters, y = 1.2), # fill out the significant letters
  #           size = 4,
  #           hjust = 0.5) +
  # geom_vline(xintercept = c(0.4, 2.6), color = "gray60") +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$Fr_Consumed)*1.15)) +   #y limits
  scale_fill_manual(values = c(rep("#0072B2", 4), "#D55E00", "#0072B2", rep("#D55E00", 6))) +
  ylab("Fraction consumed leaf area") +    #y-label with Mb in italic
  xlab("Accession") +
  # ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```
```{r, fig.width=7, fig.height=4}
print(p)
```
```{r, fig.width=7, fig.height=4}
maxdat <- aggregate(x = data$Fr_Consumed, by = list(data$Accession), FUN = max)
colnames(maxdat) <- c("Accession", "Max")
data_S <- merge(data_S, maxdat, by = "Accession")
data_S$Accession <- factor(data_S$Accession, levels = c("DD", "OH", "HE", "CD", "KT", "KI", "MC", "GO1", "IT1", "RI", "VL", "CY"))

p + geom_text(data = data_S,
              aes(x = Accession, y = Max + 0.01),
              label = data_S$Catnr4r)

p + geom_text(data = data_S,
              aes(x = Accession, y = 0.105),
              label = data_S$Catnr4r)

p + geom_text(data = data_S,
              aes(x = Accession, y = c(rep(0.003, 2), 0.017, rep(0.003, 9)),
              label = Catnr4r))

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 10,
         height = 5,
         units = "cm",
         dpi = 300),
  x = c("Pilot1_Leafdamagefraction.svg", "Pilot1_Leafdamagefraction.png"))
```



