---
title: "12 accession Exp - Caterpillar Survival FP"
author: "Kris"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---

# Start
## Import pre-made functions
```{r, warning = FALSE, message = FALSE}
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R") #Functions for comparing distributions, models and model assumptions
#source("C:/Users/K/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R") #Functions for comparing distributions, models and model assumptions
```

## Libraries
```{r, warning = FALSE, message = FALSE}
library(plyr) #for summary of the data
packageVersion("plyr")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(ggh4x) #for coloring strip facet wrap
packageVersion("ggh4x")
library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")
```

## Processing full data set
### load data
```{r}
#data <- read.csv("../Data/Caterpillar Biomass/Caterpillar_biomass_FP_R.csv")
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Metadata FP/Meta_Data_FP.csv")
```

### Removing NAs and sterile treatment 
```{r}
data <- data[complete.cases(data$Survival_fraction), ]
data <- data[data$Soil_conditioning != "S", ]
```
  
### Remove Batch 1 for all accessions
```{r}
#cn <- c(which(data$Accession == "VL" & data$Batch == 1), which(data$Accession == "HM" & data$Batch == 1))
data <- data[data$Batch != 1, ]
```
  
### Defining factors
```{r}
data$Soil_conditioning <- as.factor(data$Soil_conditioning)
#data$Batch <- as.factor(data$Batch)
data$Block_g <- as.factor(data$Block_g)
data$Domestication <- factor(data$Domestication, levels = c("Wild", "Cultivated"))
#data$Origin <- as.factor(data$Origin)
```

# Exploring data
```{r}
str(data[ , c(1:10, 57:62)])
head(data[ , c(1:10, 57:62)], 12)
summary(data[ , c(1:10, 57:62)]) #check for outliers
```

## Summary table
```{r}
DataSum <- ddply(data, c("Domestication", "Soil_conditioning"), summarise,
      N = sum(!is.na(Survival_fraction)), #number of samples
      Mis = sum(is.na(Survival_fraction)), #number of missing values
      Mean = round(mean(Survival_fraction, na.rm = T), 3), #mean
      Median = round(median(Survival_fraction, na.rm = T), 3), #median
      SD = round(sd(Survival_fraction, na.rm = T), 3), #standard deviation
      SE = round(sd(Survival_fraction, na.rm = T) / sqrt(N), 5), #standard error
      LCI = round(Mean - (2*SE), 3), #lower confidence interval
      HCI = round(Mean + (2*SE), 3)) # upper confidence interval
print(DataSum)
```

## Data plot
### Batches together
```{r}
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#E69F00", 1), rep("#D55E00", 1))))

ggplot(data = data,
       mapping = aes(y = Survival_fraction, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Survival_fraction, x = Soil_conditioning), #Error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Domestication, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, max(1.05))) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Fraction surviving caterpillars") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```

### Batches split
```{r, fig.width=10}
strip1 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#E69F00", 2), rep("#D55E00", 2), rep("gray", 4))))

ggplot(data = data,
       mapping = aes(y = Survival_fraction, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Survival_fraction, x = Soil_conditioning), #Error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Domestication + Batch, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip1
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, max(1.05))) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Fraction surviving caterpillars") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```

# Statistical analysis
## Testing only batch 2
### Defining model
```{r}
glmb1 <- glmmTMB(data = data, formula = cbind(Survival, Dead) ~ Domestication*Soil_conditioning, family = binomial(link = "logit"))
glmbb1 <- glmmTMB(data = data, formula = cbind(Survival, Dead) ~ Domestication*Soil_conditioning, family = betabinomial(link = "logit"))

glmb2 <- glmmTMB(data = data, formula = cbind(Survival, Dead) ~ Domestication*Soil_conditioning + (1 | Accession), family = binomial(link = "logit"))
#glmbb2 <- glmmTMB(data = data, formula = cbind(Survival, Dead) ~ Domestication*Soil_conditioning + (1 | Accession), family = betabinomial(link = "logit"))

glmb3 <- glmmTMB(data = data, formula = cbind(Survival, Dead) ~ Domestication*Soil_conditioning + (1 | Accession) + (1 | Block_g), family = binomial(link = "logit"))
glmbb3 <- glmmTMB(data = data, formula = cbind(Survival, Dead) ~ Domestication*Soil_conditioning + (1 | Block_g), family = betabinomial(link = "logit"))
```

### Model output
```{r}
# summary(glmb1)
# summary(glmbb1)
# Anova(glmb1)
# Anova(glmbb1)
# 
# summary(glmb2)
summary(glmb3)
# Anova(glmb2)
Anova(glmb3)

an_glmb3 <- Anova(glmb3)
Stat_catsurvival_dom <- as.data.frame(an_glmb3)
write.csv(Stat_catsurvival_dom, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_catsurvival_dom.csv")
```

### Model assumptions
```{r, fig.width=10, fig.height=7}
DHARMa.sum(Model = glmb1)
DHARMa.sum(Model = glmbb1)

DHARMa.sum(Model = glmb2)
#DHARMa.sum(Model = glmbb2)

DHARMa.sum(Model = glmb3)
DHARMa.sum(Model = glmbb3)
```
   
<!-- ### Post hoc test -->
<!-- ```{r} -->
<!-- emmeans(glmbb1, pairwise ~ Domestication) -->
<!-- ph <- cld(emmeans(glmbb1, pairwise ~ Domestication), Letters = letters) -->
<!-- print(ph) -->
<!-- ``` -->
   
# Final figure
### Data preperation
```{r, fig.width=10, fig.height=4}
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#0072B2", 1), rep("#D55E00", 1))),
                       text_x = elem_list_text(colour = c(rep("white", 1), rep("black", 1))))

Stats <- paste0("GLMM: Dom p < ", 0.026, "; Soil p = ", 0.088, "; \nInteraction p = ", 0.230)

DataSum$Letters <- c("a", "", "b", "")
```

### Figure
```{r, fig.width=3.5, fig.height=4}
ggplot(data = data,
       mapping = aes(y = Survival_fraction, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Survival_fraction, x = Soil_conditioning), #Error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap2(~ Domestication, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 1.11, xend = 2.2, yend = 1.11)) +
  geom_text(data = DataSum,
            aes(label = Letters, y = 1.15, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5,
            #position = position_dodge(0.5)
           ) +
  #geom_text(data = DataSum,
   #         aes(label = lab2, y = 1.1), # fill out the significant letters
    #        size = 8,
     #       hjust = -1.1,
            #position = position_dodge(0.5)
      #     ) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     breaks = seq(0, 1, 0.2), #steps on the y-axis
                     limits = c(0, 1.2)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Fraction surviving caterpillars") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```

```{r}
colnames(DataSum)[5] <- "Survival_fraction"
```

```{r, fig.width=3.2, fig.height=4}
ggplot(data = data,
       aes(y = Survival_fraction, x = Soil_conditioning)) + #Defines the data
  geom_errorbar(data = DataSum, 
                aes(ymin = Survival_fraction - SE, ymax = Survival_fraction + SE, color = Soil_conditioning), 
                linewidth = 1,
                width = 0.3) +
  geom_point(data = DataSum,
             aes(y = Survival_fraction, x = Soil_conditioning, color = Soil_conditioning),
             size = 2) +
  facet_wrap2(~ Domestication, #facet wrap2 needed to color strip
             nrow = 1,
             strip = strip0
             #labeller = labeller(CaterpillarNr = c("4" = "4 Cat", "8" = "8 Cat"))
             ) +
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  geom_segment(aes(x = 0.8, y = 1.05, xend = 2.2, yend = 1.05)) +
  geom_text(data = DataSum,
            aes(label = Letters, y = 1.1, x = 1.5), # fill out the significant letters
            size = 4,
            #position = position_dodge(0.5),
            hjust = 0.5) +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, 1.15)) +   #y limits
  scale_color_manual(values = c("Co" = "#882255", "Mb" = "#009E73")) +  
  ylab("Fraction surviving caterpillars") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
    theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 3.4,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_Catsurvival_Dom.svg", "PF_Catsurvival_Dom.png"))
```