---
title: "Root biomass - 12 accession experiment FP"
author: "Kris"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---

# Introduction

   

# Start
## Import pre-made functions
```{r, warning = FALSE, message = FALSE}
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/R_Scripts/SelfMadeFunctions.R") #Functions for comparing distributions, models and model assumptions
```
    
## Libraries
```{r, warning = FALSE, message = FALSE}
R.version$version.string # prints R version

library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(ggh4x) #for coloring strip facet wrap
packageVersion("ggh4x")
library(plyr) #for summary table
packageVersion("plyr")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")
```

## Loading data and processing
#### Loading data
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Data/Metadata FP/Meta_Data_FP.csv")
```

#### Remove NAs
```{r}
data <- data[complete.cases(data[ , "Root_Weight"]),]
```

### Removing sterile treatment and Batch 1
```{r}
data <- data[data$Soil_conditioning != "S", ]
data <- data[data$Batch != 1, ]
```

#### Defining factors
```{r}
data$Soil_conditioning <- as.factor(data$Soil_conditioning)
data$Accession <- factor(data$Accession, levels = c("OH", "DD", "HE", "KI", "VL", "CD", "RI", "KT", "MC", "HM", "IT1", "GO1"))
data$Domestication <- factor(data$Domestication, levels = c("Wild", "Cultivated"))
```

# Data analysis
## Data exploration
```{r}
str(data[c(1:8, 65:67)])
head(data[c(1:8, 65:67)], 12)
summary(data[c(1:8, 65:67)]) #check for outliers
table(data$Block_g, data$Accession, data$Soil_conditioning)
``` 

```{r}
DataSum1 <- ddply(data, c("Accession", "Soil_conditioning"), summarise,
      N = sum(!is.na(Root_Weight)), #number of samples
      Mis = sum(is.na(Root_Weight)), #number of missing values
      Mean = round(mean(Root_Weight, na.rm=T),3), #mean
      Median = round(median(Root_Weight, na.rm=T),3), #median
      SD = round(sd(Root_Weight, na.rm=T),3), #standard deviation
      SE = round(sd(Root_Weight, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval
DataSum1

ddply(data, c("Soil_conditioning"), summarise,
      N = sum(!is.na(Root_Weight)), #number of samples
      Mis = sum(is.na(Root_Weight)), #number of missing values
      Mean = round(mean(Root_Weight, na.rm=T),3), #mean
      Median = round(median(Root_Weight, na.rm=T),3), #median
      SD = round(sd(Root_Weight, na.rm=T),3), #standard deviation
      SE = round(sd(Root_Weight, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval
```

```{r}
DataSum2 <- ddply(data, c("Domestication", "Soil_conditioning"), summarise,
      N = sum(!is.na(Root_Weight)), #number of samples
      Mis = sum(is.na(Root_Weight)), #number of missing values
      Mean = round(mean(Root_Weight, na.rm=T),3), #mean
      Median = round(median(Root_Weight, na.rm=T),3), #median
      SD = round(sd(Root_Weight, na.rm=T),3), #standard deviation
      SE = round(sd(Root_Weight, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval
DataSum2

ddply(data, c("Domestication"), summarise,
      N = sum(!is.na(Root_Weight)), #number of samples
      Mis = sum(is.na(Root_Weight)), #number of missing values
      Mean = round(mean(Root_Weight, na.rm=T),3), #mean
      Median = round(median(Root_Weight, na.rm=T),3), #median
      SD = round(sd(Root_Weight, na.rm=T),3), #standard deviation
      SE = round(sd(Root_Weight, na.rm=T) / sqrt(N),5), #standard error
      LCI = round(Mean - (2*SE),3), #lower confidence interval
      HCI = round(Mean + (2*SE),3)) # upper confidence interval
```

## Plots
```{r}
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#E69F00", 5), rep("#D55E00", 7))))
```

### Plot all accessions
```{r, fig.width=10, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = Root_Weight, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Root_Weight, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$Root_Weight)*1.1)) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) +
  ylab("Root biomass (g)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  #ggtitle("Infestation time") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```

### Plot domestication
```{r}
strip1 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#E69F00", 1), rep("#D55E00", 1))))
```
```{r, fig.width = 3, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = Root_Weight, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Root_Weight, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Domestication, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip1
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$Root_Weight)*1.1)) +   #y limits
  scale_fill_manual(values = c("#56B4E9", "#009E73")) +
  ylab("Root biomass (g)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  #ggtitle("Infestation time") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 9, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "none")
```

# Statistical analysis Accessions
### Model
```{r}
lm1 <- glmmTMB(data = data, formula = Root_Weight ~ Accession*Soil_conditioning, family = gaussian(link = "identity"))
lm3 <- glmmTMB(data = data, formula = log(Root_Weight) ~ Accession*Soil_conditioning, family = gaussian(link = "identity"))
lm5 <- glmmTMB(data = data, formula = sqrt(Root_Weight) ~ Accession*Soil_conditioning, family = gaussian(link = "identity"))

glm1 <- glmmTMB(data = data, formula = Root_Weight ~ Accession*Soil_conditioning, family = Gamma(link = "log"))
glm3 <- glmmTMB(data = data, formula = log(Root_Weight + 1) ~ Accession*Soil_conditioning, family = Gamma(link = "log"))
glm5 <- glmmTMB(data = data, formula = sqrt(Root_Weight) ~ Accession*Soil_conditioning, family = Gamma(link = "log"))
```

### Model output
```{r}
summary(lm1)
Anova(lm1)
summary(lm3)
Anova(lm3)
summary(lm5)
Anova(lm5)

summary(glm1)
Anova(glm1)
summary(glm3)
Anova(glm3)
summary(glm5)
Anova(glm5)

an_lm1 <- Anova(lm1)
Stat_rootbiomass_acc <- as.data.frame(an_lm1)
write.csv(Stat_rootbiomass_acc, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/Statistical output/Stat_rootbiomass_acc.csv")
```

### Model assumptions
```{r, fig.width=10, fig.height=7}
DHARMa.sum(Model = lm1)
DHARMa.sum(Model = lm3)
DHARMa.sum(Model = lm5)

DHARMa.sum(Model = glm1)
DHARMa.sum(Model = glm3)
DHARMa.sum(Model = glm5)
```

### Post hoc test
```{r}
EMM <- emmeans(lm1, pairwise ~ Soil_conditioning | Accession, adjust = "tukey", type = "response") # only p value correction within each group. No correction as ther is only one comparison
EMM$contrasts
EMM_Corr <- rbind(EMM$contrasts)
EMM_Corr # Bonferoni correction

EMS <- rbind(cld(EMM, Letters = letters))
EMS

EMMA2 <- emmeans(lm1, pairwise ~ Accession, adjust = "tukey", type = "response") # more stringed
EMA <- cld(EMMA2, Letters = letters)
EMA
```

# Final plots
## All accessions
### Preparation
```{r}
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#0072B2", 5), rep("#D55E00", 7))),
                        text_x = elem_list_text(colour = c(rep("white", 5), rep("black", 7))))

# Statistical info main effects
Stats <- paste0("\nLM: Accession p < 0.001; Soil treatment p = 0.040; Interaction p < 0.001")

# Significant letters in right order
EMA <- EMA[order(EMA$Accession), ]

# Alternate letters and blanks
Blank <- rep("", 12)
Letters <- c()

for (i in 1:12) {
  Letters[length(Letters) + 1] <- EMA$.group[i]
  Letters[length(Letters) + 1] <- Blank[i]
}

# Remove spaces around letters
Letters <- gsub(pattern = " ", replacement = "", x = Letters)
Letters

# Significance star for difference in soil conditioning
Star <- c(rep(c("", "*"), 2), rep("", 3), "**", rep("", 3), "**", "", ".\n", rep("", 8), ".\n", "")

DataSum <- data.frame(Accession = rep(EMA$Accession, each = 2),
                      Soil_conditioning = rep(c("Co", "Mb"), 12),
                      Letters = Letters,
                      Star = Star)
```

### Plot
```{r, fig.width=10, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = Root_Weight, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Root_Weight, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  geom_segment(aes(x = 0.8, y = 1.15, xend = 2.2, yend = 1.15)) +
  geom_text(data = DataSum,
            aes(label = Letters, y = 1.2, x = 1.5), # fill out the significant letters
            size = 4,
            hjust = 0.5) +
  geom_text(data = DataSum,
            aes(label = Star, y = 1.03, x = 1.5), # fill out the significant letters
            size = 7,
            hjust = 0.5) +  
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     breaks = seq(0, 1, 0.25), #steps on the y-axis
                     limits = c(0, 1.25)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("Root biomass (g)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 10,
         height = 4,
         units = "cm",
         dpi = 300),
  x = c("PF_RootBiomass_Acc.svg", "PF_RootBiomass_Acc.png"))
```

### Stats for Soil conditioning
```{r}
strip0 <- strip_themed(background_x = elem_list_rect(fill = c(rep("#0072B2", 5), rep("#D55E00", 7))),
                        text_x = elem_list_text(colour = c(rep("white", 5), rep("black", 7))))

# Significant letters in right order
EMS <- EMS[order(EMS$Soil_conditioning), ]
EMS <- EMS[order(EMS$Accession), ]

# Rename letters column
colnames(EMS)[8] <- "Letters"

# Remove spaces around letters
EMS$Letters <- gsub(pattern = " ", replacement = "", x = EMS$Letters)
EMS$Letters

# Calculate max
Test <- aggregate(data$Root_Weight, by = list(data$Soil_conditioning, data$Accession), FUN = max)

# Make a data frame with all info needed
Test$Letters <- EMS$Letters
colnames(Test)[1:3] <- c("Soil_conditioning", "Accession", "Max")
```

### Plot
```{r, fig.width=10, fig.height=4}
ggplot(data = data, 
       mapping = aes(y = Root_Weight, x = Soil_conditioning, fill = Soil_conditioning)) + #Defines the data
  stat_boxplot(aes(y = Root_Weight, x = Soil_conditioning),     #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar", 
               linetype = 1, 
               width = 0.2) +
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot
  facet_wrap2(~ Accession, # Split data into three graphs, a graph for each caterpillar number
             nrow = 1,
             strip = strip0
             #labeller = labeller(Time = c("17" = "17 days"))
             ) + 
  geom_jitter(width = 0.25,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean, 
               geom = "point", 
               size = 1.5, 
               color = "gray30", 
               fill = "white",
               shape = 23) + #Shows the mean as a white circle
  #geom_segment(aes(x = 0.8, y = 1.4, xend = 2.2, yend = 1.4)) +
  geom_text(data = Test,
            aes(label = Letters, y = 1.2), # fill out the significant letters
            size = 4,
            hjust = 0.5) +
  geom_vline(xintercept = c(0.4, 2.6), color = "gray60") +
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$Root_Weight)*1.15)) +   #y limits
  scale_fill_manual(values = c("#882255", "#009E73")) +
  ylab("Root biomass (g)") +    #y-label with Mb in italic
  xlab("Soil treatment") +
  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```
