---
title: "FP_05_differential_abundance_Summary_Results - Data preperation - Only batch 2"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b635.html)
   
   
# 5.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(DESeq2)
packageVersion("DESeq2")
library(stringr)
packageVersion("stringr")
library(dplyr)
packageVersion("dplyr")
```


# 5.3 Create lists of DA ASVs
## 5.3.1 Accession ASV level
### Load DA ASVs
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Acc_Bac.RData")
```

### Extract DA ASVs
```{r}
## From each data frame, extract the ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASVs <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_ASVs <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))

# Within each plant concatenate the identified asvs
Acc_Bac_Total_DA_ASV <- mapply(function(s,t,w,x,y,z){
  input <- c(s,t,w,x,y,z)},
  s = Acc_Bac_DESeqLRT_ASVs,
  t = Acc_Bac_DESeqWald_ASVs,
  w = Acc_Bac_zinbLRT_ASVs,
  x = Acc_Bac_zinbWald_ASVs,
  y = Acc_Bac_ancomWZ_ASVs,
  z = Acc_Bac_ancomNoZ_ASVs,
  SIMPLIFY = FALSE)
```

### Total number of ASVs
```{r}
# Total number of DA ASVs
lapply(Acc_Bac_Total_DA_ASV, function(x) length(x))

save(Acc_Bac_Total_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_Total_DA_ASV.RData")

# Remove duplicated asvs names from that list
Acc_Bac_NoDup_DA_ASV <- lapply(Acc_Bac_Total_DA_ASV, function(x)
  unique(x))

# Number of DA ASVs without duplicates
lapply(Acc_Bac_NoDup_DA_ASV, function(x) length(x))

save(Acc_Bac_NoDup_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV.RData")

# Extracting taxonomy
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps.RData")

# Extract taxonomy of DA ASVs
FP_unnormalized_bac_ps_Tax <- as.data.frame(tax_table(FP_unnormalized_bac_ps))
Acc_Bac_NoDup_DA_ASV_Tax <- lapply(Acc_Bac_NoDup_DA_ASV, function(x) FP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_NoDup_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_NoDup_DA_ASV_Tax_Table <- lapply(Acc_Bac_NoDup_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_NoDup_DA_ASV_Tax_Table <- lapply(Acc_Bac_NoDup_DA_ASV_Tax_Table, function(ranklist) {
                                        lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_NoDup_DA_ASV_Tax_Table$HM$Class
Acc_Bac_NoDup_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_NoDup_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring twice
```{r}
# Filter to ASVs detected by at least 2 tests
Acc_Bac_NumbOcc_DA_ASV <- lapply(Acc_Bac_Total_DA_ASV, function(x) table(unlist(x)))
Acc_Bac_TwoTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 2]))
Acc_Bac_TwoTimes_DA_ASV


# Number of DA ASVs occurring in two tests
lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_TwoTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_TwoTimes_DA_ASV_Tax <- lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) FP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_TwoTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_TwoTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_TwoTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_TwoTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_TwoTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_TwoTimes_DA_ASV_Tax_Table$HM$Class
Acc_Bac_TwoTimes_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_TwoTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring three times
```{r}
# Filter to ASVs detected by at least 3 tests
Acc_Bac_NumbOcc_DA_ASV <- lapply(Acc_Bac_Total_DA_ASV, function(x) table(unlist(x)))
Acc_Bac_ThreeTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 3]))
Acc_Bac_ThreeTimes_DA_ASV

# Number of DA ASVs occurring in three tests
lapply(Acc_Bac_ThreeTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_ThreeTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_ThreeTimes_DA_ASV_Tax <- lapply(Acc_Bac_ThreeTimes_DA_ASV, function(x) FP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_ThreeTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_ThreeTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_ThreeTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table$HM$Class
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_ThreeTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring four times
```{r}
# Filter to ASVs detected by at least 4 tests
Acc_Bac_FourTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 4]))
Acc_Bac_FourTimes_DA_ASV

# Number of DA ASVs occurring in four tests
lapply(Acc_Bac_FourTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_FourTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FourTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_FourTimes_DA_ASV_Tax <- lapply(Acc_Bac_FourTimes_DA_ASV, function(x) FP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_FourTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FourTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_FourTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_FourTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_FourTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_FourTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_FourTimes_DA_ASV_Tax_Table$HM$Class
Acc_Bac_FourTimes_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_FourTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FourTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring five times
```{r}
# Filter to ASVs detected by at least 5 tests
Acc_Bac_FiveTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 5]))
Acc_Bac_FiveTimes_DA_ASV

# Number of DA ASVs occurring in Five tests
lapply(Acc_Bac_FiveTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_FiveTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FiveTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_FiveTimes_DA_ASV_Tax <- lapply(Acc_Bac_FiveTimes_DA_ASV, function(x) FP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_FiveTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FiveTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_FiveTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_FiveTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_FiveTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_FiveTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_FiveTimes_DA_ASV_Tax_Table$HM$Class
Acc_Bac_FiveTimes_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_FiveTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_FiveTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring six times
```{r}
# Filter to ASVs detected by at least 6 tests
Acc_Bac_SixTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 6]))
Acc_Bac_SixTimes_DA_ASV

# Number of DA ASVs occurring in Six tests
lapply(Acc_Bac_SixTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_SixTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_SixTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_SixTimes_DA_ASV_Tax <- lapply(Acc_Bac_SixTimes_DA_ASV, function(x) FP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_SixTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_SixTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_SixTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_SixTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_SixTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_SixTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Print output
Acc_Bac_SixTimes_DA_ASV_Tax_Table

save(Acc_Bac_SixTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_SixTimes_DA_ASV_Tax_Table.RData")
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.3.2 Accession Family level
### Load DA Family
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_FamilyLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_FamilyLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_FamilyLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_FamilyLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_FamilyLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_FamilyLevel/fdr_ancomNoZ_Acc_Bac.RData")

## Load Phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

### Extract DA Family
#### Extracting data from tests
```{r}
# From each data frame, extract the Family or ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASV_Family <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASV_Family <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASV_Family <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASV_Family <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_Family <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_Family <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))
```

#### Merging tests and transposing list
```{r}
Acc_DESeq_DA_ASV_Family_l <- list(DESeqLRT = Acc_Bac_DESeqLRT_ASV_Family, DESeqWald = Acc_Bac_DESeqWald_ASV_Family, zinbLRT = Acc_Bac_zinbLRT_ASV_Family, zinbWald = Acc_Bac_zinbWald_ASV_Family)

# transpose list
Acc <- names(Acc_DESeq_DA_ASV_Family_l[[1]])
Acc_DESeq_DA_ASV_Family_l_t <- vector("list", length(Acc))
names(Acc_DESeq_DA_ASV_Family_l_t) <- Acc

for (cd in Acc) {
  Acc_DESeq_DA_ASV_Family_l_t[[cd]] <- lapply(Acc_DESeq_DA_ASV_Family_l, function(m) m[[cd]])
}

## DA Family Ancom
Acc_Ancom_DA_Family_l <- list(ancomWZ = Acc_Bac_ancomWZ_Family, ancomNoZ = Acc_Bac_ancomNoZ_Family)

# transpose list
Acc_Ancom_DA_Family_l_t <- vector("list", length(Acc))
names(Acc_Ancom_DA_Family_l_t) <- Acc

for (cd in Acc) {
  Acc_Ancom_DA_Family_l_t[[cd]] <- lapply(Acc_Ancom_DA_Family_l, function(m) m[[cd]])
}
```

### Extract taxomomy
#### DESeq
```{r}
tax <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
tax$ASV <- rownames(tax)

Acc_NoDup_DESeq_DA_ASV_Family_Tax <- lapply(Acc_DESeq_DA_ASV_Family_l_t, function(x) {
  lapply(x, function(y) {
    unique(tax[tax$ASV %in% y, c("Family", "Order", "Class", "Phylum")])
    })})
```

#### Acnom
```{r}
#lapply(Acc_Ancom_DA_Family_l, function(x) length(x))

Acc_NoDup_Ancom_DA_Family_tax <- lapply(Acc_Ancom_DA_Family_l_t, function(y) {
  lapply(y, function(x) {
    # Take unique names in case of duplicates
    DA_tax <- unique(x)
    
    # Split data between long and short taxonomic names
    full_tax_strings   <- DA_tax[grepl("k__", DA_tax)]
    family_only_strings <- DA_tax[!grepl("k__", DA_tax)]
    
    # Extract taxomomy from long names
    tax_full <- tibble(Taxon = full_tax_strings) %>%
      mutate(
        Phylum = str_extract(Taxon, "p__.*?(?=_c__|$)"),
        Class  = str_extract(Taxon, "c__.*?(?=_o__|$)"),
        Order  = str_extract(Taxon, "o__.*?(?=_f__|$)"),
        Family = str_extract(Taxon, "f__.*?(?=_g__|$)")
      ) %>% select(Phylum, Class, Order, Family)
    
    # Extract taxonomy from ps object for short names
    tax_ps <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
    #family_only <- sub("^f__", "", family_only_strings)
    
    tax_family_map <- unique(tax_ps[tax_ps$Family %in% family_only_strings, c("Family", "Order", "Class", "Phylum")])
    #nrow(tax_family_map)
    
    # Remove left over Incertae Sedis from short names object
    tax_family_map <- tax_family_map %>%
      filter(Family != "f__Incertae_Sedis")
   
    # Combine long and short names
    tax_combined <- as.data.frame(bind_rows(tax_full, tax_family_map))
})})

#lapply(Acc_NoDup_Ancom_DA_Family_tax, function(x) nrow(x))
```

#### Merge Ancom and DESeq
```{r}
Acc_NoDup_DESeq_DA_ASV_Family_Tax_m <- lapply(Acc_NoDup_DESeq_DA_ASV_Family_Tax, function(x) do.call(rbind, x))
Acc_NoDup_Ancom_DA_Family_tax_m <- lapply(Acc_NoDup_Ancom_DA_Family_tax, function(x) do.call(rbind, x))

# Merge and filter on unique family
Acc_Total_Merged_DA_Family <- mapply(function(y,z){
  input <- rbind(y,z)},
  y = Acc_NoDup_DESeq_DA_ASV_Family_Tax_m,
  z = Acc_NoDup_Ancom_DA_Family_tax_m,
  SIMPLIFY = FALSE)

lapply(Acc_Total_Merged_DA_Family, function(x) nrow(x))
lapply(Acc_Total_Merged_DA_Family, function(x) table(x$Family))

save(Acc_Total_Merged_DA_Family, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_FamilyLevel/Acc_Total_Merged_DA_Family.RData")
```

### Remove duplicat Family
```{r}
Acc_NoDup_Merged_DA_Family <- lapply(Acc_Total_Merged_DA_Family, function(x) unique(x))
lapply(Acc_NoDup_Merged_DA_Family, function(x) nrow(x))
lapply(Acc_NoDup_Merged_DA_Family, function(x) table(x$Family))

save(Acc_NoDup_Merged_DA_Family, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_FamilyLevel/Acc_NoDup_Merged_DA_Family.RData")
```

### DA Family occurring twice
```{r}
# Filter to Family detected by at least 2 tests
Acc_TwoTimes_Merged_DA_Family <- lapply(Acc_Total_Merged_DA_Family, function(x) {
    dup <- x[duplicated(x) | duplicated(x, fromLast = TRUE), , drop = FALSE]
    unique(dup)
  })

# Number of DA Family occurring in two tests
lapply(Acc_TwoTimes_Merged_DA_Family, function(x) nrow(x))

save(Acc_TwoTimes_Merged_DA_Family, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_FamilyLevel/Acc_TwoTimes_Merged_DA_Family.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


## 5.3.3 Accession Order level
### Load DA Order
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_OrderLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_OrderLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_OrderLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_OrderLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_OrderLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_OrderLevel/fdr_ancomNoZ_Acc_Bac.RData")

## Load Phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

### Extract DA Order
#### Extracting data from tests
```{r}
# From each data frame, extract the Order or ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASV_Order <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASV_Order <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASV_Order <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASV_Order <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_Order <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_Order <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))
```

#### Merging tests and transposing list
```{r}
Acc_DESeq_DA_ASV_Order_l <- list(DESeqLRT = Acc_Bac_DESeqLRT_ASV_Order, DESeqWald = Acc_Bac_DESeqWald_ASV_Order, zinbLRT = Acc_Bac_zinbLRT_ASV_Order, zinbWald = Acc_Bac_zinbWald_ASV_Order)

# transpose list
Acc <- names(Acc_DESeq_DA_ASV_Order_l[[1]])
Acc_DESeq_DA_ASV_Order_l_t <- vector("list", length(Acc))
names(Acc_DESeq_DA_ASV_Order_l_t) <- Acc

for (cd in Acc) {
  Acc_DESeq_DA_ASV_Order_l_t[[cd]] <- lapply(Acc_DESeq_DA_ASV_Order_l, function(m) m[[cd]])
}

## DA Order Ancom
Acc_Ancom_DA_Order_l <- list(ancomWZ = Acc_Bac_ancomWZ_Order, ancomNoZ = Acc_Bac_ancomNoZ_Order)

# transpose list
Acc_Ancom_DA_Order_l_t <- vector("list", length(Acc))
names(Acc_Ancom_DA_Order_l_t) <- Acc

for (cd in Acc) {
  Acc_Ancom_DA_Order_l_t[[cd]] <- lapply(Acc_Ancom_DA_Order_l, function(m) m[[cd]])
}
```

### Extract taxomomy
#### DESeq
```{r}
tax <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
tax$ASV <- rownames(tax)

Acc_NoDup_DESeq_DA_ASV_Order_Tax <- lapply(Acc_DESeq_DA_ASV_Order_l_t, function(x) {
  lapply(x, function(y) {
    unique(tax[tax$ASV %in% y, c("Order", "Class", "Phylum")])
    })})
```

#### Acnom
```{r}
#lapply(Acc_Ancom_DA_Order_l, function(x) length(x))

Acc_NoDup_Ancom_DA_Order_tax <- lapply(Acc_Ancom_DA_Order_l_t, function(y) {
  lapply(y, function(x) {
    # Take unique names in case of duplicates
    DA_tax <- unique(x)
    
    # Split data between long and short taxonomic names
    full_tax_strings   <- DA_tax[grepl("k__", DA_tax)]
    family_only_strings <- DA_tax[!grepl("k__", DA_tax)]
    
    # Extract taxomomy from long names
    tax_full <- tibble(Taxon = full_tax_strings) %>%
      mutate(
        Phylum = str_extract(Taxon, "p__.*?(?=_c__|$)"),
        Class  = str_extract(Taxon, "c__.*?(?=_o__|$)"),
        Order  = str_extract(Taxon, "o__.*?(?=_f__|$)")
      ) %>% select(Phylum, Class, Order)
    
    # Extract taxonomy from ps object for short names
    tax_ps <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
    #family_only <- sub("^f__", "", family_only_strings)
    
    tax_family_map <- unique(tax_ps[tax_ps$Order %in% family_only_strings, c("Order", "Class", "Phylum")])
    #nrow(tax_family_map)
    
    # Remove left over Incertae Sedis from short names object
    tax_family_map <- tax_family_map %>%
      filter(Order != "f__Incertae_Sedis")
   
    # Combine long and short names
    tax_combined <- as.data.frame(bind_rows(tax_full, tax_family_map))
})})

#lapply(Acc_NoDup_Ancom_DA_Order_tax, function(x) nrow(x))
```

#### Merge Ancom and DESeq
```{r}
Acc_NoDup_DESeq_DA_ASV_Order_Tax_m <- lapply(Acc_NoDup_DESeq_DA_ASV_Order_Tax, function(x) do.call(rbind, x))
Acc_NoDup_Ancom_DA_Order_tax_m <- lapply(Acc_NoDup_Ancom_DA_Order_tax, function(x) do.call(rbind, x))

# Merge and filter on unique family
Acc_Total_Merged_DA_Order <- mapply(function(y,z){
  input <- rbind(y,z)},
  y = Acc_NoDup_DESeq_DA_ASV_Order_Tax_m,
  z = Acc_NoDup_Ancom_DA_Order_tax_m,
  SIMPLIFY = FALSE)

lapply(Acc_Total_Merged_DA_Order, function(x) nrow(x))
lapply(Acc_Total_Merged_DA_Order, function(x) table(x$Order))

save(Acc_Total_Merged_DA_Order, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_OrderLevel/Acc_Total_Merged_DA_Order.RData")
```

### Remove duplicat Order
```{r}
Acc_NoDup_Merged_DA_Order <- lapply(Acc_Total_Merged_DA_Order, function(x) unique(x))
lapply(Acc_NoDup_Merged_DA_Order, function(x) nrow(x))
lapply(Acc_NoDup_Merged_DA_Order, function(x) table(x$Order))

save(Acc_NoDup_Merged_DA_Order, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_OrderLevel/Acc_NoDup_Merged_DA_Order.RData")
```

### DA Order occurring twice
```{r}
# Filter to Order detected by at least 2 tests
Acc_TwoTimes_Merged_DA_Order <- lapply(Acc_Total_Merged_DA_Order, function(x) {
    dup <- x[duplicated(x) | duplicated(x, fromLast = TRUE), , drop = FALSE]
    unique(dup)
  })

# Number of DA Order occurring in two tests
lapply(Acc_TwoTimes_Merged_DA_Order, function(x) nrow(x))

save(Acc_TwoTimes_Merged_DA_Order, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_OrderLevel/Acc_TwoTimes_Merged_DA_Order.RData")
```
```{r}
# Clean environment
rm(list = ls())
```

## 5.3.4 Accession Class level
### Load DA Class
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ClassLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ClassLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ClassLevel/fdr_ancomNoZ_Acc_Bac.RData")

## Load Phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

### Extract DA Class
#### Extracting data from tests
```{r}
# From each data frame, extract the Class or ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASV_Class <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASV_Class <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASV_Class <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASV_Class <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_Class <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_Class <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))
```

#### Merging tests and transposing list
```{r}
Acc_DESeq_DA_ASV_Class_l <- list(DESeqLRT = Acc_Bac_DESeqLRT_ASV_Class, DESeqWald = Acc_Bac_DESeqWald_ASV_Class, zinbLRT = Acc_Bac_zinbLRT_ASV_Class, zinbWald = Acc_Bac_zinbWald_ASV_Class)

# transpose list
Acc <- names(Acc_DESeq_DA_ASV_Class_l[[1]])
Acc_DESeq_DA_ASV_Class_l_t <- vector("list", length(Acc))
names(Acc_DESeq_DA_ASV_Class_l_t) <- Acc

for (cd in Acc) {
  Acc_DESeq_DA_ASV_Class_l_t[[cd]] <- lapply(Acc_DESeq_DA_ASV_Class_l, function(m) m[[cd]])
}

## DA Class Ancom
Acc_Ancom_DA_Class_l <- list(ancomWZ = Acc_Bac_ancomWZ_Class, ancomNoZ = Acc_Bac_ancomNoZ_Class)

# transpose list
Acc_Ancom_DA_Class_l_t <- vector("list", length(Acc))
names(Acc_Ancom_DA_Class_l_t) <- Acc

for (cd in Acc) {
  Acc_Ancom_DA_Class_l_t[[cd]] <- lapply(Acc_Ancom_DA_Class_l, function(m) m[[cd]])
}
```

### Extract taxomomy
#### DESeq
```{r}
tax <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
tax$ASV <- rownames(tax)

Acc_NoDup_DESeq_DA_ASV_Class_Tax <- lapply(Acc_DESeq_DA_ASV_Class_l_t, function(x) {
  lapply(x, function(y) {
    unique(tax[tax$ASV %in% y, c("Class", "Phylum")])
    })})
```

#### Acnom
```{r}
#lapply(Acc_Ancom_DA_Class_l, function(x) length(x))

Acc_NoDup_Ancom_DA_Class_tax <- lapply(Acc_Ancom_DA_Class_l_t, function(y) {
  lapply(y, function(x) {
    # Take unique names in case of duplicates
    DA_tax <- unique(x)
    
    # Split data between long and short taxonomic names
    full_tax_strings   <- DA_tax[grepl("k__", DA_tax)]
    family_only_strings <- DA_tax[!grepl("k__", DA_tax)]
    
    # Extract taxomomy from long names
    tax_full <- tibble(Taxon = full_tax_strings) %>%
      mutate(
        Phylum = str_extract(Taxon, "p__.*?(?=_c__|$)"),
        Class  = str_extract(Taxon, "c__.*?(?=_o__|$)")
      ) %>% select(Phylum, Class)
    
    # Extract taxonomy from ps object for short names
    tax_ps <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
    #family_only <- sub("^f__", "", family_only_strings)
    
    tax_family_map <- unique(tax_ps[tax_ps$Class %in% family_only_strings, c("Class", "Phylum")])
    #nrow(tax_family_map)
    
    # Remove left over Incertae Sedis from short names object
    tax_family_map <- tax_family_map %>%
      filter(Class != "f__Incertae_Sedis")
   
    # Combine long and short names
    tax_combined <- as.data.frame(bind_rows(tax_full, tax_family_map))
})})

#lapply(Acc_NoDup_Ancom_DA_Class_tax, function(x) nrow(x))
```

#### Merge Ancom and DESeq
```{r}
Acc_NoDup_DESeq_DA_ASV_Class_Tax_m <- lapply(Acc_NoDup_DESeq_DA_ASV_Class_Tax, function(x) do.call(rbind, x))
Acc_NoDup_Ancom_DA_Class_tax_m <- lapply(Acc_NoDup_Ancom_DA_Class_tax, function(x) do.call(rbind, x))

# Merge and filter on unique family
Acc_Total_Merged_DA_Class <- mapply(function(y,z){
  input <- rbind(y,z)},
  y = Acc_NoDup_DESeq_DA_ASV_Class_Tax_m,
  z = Acc_NoDup_Ancom_DA_Class_tax_m,
  SIMPLIFY = FALSE)

lapply(Acc_Total_Merged_DA_Class, function(x) nrow(x))
lapply(Acc_Total_Merged_DA_Class, function(x) table(x$Class))

save(Acc_Total_Merged_DA_Class, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_Total_Merged_DA_Class.RData")
```

### Remove duplicat Class
```{r}
Acc_NoDup_Merged_DA_Class <- lapply(Acc_Total_Merged_DA_Class, function(x) unique(x))
lapply(Acc_NoDup_Merged_DA_Class, function(x) nrow(x))
lapply(Acc_NoDup_Merged_DA_Class, function(x) table(x$Class))

save(Acc_NoDup_Merged_DA_Class, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_NoDup_Merged_DA_Class.RData")
```

### DA Class occurring twice
```{r}
# Filter to Class detected by at least 2 tests
Acc_TwoTimes_Merged_DA_Class <- lapply(Acc_Total_Merged_DA_Class, function(x) {
    dup <- x[duplicated(x) | duplicated(x, fromLast = TRUE), , drop = FALSE]
    unique(dup)
  })

# Number of DA Class occurring in two tests
lapply(Acc_TwoTimes_Merged_DA_Class, function(x) nrow(x))

save(Acc_TwoTimes_Merged_DA_Class, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_TwoTimes_Merged_DA_Class.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


## 5.3.5 Accession Phylum level
### Load DA Phylum
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_PhylumLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_PhylumLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_PhylumLevel/fdr_ancomNoZ_Acc_Bac.RData")

## Load Phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

### Extract DA Phylum
#### Extracting data from tests
```{r}
# From each data frame, extract the Phylum or ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASV_Phylum <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASV_Phylum <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASV_Phylum <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASV_Phylum <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_Phylum <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_Phylum <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))
```

#### Merging tests and transposing list
```{r}
Acc_DESeq_DA_ASV_Phylum_l <- list(DESeqLRT = Acc_Bac_DESeqLRT_ASV_Phylum, DESeqWald = Acc_Bac_DESeqWald_ASV_Phylum, zinbLRT = Acc_Bac_zinbLRT_ASV_Phylum, zinbWald = Acc_Bac_zinbWald_ASV_Phylum)

# transpose list
Acc <- names(Acc_DESeq_DA_ASV_Phylum_l[[1]])
Acc_DESeq_DA_ASV_Phylum_l_t <- vector("list", length(Acc))
names(Acc_DESeq_DA_ASV_Phylum_l_t) <- Acc

for (cd in Acc) {
  Acc_DESeq_DA_ASV_Phylum_l_t[[cd]] <- lapply(Acc_DESeq_DA_ASV_Phylum_l, function(m) m[[cd]])
}

## DA Phylum Ancom
Acc_Ancom_DA_Phylum_l <- list(ancomWZ = Acc_Bac_ancomWZ_Phylum, ancomNoZ = Acc_Bac_ancomNoZ_Phylum)

# transpose list
Acc_Ancom_DA_Phylum_l_t <- vector("list", length(Acc))
names(Acc_Ancom_DA_Phylum_l_t) <- Acc

for (cd in Acc) {
  Acc_Ancom_DA_Phylum_l_t[[cd]] <- lapply(Acc_Ancom_DA_Phylum_l, function(m) m[[cd]])
}
```

### Extract taxomomy
#### DESeq
```{r}
tax <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
tax$ASV <- rownames(tax)

Acc_NoDup_DESeq_DA_ASV_Phylum_Tax <- lapply(Acc_DESeq_DA_ASV_Phylum_l_t, function(x) {
  lapply(x, function(y) {
    unique(tax[tax$ASV %in% y, c("Phylum")])
    })})
```

#### Acnom
```{r}
#lapply(Acc_Ancom_DA_Phylum_l, function(x) length(x))

Acc_NoDup_Ancom_DA_Phylum_tax <- lapply(Acc_Ancom_DA_Phylum_l_t, function(y) {
  lapply(y, function(x) {
    # Take unique names in case of duplicates
    DA_tax <- unique(x)
    
    # Split data between long and short taxonomic names
    full_tax_strings   <- DA_tax[grepl("k__", DA_tax)]
    family_only_strings <- DA_tax[!grepl("k__", DA_tax)]
    
    # Extract taxomomy from long names
    tax_full <- tibble(Taxon = full_tax_strings) %>%
      mutate(
        Phylum = str_extract(Taxon, "p__.*?(?=_c__|$)"),
        Class  = str_extract(Taxon, "c__.*?(?=_o__|$)")
      ) %>% select(Phylum, Class)
    
    # Extract taxonomy from ps object for short names
    tax_ps <- as.data.frame(tax_table(FP_unnormalized_bac_ps_ForDA_clean))
    #family_only <- sub("^f__", "", family_only_strings)
    
    tax_family_map <- unique(tax_ps[tax_ps$Phylum %in% family_only_strings, c("Class", "Phylum")])
    #nrow(tax_family_map)
    
    # Remove left over Incertae Sedis from short names object
    tax_family_map <- tax_family_map %>%
      filter(Phylum != "f__Incertae_Sedis")
   
    # Combine long and short names
    tax_combined <- as.data.frame(bind_rows(tax_full, tax_family_map))
    tax_combined <- tax_combined$Phylum
})})

#lapply(Acc_NoDup_Ancom_DA_Phylum_tax, function(x) nrow(x))
```

#### Merge Ancom and DESeq
```{r}
Acc_NoDup_DESeq_DA_ASV_Phylum_Tax_m <- lapply(Acc_NoDup_DESeq_DA_ASV_Phylum_Tax, function(x) unlist(x, use.names = FALSE))
Acc_NoDup_Ancom_DA_Phylum_tax_m <- lapply(Acc_NoDup_Ancom_DA_Phylum_tax, function(x) unlist(x, use.names = FALSE))

# Merge and filter on unique family
Acc_Total_Merged_DA_Phylum <- mapply(function(y,z){
  input <- c(y,z)},
  y = Acc_NoDup_DESeq_DA_ASV_Phylum_Tax_m,
  z = Acc_NoDup_Ancom_DA_Phylum_tax_m,
  SIMPLIFY = FALSE)

lapply(Acc_Total_Merged_DA_Phylum, function(x) length(x))
lapply(Acc_Total_Merged_DA_Phylum, function(x) table(x))

save(Acc_Total_Merged_DA_Phylum, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_Total_Merged_DA_Phylum.RData")
```

### Remove duplicat Phylum
```{r}
Acc_NoDup_Merged_DA_Phylum <- lapply(Acc_Total_Merged_DA_Phylum, function(x) unique(x))
lapply(Acc_NoDup_Merged_DA_Phylum, function(x) length(x))
lapply(Acc_NoDup_Merged_DA_Phylum, function(x) table(x))

save(Acc_NoDup_Merged_DA_Phylum, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_NoDup_Merged_DA_Phylum.RData")
```

### DA Phylum occurring twice
```{r}
# Filter to Phylum detected by at least 2 tests
Acc_TwoTimes_Merged_DA_Phylum <- lapply(Acc_Total_Merged_DA_Phylum, function(x) {
    dup <- x[duplicated(x) | duplicated(x, fromLast = TRUE), drop = FALSE]
    unique(dup)
  })

# Number of DA Phylum occurring in two tests
lapply(Acc_TwoTimes_Merged_DA_Phylum, function(x) length(x))

save(Acc_TwoTimes_Merged_DA_Phylum, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_TwoTimes_Merged_DA_Phylum.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


## 5.3.6 All data ASV level
### Load DA ASVs
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbWald_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_DESeq_ASVLevel/sigtab_zinbLRT_All_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomWZ_All_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/fdr_ancomNoZ_All_Bac.RData")
```

### Extract DA ASVs
```{r}
## From each data frame, extract the ASV IDs caught by each test
All_Bac_DESeqLRT_ASVs <- rownames(sigtab_stddds2_LRT_All_Bac)
All_Bac_DESeqWald_ASVs <- rownames(sigtab_stddds2_Wald_All_Bac)
All_Bac_zinbLRT_ASVs <- rownames(sigtab_zinbLRT_All_Bac)
All_Bac_zinbWald_ASVs <- rownames(sigtab_zinbWald_All_Bac)
All_Bac_ancomWZ_ASVs <- as.vector(fdr_ancomWZ_All_Bac[ ,'Species'])
All_Bac_ancomNoZ_ASVs <- as.vector(fdr_ancomNoZ_All_Bac[ ,'Species'])

All_Bac_Total_DA_ASV <- c(All_Bac_DESeqLRT_ASVs, All_Bac_DESeqWald_ASVs, All_Bac_zinbLRT_ASVs, All_Bac_zinbWald_ASVs, All_Bac_ancomWZ_ASVs, All_Bac_ancomNoZ_ASVs)
```

### Total number of ASVs
```{r}
# Total number of DA ASVs
length(All_Bac_Total_DA_ASV)

save(All_Bac_Total_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_Total_DA_ASV.RData")

# Remove duplicated asvs names from that list
All_Bac_NoDup_DA_ASV <- unique(All_Bac_Total_DA_ASV)

# Number of DA ASVs without duplicates
length(All_Bac_NoDup_DA_ASV)

save(All_Bac_NoDup_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_NoDup_DA_ASV.RData")

# Extracting taxonomy
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_unnormalized_bac_ps.RData")

# Extract taxonomy of DA ASVs
FP_unnormalized_bac_ps_Tax <- as.data.frame(tax_table(FP_unnormalized_bac_ps))
All_Bac_NoDup_DA_ASV_Tax <- FP_unnormalized_bac_ps_Tax[All_Bac_NoDup_DA_ASV, ]

save(All_Bac_NoDup_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_NoDup_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
All_Bac_NoDup_DA_ASV_Tax_Table <- lapply(All_Bac_NoDup_DA_ASV_Tax[ , 2:6], function(df) table(df))
All_Bac_NoDup_DA_ASV_Tax_Table <- lapply(All_Bac_NoDup_DA_ASV_Tax_Table, as.data.frame)

# Examples
All_Bac_NoDup_DA_ASV_Tax_Table$Class

save(All_Bac_NoDup_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_NoDup_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring twice
```{r}
# Filter to ASVs detected by at least 2 tests
All_Bac_NumbOcc_DA_ASV <- table(unlist(All_Bac_Total_DA_ASV))
All_Bac_TwoTimes_DA_ASV <- names(All_Bac_NumbOcc_DA_ASV[All_Bac_NumbOcc_DA_ASV >= 2])
All_Bac_TwoTimes_DA_ASV

# Number of DA ASVs occurring in two tests
length(All_Bac_TwoTimes_DA_ASV)

save(All_Bac_TwoTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
All_Bac_TwoTimes_DA_ASV_Tax <- FP_unnormalized_bac_ps_Tax[All_Bac_TwoTimes_DA_ASV, ]

save(All_Bac_TwoTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
All_Bac_TwoTimes_DA_ASV_Tax_Table <- lapply(All_Bac_TwoTimes_DA_ASV_Tax[ , 2:6], function(df) table(df))
All_Bac_TwoTimes_DA_ASV_Tax_Table <- lapply(All_Bac_TwoTimes_DA_ASV_Tax_Table, as.data.frame)

# Examples
All_Bac_TwoTimes_DA_ASV_Tax_Table$Class

save(All_Bac_TwoTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring three times
```{r}
# Filter to ASVs detected by at least 3 tests
All_Bac_ThreeTimes_DA_ASV <- names(All_Bac_NumbOcc_DA_ASV[All_Bac_NumbOcc_DA_ASV >= 3])
All_Bac_ThreeTimes_DA_ASV

# Number of DA ASVs occurring in three tests
length(All_Bac_ThreeTimes_DA_ASV)
save(All_Bac_ThreeTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_ThreeTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
All_Bac_ThreeTimes_DA_ASV_Tax <- FP_unnormalized_bac_ps_Tax[All_Bac_ThreeTimes_DA_ASV, ]

save(All_Bac_ThreeTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_ThreeTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
All_Bac_ThreeTimes_DA_ASV_Tax_Table <- lapply(All_Bac_ThreeTimes_DA_ASV_Tax[ , 2:6], function(df) table(df))
All_Bac_ThreeTimes_DA_ASV_Tax_Table <- lapply(All_Bac_ThreeTimes_DA_ASV_Tax_Table, as.data.frame)

# Examples
All_Bac_ThreeTimes_DA_ASV_Tax_Table

save(All_Bac_ThreeTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_ThreeTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring four times
```{r}
# Filter to ASVs detected by at least 3 tests
All_Bac_FourTimes_DA_ASV <- names(All_Bac_NumbOcc_DA_ASV[All_Bac_NumbOcc_DA_ASV >= 4])
All_Bac_FourTimes_DA_ASV

# Number of DA ASVs occurring in three tests
length(All_Bac_FourTimes_DA_ASV)
save(All_Bac_FourTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_FourTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
All_Bac_FourTimes_DA_ASV_Tax <- FP_unnormalized_bac_ps_Tax[All_Bac_FourTimes_DA_ASV, ]

save(All_Bac_FourTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_FourTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
All_Bac_FourTimes_DA_ASV_Tax_Table <- lapply(All_Bac_FourTimes_DA_ASV_Tax[ , 2:6], function(df) table(df))
All_Bac_FourTimes_DA_ASV_Tax_Table <- lapply(All_Bac_FourTimes_DA_ASV_Tax_Table, as.data.frame)

# Examples
All_Bac_FourTimes_DA_ASV_Tax_Table

save(All_Bac_FourTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_FourTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring five times
```{r}
# Filter to ASVs detected by at least 3 tests
All_Bac_FiveTimes_DA_ASV <- names(All_Bac_NumbOcc_DA_ASV[All_Bac_NumbOcc_DA_ASV >= 5])
All_Bac_FiveTimes_DA_ASV

# Number of DA ASVs occurring in three tests
length(All_Bac_FiveTimes_DA_ASV)
save(All_Bac_FiveTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_FiveTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
All_Bac_FiveTimes_DA_ASV_Tax <- FP_unnormalized_bac_ps_Tax[All_Bac_FiveTimes_DA_ASV, ]

save(All_Bac_FiveTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_FiveTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
All_Bac_FiveTimes_DA_ASV_Tax_Table <- lapply(All_Bac_FiveTimes_DA_ASV_Tax[ , 2:6], function(df) table(df))
All_Bac_FiveTimes_DA_ASV_Tax_Table <- lapply(All_Bac_FiveTimes_DA_ASV_Tax_Table, as.data.frame)

# Examples
All_Bac_FiveTimes_DA_ASV_Tax_Table

save(All_Bac_FiveTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_FiveTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring six times
```{r}
# Filter to ASVs detected by at least 3 tests
All_Bac_SixTimes_DA_ASV <- names(All_Bac_NumbOcc_DA_ASV[All_Bac_NumbOcc_DA_ASV >= 6])
All_Bac_SixTimes_DA_ASV

# Number of DA ASVs occurring in three tests
length(All_Bac_SixTimes_DA_ASV)
save(All_Bac_SixTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_SixTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
All_Bac_SixTimes_DA_ASV_Tax <- FP_unnormalized_bac_ps_Tax[All_Bac_SixTimes_DA_ASV, ]

save(All_Bac_SixTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_SixTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
All_Bac_SixTimes_DA_ASV_Tax_Table <- lapply(All_Bac_SixTimes_DA_ASV_Tax[ , 2:6], function(df) table(df))
All_Bac_SixTimes_DA_ASV_Tax_Table <- lapply(All_Bac_SixTimes_DA_ASV_Tax_Table, as.data.frame)

# Examples
All_Bac_SixTimes_DA_ASV_Tax_Table

save(All_Bac_SixTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_SixTimes_DA_ASV_Tax_Table.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


## 5.3.5 All data Class and Phylum level
No DA classes and phyla detected when analysing whole data.
