---
title: "FP_07_differential_abundance_Summary_Results Heatmaps logfoldchange"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b635.html)
- [Excamples of pheatmap package](https://davetang.github.io/muse/pheatmap.html)   
   
   
# 5.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(pheatmap)
packageVersion("pheatmap")
library(dplyr)
packageVersion("dplyr")
library(stringr)
packageVersion("stringr")
```

# 5.7 Heatmap log2fold change
## 5.7.1 Accessions seperate at ASV level
### Heatmap At ASV level DA ASVs occuring twice
#### Load data
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
```

#### Remove RI
```{r}
Acc_Bac_TwoTimes_DA_ASV <- Acc_Bac_TwoTimes_DA_ASV[names(Acc_Bac_TwoTimes_DA_ASV) != "RI"]
```

#### Combine DA ASVs per accession
```{r}
# Number of DA ASVs
lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) length(x))
length(unique(unlist(Acc_Bac_TwoTimes_DA_ASV)))

# Filter log2fold change for DA ASVs per accession
lfc_DA_specific <- lapply(names(Acc_Bac_TwoTimes_DA_ASV), function(name) {
  da_asvs <- Acc_Bac_TwoTimes_DA_ASV[[name]]
  lfc_table <- ancomWZ_Bac_Acc[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name)
  lfc_table
})

names(lfc_DA_specific) <- names(Acc_Bac_TwoTimes_DA_ASV)

# merge all accessions together
AllAcc_lfc <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific)
rownames(AllAcc_lfc) <- AllAcc_lfc$ASV
nrow(AllAcc_lfc)
AllAcc_lfc$ASV <- NULL
```

#### Prepare taxonomy column class and phylum
```{r}
# Put all unique DA ASVs in one vector
Acc_Bac_TwoTimes_DA_ASV <- unique(unlist(Acc_Bac_TwoTimes_DA_ASV, use.names = FALSE))

# Only keep DA in ps object
FP_unnormalized_bac_ps_ForDA_clean <- prune_taxa(Acc_Bac_TwoTimes_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df_clean <- psmelt(FP_unnormalized_bac_ps_ForDA_clean)
head(df_clean[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df_clean[!duplicated(df_clean$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Keep only those genera present in your heat map matrix
tax <- tax[match(rownames(AllAcc_lfc), tax$OTU), ]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)

# Add genus to matrix
AllAcc_lfc2 <- AllAcc_lfc
AllAcc_lfc2$ASV <- rownames(AllAcc_lfc2)
tax$ASV <- rownames(tax)
AllAcc_lfc2 <- merge(tax, AllAcc_lfc2, ID = "ASV")
rownames(AllAcc_lfc2) <- paste(AllAcc_lfc2$ASV, AllAcc_lfc2$Genus, sep = "_")
AllAcc_lfc2$Phylum <- sub("p__", "", AllAcc_lfc2$Phylum)
AllAcc_lfc2$Class <- sub("c__", "", AllAcc_lfc2$Class)

AllAcc_lfc3_df <- AllAcc_lfc2
tax2 <- AllAcc_lfc2[1:6]
AllAcc_lfc2 <- AllAcc_lfc2[7:10]
head(AllAcc_lfc2)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7", 
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000", 
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB", 
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD", 
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

classes <- unique(tax$Class)
class_colors <- setNames(colors35[1:length(classes)], classes)
class_colors <- list(Class = class_colors)

classes2 <- unique(tax2$Class)
class_colors2 <- setNames(colors35[1:length(classes2)], classes2)
class_colors2 <- list(Class = class_colors2)

phylums <- unique(tax$Phylum)
phylums_colors <- setNames(colors35[1:length(phylums)], phylums)
phylums_colors <- list(Phylum = phylums_colors)

# Make color pallet for Phylum for genus in plot
phylums2 <- unique(tax2$Phylum)
phylums_colors2 <- setNames(colors35[1:length(phylums2)], phylums2)
phylums_colors2 <- list(Phylum = phylums_colors2)

phylum_class_colors <- c(phylums_colors2, class_colors2)

# Turn data frame into a matrix
AllAcc_lfc <- as.matrix(AllAcc_lfc)
AllAcc_lfc2 <- as.matrix(AllAcc_lfc2)

# replace NAs by 0s
AllAcc_lfc_clu <- AllAcc_lfc
AllAcc_lfc2_clu <- AllAcc_lfc2
AllAcc_lfc_clu[is.na(AllAcc_lfc_clu)] <- 0
AllAcc_lfc2_clu[is.na(AllAcc_lfc2_clu)] <- 0

# counting number of ASVs per accession
ASV_numb <- lapply(lfc_DA_specific, function(x) nrow(x))
ASV_numb_df <- data.frame(Accession = names(ASV_numb), `Number of ASVs` = unlist(ASV_numb))
ASV_numb_df$Accession <- NULL
colnames(ASV_numb_df) <- "Number of ASVs"
ASV_numb_df

# counting number of enriched and depleted ASVs per accession
ASV_numb_pn <- lapply(lfc_DA_specific, function(x) {
  lfc <- x[, 1]
  lfc_pos <- lfc[lfc > 0]
  lfc_neg <- lfc[lfc < 0]
  output <- c(Pos = length(lfc_pos), Neg = length(lfc_neg))
})
ASV_numb_pn

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=8, fig.width=6}
pheatmap(AllAcc_lfc,
          clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = FALSE,
          annotation_row = tax["Phylum"],
          annotation_colors = phylums_colors,
          annotation_col = ASV_numb_df)
```
```{r, fig.height=34, fig.width=8}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_row = 6,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax2[5:6],
          annotation_colors = phylum_class_colors,
          annotation_col = ASV_numb_df)
```
```{r, fig.height=9, fig.width=6}
pheatmap(AllAcc_lfc,
          clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax["Class"],
          annotation_colors = class_colors,
          annotation_col = ASV_numb_df)
```

#### Most occuring taxa
```{r}
# Order by phylum
tab_phyl <- as.data.frame(table(AllAcc_lfc3_df$Phylum))
tab_phyl <- tab_phyl[order(tab_phyl$Freq, decreasing = TRUE), ]

# select top 7 phyla
top7_phyla <- as.character(tab_phyl$Var1[1:7])

# Order by Class
tab_cla <- as.data.frame(table(AllAcc_lfc3_df$Class))
tab_cla <- tab_cla[order(tab_cla$Freq, decreasing = TRUE), ]

# select top 7 classes
top7_cla <- as.character(tab_cla$Var1[1:7])

# replace other classes and phyla with "Other"
AllAcc_lfc3 <- AllAcc_lfc3_df
AllAcc_lfc3$Phylum <- ifelse(AllAcc_lfc3$Phylum %in% top7_phyla,
                             AllAcc_lfc3$Phylum,
                             "Other")
AllAcc_lfc3$Class <- ifelse(AllAcc_lfc3$Class %in% top7_cla,
                             AllAcc_lfc3$Class,
                             "Other")

# put other last for plotting
AllAcc_lfc3$Phylum <- factor(AllAcc_lfc3$Phylum,
                             levels = c(top7_phyla, "Other"))
AllAcc_lfc3$Class <- factor(AllAcc_lfc3$Class,
                             levels = c(top7_cla, "Other"))

# Subset data frames
tax3 <- AllAcc_lfc3[1:6]
AllAcc_lfc3 <- AllAcc_lfc3[7:10]

# Turn data frame into a matrix
AllAcc_lfc3 <- as.matrix(AllAcc_lfc3)

# Define colors annotation
phylums3 <- levels(tax3$Phylum)
phylums_colors3 <- setNames(colors35[1:(length(phylums3)-1)], phylums3[-length(phylums3)])
phylums_colors3 <- c(phylums_colors3, Other = "#BBBBBB")
phylums_colors3 <- list(Phylum = phylums_colors3)

classes3 <- levels(tax3$Class)
classes_colors3 <- setNames(colors35[1:(length(classes3)-1)], classes3[-length(classes3)])
classes_colors3 <- c(classes_colors3, Other = "#BBBBBB")
classes_colors3 <- list(Class = classes_colors3)

phylum_class_colors3 <- c(phylums_colors3, classes_colors3)

# replace NAs by 0s
AllAcc_lfc3_clu <- AllAcc_lfc3
AllAcc_lfc3_clu[is.na(AllAcc_lfc3_clu)] <- 0
```


##### Heatmap 
```{r, fig.height=8, fig.width=6}
pheatmap(AllAcc_lfc3,
          clustering_distance_rows = dist(AllAcc_lfc3_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc3_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = FALSE,
          annotation_row = tax3["Phylum"],
          annotation_colors = phylums_colors3,
          annotation_col = ASV_numb_df)
```
```{r, fig.height=8, fig.width=6}
p <- pheatmap(AllAcc_lfc3,
          clustering_distance_rows = dist(AllAcc_lfc3_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc3_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = FALSE,
          annotation_row = tax3["Class"],
          annotation_colors = classes_colors3,
          annotation_col = ASV_numb_df)
p 

files <- c("FP_DAHeatmap_ASV.svg", "FP_DAHeatmap_ASV.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 12/2.54, height = 18/2.54)
  } else {
    png(file_path, width = 12, height = 18, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)

# Higher plot
files <- c("FP_DAHeatmap_ASV_high.svg", "FP_DAHeatmap_ASV_high.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 12/2.54, height = 21/2.54)
  } else {
    png(file_path, width = 12, height = 21, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=8, fig.width=6}
pheatmap(AllAcc_lfc3,
          clustering_distance_rows = dist(AllAcc_lfc3_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc3_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = FALSE,
          annotation_row = tax3[5:6],
          annotation_colors = phylum_class_colors3,
          annotation_col = ASV_numb_df)
```

##### Heatmap ordered by phylum
```{r}
# Create ordering index
ord <- order(tax3$Phylum, tax3$Class, tax3$Order, tax3$Family)

# Reorder matrix
AllAcc_lfc3 <- AllAcc_lfc3[ord, ]

# Reorder annotation accordingly
tax3 <- tax3[ord, , drop = FALSE]
```

```{r, fig.height=8, fig.width=6}
pheatmap(AllAcc_lfc3,
          #clustering_distance_rows = dist(AllAcc_lfc3_clu),
          cluster_rows = FALSE,
          clustering_distance_cols = dist(t(AllAcc_lfc3_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = FALSE,
          annotation_row = tax3["Phylum"],
          annotation_colors = phylums_colors3,
          annotation_col = ASV_numb_df)
```
```{r, fig.height=8, fig.width=6}
pheatmap(AllAcc_lfc3,
          #clustering_distance_rows = dist(AllAcc_lfc3_clu),
          cluster_rows = FALSE,
          clustering_distance_cols = dist(t(AllAcc_lfc3_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = FALSE,
          annotation_row = tax3["Class"],
          annotation_colors = classes_colors3,
          annotation_col = ASV_numb_df)
```

#### Heatmap orderded by different taxonomic levels
##### Phylum
```{r, fig.height=8, fig.width=6}
# Create ordering index
ord <- order(tax$Phylum, tax$Class, tax$Order, tax$Family)

# Reorder matrix
AllAcc_lfc <- AllAcc_lfc[ord, ]

# Reorder annotation accordingly
tax <- tax[ord, , drop = FALSE]

# Heatmap
pheatmap(AllAcc_lfc,
         cluster_rows = FALSE,
          #clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax["Phylum"],
          annotation_colors = class_colors,
          annotation_col = ASV_numb_df)
```

##### Class
```{r, fig.height=10, fig.width=6}
# Make sure legend is in same direction
tax$Class <- factor(tax$Class, levels = unique(tax$Class))

# Heatmap
pheatmap(AllAcc_lfc,
         cluster_rows = FALSE,
          #clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax["Class"],
          annotation_colors = phylums_colors, # if using class_colors, reorder color vector as well
          annotation_col = ASV_numb_df)
```

##### Order
```{r, fig.height=14.5, fig.width=6}
# Make sure legend is in same direction
tax$Order <- factor(tax$Order, levels = unique(tax$Order))

# Heatmap
pheatmap(AllAcc_lfc,
         cluster_rows = FALSE,
          #clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax["Order"],
          annotation_colors = class_colors,
          #annotation_col = ASV_numb_df
         )
```

##### Family
```{r, fig.height=23, fig.width=6}
# Make sure legend is in same direction
tax$Family <- factor(tax$Family, levels = unique(tax$Family))

# Heatmap
pheatmap(AllAcc_lfc,
         cluster_rows = FALSE,
          #clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax["Family"],
          annotation_colors = class_colors,
          #annotation_col = ASV_numb_df
         )
```

##### All together
```{r, fig.height=10, fig.width=6}
# Reorder color class vector
class_colors <- class_colors[levels(tax$Family)]

# Heatmap
pheatmap(AllAcc_lfc,
         cluster_rows = FALSE,
          #clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax[2:5],
          annotation_colors = class_colors,
          #annotation_col = ASV_numb_df
         )
```
```{r, fig.height=46, fig.width=6}
# Heatmap
pheatmap(AllAcc_lfc,
         cluster_rows = FALSE,
          #clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          show_rownames = FALSE,
          fontsize_row = 3,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90, 
          annotation_row = tax[2:5],
          annotation_colors = class_colors,
          #annotation_col = ASV_numb_df
         )
```
```{r}
# Clean environment
rm(list = ls())
```


## 5.7.6 DA overlap HM and VL
### ASV level
#### Loading data
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub//Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")

# Extracting overlapping ASVs HM and VL
Acc_Bac_HM_VL_DA_ASV <- c(Acc_Bac_TwoTimes_DA_ASV$VL, Acc_Bac_TwoTimes_DA_ASV$HM)
Acc_Bac_NumbOcc_DA_ASV <- table(Acc_Bac_HM_VL_DA_ASV)
Acc_Bac_HMVLoverlap_DA_ASV <- names(Acc_Bac_NumbOcc_DA_ASV[Acc_Bac_NumbOcc_DA_ASV == 2])
Acc_Bac_HMVLoverlap_DA_ASV

# Select overlap of GO1 and CD with ASV list of HM and VL
ASVs_GO1 <- intersect(Acc_Bac_HMVLoverlap_DA_ASV, Acc_Bac_TwoTimes_DA_ASV$GO1)
ASVs_CD <- intersect(Acc_Bac_HMVLoverlap_DA_ASV, Acc_Bac_TwoTimes_DA_ASV$CD)
ASV_list <- list(HM = Acc_Bac_HMVLoverlap_DA_ASV, VL = Acc_Bac_HMVLoverlap_DA_ASV, GO1 = ASVs_GO1, CD = ASVs_CD)

# Filter log2fold change for DA ASVs per accession
lfc_DA_specific <- lapply(names(ASV_list), function(name) {
  da_asvs <- ASV_list[[name]]
  lfc_table <- ancomWZ_Bac_Acc[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name)
  lfc_table
})

names(lfc_DA_specific) <- names(ASV_list)

# merge all accessions together
lfc_table <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific)
rownames(lfc_table) <- lfc_table$ASV
nrow(lfc_table)
lfc_table$ASV <- NULL

# Change order accessions
lfc_table <- lfc_table[c(3, 1, 2, 4)]

# check table
head(lfc_table)
```

#### Prepare taxonomy column class and phylum
```{r}
# Only keep DA in ps object
FP_unnormalized_bac_ps_ForDA_clean <- prune_taxa(Acc_Bac_HMVLoverlap_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df_clean <- psmelt(FP_unnormalized_bac_ps_ForDA_clean)
head(df_clean[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df_clean[!duplicated(df_clean$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Keep only those genera present in your heat map matrix
tax <- tax[match(rownames(lfc_table), tax$OTU), ]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)

# Add genus in row names to new data frame
lfc_table$ASV <- rownames(lfc_table)
tax$ASV <- rownames(tax)
lfc_table <- merge(tax, lfc_table, ID = "ASV")

# Add higher taxonomic level if genus is NA or Incertae Sedis
taxon <- lfc_table$Genus
taxon[is.na(taxon)] <- lfc_table$Family[is.na(taxon)]
taxon[is.na(taxon)] <- lfc_table$Order[is.na(taxon)]

taxon[taxon == "g__Incertae_Sedis"] <- lfc_table$Family[taxon == "g__Incertae_Sedis"]
taxon[taxon == "f__Incertae_Sedis"] <- lfc_table$Order[taxon == "f__Incertae_Sedis"]
taxon[taxon == "o__Incertae_Sedis"] <- lfc_table$Class[taxon == "o__Incertae_Sedis"]

taxon[taxon == "g__Unclassified"] <- lfc_table$Family[taxon == "g__Unclassified"]
taxon[taxon == "f__Unclassified"] <- lfc_table$Order[taxon == "f__Unclassified"]
taxon[taxon == "o__Unclassified"] <- lfc_table$Class[taxon == "o__Unclassified"]

rownames(lfc_table) <- paste(lfc_table$ASV, taxon, sep = "_")

lfc_table$Phylum <- sub("p__", "", lfc_table$Phylum)
lfc_table$Class <- sub("c__", "", lfc_table$Class)

# make new dfs
tax <- lfc_table[1:6]
lfc_table <- lfc_table[c(8,9,10,7)]
head(lfc_table)
head(tax)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7", 
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000", 
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB", 
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD", 
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

classes <- unique(tax$Class)
class_colors <- setNames(colors35[1:length(classes)], classes)
class_colors <- list(Class = class_colors)

phylums <- unique(tax$Phylum)
phylums_colors <- setNames(colors35[1:length(phylums)], phylums)
phylums_colors <- list(Phylum = phylums_colors)

phylum_class_colors <- c(phylums_colors, class_colors)

# Row labels of genus in italic 
rn <- rownames(lfc_table)

## Create expression labels
row_labels <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels <- as.expression(row_labels)

# Turn data frame into a matrix
lfc_table <- as.matrix(lfc_table)
```

#### Heatmap HM and VL
```{r, fig.height=6, fig.width=6.3}
max_abs <- max(abs(lfc_table), na.rm = TRUE)

pheatmap(lfc_table[, 1:2],
          cluster_rows = TRUE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          #breaks = breaks,
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Phylum"],
          annotation_colors = phylums_colors)
```
```{r, fig.height=6, fig.width=6.3}
pheatmap(lfc_table[, 1:2],
          cluster_rows = TRUE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Class"],
          annotation_colors = class_colors)
```
```{r, fig.height=6, fig.width=6.3}
pheatmap(lfc_table[, 1:2],
          cluster_rows = TRUE,
          cluster_cols = FALSE,
         color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax[5:6],
          annotation_colors = phylum_class_colors)
```

#### Heatmap four accessions
```{r, fig.height=6, fig.width=7.5}
max_abs <- max(abs(lfc_table), na.rm = TRUE)

pheatmap(lfc_table,
          cluster_rows = TRUE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          #breaks = breaks,
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Phylum"],
          annotation_colors = phylums_colors)
```
```{r, fig.height=6, fig.width=7.5}
p <- pheatmap(lfc_table,
          cluster_rows = TRUE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Class"],
          annotation_colors = class_colors,
          labels_row = row_labels)
p 

files <- c("FP_DAHeatmap_ASV_overlap_HMVL.svg", "FP_DAHeatmap_ASV_overlap_HMVL.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 18/2.54, height = 12/2.54)
  } else {
    png(file_path, width = 18, height = 12, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=6, fig.width=7.5}
pheatmap(lfc_table,
          cluster_rows = TRUE,
          cluster_cols = FALSE,
         color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax[5:6],
          annotation_colors = phylum_class_colors)
```
```{r, fig.height=6, fig.width=6.4}
# Create ordering index
ord <- order(tax$Phylum, tax$Class, tax$Order, tax$Family, tax$Genus)

# Reorder matrix
lfc_table <- lfc_table[ord, ]

# Reorder annotation accordingly
tax <- tax[ord, , drop = FALSE]

# Reorder color class vector
phylums_colors <- phylums_colors[levels(tax$Phylum)]

# Make sure legend is in same direction
tax$Phylum <- factor(tax$Phylum, levels = unique(tax$Phylum))
tax$Class <- factor(tax$Class, levels = unique(tax$Class))
tax$Order <- factor(tax$Order, levels = unique(tax$Order))
tax$Family <- factor(tax$Family, levels = unique(tax$Family))
tax$Genus <- factor(tax$Genus, levels = unique(tax$Genus))

# Heatmap
pheatmap(lfc_table,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
         color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax[3:6],
          annotation_colors = phylums_colors)
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.7.8 Check specific taxonomic groups
### Bacilli
Heatmap At ASV level DA ASVs occuring twice.   

#### Load data
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
```

#### Remove RI
```{r}
Acc_Bac_TwoTimes_DA_ASV <- Acc_Bac_TwoTimes_DA_ASV[names(Acc_Bac_TwoTimes_DA_ASV) != "RI"]
```

#### Combine DA ASVs per accession
```{r}
# Number of DA ASVs
lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) length(x))
length(unique(unlist(Acc_Bac_TwoTimes_DA_ASV)))

# Filter log2fold change for DA ASVs per accession
lfc_DA_specific <- lapply(names(Acc_Bac_TwoTimes_DA_ASV), function(name) {
  da_asvs <- Acc_Bac_TwoTimes_DA_ASV[[name]]
  lfc_table <- ancomWZ_Bac_Acc[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name)
  lfc_table
})

names(lfc_DA_specific) <- names(Acc_Bac_TwoTimes_DA_ASV)

# merge all accessions together
AllAcc_lfc <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific)
rownames(AllAcc_lfc) <- AllAcc_lfc$ASV
nrow(AllAcc_lfc)
#AllAcc_lfc$ASV <- NULL

# Change order accessions
AllAcc_lfc <- AllAcc_lfc[c(1,4,2,3,5)]
```

#### Prepare taxonomy column class and phylum
```{r}
# Put all unique DA ASVs in one vector
Acc_Bac_TwoTimes_DA_ASV <- unique(unlist(Acc_Bac_TwoTimes_DA_ASV, use.names = FALSE))

# Only keep DA in ps object
FP_unnormalized_bac_ps_ForDA_clean <- prune_taxa(Acc_Bac_TwoTimes_DA_ASV, FP_unnormalized_bac_ps_ForDA_clean)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df_clean <- psmelt(FP_unnormalized_bac_ps_ForDA_clean)
head(df_clean[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df_clean[!duplicated(df_clean$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Keep only those genera present in your heat map matrix
tax <- tax[match(rownames(AllAcc_lfc), tax$OTU), ]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)
```

#### Subset Bacilli
```{r}
# Subset taxonomy table
tax_bas <- tax[tax$Class == "c__Bacilli", ]
tax_bas

# Extract ASV numbers Bacilli ASVs
ASVs_bas <- rownames(tax_bas)

# Subset lfc matrix with Bacilli
AllAcc_lfc_bas <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_bas, ]
AllAcc_lfc_bas
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_bas$ASV <- rownames(AllAcc_lfc_bas)
tax_bas$ASV <- rownames(tax_bas)
AllAcc_lfc_bas <- merge(tax_bas, AllAcc_lfc_bas, ID = "ASV")

# Add higher taxon_basomic level if genus is NA or Incertae Sedis
taxon_bas <- AllAcc_lfc_bas$Genus
taxon_bas[is.na(taxon_bas)] <- AllAcc_lfc_bas$Family[is.na(taxon_bas)]
taxon_bas[is.na(taxon_bas)] <- AllAcc_lfc_bas$Order[is.na(taxon_bas)]

taxon_bas[taxon_bas == "g__Incertae_Sedis"] <- AllAcc_lfc_bas$Family[taxon_bas == "g__Incertae_Sedis"]
taxon_bas[taxon_bas == "f__Incertae_Sedis"] <- AllAcc_lfc_bas$Order[taxon_bas == "f__Incertae_Sedis"]
taxon_bas[taxon_bas == "o__Incertae_Sedis"] <- AllAcc_lfc_bas$Class[taxon_bas == "o__Incertae_Sedis"]

taxon_bas[taxon_bas == "g__Unclassified"] <- AllAcc_lfc_bas$Family[taxon_bas == "g__Unclassified"]
taxon_bas[taxon_bas == "f__Unclassified"] <- AllAcc_lfc_bas$Order[taxon_bas == "f__Unclassified"]
taxon_bas[taxon_bas == "o__Unclassified"] <- AllAcc_lfc_bas$Class[taxon_bas == "o__Unclassified"]

rownames(AllAcc_lfc_bas) <- paste(AllAcc_lfc_bas$ASV, taxon_bas, sep = "_")

AllAcc_lfc_bas$Phylum <- sub("p__", "", AllAcc_lfc_bas$Phylum)
AllAcc_lfc_bas$Class <- sub("c__", "", AllAcc_lfc_bas$Class)
AllAcc_lfc_bas$Order <- sub("o__", "", AllAcc_lfc_bas$Order)
AllAcc_lfc_bas$Family <- sub("f__", "", AllAcc_lfc_bas$Family)

AllAcc_lfc_bas3_df <- AllAcc_lfc_bas
tax_bas2 <- AllAcc_lfc_bas[1:6]
AllAcc_lfc_bas <- AllAcc_lfc_bas[7:10]
head(AllAcc_lfc_bas)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_bas2$Family)
Family_colors2_bas2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_bas2 <- list(Family = Family_colors2_bas2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_bas2$Order)
Order_colors2_bas2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_bas2 <- list(Order = Order_colors2_bas2)

Order_Family_colors2_bas <- c(Order_colors2_bas2, Family_colors2_bas2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_bas)

## Create expression labels
row_labels_bas <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_bas <- as.expression(row_labels_bas)

# Turn data frame into a matrix
AllAcc_lfc_bas <- as.matrix(AllAcc_lfc_bas)

# replace NAs by 0s
AllAcc_lfc_bas_clu <- AllAcc_lfc_bas
AllAcc_lfc_bas_clu[is.na(AllAcc_lfc_bas_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_bas), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=5, fig.width=8}
pheatmap(AllAcc_lfc_bas,
          clustering_distance_rows = dist(AllAcc_lfc_bas_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_bas_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_bas2[3:4],
          annotation_colors = Order_Family_colors2_bas,
         labels_row = row_labels_bas)
```

#### Heatmap order by taxonomy
```{r, fig.height=4.5, fig.width=6}
# Create ordering index
ord <- order(tax_bas2$Phylum, tax_bas2$Class, tax_bas2$Order, tax_bas2$Family, tax_bas2$Genus)

# Reorder matrix
AllAcc_lfc_bas <- AllAcc_lfc_bas[ord, ]

# Reorder labels rows
row_labels_bas <- row_labels_bas[ord]

# Reorder annotation accordingly
tax_bas2 <- tax_bas2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_bas2$Phylum <- factor(tax_bas2$Phylum, levels = unique(tax_bas2$Phylum))
tax_bas2$Class <- factor(tax_bas2$Class, levels = unique(tax_bas2$Class))
tax_bas2$Order <- factor(tax_bas2$Order, levels = unique(tax_bas2$Order))
tax_bas2$Family <- factor(tax_bas2$Family, levels = unique(tax_bas2$Family))
tax_bas2$Genus <- factor(tax_bas2$Genus, levels = unique(tax_bas2$Genus))

# Reorder color class vector
Order_Family_colors2_bas$Order <- Order_Family_colors2_bas$Order[levels(tax_bas2$Order)]
Order_Family_colors2_bas$Family <- Order_Family_colors2_bas$Family[levels(tax_bas2$Family)]
```

```{r, fig.height=4.5, fig.width=6}
# Heatmap
p <- pheatmap(AllAcc_lfc_bas,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_bas2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_bas2[3:4],
          annotation_colors = Order_Family_colors2_bas,
         labels_row = row_labels_bas)
p 

files <- c("FP_DAHeatmap_ASV_Bacilli.svg", "FP_DAHeatmap_ASV_Bacili.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 7/2.54)
  } else {
    png(file_path, width = 20, height = 7, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=4.5, fig.width=6}
rm(p)
```

## Gammaproteobacteria
#### Subset Gammaproteobacteria
```{r}
# Subset taxonomy table
tax_gam <- tax[tax$Class == "c__Gammaproteobacteria", ]
tax_gam

# Extract ASV numbers Bacilli ASVs
ASVs_gam <- rownames(tax_gam)

# Subset lfc matrix with Bacilli
AllAcc_lfc_gam <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_gam, ]
AllAcc_lfc_gam
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_gam$ASV <- rownames(AllAcc_lfc_gam)
tax_gam$ASV <- rownames(tax_gam)
AllAcc_lfc_gam <- merge(tax_gam, AllAcc_lfc_gam, ID = "ASV")

# Add higher taxon_gamomic level if genus is NA or Incertae Sedis
taxon_gam <- AllAcc_lfc_gam$Genus
taxon_gam[is.na(taxon_gam)] <- AllAcc_lfc_gam$Family[is.na(taxon_gam)]
taxon_gam[is.na(taxon_gam)] <- AllAcc_lfc_gam$Order[is.na(taxon_gam)]

taxon_gam[taxon_gam == "g__Incertae_Sedis"] <- AllAcc_lfc_gam$Family[taxon_gam == "g__Incertae_Sedis"]
taxon_gam[taxon_gam == "f__Incertae_Sedis"] <- AllAcc_lfc_gam$Order[taxon_gam == "f__Incertae_Sedis"]
taxon_gam[taxon_gam == "o__Incertae_Sedis"] <- AllAcc_lfc_gam$Class[taxon_gam == "o__Incertae_Sedis"]

taxon_gam[taxon_gam == "g__Unclassified"] <- AllAcc_lfc_gam$Family[taxon_gam == "g__Unclassified"]
taxon_gam[taxon_gam == "f__Unclassified"] <- AllAcc_lfc_gam$Order[taxon_gam == "f__Unclassified"]
taxon_gam[taxon_gam == "o__Unclassified"] <- AllAcc_lfc_gam$Class[taxon_gam == "o__Unclassified"]

rownames(AllAcc_lfc_gam) <- paste(AllAcc_lfc_gam$ASV, taxon_gam, sep = "_")

AllAcc_lfc_gam$Phylum <- sub("p__", "", AllAcc_lfc_gam$Phylum)
AllAcc_lfc_gam$Class <- sub("c__", "", AllAcc_lfc_gam$Class)
AllAcc_lfc_gam$Order <- sub("o__", "", AllAcc_lfc_gam$Order)
AllAcc_lfc_gam$Family <- sub("f__", "", AllAcc_lfc_gam$Family)

AllAcc_lfc_gam3_df <- AllAcc_lfc_gam
tax_gam2 <- AllAcc_lfc_gam[1:6]
AllAcc_lfc_gam <- AllAcc_lfc_gam[7:10]
head(AllAcc_lfc_gam)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_gam2$Family)
Family_colors2_gam2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_gam2 <- list(Family = Family_colors2_gam2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_gam2$Order)
Order_colors2_gam2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_gam2 <- list(Order = Order_colors2_gam2)

Order_Family_colors2_gam <- c(Order_colors2_gam2, Family_colors2_gam2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_gam)

## Create expression labels
row_labels_gam <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_gam <- as.expression(row_labels_gam)

# Turn data frame into a matrix
AllAcc_lfc_gam <- as.matrix(AllAcc_lfc_gam)

# replace NAs by 0s
AllAcc_lfc_gam_clu <- AllAcc_lfc_gam
AllAcc_lfc_gam_clu[is.na(AllAcc_lfc_gam_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_gam), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=12, fig.width=8}
pheatmap(AllAcc_lfc_gam,
          clustering_distance_rows = dist(AllAcc_lfc_gam_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_gam_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          fontsize_row = 7,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_gam2[3:4],
          annotation_colors = Order_Family_colors2_gam,
         labels_row = row_labels_gam)
```

#### Heatmap order by taxonomy
```{r, fig.height=20, fig.width=10}
# Create ordering index
ord <- order(tax_gam2$Phylum, tax_gam2$Class, tax_gam2$Order, tax_gam2$Family, tax_gam2$Genus)

# Reorder matrix
AllAcc_lfc_gam <- AllAcc_lfc_gam[ord, ]

# Reorder labels rows
row_labels_gam <- row_labels_gam[ord]

# Reorder annotation accordingly
tax_gam2 <- tax_gam2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_gam2$Phylum <- factor(tax_gam2$Phylum, levels = unique(tax_gam2$Phylum))
tax_gam2$Class <- factor(tax_gam2$Class, levels = unique(tax_gam2$Class))
tax_gam2$Order <- factor(tax_gam2$Order, levels = unique(tax_gam2$Order))
tax_gam2$Family <- factor(tax_gam2$Family, levels = unique(tax_gam2$Family))
tax_gam2$Genus <- factor(tax_gam2$Genus, levels = unique(tax_gam2$Genus))

# Reorder color class vector
Order_Family_colors2_gam$Order <- Order_Family_colors2_gam$Order[levels(tax_gam2$Order)]
Order_Family_colors2_gam$Family <- Order_Family_colors2_gam$Family[levels(tax_gam2$Family)]
```

```{r, fig.height=12, fig.width=6.4}
# Heatmap
p <- pheatmap(AllAcc_lfc_gam,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_gam2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_gam2[3:4],
          annotation_colors = Order_Family_colors2_gam,
         labels_row = row_labels_gam)
p 

files <- c("FP_DAHeatmap_ASV_Gammaproteobacteria.svg", "FP_DAHeatmap_ASV_Gammaproteobacteria.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 34/2.54)
  } else {
    png(file_path, width = 20, height = 34, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```

## Alphaproteobacteria
#### Subset Alphaproteobacteria
```{r}
# Subset taxonomy table
tax_alp <- tax[tax$Class == "c__Alphaproteobacteria", ]
tax_alp

# Extract ASV numbers Bacilli ASVs
ASVs_alp <- rownames(tax_alp)

# Subset lfc matrix with Bacilli
AllAcc_lfc_alp <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_alp, ]
AllAcc_lfc_alp
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_alp$ASV <- rownames(AllAcc_lfc_alp)
tax_alp$ASV <- rownames(tax_alp)
AllAcc_lfc_alp <- merge(tax_alp, AllAcc_lfc_alp, ID = "ASV")

# Add higher taxon_alpomic level if genus is NA or Incertae Sedis
taxon_alp <- AllAcc_lfc_alp$Genus
taxon_alp[is.na(taxon_alp)] <- AllAcc_lfc_alp$Family[is.na(taxon_alp)]
taxon_alp[is.na(taxon_alp)] <- AllAcc_lfc_alp$Order[is.na(taxon_alp)]

taxon_alp[taxon_alp == "g__Incertae_Sedis"] <- AllAcc_lfc_alp$Family[taxon_alp == "g__Incertae_Sedis"]
taxon_alp[taxon_alp == "f__Incertae_Sedis"] <- AllAcc_lfc_alp$Order[taxon_alp == "f__Incertae_Sedis"]
taxon_alp[taxon_alp == "o__Incertae_Sedis"] <- AllAcc_lfc_alp$Class[taxon_alp == "o__Incertae_Sedis"]

taxon_alp[taxon_alp == "g__Unclassified"] <- AllAcc_lfc_alp$Family[taxon_alp == "g__Unclassified"]
taxon_alp[taxon_alp == "f__Unclassified"] <- AllAcc_lfc_alp$Order[taxon_alp == "f__Unclassified"]
taxon_alp[taxon_alp == "o__Unclassified"] <- AllAcc_lfc_alp$Class[taxon_alp == "o__Unclassified"]

rownames(AllAcc_lfc_alp) <- paste(AllAcc_lfc_alp$ASV, taxon_alp, sep = "_")

AllAcc_lfc_alp$Phylum <- sub("p__", "", AllAcc_lfc_alp$Phylum)
AllAcc_lfc_alp$Class <- sub("c__", "", AllAcc_lfc_alp$Class)
AllAcc_lfc_alp$Order <- sub("o__", "", AllAcc_lfc_alp$Order)
AllAcc_lfc_alp$Family <- sub("f__", "", AllAcc_lfc_alp$Family)

AllAcc_lfc_alp3_df <- AllAcc_lfc_alp
tax_alp2 <- AllAcc_lfc_alp[1:6]
AllAcc_lfc_alp <- AllAcc_lfc_alp[7:10]
head(AllAcc_lfc_alp)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_alp2$Family)
Family_colors2_alp2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_alp2 <- list(Family = Family_colors2_alp2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_alp2$Order)
Order_colors2_alp2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_alp2 <- list(Order = Order_colors2_alp2)

Order_Family_colors2_alp <- c(Order_colors2_alp2, Family_colors2_alp2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_alp)

## Create expression labels
row_labels_alp <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_alp <- as.expression(row_labels_alp)

# Turn data frame into a matrix
AllAcc_lfc_alp <- as.matrix(AllAcc_lfc_alp)

# replace NAs by 0s
AllAcc_lfc_alp_clu <- AllAcc_lfc_alp
AllAcc_lfc_alp_clu[is.na(AllAcc_lfc_alp_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_alp), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=12, fig.width=8}
pheatmap(AllAcc_lfc_alp,
          clustering_distance_rows = dist(AllAcc_lfc_alp_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_alp_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_alp2[3:4],
          annotation_colors = Order_Family_colors2_alp,
         labels_row = row_labels_alp)
```

#### Heatmap order by taxonomy
```{r, fig.height=12, fig.width=8}
# Create ordering index
ord <- order(tax_alp2$Phylum, tax_alp2$Class, tax_alp2$Order, tax_alp2$Family, tax_alp2$Genus)

# Reorder matrix
AllAcc_lfc_alp <- AllAcc_lfc_alp[ord, ]

# Reorder labels rows
row_labels_alp <- row_labels_alp[ord]

# Reorder annotation accordingly
tax_alp2 <- tax_alp2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_alp2$Phylum <- factor(tax_alp2$Phylum, levels = unique(tax_alp2$Phylum))
tax_alp2$Class <- factor(tax_alp2$Class, levels = unique(tax_alp2$Class))
tax_alp2$Order <- factor(tax_alp2$Order, levels = unique(tax_alp2$Order))
tax_alp2$Family <- factor(tax_alp2$Family, levels = unique(tax_alp2$Family))
tax_alp2$Genus <- factor(tax_alp2$Genus, levels = unique(tax_alp2$Genus))

# Reorder color class vector
Order_Family_colors2_alp$Order <- Order_Family_colors2_alp$Order[levels(tax_alp2$Order)]
Order_Family_colors2_alp$Family <- Order_Family_colors2_alp$Family[levels(tax_alp2$Family)]
```

```{r, fig.height=12, fig.width=9}
# Heatmap
p <- pheatmap(AllAcc_lfc_alp,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_alp2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_alp2[3:4],
          annotation_colors = Order_Family_colors2_alp,
         labels_row = row_labels_alp)
p 

files <- c("FP_DAHeatmap_ASV_Alphaproteobacteria.svg", "FP_DAHeatmap_ASV_Alphaproteobacteria.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 28/2.54)
  } else {
    png(file_path, width = 20, height = 28, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```

## Actinobacteria
#### Subset Actinobacteria
```{r}
# Subset taxonomy table
tax_act <- tax[tax$Class == "c__Actinobacteria", ]
tax_act

# Extract ASV numbers Bacilli ASVs
ASVs_act <- rownames(tax_act)

# Subset lfc matrix with Bacilli
AllAcc_lfc_act <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_act, ]
AllAcc_lfc_act
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_act$ASV <- rownames(AllAcc_lfc_act)
tax_act$ASV <- rownames(tax_act)
AllAcc_lfc_act <- merge(tax_act, AllAcc_lfc_act, ID = "ASV")

# Add higher taxon_actomic level if genus is NA or Incertae Sedis
taxon_act <- AllAcc_lfc_act$Genus
taxon_act[is.na(taxon_act)] <- AllAcc_lfc_act$Family[is.na(taxon_act)]
taxon_act[is.na(taxon_act)] <- AllAcc_lfc_act$Order[is.na(taxon_act)]

taxon_act[taxon_act == "g__Incertae_Sedis"] <- AllAcc_lfc_act$Family[taxon_act == "g__Incertae_Sedis"]
taxon_act[taxon_act == "f__Incertae_Sedis"] <- AllAcc_lfc_act$Order[taxon_act == "f__Incertae_Sedis"]
taxon_act[taxon_act == "o__Incertae_Sedis"] <- AllAcc_lfc_act$Class[taxon_act == "o__Incertae_Sedis"]

taxon_act[taxon_act == "g__Unclassified"] <- AllAcc_lfc_act$Family[taxon_act == "g__Unclassified"]
taxon_act[taxon_act == "f__Unclassified"] <- AllAcc_lfc_act$Order[taxon_act == "f__Unclassified"]
taxon_act[taxon_act == "o__Unclassified"] <- AllAcc_lfc_act$Class[taxon_act == "o__Unclassified"]

rownames(AllAcc_lfc_act) <- paste(AllAcc_lfc_act$ASV, taxon_act, sep = "_")

AllAcc_lfc_act$Phylum <- sub("p__", "", AllAcc_lfc_act$Phylum)
AllAcc_lfc_act$Class <- sub("c__", "", AllAcc_lfc_act$Class)
AllAcc_lfc_act$Order <- sub("o__", "", AllAcc_lfc_act$Order)
AllAcc_lfc_act$Family <- sub("f__", "", AllAcc_lfc_act$Family)

AllAcc_lfc_act3_df <- AllAcc_lfc_act
tax_act2 <- AllAcc_lfc_act[1:6]
AllAcc_lfc_act <- AllAcc_lfc_act[7:10]
head(AllAcc_lfc_act)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_act2$Family)
Family_colors2_act2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_act2 <- list(Family = Family_colors2_act2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_act2$Order)
Order_colors2_act2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_act2 <- list(Order = Order_colors2_act2)

Order_Family_colors2_act <- c(Order_colors2_act2, Family_colors2_act2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_act)

## Create expression labels
row_labels_act <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_act <- as.expression(row_labels_act)

# Turn data frame into a matrix
AllAcc_lfc_act <- as.matrix(AllAcc_lfc_act)

# replace NAs by 0s
AllAcc_lfc_act_clu <- AllAcc_lfc_act
AllAcc_lfc_act_clu[is.na(AllAcc_lfc_act_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_act), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=9, fig.width=8}
pheatmap(AllAcc_lfc_act,
          clustering_distance_rows = dist(AllAcc_lfc_act_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_act_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          fontsize_row = 7,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_act2[3:4],
          annotation_colors = Order_Family_colors2_act,
         labels_row = row_labels_act)
```

#### Heatmap order by taxonomy
```{r, fig.height=8, fig.width=7}
# Create ordering index
ord <- order(tax_act2$Phylum, tax_act2$Class, tax_act2$Order, tax_act2$Family, tax_act2$Genus)

# Reorder matrix
AllAcc_lfc_act <- AllAcc_lfc_act[ord, ]

# Reorder labels rows
row_labels_act <- row_labels_act[ord]

# Reorder annotation accordingly
tax_act2 <- tax_act2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_act2$Phylum <- factor(tax_act2$Phylum, levels = unique(tax_act2$Phylum))
tax_act2$Class <- factor(tax_act2$Class, levels = unique(tax_act2$Class))
tax_act2$Order <- factor(tax_act2$Order, levels = unique(tax_act2$Order))
tax_act2$Family <- factor(tax_act2$Family, levels = unique(tax_act2$Family))
tax_act2$Genus <- factor(tax_act2$Genus, levels = unique(tax_act2$Genus))

# Reorder color class vector
Order_Family_colors2_act$Order <- Order_Family_colors2_act$Order[levels(tax_act2$Order)]
Order_Family_colors2_act$Family <- Order_Family_colors2_act$Family[levels(tax_act2$Family)]
```

```{r, fig.height=8, fig.width=7}
# Heatmap
p <- pheatmap(AllAcc_lfc_act,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_act2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_act2[3:4],
          annotation_colors = Order_Family_colors2_act,
         labels_row = row_labels_act)
p 

files <- c("FP_DAHeatmap_ASV_Actinobacteria.svg", "FP_DAHeatmap_ASV_Actinobacteria.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 20/2.54)
  } else {
    png(file_path, width = 20, height = 20, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```

## Gemmatimonadia
#### Subset Gemmatimonadia
```{r}
# Subset taxonomy table
tax_gem <- tax[tax$Class == "c__Gemmatimonadia", ]
tax_gem

# Extrgem ASV numbers Bacilli ASVs
ASVs_gem <- rownames(tax_gem)

# Subset lfc matrix with Bacilli
AllAcc_lfc_gem <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_gem, ]
AllAcc_lfc_gem
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_gem$ASV <- rownames(AllAcc_lfc_gem)
tax_gem$ASV <- rownames(tax_gem)
AllAcc_lfc_gem <- merge(tax_gem, AllAcc_lfc_gem, ID = "ASV")

# Add higher taxon_gemomic level if genus is NA or Incertae Sedis
taxon_gem <- AllAcc_lfc_gem$Genus
taxon_gem[is.na(taxon_gem)] <- AllAcc_lfc_gem$Family[is.na(taxon_gem)]
taxon_gem[is.na(taxon_gem)] <- AllAcc_lfc_gem$Order[is.na(taxon_gem)]

taxon_gem[taxon_gem == "g__Incertae_Sedis"] <- AllAcc_lfc_gem$Family[taxon_gem == "g__Incertae_Sedis"]
taxon_gem[taxon_gem == "f__Incertae_Sedis"] <- AllAcc_lfc_gem$Order[taxon_gem == "f__Incertae_Sedis"]
taxon_gem[taxon_gem == "o__Incertae_Sedis"] <- AllAcc_lfc_gem$Class[taxon_gem == "o__Incertae_Sedis"]

taxon_gem[taxon_gem == "g__Unclassified"] <- AllAcc_lfc_gem$Family[taxon_gem == "g__Unclassified"]
taxon_gem[taxon_gem == "f__Unclassified"] <- AllAcc_lfc_gem$Order[taxon_gem == "f__Unclassified"]
taxon_gem[taxon_gem == "o__Unclassified"] <- AllAcc_lfc_gem$Class[taxon_gem == "o__Unclassified"]

rownames(AllAcc_lfc_gem) <- paste(AllAcc_lfc_gem$ASV, taxon_gem, sep = "_")

AllAcc_lfc_gem$Phylum <- sub("p__", "", AllAcc_lfc_gem$Phylum)
AllAcc_lfc_gem$Class <- sub("c__", "", AllAcc_lfc_gem$Class)
AllAcc_lfc_gem$Order <- sub("o__", "", AllAcc_lfc_gem$Order)
AllAcc_lfc_gem$Family <- sub("f__", "", AllAcc_lfc_gem$Family)

AllAcc_lfc_gem3_df <- AllAcc_lfc_gem
tax_gem2 <- AllAcc_lfc_gem[1:6]
AllAcc_lfc_gem <- AllAcc_lfc_gem[7:10]
head(AllAcc_lfc_gem)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_gem2$Family)
Family_colors2_gem2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_gem2 <- list(Family = Family_colors2_gem2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_gem2$Order)
Order_colors2_gem2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_gem2 <- list(Order = Order_colors2_gem2)

Order_Family_colors2_gem <- c(Order_colors2_gem2, Family_colors2_gem2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_gem)

## Create expression labels
row_labels_gem <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_gem <- as.expression(row_labels_gem)

# Turn data frame into a matrix
AllAcc_lfc_gem <- as.matrix(AllAcc_lfc_gem)

# replace NAs by 0s
AllAcc_lfc_gem_clu <- AllAcc_lfc_gem
AllAcc_lfc_gem_clu[is.na(AllAcc_lfc_gem_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_gem), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=6, fig.width=6}
pheatmap(AllAcc_lfc_gem,
          clustering_distance_rows = dist(AllAcc_lfc_gem_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_gem_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          fontsize_row = 7,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_gem2[3:4],
          annotation_colors = Order_Family_colors2_gem,
         labels_row = row_labels_gem)
```

#### Heatmap order by taxonomy
```{r, fig.height=6, fig.width=5}
# Create ordering index
ord <- order(tax_gem2$Phylum, tax_gem2$Class, tax_gem2$Order, tax_gem2$Family, tax_gem2$Genus)

# Reorder matrix
AllAcc_lfc_gem <- AllAcc_lfc_gem[ord, ]

# Reorder labels rows
row_labels_gem <- row_labels_gem[ord]

# Reorder annotation accordingly
tax_gem2 <- tax_gem2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_gem2$Phylum <- factor(tax_gem2$Phylum, levels = unique(tax_gem2$Phylum))
tax_gem2$Class <- factor(tax_gem2$Class, levels = unique(tax_gem2$Class))
tax_gem2$Order <- factor(tax_gem2$Order, levels = unique(tax_gem2$Order))
tax_gem2$Family <- factor(tax_gem2$Family, levels = unique(tax_gem2$Family))
tax_gem2$Genus <- factor(tax_gem2$Genus, levels = unique(tax_gem2$Genus))

# Reorder color class vector
Order_Family_colors2_gem$Order <- Order_Family_colors2_gem$Order[levels(tax_gem2$Order)]
Order_Family_colors2_gem$Family <- Order_Family_colors2_gem$Family[levels(tax_gem2$Family)]
```

```{r, fig.height=8, fig.width=7}
# Heatmap
p <- pheatmap(AllAcc_lfc_gem,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_gem2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_gem2[3:4],
          annotation_colors = Order_Family_colors2_gem,
         labels_row = row_labels_gem)
p 

files <- c("FP_DAHeatmap_ASV_Gemmatimonadia.svg", "FP_DAHeatmap_ASV_Gemmatimonadia.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 17/2.54)
  } else {
    png(file_path, width = 20, height = 17, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```

## Thermoleophilia
#### Subset Thermoleophilia
```{r}
# Subset taxonomy table
tax_the <- tax[tax$Class == "c__Thermoleophilia", ]
tax_the

# Extrthe ASV numbers Bacilli ASVs
ASVs_the <- rownames(tax_the)

# Subset lfc matrix with Bacilli
AllAcc_lfc_the <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_the, ]
AllAcc_lfc_the
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_the$ASV <- rownames(AllAcc_lfc_the)
tax_the$ASV <- rownames(tax_the)
AllAcc_lfc_the <- merge(tax_the, AllAcc_lfc_the, ID = "ASV")

# Add higher taxon_theomic level if genus is NA or Incertae Sedis
taxon_the <- AllAcc_lfc_the$Genus
taxon_the[is.na(taxon_the)] <- AllAcc_lfc_the$Family[is.na(taxon_the)]
taxon_the[is.na(taxon_the)] <- AllAcc_lfc_the$Order[is.na(taxon_the)]

taxon_the[taxon_the == "g__Incertae_Sedis"] <- AllAcc_lfc_the$Family[taxon_the == "g__Incertae_Sedis"]
taxon_the[taxon_the == "f__Incertae_Sedis"] <- AllAcc_lfc_the$Order[taxon_the == "f__Incertae_Sedis"]
taxon_the[taxon_the == "o__Incertae_Sedis"] <- AllAcc_lfc_the$Class[taxon_the == "o__Incertae_Sedis"]

taxon_the[taxon_the == "g__Unclassified"] <- AllAcc_lfc_the$Family[taxon_the == "g__Unclassified"]
taxon_the[taxon_the == "f__Unclassified"] <- AllAcc_lfc_the$Order[taxon_the == "f__Unclassified"]
taxon_the[taxon_the == "o__Unclassified"] <- AllAcc_lfc_the$Class[taxon_the == "o__Unclassified"]

rownames(AllAcc_lfc_the) <- paste(AllAcc_lfc_the$ASV, taxon_the, sep = "_")

AllAcc_lfc_the$Phylum <- sub("p__", "", AllAcc_lfc_the$Phylum)
AllAcc_lfc_the$Class <- sub("c__", "", AllAcc_lfc_the$Class)
AllAcc_lfc_the$Order <- sub("o__", "", AllAcc_lfc_the$Order)
AllAcc_lfc_the$Family <- sub("f__", "", AllAcc_lfc_the$Family)

AllAcc_lfc_the3_df <- AllAcc_lfc_the
tax_the2 <- AllAcc_lfc_the[1:6]
AllAcc_lfc_the <- AllAcc_lfc_the[7:10]
head(AllAcc_lfc_the)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_the2$Family)
Family_colors2_the2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_the2 <- list(Family = Family_colors2_the2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_the2$Order)
Order_colors2_the2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_the2 <- list(Order = Order_colors2_the2)

Order_Family_colors2_the <- c(Order_colors2_the2, Family_colors2_the2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_the)

## Create expression labels
row_labels_the <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_the <- as.expression(row_labels_the)

# Turn data frame into a matrix
AllAcc_lfc_the <- as.matrix(AllAcc_lfc_the)

# replace NAs by 0s
AllAcc_lfc_the_clu <- AllAcc_lfc_the
AllAcc_lfc_the_clu[is.na(AllAcc_lfc_the_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_the), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=5, fig.width=6}
pheatmap(AllAcc_lfc_the,
          clustering_distance_rows = dist(AllAcc_lfc_the_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_the_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          fontsize_row = 7,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_the2[3:4],
          annotation_colors = Order_Family_colors2_the,
         labels_row = row_labels_the)
```

#### Heatmap order by taxonomy
```{r, fig.height=5, fig.width=5}
# Create ordering index
ord <- order(tax_the2$Phylum, tax_the2$Class, tax_the2$Order, tax_the2$Family, tax_the2$Genus)

# Reorder matrix
AllAcc_lfc_the <- AllAcc_lfc_the[ord, ]

# Reorder labels rows
row_labels_the <- row_labels_the[ord]

# Reorder annotation accordingly
tax_the2 <- tax_the2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_the2$Phylum <- factor(tax_the2$Phylum, levels = unique(tax_the2$Phylum))
tax_the2$Class <- factor(tax_the2$Class, levels = unique(tax_the2$Class))
tax_the2$Order <- factor(tax_the2$Order, levels = unique(tax_the2$Order))
tax_the2$Family <- factor(tax_the2$Family, levels = unique(tax_the2$Family))
tax_the2$Genus <- factor(tax_the2$Genus, levels = unique(tax_the2$Genus))

# Reorder color class vector
Order_Family_colors2_the$Order <- Order_Family_colors2_the$Order[levels(tax_the2$Order)]
Order_Family_colors2_the$Family <- Order_Family_colors2_the$Family[levels(tax_the2$Family)]
```

```{r, fig.height=5, fig.width=6}
# Heatmap
p <- pheatmap(AllAcc_lfc_the,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_the2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_the2[3:4],
          annotation_colors = Order_Family_colors2_the,
         labels_row = row_labels_the)
p 

files <- c("FP_DAHeatmap_ASV_Thermoleophilia.svg", "FP_DAHeatmap_ASV_Thermoleophilia.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 9/2.54)
  } else {
    png(file_path, width = 20, height = 9, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```

## Ktedonobacteria
#### Subset Ktedonobacteria
```{r}
# Subset taxonomy table
tax_kte <- tax[tax$Class == "c__Ktedonobacteria", ]
tax_kte

# Extrkte ASV numbers Bacilli ASVs
ASVs_kte <- rownames(tax_kte)

# Subset lfc matrix with Bacilli
AllAcc_lfc_kte <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_kte, ]
AllAcc_lfc_kte
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_kte$ASV <- rownames(AllAcc_lfc_kte)
tax_kte$ASV <- rownames(tax_kte)
AllAcc_lfc_kte <- merge(tax_kte, AllAcc_lfc_kte, ID = "ASV")

# Add higher taxon_kteomic level if genus is NA or Incertae Sedis
taxon_kte <- AllAcc_lfc_kte$Genus
taxon_kte[is.na(taxon_kte)] <- AllAcc_lfc_kte$Family[is.na(taxon_kte)]
taxon_kte[is.na(taxon_kte)] <- AllAcc_lfc_kte$Order[is.na(taxon_kte)]

taxon_kte[taxon_kte == "g__Incertae_Sedis"] <- AllAcc_lfc_kte$Family[taxon_kte == "g__Incertae_Sedis"]
taxon_kte[taxon_kte == "f__Incertae_Sedis"] <- AllAcc_lfc_kte$Order[taxon_kte == "f__Incertae_Sedis"]
taxon_kte[taxon_kte == "o__Incertae_Sedis"] <- AllAcc_lfc_kte$Class[taxon_kte == "o__Incertae_Sedis"]

taxon_kte[taxon_kte == "g__Unclassified"] <- AllAcc_lfc_kte$Family[taxon_kte == "g__Unclassified"]
taxon_kte[taxon_kte == "f__Unclassified"] <- AllAcc_lfc_kte$Order[taxon_kte == "f__Unclassified"]
taxon_kte[taxon_kte == "o__Unclassified"] <- AllAcc_lfc_kte$Class[taxon_kte == "o__Unclassified"]

rownames(AllAcc_lfc_kte) <- paste(AllAcc_lfc_kte$ASV, taxon_kte, sep = "_")

AllAcc_lfc_kte$Phylum <- sub("p__", "", AllAcc_lfc_kte$Phylum)
AllAcc_lfc_kte$Class <- sub("c__", "", AllAcc_lfc_kte$Class)
AllAcc_lfc_kte$Order <- sub("o__", "", AllAcc_lfc_kte$Order)
AllAcc_lfc_kte$Family <- sub("f__", "", AllAcc_lfc_kte$Family)

AllAcc_lfc_kte3_df <- AllAcc_lfc_kte
tax_kte2 <- AllAcc_lfc_kte[1:6]
AllAcc_lfc_kte <- AllAcc_lfc_kte[7:10]
head(AllAcc_lfc_kte)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_kte2$Family)
Family_colors2_kte2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_kte2 <- list(Family = Family_colors2_kte2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_kte2$Order)
Order_colors2_kte2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_kte2 <- list(Order = Order_colors2_kte2)

Order_Family_colors2_kte <- c(Order_colors2_kte2, Family_colors2_kte2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_kte)

## Create expression labels
row_labels_kte <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_kte <- as.expression(row_labels_kte)

# Turn data frame into a matrix
AllAcc_lfc_kte <- as.matrix(AllAcc_lfc_kte)

# replace NAs by 0s
AllAcc_lfc_kte_clu <- AllAcc_lfc_kte
AllAcc_lfc_kte_clu[is.na(AllAcc_lfc_kte_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_kte), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=5, fig.width=7}
pheatmap(AllAcc_lfc_kte,
          clustering_distance_rows = dist(AllAcc_lfc_kte_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_kte_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_kte2[3:4],
          annotation_colors = Order_Family_colors2_kte,
         labels_row = row_labels_kte)
```

#### Heatmap order by taxonomy
```{r, fig.height=4, fig.width=6}
# Create ordering index
ord <- order(tax_kte2$Phylum, tax_kte2$Class, tax_kte2$Order, tax_kte2$Family, tax_kte2$Genus)

# Reorder matrix
AllAcc_lfc_kte <- AllAcc_lfc_kte[ord, ]

# Reorder labels rows
row_labels_kte <- row_labels_kte[ord]

# Reorder annotation accordingly
tax_kte2 <- tax_kte2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_kte2$Phylum <- factor(tax_kte2$Phylum, levels = unique(tax_kte2$Phylum))
tax_kte2$Class <- factor(tax_kte2$Class, levels = unique(tax_kte2$Class))
tax_kte2$Order <- factor(tax_kte2$Order, levels = unique(tax_kte2$Order))
tax_kte2$Family <- factor(tax_kte2$Family, levels = unique(tax_kte2$Family))
tax_kte2$Genus <- factor(tax_kte2$Genus, levels = unique(tax_kte2$Genus))

# Reorder color class vector
Order_Family_colors2_kte$Order <- Order_Family_colors2_kte$Order[levels(tax_kte2$Order)]
Order_Family_colors2_kte$Family <- Order_Family_colors2_kte$Family[levels(tax_kte2$Family)]
```

```{r, fig.height=4, fig.width=6}
# Heatmap
p <- pheatmap(AllAcc_lfc_kte,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_kte2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_kte2[3:4],
          annotation_colors = Order_Family_colors2_kte,
         labels_row = row_labels_kte)
p 

files <- c("FP_DAHeatmap_ASV_Ktedonobacteria.svg", "FP_DAHeatmap_ASV_Ktedonobacteria.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 7.5/2.54)
  } else {
    png(file_path, width = 20, height = 7.5, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```

## Other Classes
#### Subset Other
```{r}
# Most abundant classes
Ab_class <- c("c__Gammaproteobacteria", "c__Alphaproteobacteria", "c__Actinobacteria", "c__Gemmatimonadia", "c__Thermoleophilia", "c__Ktedonobacteria", "c__Bacilli")

# Subset taxonomy table
tax_Other <- tax[!(tax$Class %in% Ab_class), ]
tax_Other

# ExtrOther ASV numbers Bacilli ASVs
ASVs_Other <- rownames(tax_Other)

# Subset lfc matrix with Bacilli
AllAcc_lfc_Other <- AllAcc_lfc[AllAcc_lfc$ASV %in% ASVs_Other, ]
AllAcc_lfc_Other
```

#### Preparing heatmap
```{r}
# Add genus to matrix
AllAcc_lfc_Other$ASV <- rownames(AllAcc_lfc_Other)
tax_Other$ASV <- rownames(tax_Other)
AllAcc_lfc_Other <- merge(tax_Other, AllAcc_lfc_Other, ID = "ASV")

# Add higher taxon_Otheromic level if genus is NA or Incertae Sedis
taxon_Other <- AllAcc_lfc_Other$Genus
taxon_Other[is.na(taxon_Other)] <- AllAcc_lfc_Other$Family[is.na(taxon_Other)]
taxon_Other[is.na(taxon_Other)] <- AllAcc_lfc_Other$Order[is.na(taxon_Other)]

taxon_Other[taxon_Other == "g__Incertae_Sedis"] <- AllAcc_lfc_Other$Family[taxon_Other == "g__Incertae_Sedis"]
taxon_Other[taxon_Other == "f__Incertae_Sedis"] <- AllAcc_lfc_Other$Order[taxon_Other == "f__Incertae_Sedis"]
taxon_Other[taxon_Other == "o__Incertae_Sedis"] <- AllAcc_lfc_Other$Class[taxon_Other == "o__Incertae_Sedis"]

taxon_Other[taxon_Other == "g__Unclassified"] <- AllAcc_lfc_Other$Family[taxon_Other == "g__Unclassified"]
taxon_Other[taxon_Other == "f__Unclassified"] <- AllAcc_lfc_Other$Order[taxon_Other == "f__Unclassified"]
taxon_Other[taxon_Other == "o__Unclassified"] <- AllAcc_lfc_Other$Class[taxon_Other == "o__Unclassified"]

rownames(AllAcc_lfc_Other) <- paste(AllAcc_lfc_Other$ASV, taxon_Other, sep = "_")

AllAcc_lfc_Other$Phylum <- sub("p__", "", AllAcc_lfc_Other$Phylum)
AllAcc_lfc_Other$Class <- sub("c__", "", AllAcc_lfc_Other$Class)
AllAcc_lfc_Other$Order <- sub("o__", "", AllAcc_lfc_Other$Order)
AllAcc_lfc_Other$Family <- sub("f__", "", AllAcc_lfc_Other$Family)

AllAcc_lfc_Other3_df <- AllAcc_lfc_Other
tax_Other2 <- AllAcc_lfc_Other[1:6]
AllAcc_lfc_Other <- AllAcc_lfc_Other[7:10]
head(AllAcc_lfc_Other)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7",
              "#D55E00", "#999999", "#88CCEE", "#44AA99", "#0072B2", "#000000",
              "#117733", "#999933", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB",
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD",
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87")

Family2 <- unique(tax_Other2$Family)
Family_colors2_Other2 <- setNames(colors35[1:length(Family2)], Family2)
Family_colors2_Other2 <- list(Family = Family_colors2_Other2)

# Make color pallet for Order for genus in plot
Order2 <- unique(tax_Other2$Order)
Order_colors2_Other2 <- setNames(colors35[1:length(Order2)], Order2)
Order_colors2_Other2 <- list(Order = Order_colors2_Other2)

Class2 <- unique(tax_Other2$Class)
Class_colors2_Other2 <- setNames(colors35[1:length(Class2)], Class2)
Class_colors2_Other2 <- list(Class = Class_colors2_Other2)

# Make color pallet for Phylum for genus in plot
Phylum2 <- unique(tax_Other2$Phylum)
Phylum_colors2_Other2 <- setNames(colors35[1:length(Phylum2)], Phylum2)
Phylum_colors2_Other2 <- list(Phylum = Phylum_colors2_Other2)

colors2_Other <- c(Phylum_colors2_Other2, Class_colors2_Other2, Order_colors2_Other2, Family_colors2_Other2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc_Other)

## Create expression labels
row_labels_Other <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels_Other <- as.expression(row_labels_Other)

# Turn data frame into a matrix
AllAcc_lfc_Other <- as.matrix(AllAcc_lfc_Other)

# replace NAs by 0s
AllAcc_lfc_Other_clu <- AllAcc_lfc_Other
AllAcc_lfc_Other_clu[is.na(AllAcc_lfc_Other_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc_Other), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -1, length.out = 10), 
                   seq(-1, 1, length.out = 20), # More breaks in this range for sensitivity
                   seq(1, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=10, fig.width=9}
pheatmap(AllAcc_lfc_Other,
          clustering_distance_rows = dist(AllAcc_lfc_Other_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_Other_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_col = 14,
          fontsize_row = 7,
          na_col = "grey85",
          angle_col = 90,
          show_rownames = TRUE,
          annotation_row = tax_Other2[3:6],
          annotation_colors = colors2_Other,
         labels_row = row_labels_Other)
```

#### Heatmap order by taxonomy
```{r, fig.height=15, fig.width=8}
# Create ordering index
ord <- order(tax_Other2$Phylum, tax_Other2$Class, tax_Other2$Order, tax_Other2$Family, tax_Other2$Genus)

# Reorder matrix
AllAcc_lfc_Other <- AllAcc_lfc_Other[ord, ]

# Reorder labels rows
row_labels_Other <- row_labels_Other[ord]

# Reorder annotation accordingly
tax_Other2 <- tax_Other2[ord, , drop = FALSE]

# Make sure legend is in same direction
tax_Other2$Phylum <- factor(tax_Other2$Phylum, levels = unique(tax_Other2$Phylum))
tax_Other2$Class <- factor(tax_Other2$Class, levels = unique(tax_Other2$Class))
tax_Other2$Order <- factor(tax_Other2$Order, levels = unique(tax_Other2$Order))
tax_Other2$Family <- factor(tax_Other2$Family, levels = unique(tax_Other2$Family))
tax_Other2$Genus <- factor(tax_Other2$Genus, levels = unique(tax_Other2$Genus))

# Reorder color class vector
colors2_Other$Phylum <- colors2_Other$Phylum[levels(tax_Other2$Phylum)]
colors2_Other$Class  <- colors2_Other$Class[levels(tax_Other2$Class)]
colors2_Other$Order  <- colors2_Other$Order[levels(tax_Other2$Order)]
colors2_Other$Family <- colors2_Other$Family[levels(tax_Other2$Family)]

# Heatmap
pheatmap(AllAcc_lfc_Other,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_Other2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_Other2[3:6],
          annotation_colors = colors2_Other,
         labels_row = row_labels_Other)
```
```{r, fig.height=15, fig.width=8}
p <- pheatmap(AllAcc_lfc_Other,
          cluster_rows = FALSE,
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant tax_Other2a (log2FC by Family)",
          cellwidth = 25,
          cellheight = 9.5,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90,
          annotation_row = tax_Other2[c(3,5)],
          annotation_colors = colors2_Other,
         labels_row = row_labels_Other)
p 

files <- c("FP_DAHeatmap_ASV_OtherClasses.svg", "FP_DAHeatmap_ASV_OtherClasses.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 20/2.54, height = 36/2.54)
  } else {
    png(file_path, width = 20, height = 31, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=12, fig.width=6.4}
rm(p)
```
