---
title: "FP_10.1_Correlations - FP microbiome and FP caterpillar performance"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

# 6.0 load libraries
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(vegan)
packageVersion("vegan")
library(ggplot2)
packageVersion("ggplot2")
library(patchwork) # for wrap_plot
packageVersion("patchwork")
library(scales)
packageVersion("scales")
```

# 6.3 ASV - Caterpillar performance correlations HM
## Load data
```{r}
# Selection of DA ASVs of VL, HM, overlap VL and HM and DA in All data. Object made in upset plot file
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Selction_DA_ASVs.RData")

# load ps object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

## Set relative abundance
```{r}
# Calculate relative abundance
FP_ps <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))
```

## Prepare data set
```{r}
# make one vector with unique ASVs
ASVs <- Selction_DA_ASVs$HM_twoDAtests

# Only select HM
FP_ps <- subset_samples(FP_ps, Accession == "HM")

# prune ps object
FP_ps <- prune_taxa(ASVs, FP_ps) #only keep taxa that are in ASVs vector
#length(intersect(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean))) # check overlap
#setdiff(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean)) # check difference

# Phyloseq object to data frame with relative abundance, meta data and taxonomy
FP_ps <- psmelt(FP_ps)

# remove NAs cat weight
FP_ps[is.na(FP_ps$Mean_Weight_Cat), 1:10]
FP_ps <- FP_ps[!is.na(FP_ps$Mean_Weight_Cat), ]
```

### Remove ASVs with only 0s
```{r}
# identify ASVs with zero abundance
Abund_sum <- aggregate(FP_ps$Abundance, by = list(FP_ps$OTU), FUN = sum)
Abund_sum[Abund_sum$x == 0, ]
Abund_zero <- Abund_sum[Abund_sum$x == 0, "Group.1"]

# Remove ASVs
FP_ps <- FP_ps[!(FP_ps$OTU %in% Abund_zero), ]
ASVs <- ASVs[!ASVs %in% Abund_zero]
```

## Plot
### Function for plot
```{r fig.height=20, fig.width=20}
Plot1 <- function(x, title) {
  ggplot(data = FP_ps[FP_ps$OTU == x, ], aes(x = Abundance, y = Mean_Weight_Cat)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    # scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    ggtitle(title) +
    xlab("ASV abundance") +
    ylab("Caterpillar weight") +
    theme(panel.background = element_rect(fill = "white"), #theme of the graph
          panel.border = element_blank(),
          axis.line = element_line(), #show line for x and y axis
          plot.title = element_text(size = 12, hjust = 0),
          axis.text = element_text(size = 8, color = "black"),
          axis.title = element_text(size = 12),
          axis.title.x = element_text(vjust = -0.5),
          axis.title.y = element_text(vjust = 2.5), #, hjust = -0
          strip.text = element_text(size = 13),
          legend.position = "right")
}
```

## Model assumptions
### Assumptions with graphs
```{r}
NormTest <- function(dat, title) {
  hist(dat, prob = TRUE, col = "grey", main = title) #distribution histogram
  lines(density(dat), col = "blue", lwd = 2)
  
  qqnorm(dat)
  qqline(dat)
  
  print(shapiro.test(dat))
}

# Caterpillar performance
par(mfrow = c(1, 2))
NormTest(dat = FP_ps[FP_ps$OTU == "bASV_42", "Mean_Weight_Cat"], title = "b")
par(mfrow = c(1, 1))
```

### Assumptions only Shapriro Wilk test
```{r}
# Abundance
Shap_res_ab <- list()
for(i in 1:length(ASVs)) {
  Shap_res_ab[[i]] <- shapiro.test(FP_ps[FP_ps$OTU == ASVs[i], "Abundance"])
}
names(Shap_res_ab) <- ASVs
Shap_res_ab
```

## Correlation tests
### Run tests
```{r}
CorTest2 <- function(x) {
  sub_data <- FP_ps[FP_ps$OTU == x, ]
  cs0 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("pearson")) # parametric
  cs1 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("spearman")) # non parametric
  cs2 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("kendall")) # non parametric
  rm(sub_data)
  
  # Make data frame with results
  Sum1 <- data.frame(Pearson_est = cs0$estimate, Pearson_stat = cs0$statistic, Pearson_pv = cs0$p.value,
                     Spearman_est = cs1$estimate, Spearman_stat = cs1$statistic, Spearman_pv = cs1$p.value, 
                     Kendall_est = cs2$estimate, Kendall_stat = cs2$statistic, Kendall_pv = cs2$p.value)
  rownames(Sum1) <- x
  return(Sum1)
}

# Run function
List_CorTest2 <- list() # make an empty list to store the plots

for (i in 1:length(ASVs)) {
  List_CorTest2[[i]] <- CorTest2(x = ASVs[i])
}

names(List_CorTest2) <- ASVs
df_CorTest2 <- do.call(rbind.data.frame, List_CorTest2)
df_CorTest2$ASV <- row.names(df_CorTest2)

# Add taxonomy to correlation results
## Extract unique ASVs
FP_ps_ASV_unique <- FP_ps[!duplicated(FP_ps$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]
colnames(FP_ps_ASV_unique)[1] <- "ASV"

## Merge two data frames
df_CorTest2 <- merge(df_CorTest2, FP_ps_ASV_unique, by = "ASV")

# p value correction
df_CorTest2$Pearson_pv_pc <- p.adjust(df_CorTest2$Pearson_pv, method = "fdr")
df_CorTest2$Spearman_pv_pc <- p.adjust(df_CorTest2$Spearman_pv, method = "fdr")
df_CorTest2$Kendall_pv_pc <- p.adjust(df_CorTest2$Kendall_pv, method = "fdr")

# print head data frame
df_CorTest2[1:10, ]
nrow(df_CorTest2)
```

### Select low p values of not corrected p values
```{r}
#Pearson_0.05 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ]
Spearman_0.05 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.05, ]
nrow(Spearman_0.05)
Kendall_0.05 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.05, ]

#Pearson_0.01 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ]
Spearman_0.01 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.01, ]
nrow(Spearman_0.01)
Kendall_0.01 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.01, ]

SpearmanCorr_0.05_HM <- Spearman_0.05[ , c("ASV", "Spearman_est", "Spearman_stat", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
SpearmanCorr_0.05_HM
SpearmanCorr_0.05_neg_HM <- Spearman_0.05[Spearman_0.05$Spearman_est <= 0, c("ASV", "Spearman_est", "Spearman_stat",  "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
SpearmanCorr_0.05_neg_HM
Spearman_0.01[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01[Spearman_0.01$Spearman_est <= 0, c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]

# ASVs with low p values
# No pearson as data is not normally distributed !
table(c(Spearman_0.05$ASV, Kendall_0.05$ASV))
table(c(Spearman_0.01$ASV, Kendall_0.01$ASV))

# Store ASVs
ASVs_0.05 <- unique(c(Spearman_0.05$ASV))
ASVs_0.05_neg <- unique(c(Spearman_0.05[Spearman_0.05$Spearman_est <= 0, "ASV"]))
ASVs_0.01 <- unique(c(Spearman_0.01$ASV))
ASVs_0.01_neg <- unique(c(Spearman_0.01[Spearman_0.01$Spearman_est <= 0, "ASV"]))
ASVs_0.01_neg <- ASVs_0.01_neg[!is.na(ASVs_0.01_neg)]

length(ASVs_0.05_neg)
length(ASVs_0.01_neg)

#write.csv(SpearmanCorr_0.05_HM, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Statistical_output/FP/SpearmanCorr_0.05_HM.csv")
#write.csv(SpearmanCorr_0.05_neg_HM, "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Statistical_output/FP/SpearmanCorr_0.05_neg_HM.csv")
```

### Select low p values of corrected p values
```{r}
#Pearson_0.05c <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ]
Spearman_0.05c <- df_CorTest2[df_CorTest2$Spearman_pv_pc <= 0.05, ]
Kendall_0.05c <- df_CorTest2[df_CorTest2$Kendall_pv_pc <= 0.05, ]

#Pearson_0.01c <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ]
Spearman_0.01c <- df_CorTest2[df_CorTest2$Spearman_pv_pc <= 0.01, ]
Kendall_0.01c <- df_CorTest2[df_CorTest2$Kendall_pv_pc <= 0.01, ]

Spearman_0.05c[ , c("ASV", "Spearman_est", "Spearman_pv_pc", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01c[ , c("ASV", "Spearman_est", "Spearman_pv_pc", "Genus", "Family", "Class", "Phylum")]

# ASVs with low p values
# No pearson as data is not normally distributed !
table(c(Spearman_0.05c$ASV, Kendall_0.05c$ASV))
table(c(Spearman_0.01c$ASV, Kendall_0.01c$ASV))
table(c(Spearman_0.05c$ASV, Kendall_0.05c$ASV, Spearman_0.01c$ASV, Kendall_0.01c$ASV))

# Store ASVs
ASVs_0.05c <- unique(c(Spearman_0.05c$ASV))
ASVs_0.01c <- unique(c(Spearman_0.01c$ASV))

DAASVs_SpCorr_ASVCatPerf_HM <- list(ASVs_0.05 = ASVs_0.05, ASVs_0.01 = ASVs_0.01, ASVs_0.05_pc = ASVs_0.05c, ASVs_0.01_pc = ASVs_0.01c, ASVs_0.05_neg = ASVs_0.05_neg, ASVs_0.01_neg = ASVs_0.01_neg)

#save(DAASVs_SpCorr_ASVCatPerf_HM, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_CorrAnlysis/DAASVs_SpCorr_ASVCatPerf_HM.RData")
```

## Plot highly correlating ASVs
No p value correction used.
```{r}
# P value < 0.05
List_Graph_0.05 <- list()
for (i in 1:length(ASVs_0.05)) {
  List_Graph_0.05[[i]] <- Plot1(x = ASVs_0.05[i], title = ASVs_0.05[i])
}

# P value < 0.01
List_Graph_0.01 <- list()
for (i in 1:length(ASVs_0.01)) {
  List_Graph_0.01[[i]] <- Plot1(x = ASVs_0.01[i], title = ASVs_0.01[i])
}

# P value < 0.05  only negative correlations
List_Graph_0.05_neg <- list()
for (i in 1:length(ASVs_0.05_neg)) {
  List_Graph_0.05_neg[[i]] <- Plot1(x = ASVs_0.05_neg[i], title = ASVs_0.05_neg[i])
}

# P value < 0.01 only negative correlations
List_Graph_0.01_neg <- list()
for (i in 1:length(ASVs_0.01_neg)) {
  List_Graph_0.01_neg[[i]] <- Plot1(x = ASVs_0.01_neg[i], title = ASVs_0.01_neg[i])
}

#Plot1(x = "bASV_2431", title = "bASV_2431")
```

### Print plots
```{r, fig.height=20, fig.width=27}
# Print plots p value < 0.05
wrap_plots(List_Graph_0.05)
```
```{r, fig.height=7, fig.width=13}
# Print plots p value < 0.01
wrap_plots(List_Graph_0.01, ncol = 5)
```
```{r, fig.height=7, fig.width=9}
# Print plots p value < 0.05 only negative correlations
wrap_plots(List_Graph_0.05_neg)
```
```{r, fig.height=2.3, fig.width=7}
# Print plots p value < 0.01 only negative correlations
wrap_plots(List_Graph_0.01_neg)
```

### Plot for publication
#### Prepare data
```{r}
# Remove Unclassified and Incertae Sedis
taxon <- SpearmanCorr_0.05_neg_HM$Genus

taxon[taxon == "g__Incertae_Sedis"] <- SpearmanCorr_0.05_neg_HM$Family[taxon == "g__Incertae_Sedis"]
taxon[taxon == "f__Incertae_Sedis"] <- SpearmanCorr_0.05_neg_HM$Order[taxon == "f__Incertae_Sedis"]
taxon[taxon == "o__Incertae_Sedis"] <- SpearmanCorr_0.05_neg_HM$Class[taxon == "o__Incertae_Sedis"]
taxon[taxon == "c__Incertae_Sedis"] <- SpearmanCorr_0.05_neg_HM$Phylum[taxon == "c__Incertae_Sedis"]

taxon[taxon == "g__Unclassified"] <- SpearmanCorr_0.05_neg_HM$Family[taxon == "g__Unclassified"]
taxon[taxon == "f__Unclassified"] <- SpearmanCorr_0.05_neg_HM$Order[taxon == "f__Unclassified"]
taxon[taxon == "o__Unclassified"] <- SpearmanCorr_0.05_neg_HM$Class[taxon == "o__Unclassified"]
taxon[taxon == "c__Unclassified"] <- SpearmanCorr_0.05_neg_HM$Phylum[taxon == "c__Unclassified"]

taxon <- sub("g__", "", taxon)
taxon <- sub("f__", "", taxon)
SpearmanCorr_0.05_neg_HM$ASV2 <- sub("bASV_", "ASV ", SpearmanCorr_0.05_neg_HM$ASV)

SpearmanCorr_0.05_neg_HM$Name <- paste0(taxon)
```

#### Function for plot
```{r fig.height=20, fig.width=20}
Plot2 <- function(x, title, stats, xstep, italic_title = FALSE) {
  # Make title italic if requested
  if (italic_title) {
    title <- bquote(italic(.(title)))
  }
  
  # Plot
  ggplot(data = FP_ps[FP_ps$OTU == x, ], aes(x = Abundance, y = Mean_Weight_Cat)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    scale_x_continuous(#expand = c(0, 0),
                       labels = scales::label_scientific(),
                       breaks = seq(0, max(FP_ps[FP_ps$OTU == x, "Abundance"]), xstep), #steps on the y-axis
                       limits = c(0, max(FP_ps[FP_ps$OTU == x, "Abundance"]))) +
    scale_y_continuous(#expand = c(0, 0),        #start graph at y = min and x = min
                       breaks = seq(0, 15, 5), #steps on the y-axis
                       limits = c(0, 15)) +   #y limits
    labs(title = title,
         subtitle = stats) +
    #xlab("ASV abundance") +
    #ylab("Caterpillar weight (mg)") +
    xlab(NULL) +
    ylab(NULL) +
    theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(),
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 14, hjust = 0),
        plot.subtitle = element_text(size = 11),
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 14),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(10, 15, 10, 15, "points"),
        legend.position = "none")
}


# plot.title = element_text(size = 12, hjust = 0),
#           axis.text = element_text(size = 8, color = "black"),
#           axis.title = element_text(size = 12),
#           axis.title.x = element_text(vjust = -0.5),
#           axis.title.y = element_text(vjust = 2.5), #, hjust = -0
#           strip.text = element_text(size = 13),
# 
# panel.border = element_blank(), 
#           axis.line = element_line(), #show line for x and y axis
#           plot.title = element_text(size = 18, hjust = 0),
#           plot.subtitle = element_text(size = 12),
#           axis.text = element_text(size = 12, color = "black"),
#           axis.title = element_text(size = 19),
#           axis.title.x = element_text(vjust = -0.5),
#           axis.title.y = element_text(vjust = 2.5), #, hjust = -0
#           legend.title = element_text(size = 16),
#           legend.text = element_text(size = 14),
#           #strip.text = element_text(size = 13),
#           #plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
#           plot.margin = margin(10, 20, 10, 20, "points"), # top . bottom .
#           legend.position = "right")
```

#### Run function
```{r}
# Breaks x axis
Steps <- c(2e-4, 5e-4, 3e-4, 4e-4, 1.5e-4, 2e-4, 1e-3, 3e-3, 5e-4, 1.5e-3, 5e-4)

# Graph correlationns P value < 0.05  only negative correlations
List_Graph_0.05_neg2 <- list()
italic_plots <- c(1:2, 4:10) # plot number of which titles need to be in italic

for (i in 1:nrow(SpearmanCorr_0.05_neg_HM)) {
  title <- SpearmanCorr_0.05_neg_HM$Name[[i]]
  stats <- paste0("(b", 
                  SpearmanCorr_0.05_neg_HM$ASV2[[i]], 
                  ")", "\nρ = ", 
                  round(SpearmanCorr_0.05_neg_HM$Spearman_est[[i]], 3), 
                  " ; p = ",
                  round(SpearmanCorr_0.05_neg_HM$Spearman_pv[[i]], 3))
  List_Graph_0.05_neg2[[i]] <- Plot2(x = SpearmanCorr_0.05_neg_HM$ASV[i], 
                                     title = title, 
                                     stats = stats, 
                                     xstep = Steps[[i]],
                                     italic_title = i %in% italic_plots)
}
```

#### Only axis titles on outer plots
```{r}
ncol <- 4
n <- length(List_Graph_0.05_neg2)

# identify positions
left_plots   <- seq(1, n, by = ncol)
bottom_plots <- (n - ncol + 1):n

# add axis titles conditionally
List_Graph_0.05_neg2_mod <- lapply(seq_along(List_Graph_0.05_neg2), function(i) {
  p <- List_Graph_0.05_neg2[[i]]
  
  if (i %in% left_plots) {
    p <- p + ylab("Caterpillar weight (mg)")
  }
  
  if (i %in% bottom_plots) {
    p <- p + xlab("ASV abundance")
  }
  return(p)
})
```

#### Print and save plot
```{r, fig.height=7, fig.width=9}
# Print plots p value < 0.05 only negative correlations
wrap_plots(List_Graph_0.05_neg2_mod, ncol = 4) &
  theme(panel.spacing = unit(8, "lines")) #Allow larger margins ggplot

mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 11.3,
         height = 10.5,
         units = "cm",
         dpi = 300),
  x = c("FP_SpearCorr_HM_0.05_neg4.svg", "FP_SpearCorr_HM_0.05_neg4.png"))
# ncol = 3: width = 8.5 and highet = 13.7
```


## Extract sequences correlating ASVs
```{r}
# Sequences_HM_DAASVs_SpCorrCatPerf_0.05_neg <- refseq(FP_unnormalized_bac_ps_ForDA_clean)[names(refseq(FP_unnormalized_bac_ps_ForDA_clean)) %in% ASVs_0.05_neg]
# Sequences_HM_DAASVs_SpCorrCatPerf_0.05_neg
# 
# Biostrings::writeXStringSet(Sequences_HM_DAASVs_SpCorrCatPerf_0.05_neg, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_CorrAnlysis/Sequences_HM_DAASVs_SpCorrCatPerf_0.05_neg.fasta")
```
```{r}
# Clean environment
rm(list = ls())
```


# 6.4 ASV - Caterpillar performance correlations VL
## Load data
```{r}
# Selection of DA ASVs of VL, HM, overlap VL and HM and DA in All data. Object made in upset plot file
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Selction_DA_ASVs.RData")

# load ps object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

## Set relative abundance
```{r}
# Calculate relative abundance
FP_ps <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))
```

## Prepare data set
```{r}
# make one vector with unique ASVs
ASVs <- Selction_DA_ASVs$VL_twoDAtests

# Only select HM
FP_ps <- subset_samples(FP_ps, Accession == "VL")

# prune ps object
FP_ps <- prune_taxa(ASVs, FP_ps) #only keep taxa that are in ASVs vector
#length(intersect(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean))) # check overlap
#setdiff(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean)) # check difference

# Phyloseq object to data frame with relative abundance, meta data and taxonomy
FP_ps <- psmelt(FP_ps)

# remove NAs cat weight
FP_ps[is.na(FP_ps$Mean_Weight_Cat), 1:10]
FP_ps <- FP_ps[!is.na(FP_ps$Mean_Weight_Cat), ]
```

### Remove ASVs with only 0s
```{r}
# identify ASVs with zero abundance
Abund_sum <- aggregate(FP_ps$Abundance, by = list(FP_ps$OTU), FUN = sum)
Abund_sum[Abund_sum$x == 0, ]
Abund_zero <- Abund_sum[Abund_sum$x == 0, "Group.1"]

# Remove ASVs
FP_ps <- FP_ps[!(FP_ps$OTU %in% Abund_zero), ]
ASVs <- ASVs[!ASVs %in% Abund_zero]
```

## Plot
### Function for plot
```{r fig.height=20, fig.width=20}
Plot1 <- function(x, title) {
  ggplot(data = FP_ps[FP_ps$OTU == x, ], aes(x = Abundance, y = Mean_Weight_Cat)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    # scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    ggtitle(title) +
    xlab("ASV abundance") +
    ylab("Caterpillar weight") +
    theme(panel.background = element_rect(fill = "white"), #theme of the graph
          panel.border = element_blank(),
          axis.line = element_line(), #show line for x and y axis
          plot.title = element_text(size = 12, hjust = 0),
          axis.text = element_text(size = 8, color = "black"),
          axis.title = element_text(size = 12),
          axis.title.x = element_text(vjust = -0.5),
          axis.title.y = element_text(vjust = 2.5), #, hjust = -0
          strip.text = element_text(size = 13),
          legend.position = "right")
}
```

### Run function
```{r fig.height=20, fig.width=20}
#Plot1(x = "bASV_42", title = "bASV_42")

List_Graph1 <- list()
for (i in 1:length(ASVs)) {
  List_Graph1[[i]] <- Plot1(x = ASVs[i], title = ASVs[i])
}
```

### Print plots
```{r, fig.height=10, fig.width=13}
# Print plots
wrap_plots(List_Graph1[1:20])
```

## Model assumptions
### Assumptions with graphs
```{r}
NormTest <- function(dat, title) {
  hist(dat, prob = TRUE, col = "grey", main = title) #distribution histogram
  lines(density(dat), col = "blue", lwd = 2)
  
  qqnorm(dat)
  qqline(dat)
  
  print(shapiro.test(dat))
}

# Caterpillar performance
par(mfrow = c(1, 2))
#NormTest(dat = FP_ps[FP_ps$OTU == "bASV_42", "Mean_Weight_Cat"], title = "b")
par(mfrow = c(1, 1))
```

### Assumptions only Shapriro Wilk test
```{r}
# Abundance
Shap_res_ab <- list()
for(i in 1:length(ASVs)) {
  Shap_res_ab[[i]] <- shapiro.test(FP_ps[FP_ps$OTU == ASVs[i], "Abundance"])
}
names(Shap_res_ab) <- ASVs
Shap_res_ab
```

## Correlation tests
### Run tests
```{r}
CorTest2 <- function(x) {
  sub_data <- FP_ps[FP_ps$OTU == x, ]
  cs0 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("pearson")) # parametric
  cs1 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("spearman")) # non parametric
  cs2 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("kendall")) # non parametric
  rm(sub_data)
  
  # Make data frame with results
  Sum1 <- data.frame(Pearson_est = cs0$estimate, Pearson_stat = cs0$statistic, Pearson_pv = cs0$p.value,
                     Spearman_est = cs1$estimate, Spearman_stat = cs1$statistic, Spearman_pv = cs1$p.value, 
                     Kendall_est = cs2$estimate, Kendall_stat = cs2$statistic, Kendall_pv = cs2$p.value)
  rownames(Sum1) <- x
  return(Sum1)
}

# Run function
List_CorTest2 <- list() # make an empty list to store the plots

for (i in 1:length(ASVs)) {
  List_CorTest2[[i]] <- CorTest2(x = ASVs[i])
}

names(List_CorTest2) <- ASVs
df_CorTest2 <- do.call(rbind.data.frame, List_CorTest2)
df_CorTest2$ASV <- row.names(df_CorTest2)

# Add taxonomy to correlation results
## Extract unique ASVs
FP_ps_ASV_unique <- FP_ps[!duplicated(FP_ps$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]
colnames(FP_ps_ASV_unique)[1] <- "ASV"

## Merge two data frames
df_CorTest2 <- merge(df_CorTest2, FP_ps_ASV_unique, by = "ASV")

# p value correction
df_CorTest2$Pearson_pv_pc <- p.adjust(df_CorTest2$Pearson_pv, method = "fdr")
df_CorTest2$Spearman_pv_pc <- p.adjust(df_CorTest2$Spearman_pv, method = "fdr")
df_CorTest2$Kendall_pv_pc <- p.adjust(df_CorTest2$Kendall_pv, method = "fdr")

# print head data frame
df_CorTest2[1:10, ]
```

### Select low p values of not corrected p values
```{r}
#Pearson_0.05 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ]
Spearman_0.05 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.05, ]
Kendall_0.05 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.05, ]

#Pearson_0.01 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ]
Spearman_0.01 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.01, ]
Kendall_0.01 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.01, ]

Spearman_0.1 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.1, ]

Spearman_0.01[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Class", "Phylum")]
Spearman_0.05[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Class", "Phylum")]

# Nearly significant ASVs
Spearman_0.1[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Class", "Phylum")]
summary(df_CorTest2$Spearman_pv)

# ASVs with low p values
# No pearson as data is not normally distributed !
table(c(Spearman_0.05$ASV, Kendall_0.05$ASV))
table(c(Spearman_0.01$ASV, Kendall_0.01$ASV))

# Store ASVs
ASVs_0.05 <- unique(c(Spearman_0.05$ASV))
ASVs_0.01 <- unique(c(Spearman_0.01$ASV))
ASVs_0.1 <- unique(c(Spearman_0.1$ASV))
```

## Plot nearly correlating ASVs
No p value correction used.
```{r}
# P value < 0.05
List_Graph_0.1 <- list()
for (i in 1:length(ASVs_0.1)) {
  List_Graph_0.1[[i]] <- Plot1(x = ASVs_0.1[i], title = ASVs_0.1[i])
}
```

### Print plots
```{r, fig.height=4, fig.width=6}
# Print plots p value < 0.05
wrap_plots(List_Graph_0.1)
```
```{r}
# Clean environment
rm(list = ls())
```


# 6.5 ASV - Caterpillar performance correlations CD
## Load data
```{r}
# Selection of DA ASVs of VL, HM, overlap VL and HM and DA in All data. Object made in upset plot file
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load ps object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

## Remove Batch 1
```{r}
FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Batch != 1)
```

## Set relative abundance
```{r}
# Calculate relative abundance
FP_ps <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))
```

## Prepare data set
```{r}
# make one vector with unique ASVs
ASVs <- Acc_Bac_TwoTimes_DA_ASV$CD

# Only select CD
FP_ps <- subset_samples(FP_ps, Accession == "CD")

# prune ps object
FP_ps <- prune_taxa(ASVs, FP_ps) #only keep taxa that are in ASVs vector
#length(intersect(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean))) # check overlap
#setdiff(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean)) # check difference

# Phyloseq object to data frame with relative abundance, meta data and taxonomy
FP_ps <- psmelt(FP_ps)

# remove NAs cat weight
FP_ps[is.na(FP_ps$Mean_Weight_Cat), 1:10]
FP_ps <- FP_ps[!is.na(FP_ps$Mean_Weight_Cat), ]
```

### Remove ASVs with only 0s
```{r}
# identify ASVs with zero abundance
Abund_sum <- aggregate(FP_ps$Abundance, by = list(FP_ps$OTU), FUN = sum)
Abund_sum[Abund_sum$x == 0, ]
Abund_zero <- Abund_sum[Abund_sum$x == 0, "Group.1"]

# Remove ASVs
FP_ps <- FP_ps[!(FP_ps$OTU %in% Abund_zero), ]
ASVs <- ASVs[!ASVs %in% Abund_zero]
```

## Plot
### Function for plot
```{r fig.height=20, fig.width=20}
Plot1 <- function(x, title) {
  ggplot(data = FP_ps[FP_ps$OTU == x, ], aes(x = Abundance, y = Mean_Weight_Cat)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    # scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    ggtitle(title) +
    xlab("ASV abundance") +
    ylab("Caterpillar weight") +
    theme(panel.background = element_rect(fill = "white"), #theme of the graph
          panel.border = element_blank(),
          axis.line = element_line(), #show line for x and y axis
          plot.title = element_text(size = 12, hjust = 0),
          axis.text = element_text(size = 8, color = "black"),
          axis.title = element_text(size = 12),
          axis.title.x = element_text(vjust = -0.5),
          axis.title.y = element_text(vjust = 2.5), #, hjust = -0
          strip.text = element_text(size = 13),
          legend.position = "right")
}
```

### Run function
```{r fig.height=20, fig.width=20}
#Plot1(x = "bASV_42", title = "bASV_42")

List_Graph1 <- list()
for (i in 1:length(ASVs)) {
  List_Graph1[[i]] <- Plot1(x = ASVs[i], title = ASVs[i])
}
```

### Print plots
```{r, fig.height=10, fig.width=12}
# Print plots
wrap_plots(List_Graph1[1:20])
```

## Model assumptions
### Assumptions with graphs
```{r}
NormTest <- function(dat, title) {
  hist(dat, prob = TRUE, col = "grey", main = title) #distribution histogram
  lines(density(dat), col = "blue", lwd = 2)
  
  qqnorm(dat)
  qqline(dat)
  
  print(shapiro.test(dat))
}

# Caterpillar performance
par(mfrow = c(1, 2))
NormTest(dat = FP_ps[FP_ps$OTU == "bASV_61", "Mean_Weight_Cat"], title = "b")
par(mfrow = c(1, 1))
```

### Assumptions only Shapriro Wilk test
```{r}
# Abundance
Shap_res_ab <- list()
for(i in 1:length(ASVs)) {
  Shap_res_ab[[i]] <- shapiro.test(FP_ps[FP_ps$OTU == ASVs[i], "Abundance"])
}
names(Shap_res_ab) <- ASVs
Shap_res_ab
```

## Correlation tests
### Run tests
```{r}
CorTest2 <- function(x) {
  sub_data <- FP_ps[FP_ps$OTU == x, ]
  cs0 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("pearson")) # parametric
  cs1 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("spearman")) # non parametric
  cs2 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("kendall")) # non parametric
  rm(sub_data)
  
  # Make data frame with results
  Sum1 <- data.frame(Pearson_est = cs0$estimate, Pearson_stat = cs0$statistic, Pearson_pv = cs0$p.value,
                     Spearman_est = cs1$estimate, Spearman_stat = cs1$statistic, Spearman_pv = cs1$p.value, 
                     Kendall_est = cs2$estimate, Kendall_stat = cs2$statistic, Kendall_pv = cs2$p.value)
  rownames(Sum1) <- x
  return(Sum1)
}

# Run function
List_CorTest2 <- list() # make an empty list to store the plots

for (i in 1:length(ASVs)) {
  List_CorTest2[[i]] <- CorTest2(x = ASVs[i])
}

names(List_CorTest2) <- ASVs
df_CorTest2 <- do.call(rbind.data.frame, List_CorTest2)
df_CorTest2$ASV <- row.names(df_CorTest2)

# Add taxonomy to correlation results
## Extract unique ASVs
FP_ps_ASV_unique <- FP_ps[!duplicated(FP_ps$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]
colnames(FP_ps_ASV_unique)[1] <- "ASV"

## Merge two data frames
df_CorTest2 <- merge(df_CorTest2, FP_ps_ASV_unique, by = "ASV")

# p value correction
df_CorTest2$Pearson_pv_pc <- p.adjust(df_CorTest2$Pearson_pv, method = "fdr")
df_CorTest2$Spearman_pv_pc <- p.adjust(df_CorTest2$Spearman_pv, method = "fdr")
df_CorTest2$Kendall_pv_pc <- p.adjust(df_CorTest2$Kendall_pv, method = "fdr")

# print head data frame
df_CorTest2[1:10, ]
```

### Select low p values of not corrected p values
```{r}
#Pearson_0.05 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ]
Spearman_0.05 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.05, ]
Kendall_0.05 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.05, ]

#Pearson_0.01 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ]
Spearman_0.01 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.01, ]
Kendall_0.01 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.01, ]

Spearman_0.05[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.05[Spearman_0.05$Spearman_est <= 0, c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01[Spearman_0.01$Spearman_est <= 0, c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]

summary(df_CorTest2$Spearman_pv)

# ASVs with low p values
# No pearson as data is not normally distributed !
table(c(Spearman_0.05$ASV, Kendall_0.05$ASV))
table(c(Spearman_0.01$ASV, Kendall_0.01$ASV))

# Store ASVs
ASVs_0.05 <- unique(c(Spearman_0.05$ASV))
ASVs_0.05_neg <- unique(c(Spearman_0.05[Spearman_0.05$Spearman_est <= 0, "ASV"]))
ASVs_0.01 <- unique(c(Spearman_0.01$ASV))
ASVs_0.01_neg <- unique(c(Spearman_0.01[Spearman_0.01$Spearman_est <= 0, "ASV"]))
ASVs_0.01_neg <- ASVs_0.01_neg[!is.na(ASVs_0.01_neg)]
```

<!-- ### Select low p values of corrected p values -->
<!-- ```{r} -->
<!-- #Pearson_0.05c <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ] -->
<!-- Spearman_0.05c <- df_CorTest2[df_CorTest2$Spearman_pv_pc <= 0.05, ] -->
<!-- Kendall_0.05c <- df_CorTest2[df_CorTest2$Kendall_pv_pc <= 0.05, ] -->

<!-- #Pearson_0.01c <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ] -->
<!-- Spearman_0.01c <- df_CorTest2[df_CorTest2$Spearman_pv_pc <= 0.01, ] -->
<!-- Kendall_0.01c <- df_CorTest2[df_CorTest2$Kendall_pv_pc <= 0.01, ] -->

<!-- Spearman_0.05c[ , c("ASV", "Spearman_est", "Spearman_pv_pc", "Genus", "Family", "Class", "Phylum")] -->
<!-- Spearman_0.01c[ , c("ASV", "Spearman_est", "Spearman_pv_pc", "Genus", "Family", "Class", "Phylum")] -->

<!-- # ASVs with low p values -->
<!-- # No pearson as data is not normally distributed ! -->
<!-- table(c(Spearman_0.05c$ASV, Kendall_0.05c$ASV)) -->
<!-- table(c(Spearman_0.01c$ASV, Kendall_0.01c$ASV)) -->
<!-- table(c(Spearman_0.05c$ASV, Kendall_0.05c$ASV, Spearman_0.01c$ASV, Kendall_0.01c$ASV)) -->

<!-- # Store ASVs -->
<!-- ASVs_0.05c <- unique(c(Spearman_0.05c$ASV)) -->
<!-- ASVs_0.01c <- unique(c(Spearman_0.01c$ASV)) -->

<!-- DAASVs_SpCorr_ASVCatPerf_CD <- list(ASVs_0.05 = ASVs_0.05, ASVs_0.01 = ASVs_0.01, ASVs_0.05_pc = ASVs_0.05c, ASVs_0.01_pc = ASVs_0.01c, ASVs_0.05_neg = ASVs_0.05_neg, ASVs_0.01_neg = ASVs_0.01_neg) -->

<!-- save(DAASVs_SpCorr_ASVCatPerf_CD, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_CorrAnlysis/DAASVs_SpCorr_ASVCatPerf_CD.RData") -->
<!-- ``` -->

<!-- ## Plot highly correlating ASVs -->
<!-- No p value correction used. -->
<!-- ```{r} -->
<!-- # P value < 0.05 -->
<!-- List_Graph_0.05 <- list() -->
<!-- for (i in 1:length(ASVs_0.05)) { -->
<!--   List_Graph_0.05[[i]] <- Plot1(x = ASVs_0.05[i], title = ASVs_0.05[i]) -->
<!-- } -->

<!-- # P value < 0.01 -->
<!-- List_Graph_0.01 <- list() -->
<!-- for (i in 1:length(ASVs_0.01)) { -->
<!--   List_Graph_0.01[[i]] <- Plot1(x = ASVs_0.01[i], title = ASVs_0.01[i]) -->
<!-- } -->

<!-- # P value < 0.05  only negative correlations -->
<!-- List_Graph_0.05_neg <- list() -->
<!-- for (i in 1:length(ASVs_0.05_neg)) { -->
<!--   List_Graph_0.05_neg[[i]] <- Plot1(x = ASVs_0.05_neg[i], title = ASVs_0.05_neg[i]) -->
<!-- } -->

<!-- # # P value < 0.01 only negative correlations -->
<!-- # List_Graph_0.01_neg <- list() -->
<!-- # for (i in 1:length(ASVs_0.01_neg)) { -->
<!-- #   List_Graph_0.01_neg[[i]] <- Plot1(x = ASVs_0.01_neg[i], title = ASVs_0.01_neg[i]) -->
<!-- # } -->
<!-- #  -->
<!-- #Plot1(x = "bASV_2431", title = "bASV_2431") -->
<!-- ``` -->

<!-- ### Print plots -->
<!-- ```{r, fig.height=4, fig.width=6} -->
<!-- # Print plots p value < 0.05 -->
<!-- wrap_plots(List_Graph_0.05) -->
<!-- ``` -->
<!-- ```{r, fig.height=2, fig.width=2} -->
<!-- # Print plots p value < 0.01 -->
<!-- wrap_plots(List_Graph_0.01) -->
<!-- ``` -->
<!-- ```{r, fig.height=2, fig.width=6} -->
<!-- # Print plots p value < 0.05 only negative correlations -->
<!-- wrap_plots(List_Graph_0.05_neg) -->
<!-- ``` -->
<!-- ```{r, fig.height=2.3, fig.width=7} -->
<!-- # Print plots p value < 0.01 only negative correlations -->
<!-- #wrap_plots(List_Graph_0.01_neg) -->
<!-- ``` -->


# 6.6 ASV - Caterpillar performance correlations GO1
## Load data
```{r}
# Selection of DA ASVs GO1
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load ps object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")
```

## Remove Batch 1
```{r}
FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Batch != 1)
```

## Set relative abundance
```{r}
# Calculate relative abundance
FP_ps <- transform_sample_counts(FP_unnormalized_bac_ps_ForDA_clean, function(x) x / sum(x))
```

## Prepare data set
```{r}
# make one vector with unique ASVs
ASVs <- Acc_Bac_TwoTimes_DA_ASV$GO1

# Only select GO1
FP_ps <- subset_samples(FP_ps, Accession == "GO1")

# prune ps object
FP_ps <- prune_taxa(ASVs, FP_ps) #only keep taxa that are in ASVs vector
#length(intersect(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean))) # check overlap
#setdiff(ASVs, taxa_names(FP_unnormalized_bac_ps_ForDA_clean)) # check difference

# Phyloseq object to data frame with relative abundance, meta data and taxonomy
FP_ps <- psmelt(FP_ps)

# remove NAs cat weight
FP_ps[is.na(FP_ps$Mean_Weight_Cat), 1:10]
FP_ps <- FP_ps[!is.na(FP_ps$Mean_Weight_Cat), ]
```

### Remove ASVs with only 0s
```{r}
# identify ASVs with zero abundance
Abund_sum <- aggregate(FP_ps$Abundance, by = list(FP_ps$OTU), FUN = sum)
Abund_sum[Abund_sum$x == 0, ]
Abund_zero <- Abund_sum[Abund_sum$x == 0, "Group.1"]

# Remove ASVs
FP_ps <- FP_ps[!(FP_ps$OTU %in% Abund_zero), ]
ASVs <- ASVs[!ASVs %in% Abund_zero]
```

## Plot
### Function for plot
```{r fig.height=20, fig.width=20}
Plot1 <- function(x, title) {
  ggplot(data = FP_ps[FP_ps$OTU == x, ], aes(x = Abundance, y = Mean_Weight_Cat)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    # scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    ggtitle(title) +
    xlab("ASV abundance") +
    ylab("Caterpillar weight") +
    theme(panel.background = element_rect(fill = "white"), #theme of the graph
          panel.border = element_blank(),
          axis.line = element_line(), #show line for x and y axis
          plot.title = element_text(size = 12, hjust = 0),
          axis.text = element_text(size = 8, color = "black"),
          axis.title = element_text(size = 12),
          axis.title.x = element_text(vjust = -0.5),
          axis.title.y = element_text(vjust = 2.5), #, hjust = -0
          strip.text = element_text(size = 13),
          legend.position = "right")
}
```

### Run function
```{r fig.height=20, fig.width=20}
#Plot1(x = "bASV_42", title = "bASV_42")

List_Graph1 <- list()
for (i in 1:length(ASVs)) {
  List_Graph1[[i]] <- Plot1(x = ASVs[i], title = ASVs[i])
}
```

### Print plots
```{r, fig.height=10, fig.width=12}
# Print plots
wrap_plots(List_Graph1[1:20])
```

## Model assumptions
### Assumptions with graphs
```{r}
NormTest <- function(dat, title) {
  hist(dat, prob = TRUE, col = "grey", main = title) #distribution histogram
  lines(density(dat), col = "blue", lwd = 2)
  
  qqnorm(dat)
  qqline(dat)
  
  print(shapiro.test(dat))
}

# Caterpillar performance
par(mfrow = c(1, 2))
NormTest(dat = FP_ps[FP_ps$OTU == "bASV_27", "Mean_Weight_Cat"], title = "b")
par(mfrow = c(1, 1))
```

### Assumptions only Shapriro Wilk test
```{r}
# Abundance
Shap_res_ab <- list()
for(i in 1:length(ASVs)) {
  Shap_res_ab[[i]] <- shapiro.test(FP_ps[FP_ps$OTU == ASVs[i], "Abundance"])
}
names(Shap_res_ab) <- ASVs
Shap_res_ab
```

## Correlation tests
### Run tests
```{r}
CorTest2 <- function(x) {
  sub_data <- FP_ps[FP_ps$OTU == x, ]
  cs0 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("pearson")) # parametric
  cs1 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("spearman")) # non parametric
  cs2 <- cor.test(sub_data$Abundance, sub_data$Mean_Weight_Cat, method = c("kendall")) # non parametric
  rm(sub_data)
  
  # Make data frame with results
  Sum1 <- data.frame(Pearson_est = cs0$estimate, Pearson_stat = cs0$statistic, Pearson_pv = cs0$p.value,
                     Spearman_est = cs1$estimate, Spearman_stat = cs1$statistic, Spearman_pv = cs1$p.value, 
                     Kendall_est = cs2$estimate, Kendall_stat = cs2$statistic, Kendall_pv = cs2$p.value)
  rownames(Sum1) <- x
  return(Sum1)
}

# Run function
List_CorTest2 <- list() # make an empty list to store the plots

for (i in 1:length(ASVs)) {
  List_CorTest2[[i]] <- CorTest2(x = ASVs[i])
}

names(List_CorTest2) <- ASVs
df_CorTest2 <- do.call(rbind.data.frame, List_CorTest2)
df_CorTest2$ASV <- row.names(df_CorTest2)

# Add taxonomy to correlation results
## Extract unique ASVs
FP_ps_ASV_unique <- FP_ps[!duplicated(FP_ps$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]
colnames(FP_ps_ASV_unique)[1] <- "ASV"

## Merge two data frames
df_CorTest2 <- merge(df_CorTest2, FP_ps_ASV_unique, by = "ASV")

# p value correction
df_CorTest2$Pearson_pv_pc <- p.adjust(df_CorTest2$Pearson_pv, method = "fdr")
df_CorTest2$Spearman_pv_pc <- p.adjust(df_CorTest2$Spearman_pv, method = "fdr")
df_CorTest2$Kendall_pv_pc <- p.adjust(df_CorTest2$Kendall_pv, method = "fdr")

# print head data frame
df_CorTest2[1:10, ]
```

### Select low p values of not corrected p values
```{r}
#Pearson_0.05 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ]
Spearman_0.05 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.05, ]
Kendall_0.05 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.05, ]

#Pearson_0.01 <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ]
Spearman_0.01 <- df_CorTest2[df_CorTest2$Spearman_pv <= 0.01, ]
Kendall_0.01 <- df_CorTest2[df_CorTest2$Kendall_pv <= 0.01, ]

Spearman_0.05[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.05[Spearman_0.05$Spearman_est <= 0, c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01[ , c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01[Spearman_0.01$Spearman_est <= 0, c("ASV", "Spearman_est", "Spearman_pv", "Genus", "Family", "Class", "Phylum")]

summary(df_CorTest2$Spearman_pv)

# ASVs with low p values
# No pearson as data is not normally distributed !
table(c(Spearman_0.05$ASV, Kendall_0.05$ASV))
table(c(Spearman_0.01$ASV, Kendall_0.01$ASV))

# Store ASVs
ASVs_0.05 <- unique(c(Spearman_0.05$ASV))
ASVs_0.05_neg <- unique(c(Spearman_0.05[Spearman_0.05$Spearman_est <= 0, "ASV"]))
ASVs_0.01 <- unique(c(Spearman_0.01$ASV))
ASVs_0.01_neg <- unique(c(Spearman_0.01[Spearman_0.01$Spearman_est <= 0, "ASV"]))
ASVs_0.01_neg <- ASVs_0.01_neg[!is.na(ASVs_0.01_neg)]
```

### Select low p values of corrected p values
```{r}
#Pearson_0.05c <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.05, ]
Spearman_0.05c <- df_CorTest2[df_CorTest2$Spearman_pv_pc <= 0.05, ]
Kendall_0.05c <- df_CorTest2[df_CorTest2$Kendall_pv_pc <= 0.05, ]

#Pearson_0.01c <- df_CorTest2[df_CorTest2$Pearson_pv_pc <= 0.01, ]
Spearman_0.01c <- df_CorTest2[df_CorTest2$Spearman_pv_pc <= 0.01, ]
Kendall_0.01c <- df_CorTest2[df_CorTest2$Kendall_pv_pc <= 0.01, ]

Spearman_0.05c[ , c("ASV", "Spearman_est", "Spearman_pv_pc", "Genus", "Family", "Class", "Phylum")]
Spearman_0.01c[ , c("ASV", "Spearman_est", "Spearman_pv_pc", "Genus", "Family", "Class", "Phylum")]

# ASVs with low p values
# No pearson as data is not normally distributed !
table(c(Spearman_0.05c$ASV, Kendall_0.05c$ASV))
table(c(Spearman_0.01c$ASV, Kendall_0.01c$ASV))
table(c(Spearman_0.05c$ASV, Kendall_0.05c$ASV, Spearman_0.01c$ASV, Kendall_0.01c$ASV))

# Store ASVs
ASVs_0.05c <- unique(c(Spearman_0.05c$ASV))
ASVs_0.01c <- unique(c(Spearman_0.01c$ASV))

DAASVs_SpCorr_ASVCatPerf_GO1 <- list(ASVs_0.05 = ASVs_0.05, ASVs_0.01 = ASVs_0.01, ASVs_0.05_pc = ASVs_0.05c, ASVs_0.01_pc = ASVs_0.01c, ASVs_0.05_neg = ASVs_0.05_neg, ASVs_0.01_neg = ASVs_0.01_neg)

#save(DAASVs_SpCorr_ASVCatPerf_GO1, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_CorrAnlysis/DAASVs_SpCorr_ASVCatPerf_GO1.RData")
```

## Plot highly correlating ASVs
No p value correction used.
```{r}
# # P value < 0.05
# List_Graph_0.05 <- list()
# for (i in 1:length(ASVs_0.05)) {
#   List_Graph_0.05[[i]] <- Plot1(x = ASVs_0.05[i], title = ASVs_0.05[i])
# }
# 
# # P value < 0.01
# List_Graph_0.01 <- list()
# for (i in 1:length(ASVs_0.01)) {
#   List_Graph_0.01[[i]] <- Plot1(x = ASVs_0.01[i], title = ASVs_0.01[i])
# }
# 
# # P value < 0.05  only negative correlations
# List_Graph_0.05_neg <- list()
# for (i in 1:length(ASVs_0.05_neg)) {
#   List_Graph_0.05_neg[[i]] <- Plot1(x = ASVs_0.05_neg[i], title = ASVs_0.05_neg[i])
# }
# 
# # P value < 0.01 only negative correlations
# List_Graph_0.01_neg <- list()
# for (i in 1:length(ASVs_0.01_neg)) {
#   List_Graph_0.01_neg[[i]] <- Plot1(x = ASVs_0.01_neg[i], title = ASVs_0.01_neg[i])
# }
# 
# #Plot1(x = "bASV_2431", title = "bASV_2431")
```

### Print plots
```{r, fig.height=20, fig.width=27}
# Print plots p value < 0.05
#wrap_plots(List_Graph_0.05)
```
```{r, fig.height=7, fig.width=13}
# Print plots p value < 0.01
#wrap_plots(List_Graph_0.01, ncol = 5)
```
```{r, fig.height=7, fig.width=9}
# Print plots p value < 0.05 only negative correlations
#wrap_plots(List_Graph_0.05_neg)
```
```{r, fig.height=2.3, fig.width=7}
# Print plots p value < 0.01 only negative correlations
#wrap_plots(List_Graph_0.01_neg)
```
