---
title: "FP_10.2 Heatmaps of correlating ASVs after DA analysis FP"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b635.html)
- [Excamples of pheatmap package](https://davetang.github.io/muse/pheatmap.html)   
   
   
# 6.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(pheatmap)
packageVersion("pheatmap")
```


# 6.3 ASVs correlating with Cat weight all 4 accession Only negative correlations without RI
DA ASVs of 4 accessions that correlate with caterpillar weight of all these accessions together. RI is left out.   
   
## 6.3.1 Correlation p value < 0.05
### Load data
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_unnormalized_bac_ps_ForDA_clean.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load correlating DA ASVs
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_CorrAnlysis/DAASVs_SpCorr_ASVCatPerf_HM.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
```

### Remove RI
```{r}
FP_unnormalized_bac_ps_ForDA_clean <- subset_samples(FP_unnormalized_bac_ps_ForDA_clean, Accession != "RI")
Acc_Bac_TwoTimes_DA_ASV <- Acc_Bac_TwoTimes_DA_ASV[names(Acc_Bac_TwoTimes_DA_ASV) != "RI"]
```

### Combine DA ASVs per accession
```{r}
ASVs <- DAASVs_SpCorr_ASVCatPerf_HM$ASVs_0.05_neg
DA_ASvs <- lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) x[x %in% ASVs])

lfc_DA_specific <- lapply(names(DA_ASvs), function(name) {
  da_asvs <- DA_ASvs[[name]]
  lfc_table <- ancomWZ_Bac_Acc[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name)
  lfc_table
})

names(lfc_DA_specific) <- names(DA_ASvs)

# merge all accessions together
AllAcc_lfc <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific)
rownames(AllAcc_lfc) <- AllAcc_lfc$ASV

# Change order accessions
AllAcc_lfc <- AllAcc_lfc[c(1,4,2,3,5)]
```

### Prepare taxonomy column class and phylum
```{r}
# Only keep DA in ps object
FP_unnormalized_bac_ps_ForDA_clean <- prune_taxa(ASVs, FP_unnormalized_bac_ps_ForDA_clean)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df_clean <- psmelt(FP_unnormalized_bac_ps_ForDA_clean)
head(df_clean[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df_clean[!duplicated(df_clean$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)

# Add genus to matrix
AllAcc_lfc2 <- AllAcc_lfc
tax$ASV <- rownames(tax)
AllAcc_lfc2 <- merge(tax, AllAcc_lfc2, ID = "ASV")

# Add higher taxonomic level if genus is NA or Incertae Sedis
taxon <- AllAcc_lfc2$Genus
taxon[is.na(taxon)] <- AllAcc_lfc2$Family[is.na(taxon)]
taxon[is.na(taxon)] <- AllAcc_lfc2$Order[is.na(taxon)]

taxon[taxon == "g__Incertae_Sedis"] <- AllAcc_lfc2$Family[taxon == "g__Incertae_Sedis"]
taxon[taxon == "f__Incertae_Sedis"] <- AllAcc_lfc2$Order[taxon == "f__Incertae_Sedis"]
taxon[taxon == "o__Incertae_Sedis"] <- AllAcc_lfc2$Class[taxon == "o__Incertae_Sedis"]
taxon[taxon == "c__Incertae_Sedis"] <- AllAcc_lfc2$Phylum[taxon == "c__Incertae_Sedis"]

taxon[taxon == "g__Unclassified"] <- AllAcc_lfc2$Family[taxon == "g__Unclassified"]
taxon[taxon == "f__Unclassified"] <- AllAcc_lfc2$Order[taxon == "f__Unclassified"]
taxon[taxon == "o__Unclassified"] <- AllAcc_lfc2$Class[taxon == "o__Unclassified"]
taxon[taxon == "c__Unclassified"] <- AllAcc_lfc2$Phylum[taxon == "c__Unclassified"]

rownames(AllAcc_lfc2) <- paste(AllAcc_lfc2$ASV, taxon, sep = "_")

AllAcc_lfc2$Phylum <- sub("p__", "", AllAcc_lfc2$Phylum)
AllAcc_lfc2$Class <- sub("c__", "", AllAcc_lfc2$Class)

# Subset data columns
tax <- AllAcc_lfc2[1:6]
AllAcc_lfc2 <- AllAcc_lfc2[7:10]
head(AllAcc_lfc2)

# Make color pallet for class
colors35 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#332288", "#CC79A7", 
              "#D55E00", "#0072B2", "#000000", "#999933",
              "#117733", "#88CCEE", "#44AA99", "#DDCC77", "#CC6677", "#882255", "#AA4499",
              "#661100", "#6699CC", "#AA4466", "#4477AA", "#EE6677", "#BBBBBB", 
              "#2E3440", "#8FBCBB", "#BF616A", "#A3BE8C", "#EBCB8B", "#B48EAD", 
              "#5E81AC", "#D08770", "#3B4252", "#7B9E87", "#999999")

# Make color pallet for Phylum for genus in plot
tax2 <- tax
phylums2 <- unique(tax2$Phylum)
phylums_colors2 <- setNames(colors35[1:length(phylums2)], phylums2)
phylums_colors2 <- list(Phylum = phylums_colors2)

classes2 <- unique(tax2$Class)
class_colors2 <- setNames(colors35[1:length(classes2)], classes2)
class_colors2 <- list(Class = class_colors2)

phylum_class_colors <- c(phylums_colors2, class_colors2)

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc2)

## Create expression labels
row_labels <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels <- as.expression(row_labels)

# Turn data frame into a matrix
AllAcc_lfc2 <- as.matrix(AllAcc_lfc2)

# replace NAs by 0s
AllAcc_lfc2_clu <- AllAcc_lfc2
AllAcc_lfc2_clu[is.na(AllAcc_lfc2_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc2), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -2, length.out = 15), 
                   seq(-2, 2, length.out = 40), # More breaks in this range for sensitivity
                   seq(2, max_abs, length.out = 15)))
```

### Heatmap 
```{r, fig.height=7, fig.width=7.5}
#6.2
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax2["Phylum"],
          annotation_colors = phylum_class_colors,
          labels_row = row_labels)
```
```{r, fig.height=7, fig.width=7.5}
#4.8
p <- pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          #clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          cluster_cols = FALSE,
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax2["Class"],
          annotation_colors = phylum_class_colors,
          labels_row = row_labels)
p 

files <- c("FP_SpearCorr_Heatmap_ASV.svg", "FP_SpearCorr_Heatmap_ASV.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 19/2.54, height = 9/2.54)
  } else {
    png(file_path, width = 19, height = 9, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=7.5, fig.width=7.5}
#6.5
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax2[5:6],
          annotation_colors = phylum_class_colors,
          labels_row = row_labels)
```
```{r}
# Clean environment
rm(list = ls())
```
