---
title: "CP_08_ Relative abundance per sample - Heatmaps"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
- [Excamples of pheatmap package](https://davetang.github.io/muse/pheatmap.html)   
   
   
# 4.0 Preperation
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(DESeq2)
packageVersion("DESeq2")
library(pheatmap)
packageVersion("pheatmap")
library(tidyr)
packageVersion("tidyr")
library(gridExtra)
packageVersion("gridExtra")
library(grid)
packageVersion("grid")

# library(metamisc) #for phyloseq_sep_variable (and more?)
# packageVersion("metamisc")
# library(zinbwave)
# packageVersion("zinbwave")
# library(UpSetR)
# packageVersion("UpSetR")
# library(tuple)
# packageVersion("tuple")
# library(ggpubr)
# packageVersion("ggpubr")
# library(stringr)
# packageVersion("stringr")
# library(ggVennDiagram)
# packageVersion("ggVennDiagram")
```


# 4.10 Heatmap relative abundance per sample
## 4.10.1 Accessions seperate
### Heatmap At ASV level DA ASVs occuring twice
#### Load data
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
```

#### Prepare data
```{r}
# Put all unique DA ASVs in one vector
Acc_Bac_TwoTimes_DA_ASV_comb <- unique(unlist(Acc_Bac_TwoTimes_DA_ASV, use.names = FALSE))

# Calculate relative abundance
CP_ps <- transform_sample_counts(CP_unnormalized_bac_ps_ForDA, function(x) x / sum(x))

# Only keep DA in ps object
CP_ps <- prune_taxa(Acc_Bac_TwoTimes_DA_ASV_comb, CP_ps)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df <- psmelt(CP_ps)
head(df[, c(1:8)], 12)

# unique accessions
Acc <- names(Acc_Bac_TwoTimes_DA_ASV)
```

#### Matrix per accession
```{r}
# Subset data per accession and DA ASVs
Rel_Ab_l <- list()
for (i in seq_along(Acc)) {
  Acc_sub <- Acc[i]
  sub_df <- df[df$Accession == Acc_sub & df$OTU %in% Acc_Bac_TwoTimes_DA_ASV[[Acc_sub]], c(1:3, 8)]
  sub_df <- sub_df[order(sub_df$Cat_treatment, sub_df$Sample), ]
  sub_df$Sample <- paste(sub_df$Sample, sub_df$Cat_treatment, sep = "_")
    #colnames(sub_df)[3] <- Acc_sub 
  Rel_Ab_l[[i]] <- sub_df
}

names(Rel_Ab_l) <- Acc

# Number of Co Samples
Numb_Co <- lapply(Rel_Ab_l, function(x) length(x[x$Cat_treatment == "Co", "Cat_treatment"])/length(unique(x$OTU)))

# Put each sample in separate column
Rel_Ab_l_wide <- lapply(Rel_Ab_l, function(df) {
  df$Cat_treatment <- NULL
  pivot_wider(df, names_from  = Sample, values_from = Abundance, values_fill = 0)
})

# Make ASVs rownames
Rel_Ab_l_wide <- lapply(Rel_Ab_l_wide, function(df) {
  df <- as.data.frame(df)
  rownames(df) <- df$OTU
  colnames(df)[1] <- "ASV"
  return(df)
})



tax <- lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) {
  sub <- df[df$OTU %in% x, c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]
  sub <- sub[!duplicated(sub$OTU), ]
  rownames(sub) <- sub$OTU
  colnames(sub)[1] <- "ASV"
  return(sub)
})

# Add genus to matrix
AllAcc_lfc2 <- Map(function(x, y) {
  merge(x, y, by = "ASV", all.y = TRUE)
}, tax, Rel_Ab_l_wide)

Rel_Ab_l_wide_df <- lapply(AllAcc_lfc2, function(x) {
  taxon <- x$Genus
  taxon[is.na(taxon)] <- x$Genus[is.na(taxon)]
  
  taxon[taxon == "g__Incertae_Sedis"] <- x$Family[taxon == "g__Incertae_Sedis"]
  taxon[taxon == "f__Incertae_Sedis"] <- x$Order[taxon == "f__Incertae_Sedis"]
  taxon[taxon == "o__Incertae_Sedis"] <- x$Class[taxon == "o__Incertae_Sedis"]
  taxon[taxon == "c__Incertae_Sedis"] <- x$Phylum[taxon == "c__Incertae_Sedis"]
  
  taxon[taxon == "g__Unclassified"] <- x$Family[taxon == "g__Unclassified"]
  taxon[taxon == "f__Unclassified"] <- x$Order[taxon == "f__Unclassified"]
  taxon[taxon == "o__Unclassified"] <- x$Class[taxon == "o__Unclassified"]
  taxon[taxon == "c__Unclassified"] <- x$Phylum[taxon == "c__Unclassified"]
  
  rownames(x) <- paste(x$ASV, taxon, sep = "_")
  return(x)
})

# Split matrix and taxonomyn and turn into matrix
tax <- lapply(Rel_Ab_l_wide_df, function(x) {
  x[1:6]
})

Rel_Ab_l_wide_m <- lapply(Rel_Ab_l_wide_df, function(x) {
  x <- x[7:ncol(x)]
  as.matrix(x)
})

Rel_Ab_l_wide_m$OH[, 1:6]
```

#### Headmap 
```{r, fig.height=4, fig.width=7}
plot_Rel_Ab <- list()
for (i in seq_along(Rel_Ab_l_wide_m)) {
  x <- Rel_Ab_l_wide_m[[i]]
  
  # make 0s gray
  x[x == 0] <- NA
  
  # breaks in headmap
  max_abs <- max(abs(x), na.rm = TRUE)
  breaks <- unique(c(#seq(0, 0.2*max_abs, length.out = 30), # More breaks in this range for sensitivity
                     seq(0*max_abs, max_abs, length.out = 60)))

  # Headmap
  plot_Rel_Ab[[i]] <- pheatmap(x,
            cluster_rows = FALSE,
            cluster_cols = FALSE,
            color = colorRampPalette(c("gold", "orange", "red", "darkred", "black"))(length(breaks) - 1),
            breaks = breaks,
            main = names(Rel_Ab_l_wide_m[i]),
            fontsize_row = 10,
            fontsize_col = 14,
            gaps_col = c(Numb_Co[[i]]),
            na_col = "grey90",
            angle_col = 90)
}
```

#### Print plots
```{r, fig.height=1.55, fig.width=7}
#1
plot_Rel_Ab[c(2)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(3)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(10)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(12)]
```
```{r, fig.height=1.8, fig.width=7}
#2
plot_Rel_Ab[c(7)]
```
```{r, fig.height=1.8, fig.width=7}
plot_Rel_Ab[c(9)]
```
```{r, fig.height=2, fig.width=7}
#3
plot_Rel_Ab[c(6)]
```
```{r, fig.height=2, fig.width=7}
plot_Rel_Ab[c(11)]
```
```{r, fig.height=2.5, fig.width=7}
#5
plot_Rel_Ab[4]
```
```{r, fig.height=3.1, fig.width=7}
#8
plot_Rel_Ab[c(5)]
```
```{r, fig.height=3.1, fig.width=7}
plot_Rel_Ab[c(8)]
```
```{r, fig.height=3.4, fig.width=7}
#9
plot_Rel_Ab[1]
```



#### All Plots on same scall
```{r, fig.height=4, fig.width=7}
# breaks in head map
max_abs <- lapply(Rel_Ab_l_wide_m, function(x) max(abs(x), na.rm = TRUE))
max_abs <- max(unlist(max_abs))
breaks <- unique(c(seq(0, 0.2*max_abs, length.out = 80),
                   seq(0.2*max_abs, 0.5*max_abs, length.out = 40), # More breaks in this range for sensitivity
                   seq(0.5*max_abs, max_abs, length.out = 5)))

plot_Rel_Ab <- list()
for (i in seq_along(Rel_Ab_l_wide_m)) {
  x <- Rel_Ab_l_wide_m[[i]]
  
  # make 0s gray
  x[x == 0] <- NA
  
  # Head map
  plot_Rel_Ab[[i]] <- pheatmap(x,
            cluster_rows = FALSE,
            cluster_cols = FALSE,
            color = colorRampPalette(c("gold", "orange", "red", "darkred", "black"))(length(breaks) - 1),
            breaks = breaks,
            main = names(Rel_Ab_l_wide_m[i]),
            fontsize_row = 10,
            fontsize_col = 14,
            na_col = "grey90",
            gaps_col = c(Numb_Co[[i]]),
            angle_col = 90)
}
```

#### Print plots
```{r, fig.height=1.55, fig.width=7}
#1
plot_Rel_Ab[c(2)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(3)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(10)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(12)]
```
```{r, fig.height=1.8, fig.width=7}
#2
plot_Rel_Ab[c(7)]
```
```{r, fig.height=1.8, fig.width=7}
plot_Rel_Ab[c(9)]
```
```{r, fig.height=2, fig.width=7}
#3
plot_Rel_Ab[c(6)]
```
```{r, fig.height=2, fig.width=7}
plot_Rel_Ab[c(11)]
```
```{r, fig.height=2.5, fig.width=7}
#5
plot_Rel_Ab[4]
```
```{r, fig.height=3.1, fig.width=7}
#8
plot_Rel_Ab[c(5)]
```
```{r, fig.height=3.1, fig.width=7}
plot_Rel_Ab[c(8)]
```
```{r, fig.height=3.4, fig.width=7}
#9
plot_Rel_Ab[1]
```

#### All Plots on same scall without legend
```{r, fig.height=4, fig.width=7}
max_abs <- lapply(Rel_Ab_l_wide_m, function(x) max(abs(x), na.rm = TRUE))
max_abs <- max(unlist(max_abs))
breaks <- unique(c(seq(0, 0.2*max_abs, length.out = 80),
                   seq(0.2*max_abs, 0.5*max_abs, length.out = 40), # More breaks in this range for sensitivity
                   seq(0.5*max_abs, max_abs, length.out = 5)))

plot_Rel_Ab <- list()
for (i in seq_along(Rel_Ab_l_wide_m)) {
  x <- Rel_Ab_l_wide_m[[i]]
  
  # make 0s gray
  x[x == 0] <- NA
  
  # Head map
  plot_Rel_Ab[[i]] <- pheatmap(x,
            cluster_rows = FALSE,
            cluster_cols = FALSE,
            color = colorRampPalette(c("gold", "orange", "red", "darkred", "black"))(length(breaks) - 1),
            breaks = breaks,
            main = names(Rel_Ab_l_wide_m[i]),
            fontsize_row = 10,
            fontsize_col = 14,
            na_col = "grey90",
            angle_col = 90,
            gaps_col = c(Numb_Co[[i]]),
            legend = FALSE)
}
```

#### Print plots
```{r, fig.height=1.55, fig.width=7}
#1
plot_Rel_Ab[c(2)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(3)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(10)]
```
```{r, fig.height=1.55, fig.width=7}
plot_Rel_Ab[c(12)]
```
```{r, fig.height=1.8, fig.width=7}
#2
plot_Rel_Ab[c(7)]
```
```{r, fig.height=1.8, fig.width=7}
plot_Rel_Ab[c(9)]
```
```{r, fig.height=2, fig.width=7}
#3
plot_Rel_Ab[c(6)]
```
```{r, fig.height=2, fig.width=7}
plot_Rel_Ab[c(11)]
```
```{r, fig.height=2.5, fig.width=7}
#5
plot_Rel_Ab[4]
```
```{r, fig.height=3.1, fig.width=7}
#8
plot_Rel_Ab[c(5)]
```
```{r, fig.height=3.1, fig.width=7}
plot_Rel_Ab[c(8)]
```
```{r, fig.height=3.4, fig.width=7}
#9
plot_Rel_Ab[1]
```

#### All Plots on same scall without legend and x labels
```{r, fig.height=5, fig.width=10}
max_abs <- lapply(Rel_Ab_l_wide_m, function(x) max(abs(x), na.rm = TRUE))
max_abs <- max(unlist(max_abs))
breaks <- unique(c(seq(0, 0.2*max_abs, length.out = 80),
                   seq(0.2*max_abs, 0.5*max_abs, length.out = 40), # More breaks in this range for sensitivity
                   seq(0.5*max_abs, max_abs, length.out = 5)))

plot_Rel_Ab <- list()
for (i in seq_along(Rel_Ab_l_wide_m)) {
  x <- Rel_Ab_l_wide_m[[i]]
  
  # make 0s gray
  x[x == 0] <- NA

  # Head map
  plot_Rel_Ab[[i]] <- pheatmap(x,
            cluster_rows = FALSE,
            cluster_cols = FALSE,
            color = colorRampPalette(c("gold", "orange", "red", "darkred", "black"))(length(breaks) - 1),
            breaks = breaks,
            #main = names(Rel_Ab_l_wide_m[i]),
            show_colnames = FALSE,
            show_rownames = TRUE,
            fontsize_row = 14,
            fontsize_col = 14,
            na_col = "grey90",
            angle_col = 90,
            gaps_col = c(Numb_Co[[i]]),
            cellwidth = 20,
            cellheight = 20,
            legend = FALSE)
}
```
```{r}
# Create a single legend
legend <- pheatmap(Rel_Ab_l_wide_m[[1]],
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   color = colorRampPalette(c("gold", "orange", "red", "darkred", "black"))(length(breaks) - 1),
                   breaks = breaks,
                   legend = TRUE,
                   show_rownames = FALSE,
                   show_colnames = FALSE)$gtable

# The legend is usually in the last grob of the gtable
legend_grob <- legend$grobs[[2]]
```


#### Print plots
```{r, fig.height=0.3, fig.width=7.5}
#1
plot_Rel_Ab[c(2)]
```
```{r, fig.height=0.3, fig.width=7.5}
plot_Rel_Ab[c(3)]
```
```{r, fig.height=0.3, fig.width=7.5}
plot_Rel_Ab[c(10)]
```
```{r, fig.height=0.3, fig.width=7.5}
plot_Rel_Ab[c(12)]
```
```{r, fig.height=0.6, fig.width=7.5}
#2
plot_Rel_Ab[c(7)]
```
```{r, fig.height=0.6, fig.width=7.5}
plot_Rel_Ab[c(9)]
```
```{r, fig.height=0.9, fig.width=7.5}
#3
plot_Rel_Ab[c(6)]
```
```{r, fig.height=0.9, fig.width=7.5}
plot_Rel_Ab[c(11)]
```
```{r, fig.height=1.5, fig.width=7.5}
#5
plot_Rel_Ab[4]
```
```{r, fig.height=2.4, fig.width=7.5}
#8
plot_Rel_Ab[c(5)]
```
```{r, fig.height=2.4, fig.width=7.5}
plot_Rel_Ab[c(8)]
```
```{r, fig.height=2.7, fig.width=7.5}
#9
plot_Rel_Ab[1]
```

```{r}
grid.newpage()
p <- grid.arrange(grobs = lapply(plot_Rel_Ab, function(x) x$gtable), ncol = 2)
```

#### Save plots
```{r}
# save all heatmaps
files <- c("CP_DAHeatmap_ASV_persample.svg", "CP_DAHeatmap_ASV_persample.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 54/2.54, height = 40/2.54)
  } else {
    png(file_path, width = 54, height = 40, units = "cm", res = 300)
  }
  grid::grid.draw(p)
  dev.off()
}, files)

# save legend
files <- c("CP_DAHeatmap_ASV_persample_legend.svg", "CP_DAHeatmap_ASV_persample_legend.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 2/2.54, height = 5.5/2.54)
  } else {
    png(file_path, width = 2, height = 5.5, units = "cm", res = 300)
  }
  grid::grid.draw(legend_grob)
  dev.off()
}, files)
```

```{r}
# library(gridExtra)
# library(grid)
# 
# heatmaps <- lapply(plot_Rel_Ab, function(x) x$gtable)
# 
# # Combine heatmaps in a single grid
# grid.newpage()
# grid.arrange(
#   grobs = c(heatmaps, list(legend_grob)),
#   ncol = 2,
#   widths = c(6,8)
# )
```
