---
title: "CP_07_differential_abundance_Summary_Results - Heatmaps"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
- [Excamples of pheatmap package](https://davetang.github.io/muse/pheatmap.html)   
   
   
# 4.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(pheatmap)
packageVersion("pheatmap")
library(dplyr)
packageVersion("dplyr")
library(stringr)
packageVersion("stringr")
```


# 4.10 Heatmap log2fold change
## 4.10.1 Accessions seperate

### Heatmap At ASV level DA ASVs occuring twice
#### Load data
```{r}
# load phyloseq object
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/CP_unnormalized_bac_ps_ForDA.RData")

# load DA ASVs
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/CP_DA_SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# load log2fold change values
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/Data/R_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
```

#### Combine DA ASVs per accession
```{r}
# Numer of DA ASVs
lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) length(x))
length(unique(unlist(Acc_Bac_TwoTimes_DA_ASV)))

# Filter log2fold change for DA ASVs per accession
lfc_DA_specific <- lapply(names(Acc_Bac_TwoTimes_DA_ASV), function(name) {
  da_asvs <- Acc_Bac_TwoTimes_DA_ASV[[name]]
  lfc_table <- ancomWZ_Bac_Acc[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Cat_treatmentMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name)
  lfc_table
})

names(lfc_DA_specific) <- names(Acc_Bac_TwoTimes_DA_ASV)

# merge all accessions together
AllAcc_lfc <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific)
rownames(AllAcc_lfc) <- AllAcc_lfc$ASV
nrow(AllAcc_lfc)
AllAcc_lfc$ASV <- NULL
```

#### Prepare taxonomy column class and phylum
```{r}
# Put all unique DA ASVs in one vector
Acc_Bac_TwoTimes_DA_ASV <- unique(unlist(Acc_Bac_TwoTimes_DA_ASV, use.names = FALSE))

# Only keep DA in ps object
CP_unnormalized_bac_ps_ForDA <- prune_taxa(Acc_Bac_TwoTimes_DA_ASV, CP_unnormalized_bac_ps_ForDA)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df <- psmelt(CP_unnormalized_bac_ps_ForDA)
head(df[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df[!duplicated(df$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Keep only those genera present in your heat map matrix
tax <- tax[match(rownames(AllAcc_lfc), tax$OTU), ]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)

# Add genus to matrix
AllAcc_lfc2 <- AllAcc_lfc
AllAcc_lfc2$ASV <- rownames(AllAcc_lfc2)
tax$ASV <- rownames(tax)
AllAcc_lfc2 <- merge(tax, AllAcc_lfc2, ID = "ASV")

taxon <- AllAcc_lfc2$Genus
taxon[is.na(taxon)] <- AllAcc_lfc2$Genus[is.na(taxon)]

taxon[taxon == "g__Incertae_Sedis"] <- AllAcc_lfc2$Family[taxon == "g__Incertae_Sedis"]
taxon[taxon == "f__Incertae_Sedis"] <- AllAcc_lfc2$Order[taxon == "f__Incertae_Sedis"]
taxon[taxon == "o__Incertae_Sedis"] <- AllAcc_lfc2$Class[taxon == "o__Incertae_Sedis"]
taxon[taxon == "c__Incertae_Sedis"] <- AllAcc_lfc2$Phylum[taxon == "c__Incertae_Sedis"]

taxon[taxon == "g__Unclassified"] <- AllAcc_lfc2$Family[taxon == "g__Unclassified"]
taxon[taxon == "f__Unclassified"] <- AllAcc_lfc2$Order[taxon == "f__Unclassified"]
taxon[taxon == "o__Unclassified"] <- AllAcc_lfc2$Class[taxon == "o__Unclassified"]
taxon[taxon == "c__Unclassified"] <- AllAcc_lfc2$Phylum[taxon == "c__Unclassified"]

rownames(AllAcc_lfc2) <- paste(AllAcc_lfc2$ASV, taxon, sep = "_")
#AllAcc_lfc2$taxon <- make.unique(paste(taxon, sep = "_"))

AllAcc_lfc2$Phylum <- sub("p__", "", AllAcc_lfc2$Phylum)
AllAcc_lfc2$Class <- sub("c__", "", AllAcc_lfc2$Class)

tax2 <- AllAcc_lfc2[1:6]
AllAcc_lfc2 <- AllAcc_lfc2[7:18]
head(AllAcc_lfc2)

# Make color pallet for class
colors15 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7",
              "#882255", "#44AA99", "#117733", "#332288", "#AA4499", "#DDCC77") # "#000000", "#999999", 

classes <- unique(tax$Class)
class_colors <- setNames(colors15[1:length(classes)], classes)
class_colors <- list(Class = class_colors)

classes2 <- unique(tax2$Class)
class_colors2 <- setNames(colors15[1:length(classes2)], classes2)
class_colors2 <- list(Class = class_colors2)

phylums <- unique(tax$Phylum)
phylums_colors <- setNames(colors15[1:length(phylums)], phylums)
phylums_colors <- list(Phylum = phylums_colors)

# Make color pallet for Phylum for genus in plot
phylums2 <- unique(tax2$Phylum)
phylums_colors2 <- setNames(colors15[1:length(phylums2)], phylums2)
phylums_colors2 <- list(Phylum = phylums_colors2)

phylum_class_colors <- c(phylums_colors2, class_colors2)

# counting number of ASVs per accession
ASV_numb <- colSums(!is.na(AllAcc_lfc))
ASV_numb

# Row labels of genus in italic 
rn <- rownames(AllAcc_lfc2)

## Create expression labels
row_labels <- sapply(rn, function(x) {
  tax_part <- sub("^[^_]+_[^_]+_", "", x) # Extract taxonomy
  if (grepl("^g__", tax_part)) { # grab all names with genus
    prefix <- sub("(.*g__).*", "\\1", x) # keep part that does not need to be in italic
    genus <- sub("^g__", "", tax_part) # extract genus
    bquote(.(prefix) * italic(.(genus))) # combine genus and part not in italic
  } else {
    x # change nothing for higher taxonomic levels
}})

## Convert to expression vector
row_labels <- as.expression(row_labels)

# Turn data frame into a matrix
AllAcc_lfc <- as.matrix(AllAcc_lfc)
AllAcc_lfc2 <- as.matrix(AllAcc_lfc2)

# replace NAs by 0s for clustering
AllAcc_lfc_clu <- AllAcc_lfc
AllAcc_lfc2_clu <- AllAcc_lfc2
AllAcc_lfc_clu[is.na(AllAcc_lfc_clu)] <- 0
AllAcc_lfc2_clu[is.na(AllAcc_lfc2_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(AllAcc_lfc), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -0.3, length.out = 10), 
                   seq(-0.3, 0.3, length.out = 30), # More breaks in this range for sensitivity
                   seq(0.3, max_abs, length.out = 10)))
```

#### Heatmap 
```{r, fig.height=8, fig.width=7}
pheatmap(AllAcc_lfc,
          clustering_distance_rows = dist(AllAcc_lfc_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_row = 7,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Phylum"],
          annotation_colors = phylums_colors)
```
```{r, fig.height=8, fig.width=8}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_row = 7,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax2["Phylum"],
          annotation_colors = phylums_colors2,
          labels_row = row_labels)
```
```{r, fig.height=8, fig.width=8}
p <- pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax2["Class"],
          annotation_colors = class_colors2,
          labels_row = row_labels)
p 

files <- c("CP_DAHeatmap_ASV.svg", "CP_DAHeatmap_ASV.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 22/2.54, height = 18/2.54)
  } else {
    png(file_path, width = 22, height = 18, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=8, fig.width=8.5}
pheatmap(AllAcc_lfc2,
         clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          fontsize_row = 9,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax2[5:6],
          annotation_colors = phylum_class_colors,
          labels_row = row_labels)
```
```{r}
# Clean environment
rm(list = ls())
```

