---
title: "CP_05_differential_abundance_Summary_Results - Data preparation"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
   
   
# 4.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(DESeq2)
packageVersion("DESeq2")
library(stringr)
packageVersion("stringr")
library(dplyr)
packageVersion("dplyr")
```

# 4.3 Create lists of DA ASVs
## 4.3.1 Accession ASV level
### Load DA ASVs
```{r}
## Load outputs DESeq
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_Wald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_stddds2_LRT_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbWald_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_DESeq_ASVLevel/sigtab_zinbLRT_Acc_Bac.RData")

## Load output Ancom
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomWZ_Acc_Bac.RData")
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/fdr_ancomNoZ_Acc_Bac.RData")
```

### Extract DA ASVs
```{r}
## From each data frame, extract the ASV IDs caught by each test
Acc_Bac_DESeqLRT_ASVs <- lapply(sigtab_stddds2_LRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_Acc_Bac, function(x) rownames(x))
Acc_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_Acc_Bac, function(x) rownames(x))
Acc_Bac_ancomWZ_ASVs <- lapply(fdr_ancomWZ_Acc_Bac, function(x) as.vector(x[,'Species']))
Acc_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_Acc_Bac, function(x) as.vector(x[,'Species']))

# Within each plant concatenate the identified asvs
Acc_Bac_Total_DA_ASV <- mapply(function(s,t,w,x,y,z){
  input <- c(s,t,w,x,y,z)},
  s = Acc_Bac_DESeqLRT_ASVs,
  t = Acc_Bac_DESeqWald_ASVs,
  w = Acc_Bac_zinbLRT_ASVs,
  x = Acc_Bac_zinbWald_ASVs,
  y = Acc_Bac_ancomWZ_ASVs,
  z = Acc_Bac_ancomNoZ_ASVs,
  SIMPLIFY = FALSE)
```

### Total number of ASVs
```{r}
# Total number of DA ASVs
lapply(Acc_Bac_Total_DA_ASV, function(x) length(x))

save(Acc_Bac_Total_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_Total_DA_ASV.RData")

# Remove duplicated asvs names from that list
Acc_Bac_NoDup_DA_ASV <- lapply(Acc_Bac_Total_DA_ASV, function(x)
  unique(x))

# Number of DA ASVs without duplicates
lapply(Acc_Bac_NoDup_DA_ASV, function(x) length(x))

save(Acc_Bac_NoDup_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV.RData")

# Extracting taxonomy
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps.RData")

# Extract taxonomy of DA ASVs
CP_unnormalized_bac_ps_Tax <- as.data.frame(tax_table(CP_unnormalized_bac_ps))
Acc_Bac_NoDup_DA_ASV_Tax <- lapply(Acc_Bac_NoDup_DA_ASV, function(x) CP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_NoDup_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects//CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_NoDup_DA_ASV_Tax_Table <- lapply(Acc_Bac_NoDup_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_NoDup_DA_ASV_Tax_Table <- lapply(Acc_Bac_NoDup_DA_ASV_Tax_Table, function(ranklist) {
                                        lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_NoDup_DA_ASV_Tax_Table$HM$Class
Acc_Bac_NoDup_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_NoDup_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring twice
```{r}
# Filter to ASVs detected by at least 2 tests
Acc_Bac_NumbOcc_DA_ASV <- lapply(Acc_Bac_Total_DA_ASV, function(x) table(unlist(x)))
Acc_Bac_TwoTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 2]))
Acc_Bac_TwoTimes_DA_ASV


# Number of DA ASVs occurring in two tests
lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_TwoTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_TwoTimes_DA_ASV_Tax <- lapply(Acc_Bac_TwoTimes_DA_ASV, function(x) CP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_TwoTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_TwoTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_TwoTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_TwoTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_TwoTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_TwoTimes_DA_ASV_Tax_Table$HM$Class
Acc_Bac_TwoTimes_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_TwoTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV_Tax_Table.RData")
```

### DA ASVs occuring three times
```{r}
# Filter to ASVs detected by at least 3 tests
Acc_Bac_NumbOcc_DA_ASV <- lapply(Acc_Bac_Total_DA_ASV, function(x) table(unlist(x)))
Acc_Bac_ThreeTimes_DA_ASV <- lapply(Acc_Bac_NumbOcc_DA_ASV, function(x) names(x[x >= 3]))
Acc_Bac_ThreeTimes_DA_ASV

# Number of DA ASVs occurring in three tests
lapply(Acc_Bac_ThreeTimes_DA_ASV, function(x) length(x))

save(Acc_Bac_ThreeTimes_DA_ASV, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV.RData")

# Extract taxonomy of DA ASVs
Acc_Bac_ThreeTimes_DA_ASV_Tax <- lapply(Acc_Bac_ThreeTimes_DA_ASV, function(x) CP_unnormalized_bac_ps_Tax[x, ])

save(Acc_Bac_ThreeTimes_DA_ASV_Tax, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV_Tax.RData")

# Make a table of occurrence at each taxonomic level
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_ThreeTimes_DA_ASV_Tax, function(df) {
  lapply(df[ , 2:6], function(col) table(col, useNA = "ifany"))
})
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table <- lapply(Acc_Bac_ThreeTimes_DA_ASV_Tax_Table, function(ranklist) {
                                            lapply(ranklist, as.data.frame)})

# Examples
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table$HM$Class
Acc_Bac_ThreeTimes_DA_ASV_Tax_Table$VL$Class

save(Acc_Bac_ThreeTimes_DA_ASV_Tax_Table, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_ThreeTimes_DA_ASV_Tax_Table.RData")
```
