---
title: "Caterpillar Weight Inoculation experiment HM"
author: "Kris"
date: "2026-05-21"
output: 
  html_document: 
      toc: true
---

# Start
## Import pre-made functions
```{r, warning = FALSE, message = FALSE}
source("C:/Users/kreek001/OneDrive - Wageningen University & Research/R_Scripts/DHARMa_functions.R") #Functions for comparing distributions, models and model assumptions
```
    
## Libraries
```{r, warning = FALSE, message = FALSE}
R.version$version.string # prints R version

library(car) #for Anova function
packageVersion("car")
library(emmeans) #for post hoc emmeans
packageVersion("emmeans")
library(multcomp) #for cld function at post hoc analysis
packageVersion("multcomp")
library(multcompView) #for cld function at post hoc analysis
packageVersion("multcompView")
library(ggplot2) #graphs
packageVersion("ggplot2")
library(plyr) #for summary table
packageVersion("plyr")
library(glmmTMB) #for doing glm with many different options
packageVersion("glmmTMB")

library(dplyr) # for area underneath curve
packageVersion("dplyr")
library(pracma)  # for trapz for area underneath curve
packageVersion("pracma")
```

## Load and process data
#### Load data set
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Inoculation experiment/Data/Caterpillar_weight_inoculationExpHM2.csv")
```

### Remove NAs
```{r}
data <- data[complete.cases(data[ , "Weight"]),]
```

### Define factors
```{r}
#data$Time <- as.factor(data$Time) #defined after plots
data$Treatment <- factor(data$Treatment, levels = c("Strain 1", "Strain 2", "Strain 3", "WCS417", "Heat-killed", "CTRL"))
```

# Explore data
## Check data
```{r}
str(data)
head(data, 12)
summary(data) #check for outliers
```

### Summary table
```{r}
dataSum <- ddply(data, c("Treatment", "Time"), summarize,
            N = sum(!is.na(Weight)), #number of samples
            Mis = sum(is.na(Weight)), #number of missing values
            Mean = round(mean(Weight, na.rm=T),3), #mean
            Median = round(median(Weight, na.rm=T),3), #median
            SD = round(sd(Weight, na.rm=T),3), #standard deviation
            SE = round(sd(Weight, na.rm=T) / sqrt(N),5), #standard error
            LCI = round(Mean - (2*SE),3), #lower confidence interval
            HCI = round(Mean + (2*SE),3)) # upper confidence interval
dataSum
```

## Plot data 
### Plot non transformed data
Check what the best line is. Now just used the standard option (loess with y ~ x).
```{r fig.width=10, fig.height=7}
Stats <- "LMM: Time: p < 0.001; Inoculation: p = 0.304; Interaction: p = 0.017"
lab <-  c(expression("Isolate 1 (" * italic("Pedococcus-Phycicoccus") * ")"),
          expression("Isolate 2 (" * italic("Agrobacterium") * ", Rhizobiaceae)"), 
          "Isolate 3 (Rhizobiaceae)", 
          expression(italic("Pseudomonas simiae") ~ "WCS417"), 
          "Heat-killed control", 
          "Control")

p <- ggplot(data = data, aes(x = Time, y = Weight, color = Treatment, shape = Treatment)) +
  geom_point(data = data,
             aes(x = Time, y = Weight, color = Treatment, shape = Treatment),
             inherit.aes = FALSE,
             alpha = 0.3,
             size = 1) +
  geom_smooth(data = dataSum,
              aes(x = Time, y = Mean, color = Treatment),
              method = "nls",
              formula = y ~ Asym / (1 + exp((xmid - x) / scal)),
              method.args = list(start = list(Asym = 1200, xmid = 18, scal = 3.5),
                                 control = nls.control(maxiter = 200)),
              se = FALSE,
              linewidth = 0.7) +
  geom_errorbar(data = dataSum,
                inherit.aes = FALSE,
                aes(x = Time, y = Mean, color = Treatment, shape = Treatment,
                    ymin = Mean - SE,
                    ymax = Mean + SE),
                width = 0.4,
                linewidth = 0.7) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(0,3,6,9,12,15,18),
                     limits = c(0, 19)) +   # x limits
  scale_y_continuous(expand = c(0, 0),    #start graph at y = min and x = min
                     limits = c(0, 1000)) +   # x limits
  scale_shape_manual(name = "Inoculation treatment",
                     values = c(0:5),
                     labels = lab) + # change label in legend part shape
  scale_color_manual(name = "Inoculation treatment",
                     values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288"), # customize colours
                     labels = lab) +
  ylab("Caterpillar weight (mg)") +
  xlab("Time (days)") +
  ggtitle(Stats) +
  guides(color = guide_legend(override.aes = list(size = 2.5, linewidth = 1))) + # increase size of dots and width of the lines in legend
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(),     #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        #strip.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 11),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = c(0.4, 0.7),
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.key = element_rect(fill = "transparent", color = NA))
```
```{r fig.width=10, fig.height=7}
print(p)
```
```{r fig.width=6, fig.height=4, warning=FALSE}
print(p)
```
```{r fig.width=5, fig.height=4, warning=FALSE}
mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 6,
         height = 5,
         units = "cm",
         dpi = 300),
  x = c("FP_InoculationExpHM_growthCurve.svg", "FP_InoculationExpHM_growthCurve.png"))
```

### Plot with errorbars
```{r}
q <- ggplot(data = dataSum, aes(x = Time, y = Mean, color = Treatment, shape = Treatment)) +
  geom_smooth(data = dataSum,
              aes(x = Time, y = Mean, color = Treatment),
              method = "nls",
              formula = y ~ Asym / (1 + exp((xmid - x) / scal)),
              method.args = list(start = list(Asym = 1200, xmid = 18, scal = 3.5),
                                 control = nls.control(maxiter = 200)),
              se = FALSE,
              linewidth = 0.7) +
  geom_errorbar(data = dataSum,
                aes(ymin = Mean - SE,
                    ymax = Mean + SE),
                width = 0.4,
                linewidth = 0.7) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(0,3,6,9,12,15,18),
                     limits = c(0, 19)) +   # x limits
  scale_y_continuous(expand = c(0, 0),    #start graph at y = min and x = min
                     limits = c(0, 750)) +   # x limits
  scale_shape_manual(name = "Inoculation treatment",
                     values = c(0:5),
                     labels = lab) + # change label in legend part shape
  scale_color_manual(name = "Inoculation treatment",
                     values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288"), # customize colours
                     labels = lab) +
  ylab("Caterpillar weight (mg)") +
  xlab("Time (days)") +
  ggtitle(Stats) +
  guides(color = guide_legend(override.aes = list(size = 2.5, linewidth = 1))) + # increase size of dots and width of the lines in legend
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(),     #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        #strip.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 11),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = c(0.4, 0.7),
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.key = element_rect(fill = "transparent", color = NA))
```
```{r fig.width=10, fig.height=7, warning=FALSE}
print(q)
```
```{r fig.width=6, fig.height=4, warning=FALSE}
print(q)
```

```{r fig.width=5, fig.height=4, warning=FALSE}
mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 6,
         height = 5,
         units = "cm",
         dpi = 300),
  x = c("FP_InoculationExpHM_growthCurve_OnlyErrorbar.svg", "FP_InoculationExpHM_growthCurve_OnlyErrorbar.png"))
```

### Plot without day 18
```{r fig.width=10, fig.height=4}
Stats <- "LMM: Time: p < 0.001; Inoculation: p = 0.323; Interaction: p = 0.233"

fit <- lm(Mean ~ poly(Time, 2, raw = TRUE)*Treatment, data = dataSum[dataSum$Time %in% c(7, 13, 16), ])
#fit <- lm(Mean ~ Time*Treatment, data = dataSum[dataSum$Time %in% c(7, 13, 16), ])
newdat <- expand.grid(Time = seq(7, 16, length.out = 200),
                     Treatment = unique(dataSum$Treatment))

newdat$pred <- predict(fit, newdata = newdat)


# install.packages("scam")
# library(scam)
# 
# fit <- scam(Mean ~ Treatment + s(Time, by = Treatment, bs = "mpi"), data = dataSum[dataSum$Time %in% c(7, 13, 16), ])

r <- ggplot(data = data[data$Time %in% c(7, 13, 16), ], aes(x = Time, y = Weight, color = Treatment, shape = Treatment)) +
  geom_point(data = data[data$Time %in% c(7, 13, 16), ],
             aes(x = Time, y = Weight, color = Treatment, shape = Treatment),
             inherit.aes = FALSE,
             alpha = 0.3,
             size = 1) +
  geom_line(data = newdat,
          aes(x = Time, y = pred),
          linewidth = 1) +
  geom_errorbar(data = dataSum[dataSum$Time %in% c(7, 13, 16), ],
                inherit.aes = FALSE,
                aes(x = Time, y = Mean, color = Treatment, shape = Treatment,
                    ymin = Mean - SE,
                    ymax = Mean + SE),
                width = 0.4,
                linewidth = 0.7) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(0,4,8,12,16),
                     limits = c(0, 17)) +   # x limits
  scale_y_continuous(expand = c(0, 0),    #start graph at y = min and x = min
                     limits = c(0, 570)) +   # x limits
  scale_shape_manual(name = "Inoculation treatment",
                     values = c(0:5),
                     labels = lab) + # change label in legend part shape
  scale_color_manual(name = "Inoculation treatment",
                     values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288"), # customize colours
                     labels = lab) +
  ylab("Caterpillar weight (mg)") +
  xlab("Time (days)") +
  ggtitle(Stats) +
  guides(color = guide_legend(override.aes = list(size = 2.5, linewidth = 1))) + # increase size of dots and width of the lines in legend
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(),     #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        #strip.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 11),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = c(0.4, 0.7),
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.key = element_rect(fill = "transparent", color = NA))

#Does not work: data = data, method = "lm", formula = data$Weight ~ data$Time * data$Soil_conditioning; aes(color = Soil_conditioning
```
```{r fig.width=10, fig.height=7}
print(r)
```
```{r fig.width=6, fig.height=4, warning=FALSE}
print(r)
```
```{r fig.width=5, fig.height=4, warning=FALSE}
mapply(function(x)
  ggsave(plot = last_plot(), 
         filename = x,
         path = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs",
         scale = 2.2,
         width = 6,
         height = 5,
         units = "cm",
         dpi = 300),
  x = c("FP_InoculationExpHM_growthCurve_7-16.svg", "FP_InoculationExpHM_growthCurve_7-16.png"))
```

### Plot log(Weight) data
```{r, warning=FALSE}
ggplot(data = data, aes(y = log(Weight), x = Time, colour = Treatment)) +
  geom_point() +
  geom_smooth()
```
```{r, warning=FALSE}
ggplot(data = data[data$Time %in% c(7, 13, 16), ], aes(y = log(Weight), x = Time, colour = Treatment)) +
  geom_point() +
  geom_smooth()
```

### Boxplots per time point
```{r, fig.width=6, fig.height=4}
ggplot(data = data,
       mapping = aes(y = Weight, x = Treatment, fill = Treatment)) + #Defines the data
  stat_boxplot(aes(y = Weight, x = Treatment), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap(~ Time,
             nrow = 1,
             #labeller = labeller(Time = c("21" = "21 days"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.15,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean,
               geom = "point",
               size = 1.5,
               color = "white",
               shape = 18) + #Shows the mean as a white circle
  # geom_segment(aes(x = 0.8, y = 900, xend = 2.2, yend = 900)) +
  # geom_text(data = DataSum,
  #           aes(label = Let, y = 400), # fill out the significant letters
  #           size = 4,
  #           #hjust = -2.5
  #           ) + # labels in middle of line
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$Weight*1.2))) +   #y limits
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Caterpillar weight (mg)") +    #y-label with Mb in italic
  xlab("Inoculation treatment") +
  #ggtitle("\nWilcox: p = 0.002") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

```{r, fig.width=2.2, fig.height=4}
ggplot(data = data[data$Time == 7, ],
       mapping = aes(y = Weight, x = Treatment, fill = Treatment)) + #Defines the data
  stat_boxplot(aes(y = Weight, x = Treatment), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  facet_wrap(~ Time,
             nrow = 1,
             #labeller = labeller(Time = c("21" = "21 days"))
             ) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.15,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean,
               geom = "point",
               size = 1.5,
               color = "white",
               shape = 18) + #Shows the mean as a white circle
  # geom_segment(aes(x = 0.8, y = 900, xend = 2.2, yend = 900)) +
  # geom_text(data = DataSum,
  #           aes(label = Let, y = 400), # fill out the significant letters
  #           size = 4,
  #           #hjust = -2.5
  #           ) + # labels in middle of line
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data[data$Time == 7, "Weight"]*1.2))) +   #y limits
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Caterpillar weight (mg)") +    #y-label with Mb in italic
  xlab("Inoculation treatment") +
  #ggtitle("\nWilcox: p = 0.002") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

# Statistics
## Define factors
```{r}
data$Time <- as.factor(data$Time)
```

## Kruskall wallis test at day 13 and 16
```{r}
kruskal.test(data = data[data$Time == 13, ], Weight ~ Treatment)
kruskal.test(data = data[data$Time == 16, ], Weight ~ Treatment)
#kruskal.test(data = data[data$Time == 18, ], Weight ~ Treatment)
```
   
## Testing model at day 13 and 16
```{r, fig.height=7}
lm0 <- glmmTMB(data = data[data$Time == 13, ], formula = Weight ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
lm00 <- glmmTMB(data = data[data$Time == 13, ], formula = Weight ~ Treatment, family = gaussian(link = "identity"))

DHARMa.sum(lm0)
DHARMa.sum(lm00)

summary(lm0)
summary(lm00)
Anova(lm0)
Anova(lm00)

lm1 <- glmmTMB(data = data[data$Time == 16, ], formula = Weight ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
lm2 <- glmmTMB(data = data[data$Time == 16, ], formula = Weight ~ Treatment, family = gaussian(link = "identity"))

DHARMa.sum(lm1)
DHARMa.sum(lm2)

summary(lm1)
summary(lm2)
Anova(lm1)
Anova(lm2)

# lm3 <- glmmTMB(data = data[data$Time == 18, ], formula = Weight ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
# lm4 <- glmmTMB(data = data[data$Time == 18, ], formula = Weight ~ Treatment, family = gaussian(link = "identity"))
# 
# DHARMa.sum(lm3)
# DHARMa.sum(lm4)
# 
# summary(lm3)
# summary(lm4)
# Anova(lm3)
# Anova(lm4)
```
```{r}
rm(lm1, lm2, lm3, lm4)
```

    
## Test (G)LM all time points
### Testing model
ar1 only works well if time points are equally distant from each other
```{r}
# lm1 <- glmmTMB(data = data, formula = Weight ~ Time*Treatment + ar1(Time + 0 | nr), family = gaussian(link = "identity"))
# lm2  <- glmmTMB(data = data, formula = Weight ~ Time*Treatment + (1 | nr), family = gaussian(link = "identity"))
# lm3  <- glmmTMB(data = data, formula = Weight ~ Time*Treatment, family = gaussian(link = "identity"))

lm4 <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + ar1(Time + 0 | nr), family = gaussian(link = "identity"))
lm5a  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + (Time | nr), family = gaussian(link = "identity"))
lm5b  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + (1 | nr), family = gaussian(link = "identity"))
lm5c  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + (Time | nr) + (1 | Block), family = gaussian(link = "identity"))
lm5d  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + (1 | nr) + (1 | Block), family = gaussian(link = "identity"))
lm6  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment, family = gaussian(link = "identity"))

# lm7 <- glmmTMB(data = data, formula = sqrt(Weight) ~ Time*Treatment + ar1(Time + 0 | nr), family = gaussian(link = "identity"))
# lm8  <- glmmTMB(data = data, formula = sqrt(Weight) ~ Time*Treatment + (1 | nr), family = gaussian(link = "identity"))
# lm9  <- glmmTMB(data = data, formula = sqrt(Weight) ~ Time*Treatment, family = gaussian(link = "identity"))
# 
# glm1 <- glmmTMB(data = data, formula = Weight ~ Time*Treatment + ar1(Time + 0 | nr), family = Gamma(link = "log"))
# glm2  <- glmmTMB(data = data, formula = Weight ~ Time*Treatment + (1 | nr), family = Gamma(link = "log"))
# glm3  <- glmmTMB(data = data, formula = Weight ~ Time*Treatment, family = Gamma(link = "log"))
# 
# glm4 <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + ar1(Time + 0 | nr), family = Gamma(link = "log"))
# glm5  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment + (1 | nr), family = Gamma(link = "log"))
# glm6  <- glmmTMB(data = data, formula = log(Weight) ~ Time*Treatment, family = Gamma(link = "log"))
# 
# glm7 <- glmmTMB(data = data, formula = sqrt(Weight) ~ Time*Treatment + ar1(Time + 0 | nr), family = Gamma(link = "log"))
# glm8  <- glmmTMB(data = data, formula = sqrt(Weight) ~ Time*Treatment + (1 | nr), family = Gamma(link = "log"))
# glm9  <- glmmTMB(data = data, formula = sqrt(Weight) ~ Time*Treatment, family = Gamma(link = "log"))
```
https://cran.r-project.org/web/packages/glmmTMB/vignettes/troubleshooting.html    
https://cran.r-project.org/web/packages/glmmTMB/vignettes/covstruct.html     
https://github.com/glmmTMB/glmmTMB/issues/986    
   
### Model assumptions: DHARMa plots
```{r, fig.height=7}
# DHARMa.sum(lm1)
# DHARMa.sum(lm2)
# DHARMa.sum(lm3)
DHARMa.sum(lm4) #good
DHARMa.sum(lm5a)
DHARMa.sum(lm5b)
DHARMa.sum(lm5c) # best model
DHARMa.sum(lm5d)
#DHARMa.sum(lm6)
# DHARMa.sum(lm7)
# DHARMa.sum(lm8)
# DHARMa.sum(lm9)
# 
# DHARMa.sum(glm1) #good, but lm4 better
# #DHARMa.sum(glm2)
# #DHARMa.sum(glm3)
# DHARMa.sum(glm4)
# DHARMa.sum(glm5)
# DHARMa.sum(glm6)
# DHARMa.sum(glm7) # also ok
# #DHARMa.sum(glm8)
# #DHARMa.sum(glm9)
```

### Model output
```{r}
summary(lm5c)
Anova(lm5c)
```

### Posthoc
```{r}
emmeans(lm5c, pairwise ~ Treatment|Time, adjust = "tukey")
```

### Test without day 18
```{r, fig.height=7}
lm5c  <- glmmTMB(data = data[data$Time %in% c(7, 13, 16), ], formula = log(Weight) ~ Time*Treatment + (Time | nr) + (1 | Block), family = gaussian(link = "identity"))

DHARMa.sum(lm5c) # best model

Anova(lm5c)
```

## Area underneath curve
```{r}
data$nr <- as.numeric(data$nr)
data$Time <- as.numeric(data$Time)
data$Treatment <- as.numeric(data$Treatment)

auc_data <- data %>%
  group_by(Treatment, Block, nr) %>%
  arrange(Time, .by_group = TRUE) %>%
  summarise(
    AUC = trapz(Time, Weight),
    .groups = "drop"
  )

auc_data$Treatment <- as.factor(auc_data$Treatment)
```

### Plot
```{r, fig.width=4, fig.height=4}
ggplot(data = auc_data,
       mapping = aes(y = AUC, x = Treatment, fill = Treatment)) + #Defines the data
  stat_boxplot(aes(y = AUC, x = Treatment), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  # facet_wrap(~ Time,
  #            nrow = 1,
  #            labeller = labeller(Time = c("21" = "21 days"))) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.15,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean,
               geom = "point",
               size = 1.5,
               color = "white",
               shape = 18) + #Shows the mean as a white circle
  # geom_segment(aes(x = 0.8, y = 900, xend = 2.2, yend = 900)) +
  # geom_text(data = DataSum,
  #           aes(label = Let, y = 400), # fill out the significant letters
  #           size = 4,
  #           #hjust = -2.5
  #           ) + # labels in middle of line
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(auc_data$AUC*1.2))) +   #y limits
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Area underneath growth curve") +    #y-label with Mb in italic
  xlab("Inoculation treatment") +
  #ggtitle("\nWilcox: p = 0.002") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

### Testing model
```{r}
lm2  <- glmmTMB(data = auc_data, formula = AUC ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
lm3  <- glmmTMB(data = auc_data, formula = AUC ~ Treatment, family = gaussian(link = "identity"))

lm5  <- glmmTMB(data = auc_data, formula = log(AUC) ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
lm6  <- glmmTMB(data = auc_data, formula = log(AUC) ~ Treatment, family = gaussian(link = "identity"))

lm8  <- glmmTMB(data = auc_data, formula = sqrt(AUC) ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
lm9  <- glmmTMB(data = auc_data, formula = sqrt(AUC) ~ Treatment, family = gaussian(link = "identity"))

glm2  <- glmmTMB(data = auc_data, formula = AUC ~ Treatment + (1 | Block), family = Gamma(link = "log"))
glm3  <- glmmTMB(data = auc_data, formula = AUC ~ Treatment, family = Gamma(link = "log"))

glm5  <- glmmTMB(data = auc_data, formula = log(AUC) ~ Treatment + (1 | Block), family = Gamma(link = "log"))
glm6  <- glmmTMB(data = auc_data, formula = log(AUC) ~ Treatment, family = Gamma(link = "log"))

glm8  <- glmmTMB(data = auc_data, formula = sqrt(AUC) ~ Treatment + (1 | Block), family = Gamma(link = "log"))
glm9  <- glmmTMB(data = auc_data, formula = sqrt(AUC) ~ Treatment, family = Gamma(link = "log"))
```
https://cran.r-project.org/web/packages/glmmTMB/vignettes/troubleshooting.html    
https://cran.r-project.org/web/packages/glmmTMB/vignettes/covstruct.html     
https://github.com/glmmTMB/glmmTMB/issues/986    
   
### Model assumptions: DHARMa plots
```{r, fig.height=7}
DHARMa.sum(lm2)
DHARMa.sum(lm3)
DHARMa.sum(lm5)
DHARMa.sum(lm6)
DHARMa.sum(lm8)
DHARMa.sum(lm9)

DHARMa.sum(glm2) #Best
DHARMa.sum(glm3)
DHARMa.sum(glm5)
DHARMa.sum(glm6)
DHARMa.sum(glm8)
DHARMa.sum(glm9)
```

### Model output
```{r}
summary(glm2)
Anova(glm2)
```

```{r}
rm(data)
```


# Growth rate
#### Load data set
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Inoculation experiment/Data/Caterpillar_weight_inoculationExpHM2_GrowthRate.csv")
```

#### Remove NAs
```{r}
data <- data[complete.cases(data[ , "GrowthRate"]),]
```

## Explore data
```{r}
str(data)
head(data, 12)
summary(data) #check for outliers
```

### Summary table
```{r}
ddply(data, c("Treatment"), summarize,
            N = sum(!is.na(ln_GrowthRate)), #number of samples
            Mis = sum(is.na(ln_GrowthRate)), #number of missing values
            Mean = round(mean(ln_GrowthRate, na.rm=T),3), #mean
            Median = round(median(ln_GrowthRate, na.rm=T),3), #median
            SD = round(sd(ln_GrowthRate, na.rm=T),3), #standard deviation
            SE = round(sd(ln_GrowthRate, na.rm=T) / sqrt(N),5), #standard error
            LCI = round(Mean - (2*SE),3), #lower confidence interval
            HCI = round(Mean + (2*SE),3)) # upper confidence interval
```

### Plot
```{r, fig.width=4, fig.height=4}
ggplot(data = data,
       mapping = aes(y = ln_GrowthRate, x = Treatment, fill = Treatment)) + #Defines the data
  stat_boxplot(aes(y = ln_GrowthRate, x = Treatment), #Shows the error bars (should be added before box plots to have bars behind box plots)
               geom = "errorbar",
               linetype = 1,
               width = 0.2) +
  # facet_wrap(~ Time,
  #            nrow = 1,
  #            labeller = labeller(Time = c("21" = "21 days"))) + # Split data into three graphs, a graph for each caterpillar number
  geom_boxplot(outliers = FALSE) + #Defines that the graph will be a box plot and do not show outliers, this is shown by jitter
  geom_jitter(width = 0.15,
              height = 0,
              color = "black",
              alpha = 0.2) + # Transparency is regulated with the alpha argument
  stat_summary(fun = mean,
               geom = "point",
               size = 1.5,
               color = "white",
               shape = 18) + #Shows the mean as a white circle
  # geom_segment(aes(x = 0.8, y = 900, xend = 2.2, yend = 900)) +
  # geom_text(data = DataSum,
  #           aes(label = Let, y = 400), # fill out the significant letters
  #           size = 4,
  #           #hjust = -2.5
  #           ) + # labels in middle of line
  scale_y_continuous(expand = c(0, 0),        #start graph at y = min and x = min
                     #breaks = seq(0, 30, 5)) #steps on the y-axis
                     limits = c(0, max(data$ln_GrowthRate*1.2))) +   #y limits
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#882255", "#332288")) + # Manually change the colors of the bars. You need to define fill in second line of the plot script to get it working; c("#009E73", "#56B4E9")
  ylab("Growth rate") +    #y-label with Mb in italic
  xlab("Inoculation treatment") +
  #ggtitle("\nWilcox: p = 0.002") +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 2.5), #, hjust = -0
        strip.text = element_text(size = 13),
        plot.margin = margin(5.5, 12, 5.5, 5.5, "points"),
        legend.position = "none")
```

## Testing model
```{r, fig.height=7}
lm1 <- glmmTMB(data = data, formula = ln_GrowthRate ~ Treatment + (1 | Block), family = gaussian(link = "identity"))
lm2 <- glmmTMB(data = data, formula = ln_GrowthRate  ~ Treatment, family = gaussian(link = "identity"))

DHARMa.sum(lm1)
DHARMa.sum(lm2)

summary(lm1)
Anova(lm1)
```


```{r}
rm(data)
```


# Behaviour
#### Load data set
```{r}
data <- read.csv("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Inoculation experiment/Data/Caterpillar_weight_inoculationExpHM2_GrowthRate.csv")
```

#### Remove NAs
```{r}
data <- data[!is.na(data$Cat_location_day.18) &
             data$Cat_location_day.18 != "", ]
```

## Explore data
```{r}
str(data)
head(data, 12)
summary(data) #check for outliers
```

```{r}
ggplot(data = data, 
       mapping = aes(y = Cat_location_day.18, x = Treatment, fill = Cat_location_day.18)) + #Defines the data2
  geom_bar(data = data, 
                   stat = "identity", 
                   position = "stack") +
  #scale_fill_manual(values = c("#56B4E9", "#009E73"), 
                  #  name = "Soil treatment", #name in the legend
                   # labels = c("CTRL", ~italic("M. brassicae\nPicture")),  #labels in the legend
                    #guide = guide_legend(override.aes = list(pattern = "none"))
                  #) + # Change legend appearance; guide = FALSE to remove this part in the legend
  xlab("Caterpillar location day 18") +
  ylab("Location") +    #y-label
#  ggtitle(Stats) +
  theme(panel.background = element_rect(fill = "white"), #theme of the graph
        panel.border = element_blank(), #no border around the graph (for border, use: element_rect(color = "gray30", fill = NA))
        axis.line = element_line(), #show line for x and y axis
        plot.title = element_text(size = 10, hjust = 0),
        axis.text = element_text(size = 13, color = "black"),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 1.5), #, hjust = -0
        strip.text = element_text(size = 13),
        legend.position = "right")
```

