---
title: "CP&FP_05_differential_abundance_Summary_Results - Upset plots"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

This script is adapted from a script form Melissa Uribe Acosta. Extra input comes from a script of Pedro Beschore da Costa.   
   
   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
   
   
# 4.0 load libraries, improve memory use and subset objects for the desired treatment comparisons
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(UpSetR)
packageVersion("UpSetR")
library(ggVennDiagram)
packageVersion("ggVennDiagram")
```


# 4.1 Comparing CP and FP at ASV level
## One test
### Load data
```{r}
# CP
## Acc ASVs detected by at least one test
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV.RData")
Acc_Bac_NoDup_DA_ASV_CP <- Acc_Bac_NoDup_DA_ASV
rm(Acc_Bac_NoDup_DA_ASV)

## Dom data ASVs detected by at least one test
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Dom_Bac_NoDup_DA_ASV.RData")
Dom_Bac_NoDup_DA_ASV_CP <- Dom_Bac_NoDup_DA_ASV
rm(Dom_Bac_NoDup_DA_ASV)

## All data ASVs detected by at least one test
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_NoDup_DA_ASV.RData")
All_Bac_NoDup_DA_ASV_CP <- All_Bac_NoDup_DA_ASV
rm(All_Bac_NoDup_DA_ASV)

# FP
## Acc ASVs detected by at least one test
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_NoDup_DA_ASV.RData")
Acc_Bac_NoDup_DA_ASV_FP <- Acc_Bac_NoDup_DA_ASV
rm(Acc_Bac_NoDup_DA_ASV)

## All data ASVs detected by at least one test
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_NoDup_DA_ASV.RData")
All_Bac_NoDup_DA_ASV_FP <- All_Bac_NoDup_DA_ASV
rm(All_Bac_NoDup_DA_ASV)
```

### Upset plots
#### All accessions one test
```{r, fig.hight = 12}
Acc_CP_FP <- c(CP = Acc_Bac_NoDup_DA_ASV_CP, FP = Acc_Bac_NoDup_DA_ASV_FP)
upset(fromList(Acc_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### Per accession one test
```{r}
VL_CP_FP <- list(CP_VL = Acc_Bac_NoDup_DA_ASV_CP$VL, FP_VL = Acc_Bac_NoDup_DA_ASV_FP$VL)
upset(fromList(VL_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

CD_CP_FP <- list(CP_CD = Acc_Bac_NoDup_DA_ASV_CP$CD, FP_CD = Acc_Bac_NoDup_DA_ASV_FP$CD)
upset(fromList(CD_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

RI_CP_FP <- list(CP_RI = Acc_Bac_NoDup_DA_ASV_CP$RI, FP_RI = Acc_Bac_NoDup_DA_ASV_FP$RI)
upset(fromList(RI_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

HM_CP_FP <- list(CP_HM = Acc_Bac_NoDup_DA_ASV_CP$HM, FP_HM = Acc_Bac_NoDup_DA_ASV_FP$HM)
upset(fromList(HM_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

GO1_CP_FP <- list(CP_GO1 = Acc_Bac_NoDup_DA_ASV_CP$GO1, FP_GO1 = Acc_Bac_NoDup_DA_ASV_FP$GO1)
upset(fromList(GO1_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### All data one test
```{r}
All_CP_FP <- list(CP = All_Bac_NoDup_DA_ASV_CP, FP = All_Bac_NoDup_DA_ASV_FP)
upset(fromList(All_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### All and domestication data one test
```{r}
All_Dom_CP_FP <- c(list(CP_All = All_Bac_NoDup_DA_ASV_CP, FP_All = All_Bac_NoDup_DA_ASV_FP), CP_dom = Dom_Bac_NoDup_DA_ASV_CP)
upset(fromList(All_Dom_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### All DA ASVs together
```{r}
Acc_Bac_NoDup_DA_ASV_CP <- unlist(unique(Acc_Bac_NoDup_DA_ASV_CP))
Acc_Bac_NoDup_DA_ASV_FP <- unlist(unique(Acc_Bac_NoDup_DA_ASV_FP))
Dom_Bac_NoDup_DA_ASV_CP <- unlist(unique(Dom_Bac_NoDup_DA_ASV_CP))
Acc_Dom_All_CP_FP <- list(Acc_CP = Acc_Bac_NoDup_DA_ASV_CP, Acc_FP = Acc_Bac_NoDup_DA_ASV_FP, All_CP = All_Bac_NoDup_DA_ASV_CP, All_FP = All_Bac_NoDup_DA_ASV_FP, Dom_CP = Dom_Bac_NoDup_DA_ASV_CP)
upset(fromList(Acc_Dom_All_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

## Two tests
### Load data
```{r}
# CP
## Acc ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
Acc_Bac_TwoTimes_DA_ASV_CP <- Acc_Bac_TwoTimes_DA_ASV
rm(Acc_Bac_TwoTimes_DA_ASV)

## Dom data ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Dom_Bac_TwoTimes_DA_ASV.RData")
Dom_Bac_TwoTimes_DA_ASV_CP <- Dom_Bac_TwoTimes_DA_ASV
rm(Dom_Bac_TwoTimes_DA_ASV)

## All data ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")
All_Bac_TwoTimes_DA_ASV_CP <- All_Bac_TwoTimes_DA_ASV
rm(All_Bac_TwoTimes_DA_ASV)

# FP
## Acc ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
Acc_Bac_TwoTimes_DA_ASV_FP <- Acc_Bac_TwoTimes_DA_ASV
rm(Acc_Bac_TwoTimes_DA_ASV)

## All data ASVs detected by at least two tests
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")
All_Bac_TwoTimes_DA_ASV_FP <- All_Bac_TwoTimes_DA_ASV
rm(All_Bac_TwoTimes_DA_ASV)
```

### Upset plots
#### All accessions two tests
```{r, fig.hight = 12}
Acc_CP_FP <- c(CP = Acc_Bac_TwoTimes_DA_ASV_CP, FP = Acc_Bac_TwoTimes_DA_ASV_FP)
upset(fromList(Acc_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### Per accession two tests
```{r}
VL_CP_FP <- list(CP_VL = Acc_Bac_TwoTimes_DA_ASV_CP$VL, FP_VL = Acc_Bac_TwoTimes_DA_ASV_FP$VL)
upset(fromList(VL_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

CD_CP_FP <- list(CP_CD = Acc_Bac_TwoTimes_DA_ASV_CP$CD, FP_CD = Acc_Bac_TwoTimes_DA_ASV_FP$CD)
upset(fromList(CD_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

RI_CP_FP <- list(CP_RI = Acc_Bac_TwoTimes_DA_ASV_CP$RI, FP_RI = Acc_Bac_TwoTimes_DA_ASV_FP$RI)
upset(fromList(RI_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

HM_CP_FP <- list(CP_HM = Acc_Bac_TwoTimes_DA_ASV_CP$HM, FP_HM = Acc_Bac_TwoTimes_DA_ASV_FP$HM)
upset(fromList(HM_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)

GO1_CP_FP <- list(CP_GO1 = Acc_Bac_TwoTimes_DA_ASV_CP$GO1, FP_GO1 = Acc_Bac_TwoTimes_DA_ASV_FP$GO1)
upset(fromList(GO1_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### All data two tests
```{r}
All_CP_FP <- list(CP = All_Bac_TwoTimes_DA_ASV_CP, FP = All_Bac_TwoTimes_DA_ASV_FP)
upset(fromList(All_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### All and domestication data one test
```{r}
All_Dom_CP_FP <- c(list(CP_All = All_Bac_TwoTimes_DA_ASV_CP, FP_All = All_Bac_TwoTimes_DA_ASV_FP), CP_dom = Dom_Bac_TwoTimes_DA_ASV_CP)
upset(fromList(All_Dom_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

#### All DA ASVs together
```{r}
Acc_Bac_TwoTimes_DA_ASV_CP <- unlist(unique(Acc_Bac_TwoTimes_DA_ASV_CP))
Acc_Bac_TwoTimes_DA_ASV_FP <- unlist(unique(Acc_Bac_TwoTimes_DA_ASV_FP))
Dom_Bac_TwoTimes_DA_ASV_CP <- unlist(unique(Dom_Bac_TwoTimes_DA_ASV_CP))
Acc_Dom_All_CP_FP <- list(Acc_CP = Acc_Bac_TwoTimes_DA_ASV_CP, Acc_FP = Acc_Bac_TwoTimes_DA_ASV_FP, All_CP = All_Bac_TwoTimes_DA_ASV_CP, All_FP = All_Bac_TwoTimes_DA_ASV_FP, Dom_CP = Dom_Bac_TwoTimes_DA_ASV_CP)
upset(fromList(Acc_Dom_All_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Overlapping ASVs CP and FP
#### Overlap per accession
```{r}
Acc_overlap_DA <- intersect(Acc_Bac_TwoTimes_DA_ASV_CP, Acc_Bac_TwoTimes_DA_ASV_FP)
Acc_overlap_DA
save(Acc_overlap_DA, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_OverlapDA/Acc_overlap_DA.RData")
```

#### Overlap in all data sets together
```{r}
Tot_overlap_DA <- intersect(unique(c(Acc_Bac_TwoTimes_DA_ASV_CP, Dom_Bac_TwoTimes_DA_ASV_CP, All_Bac_TwoTimes_DA_ASV_CP)), unique(c(Acc_Bac_TwoTimes_DA_ASV_FP, All_Bac_TwoTimes_DA_ASV_FP)))
Tot_overlap_DA
save(Tot_overlap_DA, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_OverlapDA/Tot_overlap_DA.RData")
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.2 Comparing CP and FP at Family level
## Two tests
### Load data
```{r}
# CP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_FamilyLevel/Acc_TwoTimes_DA_Family_lfctax.RData")
lfctax_CP <- Acc_TwoTimes_DA_Family_lfctax
rm(Acc_TwoTimes_DA_Family_lfctax)

# FP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_FamilyLevel/Acc_TwoTimes_DA_Family_lfctax.RData")
lfctax_FP <- Acc_TwoTimes_DA_Family_lfctax
rm(Acc_TwoTimes_DA_Family_lfctax)
```

### Upset plots
#### All accessions two tests
```{r, fig.hight = 12}
Acc_CP_FP <- list(CP = lfctax_CP$taxon, FP = lfctax_FP$taxon)
upset(fromList(Acc_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.3 Comparing CP and FP at Order level
## Two tests
### Load data
```{r}
# CP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_OrderLevel/Acc_TwoTimes_DA_Order_lfctax.RData")
lfctax_CP <- Acc_TwoTimes_DA_Order_lfctax
rm(Acc_TwoTimes_DA_Order_lfctax)

# FP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_OrderLevel/Acc_TwoTimes_DA_Order_lfctax.RData")
lfctax_FP <- Acc_TwoTimes_DA_Order_lfctax
rm(Acc_TwoTimes_DA_Order_lfctax)
```

### Upset plots
#### All accessions two tests
```{r, fig.hight = 12}
Acc_CP_FP <- list(CP = lfctax_CP$taxon, FP = lfctax_FP$taxon)
upset(fromList(Acc_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.4 Comparing CP and FP at Class level
## Two tests
### Load data
```{r}
# CP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_TwoTimes_DA_Class_lfctax.RData")
lfctax_CP <- Acc_TwoTimes_DA_Class_lfctax
rm(Acc_TwoTimes_DA_Class_lfctax)

# FP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ClassLevel/Acc_TwoTimes_DA_Class_lfctax.RData")
lfctax_FP <- Acc_TwoTimes_DA_Class_lfctax
rm(Acc_TwoTimes_DA_Class_lfctax)
```

### Upset plots
#### All accessions two tests
```{r, fig.hight = 12}
Acc_CP_FP <- list(CP = lfctax_CP$taxon, FP = lfctax_FP$taxon)
upset(fromList(Acc_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.5 Comparing CP and FP at Phylum level
## Two tests
### Load data
```{r}
# CP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_TwoTimes_DA_Phylum_lfctax.RData")
lfctax_CP <- Acc_TwoTimes_DA_Phylum_lfctax
rm(Acc_TwoTimes_DA_Phylum_lfctax)

# FP
## Acc ASVs detected by at least two tests
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_PhylumLevel/Acc_TwoTimes_DA_Phylum_lfctax.RData")
lfctax_FP <- Acc_TwoTimes_DA_Phylum_lfctax
rm(Acc_TwoTimes_DA_Phylum_lfctax)
```

### Upset plots
#### All accessions two tests
```{r, fig.hight = 12}
Acc_CP_FP <- list(CP = lfctax_CP$Phylum, FP = lfctax_FP$Phylum)
upset(fromList(Acc_CP_FP), order.by = "freq", nsets = 20, nintersects = 100, set_size.show = TRUE)
```

### Overlap
```{r}
intersect(lfctax_CP$Phylum, lfctax_FP$Phylum)
```
```{r}
# Clean environment
rm(list = ls())
```


# 4.6 Venn diagram DA
## Accession
```{r}
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DifferentialAbundance/SummaryFiles/Acc_Bac_conca_NoAn_nodup.RData")

simple_Venn <- ggVennDiagram(Acc_Bac_conca_NoAn_nodup)
print(simple_Venn)

# ASVs in both VL adn HM
ASV_VL_HM <- unlist(Acc_Bac_conca_NoAn_nodup[c("HM", "VL")])
Duplicates_DA_VL_HM <- ASV_VL_HM[duplicated(ASV_VL_HM)]
print(Duplicates_DA_VL_HM)

save(Duplicates_DA_VL_HM, file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DifferentialAbundance/SummaryFiles/Duplicates_DA_VL_HM.RData")
```
