---
title: "CP_FP_05 - Heatmaps of ASVs that are DA in CP and in FP"
author: "Kris de Kreek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

   
Tutorials   
- [Heatmap for final plot](http://rstudio-pubs-static.s3.amazonaws.com/288398_185f2889a5f641c6b9aa7b14fa15b634.html)
- [Excamples of pheatmap package](https://davetang.github.io/muse/pheatmap.html)   
   
   
# 5.0 load libraries
### Load libraries
```{r, warning = FALSE, message=FALSE}
R.version$version.string # prints R version

library(phyloseq)
packageVersion("phyloseq")
library(pheatmap)
packageVersion("pheatmap")
```


# 5.1 Heatmap overlapping ASVs Acc
## 5.1.1 Load data
```{r}
# load phyloseq object CP-FP together
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_together/unnormalized_bac_ps.RData")

# load DA ASVs CP for Acc
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
Acc_Bac_TwoTimes_DA_ASV_CP <- Acc_Bac_TwoTimes_DA_ASV
rm(Acc_Bac_TwoTimes_DA_ASV)

# load DA ASVs FP for Acc
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
Acc_Bac_TwoTimes_DA_ASV_FP <- Acc_Bac_TwoTimes_DA_ASV
rm(Acc_Bac_TwoTimes_DA_ASV)

# ASVs overlapping in DA tests CP and FP for Acc
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_OverlapDA/Acc_overlap_DA.RData")

# load logfold change values CP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
ancomWZ_Bac_Acc_CP <- ancomWZ_Bac_Acc
rm(ancomWZ_Bac_Acc)

# load logfold change values FP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
ancomWZ_Bac_Acc_FP <- ancomWZ_Bac_Acc
rm(ancomWZ_Bac_Acc)
```

## 5.1.2 Extract logfold change
### CP
```{r}
# Filter logfold change for DA ASVs per accession
lfc_DA_specific_CP <- lapply(names(ancomWZ_Bac_Acc_CP), function(name) {
  da_asvs <- intersect(Acc_overlap_DA, Acc_Bac_TwoTimes_DA_ASV_CP[[name]]) #Only ASVs that are shared between CP and FP and were DA in CP
  lfc_table <- ancomWZ_Bac_Acc_CP[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Cat_treatmentMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name) #, "_CP"
  lfc_table
})

names(lfc_DA_specific_CP) <- names(lfc_DA_specific_CP)

# merge all accessions together
AllAcc_lfc_CP <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific_CP)
rownames(AllAcc_lfc_CP) <- AllAcc_lfc_CP$ASV
AllAcc_lfc_CP
```

### FP
```{r}
# Filter logfold change for DA ASVs per accession
lfc_DA_specific_FP <- lapply(names(ancomWZ_Bac_Acc_FP), function(name) {
  da_asvs <- intersect(Acc_overlap_DA, Acc_Bac_TwoTimes_DA_ASV_FP[[name]]) #Only ASVs that are shared between CP and FP and were DA in FP
  lfc_table <- ancomWZ_Bac_Acc_FP[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(" ", name)
  lfc_table
})

names(lfc_DA_specific_FP) <- names(lfc_DA_specific_FP)

# merge all accessions together
AllAcc_lfc_FP <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific_FP)
rownames(AllAcc_lfc_FP) <- AllAcc_lfc_FP$ASV

# Remove RI
AllAcc_lfc_FP <- AllAcc_lfc_FP[ , -4]
AllAcc_lfc_FP
```

### Merge CP and FP
```{r}
AllAcc_lfc <- merge(AllAcc_lfc_CP, AllAcc_lfc_FP, by = "ASV")
rownames(AllAcc_lfc) <- AllAcc_lfc$ASV
```

### Adding taxonomy
```{r}
# Only keep DA in ps object
ps <- prune_taxa(Acc_overlap_DA, unnormalized_bac_ps)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df_clean <- psmelt(ps)
head(df_clean[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df_clean[!duplicated(df_clean$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)

# Add genus to matrix
AllAcc_lfc2 <- AllAcc_lfc
tax$ASV <- rownames(tax)
AllAcc_lfc2 <- merge(tax, AllAcc_lfc2, ID = "ASV")

# Add higher taxonomic level if genus is NA or Incertae Sedis
taxon <- AllAcc_lfc2$Genus
taxon[is.na(taxon)] <- AllAcc_lfc2$Family[is.na(taxon)]
taxon[is.na(taxon)] <- AllAcc_lfc2$Order[is.na(taxon)]

taxon[taxon == "g__Incertae_Sedis"] <- AllAcc_lfc2$Family[taxon == "g__Incertae_Sedis"]
taxon[taxon == "f__Incertae_Sedis"] <- AllAcc_lfc2$Order[taxon == "f__Incertae_Sedis"]
rownames(AllAcc_lfc2) <- paste(AllAcc_lfc2$ASV, taxon, sep = "_")

AllAcc_lfc2$Phylum <- sub("p__", "", AllAcc_lfc2$Phylum)
AllAcc_lfc2$Class <- sub("c__", "", AllAcc_lfc2$Class)

# Subset data columns
tax <- AllAcc_lfc2[1:6]
AllAcc_lfc2 <- AllAcc_lfc2[7:22]
head(AllAcc_lfc2)

# Make color pallet for class
colors15 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7",
              "#882255", "#44AA99", "#117733", "#332288", "#AA4499", "#DDCC77") # "#000000", "#999999", 

# Define colors
classes2 <- unique(tax$Class)
class_colors2 <- setNames(colors15[1:length(classes2)], classes2)
class_colors2 <- list(Class = class_colors2)

phylums2 <- unique(tax$Phylum)
phylums_colors2 <- setNames(colors15[1:length(phylums2)], phylums2)
phylums_colors2 <- list(Phylum = phylums_colors2)

colors <- c(phylums_colors2, class_colors2)

# Turn data frame into a matrix
AllAcc_lfc$ASV <- NULL
AllAcc_lfc <- as.matrix(AllAcc_lfc)
AllAcc_lfc2 <- as.matrix(AllAcc_lfc2)

# replace NAs by 0s
AllAcc_lfc_clu <- AllAcc_lfc
AllAcc_lfc2_clu <- AllAcc_lfc2
AllAcc_lfc_clu[is.na(AllAcc_lfc_clu)] <- 0
AllAcc_lfc2_clu[is.na(AllAcc_lfc2_clu)] <- 0
```

### Heatmap
```{r, fig.height=3, fig.width=12}
max_abs <- max(abs(AllAcc_lfc2), na.rm = TRUE)
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
         #cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          gaps_col = c(12),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Phylum"],
          annotation_colors = colors)
```
```{r, fig.height=2.3, fig.width=12}
max_abs <- max(abs(AllAcc_lfc2), na.rm = TRUE)
p <- pheatmap(AllAcc_lfc2,
          #clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          gaps_col = c(12),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Class"],
          annotation_colors = colors)
print(p)

files <- c("CP_FP_DAHeatmap_Overlap_ASV.svg", "CP_FP_DAHeatmap_Overlap_ASV.png")

mapply(function(x){
  file_path <- file.path("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/Result/Final graphs", x)

  if (grepl("svg$", x)) {
    svg(file_path, width = 24/2.54, height = 3.2/2.54)
  } else {
    png(file_path, width = 24, height = 3.2, units = "cm", res = 300)
  }
  grid::grid.draw(p$gtable)
  dev.off()
}, files)
```
```{r, fig.height=3, fig.width=12}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
         #cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[5:6],
          annotation_colors = colors)
```
```{r, fig.height=3, fig.width=12}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          #clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
         cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[5:6],
          annotation_colors = colors)
```
```{r, fig.height=4, fig.width=12}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
         #cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[3:6],
          annotation_colors = colors)
```
```{r, fig.height=2.3, fig.width=12}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          #clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          gaps_col = c(12),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[5:6],
          annotation_colors = colors)
```
```{r, fig.height=3.5, fig.width=12}
pheatmap(AllAcc_lfc2,
          clustering_distance_rows = dist(AllAcc_lfc2_clu),
          #clustering_distance_cols = dist(t(AllAcc_lfc2_clu)),
          cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("blue", "white", "red"))(50),
          breaks = seq(-max_abs, max_abs, length.out = 51),
          gaps_col = c(12),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[3:6],
          annotation_colors = colors)
```
```{r}
# Clean environment
rm(list = ls())
```


# 5.2 Heatmap overlapping ASVs All
ASVs overlapping in DA tests CP and FP for Acc, Dom and All data

## 5.2.1 Load data
```{r}
# load phyloseq object CP-FP together
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_together/unnormalized_bac_ps.RData")

# load DA ASVs CP for Acc
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
Acc_Bac_TwoTimes_DA_ASV_CP <- Acc_Bac_TwoTimes_DA_ASV
rm(Acc_Bac_TwoTimes_DA_ASV)

# load DA ASVs FP for Acc
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/Acc_Bac_TwoTimes_DA_ASV.RData")
Acc_Bac_TwoTimes_DA_ASV_FP <- Acc_Bac_TwoTimes_DA_ASV
rm(Acc_Bac_TwoTimes_DA_ASV)

# load DA ASVs CP for Dom
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/Dom_Bac_TwoTimes_DA_ASV.RData")
Dom_Bac_TwoTimes_DA_ASV_CP <- Dom_Bac_TwoTimes_DA_ASV
rm(Dom_Bac_TwoTimes_DA_ASV)

# load DA ASVs CP for All
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")
All_Bac_TwoTimes_DA_ASV_CP <- All_Bac_TwoTimes_DA_ASV
rm(All_Bac_TwoTimes_DA_ASV)

# load DA ASVs FP for All
load("C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_SummaryFiles/SummaryFiles_ASVLevel/All_Bac_TwoTimes_DA_ASV.RData")
All_Bac_TwoTimes_DA_ASV_FP <- All_Bac_TwoTimes_DA_ASV
rm(All_Bac_TwoTimes_DA_ASV)

# ASVs overlapping in DA tests CP and FP for Acc, Dom and All data
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_OverlapDA/Tot_overlap_DA.RData")

# load logfold change values Acc CP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
ancomWZ_Bac_Acc_CP <- ancomWZ_Bac_Acc
rm(ancomWZ_Bac_Acc)

# load logfold change values Acc FP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_Acc.RData")
ancomWZ_Bac_Acc_FP <- ancomWZ_Bac_Acc
rm(ancomWZ_Bac_Acc)

# load logfold change values Dom CP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_Dom.RData")
ancomWZ_Bac_Dom_CP <- ancomWZ_Bac_Dom
rm(ancomWZ_Bac_Dom)

# load logfold change values All CP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_DA_Ancom_ASVLevel/ancomWZ_Bac_All.RData")
ancomWZ_Bac_All_CP <- ancomWZ_Bac_All
rm(ancomWZ_Bac_All)

# load logfold change values All FP
load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/FP_DA_Ancom_ASVLevel/ancomWZ_Bac_All.RData")
ancomWZ_Bac_All_FP <- ancomWZ_Bac_All
rm(ancomWZ_Bac_All)
```

## 5.2.2 Extract logfold change
### CP Acc
```{r}
# Filter logfold change for DA ASVs per accession
lfc_DA_specific_CP_Acc <- lapply(names(ancomWZ_Bac_Acc_CP), function(name) {
  da_asvs <- intersect(Tot_overlap_DA, Acc_Bac_TwoTimes_DA_ASV_CP[[name]]) #Only ASVs that are shared between CP and FP and were DA in CP
  lfc_table <- ancomWZ_Bac_Acc_CP[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Cat_treatmentMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name, "_CP")
  lfc_table
})

names(lfc_DA_specific_CP_Acc) <- names(ancomWZ_Bac_Acc_CP)

# merge all accessions together
AllAcc_lfc_CP <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific_CP_Acc)
rownames(AllAcc_lfc_CP) <- AllAcc_lfc_CP$ASV
AllAcc_lfc_CP
```

### FP Acc
```{r}
# Filter logfold change for DA ASVs per accession
lfc_DA_specific_FP_Acc <- lapply(names(ancomWZ_Bac_Acc_FP), function(name) {
  da_asvs <- intersect(Tot_overlap_DA, Acc_Bac_TwoTimes_DA_ASV_FP[[name]]) #Only ASVs that are shared between CP and FP and were DA in FP
  lfc_table <- ancomWZ_Bac_Acc_FP[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name, "_FP")
  lfc_table
})

names(lfc_DA_specific_FP_Acc) <- names(ancomWZ_Bac_Acc_FP)

# merge all accessions together
AllAcc_lfc_FP <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific_FP_Acc)
rownames(AllAcc_lfc_FP) <- AllAcc_lfc_FP$ASV
AllAcc_lfc_FP
```

### CP Dom
```{r}
# Filter logfold change for DA ASVs per accession
lfc_DA_specific_CP_Dom <- lapply(names(ancomWZ_Bac_Dom_CP), function(name) {
  da_asvs <- intersect(Tot_overlap_DA, Dom_Bac_TwoTimes_DA_ASV_CP[[name]]) #Only ASVs that are shared between CP and FP and were DA in CP
  lfc_table <- ancomWZ_Bac_Dom_CP[[name]]$res$lfc
  rownames(lfc_table) <- lfc_table$taxon
  lfc_table <- lfc_table[lfc_table$taxon %in% da_asvs, "Cat_treatmentMb", drop = FALSE]
  lfc_table$ASV <- rownames(lfc_table)
  colnames(lfc_table)[1] <- paste0(name, "_CP")
  lfc_table
})

names(lfc_DA_specific_CP_Dom) <- names(ancomWZ_Bac_Dom_CP)

# merge all accessions together
AllDom_lfc_CP <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), lfc_DA_specific_CP_Dom)
rownames(AllDom_lfc_CP) <- AllDom_lfc_CP$ASV
AllDom_lfc_CP
```

### CP All
```{r}
# Filter logfold change for DA ASVs per accession
da_asvs <- intersect(Tot_overlap_DA, All_Bac_TwoTimes_DA_ASV_CP) #Only ASVs that are shared between CP and FP and were DA in CP
lfc_DA_specific_CP_All <- ancomWZ_Bac_All_CP$res$lfc
rownames(lfc_DA_specific_CP_All) <- lfc_DA_specific_CP_All$taxon
lfc_DA_specific_CP_All <- lfc_DA_specific_CP_All[lfc_DA_specific_CP_All$taxon %in% da_asvs, "Cat_treatmentMb", drop = FALSE]
lfc_DA_specific_CP_All$ASV <- rownames(lfc_DA_specific_CP_All)
colnames(lfc_DA_specific_CP_All)[1] <- paste0("All_CP")
lfc_DA_specific_CP_All
```
```{r}
rm(da_asvs)
```

### FP All
```{r}
# Filter logfold change for DA ASVs per accession
da_asvs <- intersect(Tot_overlap_DA, All_Bac_TwoTimes_DA_ASV_FP) #Only ASVs that are shared between FP and FP and were DA in FP
lfc_DA_specific_FP_All <- ancomWZ_Bac_All_FP$res$lfc
rownames(lfc_DA_specific_FP_All) <- lfc_DA_specific_FP_All$taxon
lfc_DA_specific_FP_All <- lfc_DA_specific_FP_All[lfc_DA_specific_FP_All$taxon %in% da_asvs, "Soil_conditioningMb", drop = FALSE]
lfc_DA_specific_FP_All$ASV <- rownames(lfc_DA_specific_FP_All)
colnames(lfc_DA_specific_FP_All)[1] <- paste0("All_FP")
lfc_DA_specific_FP_All
```

### Merge CP and FP
```{r}
All_lfc_l <- list(AllAcc_lfc_CP, AllAcc_lfc_FP, AllDom_lfc_CP, lfc_DA_specific_CP_All, lfc_DA_specific_FP_All)
All_lfc_df <- Reduce(function(x, y) merge(x, y, by = "ASV", all = TRUE), All_lfc_l)
rownames(All_lfc_df) <- All_lfc_df$ASV
```

### Adding taxonomy
```{r}
# Only keep DA in ps object
ps <- prune_taxa(Tot_overlap_DA, unnormalized_bac_ps)

# Phyloseq object to data frame with relative abundance, meta data and taxonomy. Each row is a ASV-sample combination
df_clean <- psmelt(ps)
head(df_clean[, c(1:9, 95:101)], 12)

# Extract unique ASV–Class-Phylum mapping
tax <- df_clean[!duplicated(df_clean$OTU), c("OTU", "Genus", "Family", "Order", "Class", "Phylum")]

# Row names must match heat map rows
rownames(tax) <- tax$OTU
tax$OTU <- NULL
head(tax)

# Add genus to matrix
All_lfc_df2 <- All_lfc_df
tax$ASV <- rownames(tax)
All_lfc_df2 <- merge(tax, All_lfc_df2, ID = "ASV")

# Add higher taxonomic level if genus is NA or Incertae Sedis
taxon <- All_lfc_df2$Genus
taxon[is.na(taxon)] <- All_lfc_df2$Family[is.na(taxon)]
taxon[is.na(taxon)] <- All_lfc_df2$Order[is.na(taxon)]

taxon[taxon == "g__Incertae_Sedis"] <- All_lfc_df2$Family[taxon == "g__Incertae_Sedis"]
taxon[taxon == "f__Incertae_Sedis"] <- All_lfc_df2$Order[taxon == "f__Incertae_Sedis"]
rownames(All_lfc_df2) <- paste(All_lfc_df2$ASV, taxon, sep = "_")

# Subset data columns
tax <- All_lfc_df2[1:6]
All_lfc_df2 <- All_lfc_df2[7:27]
head(All_lfc_df2)

# Make color pallet for class
classes <- unique(tax$Class)
class_colors <- setNames(rainbow(length(classes)), classes)
class_colors <- list(Class = class_colors)

# Make color pallet for class
phylums <- unique(tax$Phylum)
phylums_colors <- setNames(rainbow(length(phylums)), phylums)
phylums_colors <- list(Phylum = phylums_colors)

# Make color pallet for Phylum for genus in plot
tax2 <- tax
rownames(tax2) <- paste(tax2$ASV, tax2$Genus, sep = "_")
phylums2 <- unique(tax2$Phylum)
phylums_colors2 <- setNames(rainbow(length(phylums2)), phylums2)
phylums_colors2 <- list(Phylum = phylums_colors2)

# Turn data frame into a matrix
All_lfc_df$ASV <- NULL
All_lfc_df <- as.matrix(All_lfc_df)
All_lfc_df2 <- as.matrix(All_lfc_df2)

# replace NAs by 0s
All_lfc_df_clu <- All_lfc_df
All_lfc_df2_clu <- All_lfc_df2
All_lfc_df_clu[is.na(All_lfc_df_clu)] <- 0
All_lfc_df2_clu[is.na(All_lfc_df2_clu)] <- 0

# breaks in heatmap
max_abs <- max(abs(All_lfc_df), na.rm = TRUE)
breaks <- unique(c(seq(-max_abs, -0.5, length.out = 10), 
                   seq(-0.5, 0.5, length.out = 30), # More breaks in this range for sensitivity
                   seq(0.5, max_abs, length.out = 10)))
```

### Heatmap
```{r, fig.height=3, fig.width=12}
pheatmap(All_lfc_df,
          clustering_distance_rows = dist(All_lfc_df_clu),
          clustering_distance_cols = dist(t(All_lfc_df_clu)),
         #cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax["Phylum"],
          annotation_colors = phylums_colors)
```
```{r, fig.height=3, fig.width=12}
pheatmap(All_lfc_df2,
          clustering_distance_rows = dist(All_lfc_df2_clu),
          clustering_distance_cols = dist(t(All_lfc_df2_clu)),
         #cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[5:6],
          annotation_colors = phylums_colors2)
```
```{r, fig.height=3, fig.width=12}
pheatmap(All_lfc_df2,
          clustering_distance_rows = dist(All_lfc_df2_clu),
          #clustering_distance_cols = dist(t(All_lfc_df2_clu)),
         cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          gaps_col = c(12,17,19,20),
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[5:6],
          annotation_colors = phylums_colors2)
```
```{r, fig.height=5, fig.width=12}
pheatmap(All_lfc_df2,
          clustering_distance_rows = dist(All_lfc_df2_clu),
          clustering_distance_cols = dist(t(All_lfc_df2_clu)),
         #cluster_cols = FALSE, 
         #cluster_rows = FALSE, 
          color = colorRampPalette(c("darkblue", "blue", "white", "red", "darkred"))(length(breaks) - 1),
          breaks = breaks,
          #main = "Differentially abundant taxa (log2FC by Family)",
          fontsize_row = 10,
          fontsize_col = 14,
          na_col = "grey90",
          angle_col = 90, 
          annotation_row = tax[3:6],
          annotation_colors = phylums_colors2)
```


<!-- ### Extract spearman estimates -->
<!-- ```{r} -->
<!-- # All unique DA ASVs -->
<!-- DA_ASVs <- ASVs_Corr_CPandFP$ASV -->
<!-- length(DA_ASVs) -->

<!-- # Reorganise lists -->
<!-- Corr_CatPerf <- CorrTestCatPerf_ASVs_Corr_CPandFP -->
<!-- Corr_GSL <- CorrTestGSLs_ASVs_Corr_CPandFP -->

<!-- # Add ASV as column -->
<!-- Corr_GSL <- lapply(Corr_GSL, function(x) { -->
<!--   x$ASV <- rownames(x) -->
<!--   x -->
<!-- }) -->

<!-- Corr_CatPerf$ASV <- rownames(Corr_CatPerf) -->

<!-- # Add GSL as column name -->
<!-- Corr_GSL <- Map(function(df, nm) { -->
<!--   colnames(df)[4] <- nm -->
<!--   df -->
<!-- }, df = Corr_GSL, nm = names(Corr_GSL)) -->

<!-- # Merge all Spearman estimetes per GSL -->
<!-- Spear_GSL_df <- Reduce(function(x, y) { -->
<!--   merge(x, y, by = "ASV", all = TRUE) -->
<!-- }, lapply(Corr_GSL, function(df) df[, c(4, 10)])) -->

<!-- # Add Cat performance -->
<!-- Sper_df <- merge(Spear_GSL_df, Corr_CatPerf[ , c(4, 10)], by = "ASV") -->
<!-- colnames(Sper_df)[11] <- "Cat_weight" -->
<!-- rownames(Sper_df) <- Sper_df$ASV -->
<!-- Sper_df -->
<!-- ``` -->

<!-- ### Prepare taxonomy -->
<!-- ```{r} -->
<!-- # Add genus to matrix -->
<!-- Sper_df2 <- merge(ASVs_Corr_CPandFP, Sper_df, by = "ASV") -->
<!-- rownames(Sper_df2) <- paste(Sper_df2$ASV, Sper_df2$Genus, sep = "_") -->
<!-- rownames(Sper_df2)[2] <- paste(Sper_df2$ASV[2], Sper_df2$Family[2], sep = "_") -->
<!-- Sper_df2 <- Sper_df2[6:15] -->
<!-- head(Sper_df2) -->

<!-- # Turn into a matrix -->
<!-- Sper_df$ASV <- NULL -->
<!-- Sper_mat <- as.matrix(Sper_df) -->
<!-- Sper_mat2 <- as.matrix(Sper_df2) -->
<!-- ``` -->

<!-- ### Heatmap  -->
<!-- ```{r, fig.height=4.5, fig.width=7} -->
<!-- max_abs <- max(abs(Sper_mat), na.rm = TRUE) -->

<!-- pheatmap(Sper_mat, -->
<!--           #clustering_distance_rows = dist(AllAcc_lfc_clu), -->
<!--           #clustering_distance_cols = dist(t(AllAcc_lfc_clu)), -->
<!--           color = colorRampPalette(c("blue", "white", "red"))(50), -->
<!--           breaks = seq(-max_abs, max_abs, length.out = 51), -->
<!--           #main = "Differentially abundant taxa (log2FC by Family)", -->
<!--           fontsize_row = 10, -->
<!--           fontsize_col = 14, -->
<!--           na_col = "grey60", -->
<!--           angle_col = 90,  -->
<!--           #annotation_row = tax2["Phylum"], -->
<!--           #annotation_colors = phylums_colors2 -->
<!--          ) -->
<!-- ``` -->
<!-- ```{r, fig.height=4.5, fig.width=8} -->
<!-- pheatmap(Sper_mat2, -->
<!--           #clustering_distance_rows = dist(AllAcc_lfc_clu), -->
<!--           #clustering_distance_cols = dist(t(AllAcc_lfc_clu)), -->
<!--           color = colorRampPalette(c("blue", "white", "red"))(50), -->
<!--           breaks = seq(-max_abs, max_abs, length.out = 51), -->
<!--           #main = "Differentially abundant taxa (log2FC by Family)", -->
<!--           fontsize_row = 10, -->
<!--           fontsize_col = 14, -->
<!--           na_col = "grey60", -->
<!--           angle_col = 90) -->
<!-- ``` -->


<!-- ## 4.4.2 Heatmap significant GSLs -->
<!-- ### Subset GSLs -->
<!-- ```{r} -->
<!-- Sper_df3 <- Sper_df[ , c("Cat_weight", "R_Neoglucobrassicin", "R_Sinigrin", "R_Gluconapin", "R_Glucoerucin", "R_Glucoraphanin")] -->
<!-- Sper_df3 -->

<!-- Sper_df4 <- Sper_df2[ , c("Cat_weight", "R_Neoglucobrassicin", "R_Sinigrin", "R_Gluconapin", "R_Glucoerucin", "R_Glucoraphanin")] -->
<!-- Sper_df4 -->

<!-- Sper_mat3 <- as.matrix(Sper_df3) -->
<!-- Sper_mat4 <- as.matrix(Sper_df4) -->
<!-- ``` -->

<!-- ### Heatmap  -->
<!-- ```{r, fig.height=4, fig.width=5.2} -->
<!-- max_abs <- max(abs(Sper_mat3), na.rm = TRUE) -->

<!-- pheatmap(Sper_mat3, -->
<!--           color = colorRampPalette(c("blue", "white", "red"))(50), -->
<!--           breaks = seq(-max_abs, max_abs, length.out = 51), -->
<!--           #main = "Differentially abundant taxa (log2FC by Family)", -->
<!--           fontsize_row = 10, -->
<!--           fontsize_col = 14, -->
<!--           na_col = "grey60", -->
<!--           angle_col = 90) -->
<!-- ``` -->
<!-- ```{r, fig.height=4, fig.width=6.2} -->
<!-- pheatmap(Sper_mat4, -->
<!--           color = colorRampPalette(c("blue", "white", "red"))(50), -->
<!--           breaks = seq(-max_abs, max_abs, length.out = 51), -->
<!--           #main = "Differentially abundant taxa (log2FC by Family)", -->
<!--           fontsize_row = 10, -->
<!--           fontsize_col = 14, -->
<!--           na_col = "grey60", -->
<!--           angle_col = 90) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- # Clean environment -->
<!-- rm(list = ls()) -->
<!-- ``` -->


<!-- # 4.5 Heatmap correlating Genuses -->
<!-- ## 4.5.1 Load data -->
<!-- ```{r} -->
<!-- # load phyloseq object -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_unnormalized_bac_ps_ForDA.RData") -->

<!-- # Correlation test GSLs CP -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_OverlapCorrelations/CorrTestGSLs_Genus_Corr_CPandFP.RData") -->

<!-- # Correlation test Cat performance FP -->
<!-- load(file = "C:/Users/kreek001/OneDrive - Wageningen University & Research/Chapters/Experimental Chapter 3/RScripts/GitHub/TwelveAccessionExperiment/MicrobiomeAnalysis/Data/Phyloseq_objects/CP_FP_OverlapCorrelations/CorrTestCatPerf_Genus_Corr_CPandFP.RData") -->
<!-- ``` -->

<!-- ## 4.5.3 Heatmap all GSLs -->
<!-- ### Extract spearman estimates -->
<!-- ```{r} -->
<!-- # All unique DA ASVs -->
<!-- ASVs_l <- lapply(CorrTestCatPerf_Genus_Corr_CPandFP, function(x) x$ASV) -->
<!-- ASVS <- sort(unique(unlist(ASVs_l))) -->
<!-- length(ASVS) -->

<!-- # Reorganise lists -->
<!-- #Corr_CatPerf <- do.call(rbind.data.frame, CorrTestCatPerf_ASVs_Corr_CPandFP) -->
<!-- #Corr_GSL <- CorrTestGSLs_ASVs_Corr_CPandFP -->

<!-- # Transposes GSL list -->
<!-- genus_names <- names(CorrTestGSLs_Genus_Corr_CPandFP[[1]]) -->
<!-- met_names   <- names(CorrTestGSLs_Genus_Corr_CPandFP) -->

<!-- GSL_l_t <- setNames( -->
<!--   lapply(genus_names, function(gen) { -->
<!--     setNames( -->
<!--       lapply(CorrTestGSLs_Genus_Corr_CPandFP, function(met) met[[gen]]), -->
<!--       met_names -->
<!--     ) -->
<!--   }), -->
<!--   genus_names -->
<!-- ) -->


<!-- # Make a df for heatmap -->
<!-- merged_genus_df <- lapply(names(GSL_l_t), function(gen) { -->

<!--   ## 1. Merge the 9 metabolite data frames -->
<!--   met_merged <- Reduce( -->
<!--     function(x, y) merge(x, y, by = "ASV", all = TRUE), -->
<!--     lapply(names(GSL_l_t[[gen]]), function(met) { -->
<!--       df <- GSL_l_t[[gen]][[met]][, c(10, 4)] -->
<!--       colnames(df) <- c("ASV", met) -->
<!--       df -->
<!--     }) -->
<!--   ) -->

<!--   ## 2. Merge with Caterpillar performance -->
<!--   gen_df <- CorrTestCatPerf_Genus_Corr_CPandFP[[gen]][, c(10, 4)] -->
<!--   colnames(gen_df)[colnames(gen_df) != "ASV"] <- "Cat_weight" -->

<!--   ## 3. Final merge -->
<!--   out <- merge(met_merged, gen_df, by = "ASV", all = TRUE) -->

<!--   ## 4. Add genus column -->
<!--   out$Genus <- gen -->

<!--   out -->
<!-- }) -->

<!-- names(merged_genus_df) <- names(GSL_l_t) -->

<!-- # Combine into one df -->
<!-- merged_df <- do.call(rbind.data.frame, merged_genus_df) -->

<!-- # Add genus to matrix -->
<!-- rownames(merged_df) <- paste(merged_df$ASV, merged_df$Genus, sep = "_") -->
<!-- merged_df <- merged_df[2:11] -->
<!-- head(merged_df) -->

<!-- # Turn into a matrix -->
<!-- merged_mat <- as.matrix(merged_df) -->

<!-- # remove NAs only for clustering -->
<!-- merged_mat_clu <- merged_mat -->
<!-- merged_mat_clu[is.na(merged_mat_clu)] <- 0 -->
<!-- ``` -->

<!-- ### Statistical info -->
<!-- ```{r} -->
<!-- # Make a df for heatmap -->
<!-- merged_genus_df_stat <- lapply(names(GSL_l_t), function(gen) { -->

<!--   ## 1. Merge the 9 metabolite data frames -->
<!--   met_merged <- Reduce( -->
<!--     function(x, y) merge(x, y, by = "ASV", all = TRUE), -->
<!--     lapply(names(GSL_l_t[[gen]]), function(met) { -->
<!--       df <- GSL_l_t[[gen]][[met]][, c(10, 6)] -->
<!--       colnames(df) <- c("ASV", met) -->
<!--       df -->
<!--     }) -->
<!--   ) -->

<!--   ## 2. Merge with Caterpillar performance -->
<!--   gen_df <- CorrTestCatPerf_Genus_Corr_CPandFP[[gen]][, c(10, 6)] -->
<!--   colnames(gen_df)[colnames(gen_df) != "ASV"] <- "Cat_weight" -->

<!--   ## 3. Final merge -->
<!--   out <- merge(met_merged, gen_df, by = "ASV", all = TRUE) -->

<!--   ## 4. Add genus column -->
<!--   out$Genus <- gen -->

<!--   out -->
<!-- }) -->

<!-- names(merged_genus_df_stat) <- names(GSL_l_t) -->

<!-- # Combine into one df -->
<!-- merged_df_stat <- do.call(rbind.data.frame, merged_genus_df_stat) -->

<!-- # Add genus to matrix -->
<!-- rownames(merged_df_stat) <- paste(merged_df_stat$ASV, merged_df_stat$Genus, sep = "_") -->
<!-- merged_df_stat <- merged_df_stat[2:11] -->
<!-- head(merged_df_stat) -->

<!-- # Turn into a matrix -->
<!-- merged_mat_stat <- as.matrix(merged_df_stat) -->
<!-- ``` -->

<!-- ### Heatmap  -->
<!-- ```{r, fig.height=15, fig.width=9} -->
<!-- max_abs <- max(abs(merged_mat), na.rm = TRUE) -->

<!-- pheatmap(merged_mat, -->
<!--           #clustering_distance_rows = dist(merged_mat_clu), -->
<!--          cluster_rows = FALSE,  -->
<!--          clustering_distance_cols = dist(t(merged_mat_clu)), -->
<!--           color = colorRampPalette(c("blue", "white", "red"))(50), -->
<!--           breaks = seq(-max_abs, max_abs, length.out = 51), -->
<!--           #main = "Differentially abundant taxa (log2FC by Family)", -->
<!--           fontsize_row = 10, -->
<!--           fontsize_col = 14, -->
<!--           na_col = "grey60", -->
<!--           angle_col = 90) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- library(ComplexHeatmap) -->
<!-- library(grid) -->

<!-- max_abs <- max(abs(merged_mat), na.rm = TRUE) -->

<!-- Heatmap(merged_mat, -->
<!--   col = colorRampPalette(c("blue", "white", "red"))(50), -->
<!--   cluster_rows = FALSE, -->
<!--   cluster_columns = TRUE, -->
<!--   name = "value", -->

<!--   cell_fun = function(j, i, x, y, w, h, fill) { -->
<!--     if (!is.na(dot_mat[i, j]) && dot_mat[i, j] == 1) { -->
<!--       grid.points(x, y, pch = 16, size = unit(2.5, "mm")) -->
<!--     } -->
<!--   } -->
<!-- ) -->

<!-- ``` -->

<!-- <!-- ## 4.4.2 Heatmap significant GSLs --> -->
<!-- <!-- ### Subset GSLs --> -->
<!-- <!-- ```{r} --> -->
<!-- <!-- Sper_df3 <- Sper_df[ , c("Cat_weight", "R_Neoglucobrassicin", "R_Sinigrin", "R_Gluconapin", "R_Glucoerucin", "R_Glucoraphanin")] --> -->
<!-- <!-- Sper_df3 --> -->

<!-- <!-- Sper_df4 <- Sper_df2[ , c("Cat_weight", "R_Neoglucobrassicin", "R_Sinigrin", "R_Gluconapin", "R_Glucoerucin", "R_Glucoraphanin")] --> -->
<!-- <!-- Sper_df4 --> -->

<!-- <!-- Sper_mat3 <- as.matrix(Sper_df3) --> -->
<!-- <!-- Sper_mat4 <- as.matrix(Sper_df4) --> -->
<!-- <!-- ``` --> -->

<!-- <!-- ### Heatmap  --> -->
<!-- <!-- ```{r, fig.height=4, fig.width=5.2} --> -->
<!-- <!-- max_abs <- max(abs(Sper_mat3), na.rm = TRUE) --> -->

<!-- <!-- pheatmap(Sper_mat3, --> -->
<!-- <!--           color = colorRampPalette(c("blue", "white", "red"))(50), --> -->
<!-- <!--           breaks = seq(-max_abs, max_abs, length.out = 51), --> -->
<!-- <!--           #main = "Differentially abundant taxa (log2FC by Family)", --> -->
<!-- <!--           fontsize_row = 10, --> -->
<!-- <!--           fontsize_col = 14, --> -->
<!-- <!--           na_col = "grey60", --> -->
<!-- <!--           angle_col = 90) --> -->
<!-- <!-- ``` --> -->
<!-- <!-- ```{r, fig.height=4, fig.width=6.2} --> -->
<!-- <!-- pheatmap(Sper_mat4, --> -->
<!-- <!--           color = colorRampPalette(c("blue", "white", "red"))(50), --> -->
<!-- <!--           breaks = seq(-max_abs, max_abs, length.out = 51), --> -->
<!-- <!--           #main = "Differentially abundant taxa (log2FC by Family)", --> -->
<!-- <!--           fontsize_row = 10, --> -->
<!-- <!--           fontsize_col = 14, --> -->
<!-- <!--           na_col = "grey60", --> -->
<!-- <!--           angle_col = 90) --> -->
<!-- <!-- ``` --> -->
